---
title: "Jannis-DW"
output: md_document
always_allow_html: true
---

# Jannis Solution for Minseo's Project



## Preperations


```{r loading}
# use the tidyverse package for data manipulation and visualization
# use piping where possible to make the code more readable and efficient

library(kableExtra)
library(tidyverse)
library(ggplot2)
library(knitr)
library(stringr)
library(dplyr)
library(tidyr)

setwd("/Users/jannisdennochweiler/R2 Course/Projects/Minseo")

data <- read_csv("Social Media Advertising.csv")

head(data)
```

## Data cleanup

#### Cleanup-Tasks:

* convert Duration values into numeric day values
* clean currency-formatted variables such as Acquisition_Cost
* convert Date into a proper date format
* split Target_Audience into separate demographic variables, such as gender and age group
* create a cleaned dataset for later manipulation and visualization

```{r cleanup}
cleandata <- data %>%
mutate(Duration = as.numeric(str_remove(Duration, " Days"))) %>%
mutate(Acquisition_Cost = as.numeric(str_sub(Acquisition_Cost, 2))) %>%
mutate(Date = as.Date(Date, format ="%m/%d/%Y")) %>%
separate(Target_Audience, into = c("gender", "age_group"), sep = " ")

# create rmd table with output of the cleaned data

kable(head(cleandata), caption = "Cleaned Data", booktabs = TRUE) %>%
  kable_styling(latex_options = "striped", position = "center")

```

## Manipulaiton

#### Question 1: Which campaign goals bring the highest engagement scores?

* grouping data by campaign goal
* create table showing average `Engagement_Score` and standard deviation

```{r manipulation}

summary <- cleandata %>%
group_by(Campaign_Goal) %>%
summarise(
avg_engagement = mean(Engagement_Score, na.rm = TRUE),
  sd_engagement = sd(Engagement_Score, na.rm = TRUE),
  number_campaigns = n()
          )
summary
```

#### Answer 1:
All kinds of campaign goals seem to perform similarly in terms of engagement scores, with the exception of "Brand Awareness" campaigns, which have a slightly lower average engagement score. However, the differences are not substantial enough to draw definitive conclusions.


#### Question 2: Which group has the highest click-through rate?

* Calculation of Click-Through rate (Clicks/Impressions)
* decomposition of age ranges into individual ranges

```{r manipulationtwo}
# this first part here is mainly AI since I had no idea how to tackle this

data_expanded <- cleandata %>%
  mutate(

    age_group = if_else(age_group == "All Ages", "18-60",age_group)) %>%
  mutate(age_group = str_replace_all(age_group, "–", "-"),
         age_group = str_replace_all(age_group, " ", ""),)%>%
  separate(age_group, into = c("min_age", "max_age"), sep = "-") %>%
  drop_na(c(min_age, max_age)) %>%
  rowwise() %>%
  mutate(age = list(seq(min_age, max_age))) %>%
  unnest(age)


data_expanded2 <- data_expanded%>%
mutate(click_through_rate = Clicks / Impressions)


agedata <- data_expanded2 %>%
group_by(age) %>%
summarise(
avg_click_through_rate = mean(click_through_rate)
)


slice_max(agedata, avg_click_through_rate, n = 1)
?slice_max

```

#### Answer 2: There are 7 ages that share the highest click through rate of 0,314. This is because all the indivudual ages' use the campaigns targetet at the age group 18-24 as well as those that are targetet towards all ages.

## Visualisation

#### Task: Create a line graph showing how average ROI changes with campaign duration across different social media platforms.

* grouping data by `Duration` and `Channel_Used`
* campaign duration on x-axis, average ROI on the y-axis

```{r Visualisation}
# Visualisation
#Create a line graph showing how average ROI changes with campaign duration across different social media platforms.

#The data should be grouped by Duration and Channel_Used.

#The final graph should show:
#campaign duration on the x-axis
#average ROI on the y-axis
#separate lines for different social media platforms

cleandata %>%
group_by(Duration, Channel_Used) %>%
ggplot(aes(x = Duration, y = ROI, color = Channel_Used)) +
  geom_line(stat = "summary", fun = "mean") +
  labs(title = "Average ROI by Campaign Duration and Social Media Platform",
       x = "Campaign Duration (Days)",
       y = "Average ROI") +
  theme_minimal() +
  theme(legend.title = element_blank())
```




#### Interpretation: 

All kinds of Channels have a nearly identical RIO except for Pinterest which is much lower. All ROIs stay constant over the whole campaign duration. Therefore, Pinterest should not be used in a campaign.



