---
title: "lea-immenkamp"
author: "Lea Immenkamp"
date: "2026-06-09"
output: md_document
---
# Social Media Advertising
Social media advertising is an important part of modern marketing and public relations. This project explores how social media advertising campaigns differ across platforms, customer segments, and campaign goals.

The focus of this project lies on cleaning, restructuring, and visualizing advertising campaign data. I am especially interested in how different advertising strategies lead to audience interaction and financial performance.

```{r library loading}
library(tidyverse)
library(dplyr)
library(tidyr)
library(stringr)
library(lubridate)
library(knitr)
library(gt)
library(purrr)
library(ggplot2)
```


```{r data loading}
socialmedia <- read.csv("Social Media Advertising.csv")
```


```{r data cleaning}
socialmedia_clean <- socialmedia %>% 
   mutate(
    Duration = as.numeric(str_extract(Duration, "\\d+"))
  ) %>% 
  
  mutate(
    Acquisition_Cost = as.numeric(
      str_replace_all(Acquisition_Cost, "[$,]", "")
    )
  ) %>%
  
  mutate(
    Date = mdy(Date)
  ) %>%
  
  mutate(
    Target_Audience = str_replace(Target_Audience, "All Ages", "All gender")
  ) %>% 
  
  separate(
    Target_Audience,
    into = c("Gender", "Age_Group"),
    sep = " (?=\\d)",
    remove = TRUE
  ) %>%
  
  mutate(
    Gender = str_trim(Gender),
    Age_Group = str_trim(Age_Group)
  ) %>% 
  
   rename_with(tolower) %>% 
  filter(!is.na(age_group))
  

# Check variable type
glimpse(socialmedia_clean)
```

# Data mainpulation

## 1. Which campaign goals bring the highest engagement scores?
```{r data engagement score}
campaign <- socialmedia_clean %>% 
  group_by(campaign_goal) %>% 
 summarise(
    M = mean(engagement_score, na.rm = TRUE),
    SD = sd(engagement_score, na.rm = TRUE),
    n = n()
  ) %>%
  arrange(desc(M)) %>% 
  rename(
    `Campaign goal` = campaign_goal
  )


campaign %>%
  gt() %>%
  tab_header(
    title = "Engagement Score Summary by Campaign Goal"
  ) %>%
  tab_footnote(
    footnote = "Note. M = Mean, SD = Standard Derivation, n = number of campaigns."
  )
  
```

## 2. Which age group has the highest click-through rate?
```{r click-through rate}
age_summary <- socialmedia_clean %>%
  mutate(click_through_rate = clicks / impressions) %>%
  separate(age_group, into = c("age_min", "age_max"), sep = "-", convert = TRUE) %>%
  mutate(
    age = map2(age_min, age_max, ~ if (is.finite(.x) & is.finite(.y)) seq(.x, .y) else NA_real_)
  ) %>%
  unnest(age) %>%
  group_by(age) %>%
  summarise(
    mean_ctr = mean(click_through_rate, na.rm = TRUE),
    number_of_campaigns = n(),
    .groups = "drop"
  ) %>%
  arrange(age)
```

Visualization of click-through rate
```{r  click-through rate visualization}
socialmedia_clean %>%
  mutate(
    click_through_rate = clicks / impressions,
    gender = tolower(gender),
    gender = ifelse(gender == "male","female", gender)
  ) %>%
  separate(age_group, c("age_min","age_max"), sep = "\\s*-\\s*", convert = TRUE, fill = "right") %>%
  separate_rows(gender, sep = ",") %>%
  mutate(
    age = map2(age_min, age_max, ~ {
      if (is.finite(.x) & is.finite(.y) & .x <= .y) seq(.x, .y) else NA_real_
    })
  ) %>%
  unnest(age) %>%
  group_by(age, gender) %>%
  summarise(ctr = mean(click_through_rate, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(age, ctr)) +
  stat_summary(fun = mean, geom = "line") +
  facet_wrap(~gender)

```

# Data visualization
```{r line graph}
roi_data <- socialmedia_clean %>%
  group_by(duration, channel_used) %>%
  summarise(
    avg_roi = mean(roi, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(roi_data, aes(x = duration, y = avg_roi, color = channel_used)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Average ROI by Campaign Duration and Platform",
    x = "Campaign Duration",
    y = "Average ROI",
    color = "Platform"
  ) +
  theme_minimal()

```



