---
title: "Solution for Michael's project"
author: "Minseo Kim"
output: md_document
---

```{r packages}

library(tidyverse)
library(dplyr)
library(knitr)
library(ggplot2)
library(stringr)
library(tidyr)

```

```{r data}
freda<-read_csv("FReDA_panel_4waves_long_labeled.csv")
```

```{r mutate negative values}
freda_2 <- freda%>%mutate(across(c(sat3,frt68,frt69,age,nkids,reldur),~replace(.x, .x<0, NA)))
```

### 1.1. What is the correlation between satisfaction with relationship with the general intention to get children and the correlation between satisfaction with relationship with the more specific intention to get children (in the next 3 years).

Answer: The correlations are both positive but also quite small. People with higher satisfaction with relationship has slightly stronger general intention to get children, and also people with higher satisfaction with relationship has slightly stronger specific intention to get children within the next 3 years. However, the correlations are weak. 

```{r correlation_sat3&frt68, sat3&frt69}
cor(freda_2$sat3, freda_2$frt68,use="complete.obs")
#0.1555075
cor(freda_2$sat3, freda_2$frt69,use="complete.obs")
#0.1458812
```

### 1.2. Plot a heat map! On the x-axis: sat3. On the y-axis frt69. I want to see how much percent of people are in each box which displays the relative frequency of people with a specific combination of relationship satisfaction and intention to get children in the next 3 years.

```{r heatmap}
heatmap_data<- freda_2%>%drop_na(sat3,frt69)%>%group_by(sat3,frt69)%>%summarise(count=n())%>%mutate(percentage=count/sum(count))
ggplot(heatmap_data, aes(x=sat3,y=frt69,fill=percentage))+geom_tile()+scale_fill_gradient(low="white",high="red")+labs(x="Satisfaction with Relationship(sat3)", y="Intention to Get Children in the Next 3 Years (frt69)", fill="Percentage")+theme_minimal()
```

### 1.3. Why do you think that people who are satisfied with their relationship the most are seemingly less likely to get children in the near future?

Answer: They are already happy with their relationship and do not feel the need to have children to make it more fulfilling. Instead, they would rather focus on their relationship with each other.

### 1.4. What statistical problems do we run into? (Hint: Look at the distribution in the variable sat3. How many percent are unsatisfied or very unsatisfied?)

Answer: Only 3.97% of the respondents are unsatisfied or very unsatisfied with their relationship, while about 50.67% answered that they are very satisfied with their relationship.
Therefore, the distribution of sat3 is strongly focused toward high satisfaction of the relationship. This can cause a statistical problem because the correlations for low satisfaction group may be unstable and less reliable. 

```{r statistial problem}
#Unsatisfied or very unsatisfied: sat3<=4
mean(freda_2$sat3<=4, na.rm=TRUE)*100
#Very satisfied: sat3>=9
mean(freda_2$sat3>=9, na.rm=TRUE)*100
```

### 1.5. See if age is affecting the correlation between relationship satisfaction and fertility intentions (use it as a control variable)

Answer: After controlling for age, the correlation between relationship satisfaction and general fertility intention went down to 0.101 from 0.156. Also for the correlation between relationship satisfaction and fertility intention in the next 3 years went down to 0.110 from 0.146. This means that age variable does effect the correlation between relationship satisfaction and fertility intentions. However, correlation still remain positive after controlling for age as well. 

```{r age as control variable}

library(ppcor)

data_frt68<-freda_2%>%drop_na(sat3, frt68, age)
pcor.test(data_frt68$sat3, data_frt68$frt68, data_frt68$age)

data_frt69<-freda_2%>%drop_na(sat3, frt69, age)
pcor.test(data_frt69$sat3, data_frt69$frt69, data_frt69$age)

```

### 1.6. Now we can add number of children (nkids) and relationship duration (reldur) as control variables. What happens to the correlation between relationship satisfaction and fertility intentions when we introduce these variables into our model?

Answer: After controlling for the number of kids and the relationship duration, the correlation between relationship satisfaction and general fertility intention went down to 0.105 from 0.156. Also for the correlation between relationship satisfaction and fertility intention in the next 3 years went down to 0.112 from 0.146. This means that number of kids and the relationship duration do effect the correlation between relationship satisfaction and fertility intentions. However, correlation still remain positive after the controlling as well. 

```{r nkids and reldur as control variables}

data_68<-freda_2%>%drop_na(sat3, frt68, nkids, reldur)
pcor.test(data_68$sat3, data_68$frt68, data_68[,c("nkids","reldur")])
data_69<-freda_2%>%drop_na(sat3, frt69, nkids, reldur)
pcor.test(data_69$sat3, data_69$frt69, data_69[,c("nkids","reldur")])
```

### 1.7. significance: Are the correlations between relationship satisfaction and fertility intentions significant? (Hint: Look at the p-value of the correlation coefficient). What is the 95% confidence interval of the correlation coefficient? What does it mean?

Answer:

[frt68]
p-value:0.000000000000304 p<.001
95 percent confidence interval: 0.1142237, 0.1962550

p-value is <.001, meaning the correlations are statistically significant. However, there is a 95% probability that the true correlation lies between 0.114 and 0.196. This means the correlation is valid, but the assosiation is relatively weak. 

[frt69]
p-value:0.00000000000000022 p<.001
95 percent confidence interval: 0.1201297, 0.1714366

p-value is <.001, meaning the correlations are statistically significant. However, there is a 95% probability that the true correlation lies between 0.120 and 0.171. This means the correlation is valid, but the assosiation is relatively weak. 

```{r p value and confidence interval}

cor.test(freda_2$sat3, freda_2$frt68,use = "complete.obs")
cor.test(freda_2$sat3, freda_2$frt69,use="complete.obs")
```

### 2.1. Which region has more conservative viewpoints regarding the role a mother should play in the family? Which region has more egalitarian viewpoints?

Answer:The average score for the statement that "Womesn should take care of the family more than their career." was approximately 2.33 in West Germany and 2.14 in East Germany. For the statement that "A child under the age of 6 is suffering, if the mother works.", the average score was approximately 2.32 in West Germany and 1.85 in East Germany. As higher values means stronger agreement with conservative viewpoints, West Germany has more conservative viewpoints while East Germany appears to be more egalitatrian. 

```{r NA elimination}
freda_3<-freda%>%mutate(across(c(val1i3, val1i5, east),~replace(.x, .x<0, NA)))
```

```{r region names}
freda_3<-freda_3%>%mutate(east=case_when(east==1~"East Germany", east==0 ~ "West Germany"))
```

```{r region comparison}
freda_3%>%group_by(east)%>%summarise(mean_val1i3=mean(val1i3, na.rm=TRUE),mean_val1i5=mean(val1i5, na.rm=TRUE))
```


### 2.2. Do a violine plot for visualization! It should visualize the regional differences between East and West Germany according to these different values on the role of woman in families. Think for yourself what kind of graph would suit best to visualize this. Also both plots should be visualized in one graph via faceting. So you can easily compare the regional differences between the two different viewpoints on the role of a mother in a family.

```{r plot data}
plot_data<-freda_3%>%drop_na(val1i3, val1i5, east)%>%pivot_longer(cols=c(val1i3, val1i5), names_to="statement", values_to="response")
```

```{r violin plot}
ggplot(plot_data, aes(x=east, y=response, fill=east))+geom_violin(trim=FALSE) +facet_wrap(~statement)+labs(x="Region", y="Response", fill="Region")+theme_minimal()
```
