---
title: "Untitled"
author: "Celine Onyemaechi"
output: md_document
---

change that markdown output for unique names is not displayed

level 1:
spider web chart
items = regions
value = fraction of whiskies (per region) rated 80 or higher
damit sind die werte prozentual

level 2:
mehrere linien im spider web mit unterschiedlichen thresholds:
rated 50 or higher
rated 70 or higher
rated 80 or higher
damit überlappen die linien nicht, aber geben den die steigerung ganz gut wieder



```{r setup, echo = FALSE}

# clean environment
rm(list = ls())
cat("\014")

# load packages 
library(tidyverse)
#install.packages("fmsb")
library(fmsb)

# read file 
setwd("~/Data-projects-with-R-and-GitHub/Projects/felix-gogollok/whisky-statistics")
data <- read.csv2("Whisky.csv") #use csv2() cause data is semicolon separated  
```


```{r clean data set, echo = FALSE}
# look at data
colnames(data)
#View(data)

#remove trailing
data <- data %>%
  rename_with(~ gsub("\\.$", "", .))

# remove unnecessary columns 
# to create the spider chart 'Destillerie', 'Region' and 'Punkte' are relevant
# remaining columns are not needed for visualization 
data <- data  %>% select(Destillerie, Region, Punkte)

# structure of the data
str(data)
#' 1307 objects (rows) and 3 variables (columns)
#' 'Desrillerie' an 'Region' are of type chraracter (strings) 
#' and 'Punkte is of type integer


```


Prepare data for spider plot
```{r data preparation, echo = FALSE}
# check summary stats 
summary(data)
#' the column 'Punkte' (ratings) is the only column with NA's 
#' summary stats for ratings (Punkte)
#' min = 55.00 , max = 96.00 -> range of 44
#' median = 79.00
#' mean = 79.00
#' total of 20 NAs


# remove NA's from ratings column
data <- data %>% drop_na(Punkte)
summary(data$Punkte) # check if NA's are removed

# look at relevant columns 

# check how many distinct regions 
length(unique(data$Region)) # there are 5 distinct regions 
unique(data$Region) # get names of the regions 

# check number of distinct distilleries
length(unique(data$Destillerie))  # 121 distilleries
unique(data$Destillerie)

```


```{r prepare data level 1 plotting}

# compute percentage of ratings >= 80

level01_data <- data %>%
  group_by(Region) %>%
  summarize(
    percantage_high_ratings = (sum(Punkte >= 80) / n()) * 100
  )

# format data to make it compatible with fmsb
radar_level01 <- as.data.frame(t(level01_data$percantage_high_ratings)) # t() to transpose data 
colnames(radar_level01) <- level01_data$Region

# add max and min to df
radar_level01 <- rbind(rep(100, ncol(radar_level01)), 
                       rep(0,   ncol(radar_level01)),
                       radar_level01)
  
```


```{r create spider chart for ratings >= 80}

radarchart(
  radar_level01,
  axistype = 1,
  caxislabels = c("0%", "20%", "40%", "60%", "80%", "100%"),
  #caxislabels = seq(0, 100, 20),  # adjust axis labels for because of percentages
  title = "Percentage of Whiskies Rated 80 or Higher by Region",
  pcol = "#EE4000",
  plwd = 2,
  cglcol = "darkgrey",
  cglwd = 1,
  cglty = 1,
  axislabcol = "black",
  maxmin = T,
  seg = 5,
  vlcex = 0.8,
  xlim = c(-1, 1),  
  ylim = c(-1, 1.2)
)

legend("topright", 
       legend = "Whiskeys rated \n 80 or higher", 
       col = "#EE4000", 
       lwd = 2, 
       bty = "n",
       cex = 0.8)

```


```{r prepare data level 2 plotting with multiple thresholds}

# compute percentage for thresholds 60, 70 and 90

level02_data <- data %>%
  group_by(Region) %>%
  summarize(
    rated_70_more = (sum(Punkte >= 70) / n()) * 100,
    rated_80_more = (sum(Punkte >= 80) / n()) * 100,
    rated_90_more = (sum(Punkte >= 90) / n()) * 100,
  )

# format data to make it compatible with fmsb
radar_level02 <- as.data.frame(t(select(level02_data, -Region))) 
colnames(radar_level02) <- level02_data$Region

# add max and min to df
radar_level02 <- rbind(rep(100, ncol(radar_level02)), 
                       rep(0,   ncol(radar_level02)),
                       radar_level02)
  
```


```{r create spider chart multiple thresholds}

# define colours for each of the three thresholds
colours <- c("#FFA500", "#4169E1", "#EE4000")
#colours to fill
fill_colours <- c(
  rgb(255, 165, 0, max = 255, alpha = 80),  # Light orange
  rgb(65, 105, 225, max = 255, alpha = 80), # Light blue
  rgb(238, 64, 0, max = 255, alpha = 80)    # Light red
)

radarchart(
  radar_level02,
  axistype = 1,
  #caxislabels = c("0%", "20%", "40%", "60%", "80%", "100%"),
  caxislabels = seq(0, 100, 20),  # adjust axis labels for because of percentages
  title = "Percentage of Whiskies Rated 80 or Higher by Region",
  pcol = colours,
  pfcol = fill_colours,
  plwd = 2,
  cglcol = "darkgrey",
  cglwd = 1,
  cglty = 1,
  axislabcol = "black",
  maxmin = T,
  seg = 5,
  vlcex = 0.8,
  xlim = c(-1, 1),  
  ylim = c(-1, 1.2)
)

legend("topright", 
       legend = c("Rated 70 or higher", "Rated 80 or higher", "Rated 90 or higher"), 
       col = colours, 
       lwd = 2, 
       bty = "n",
       cex = 0.8)

```

