---
title: "Niclas38"
output: md_document
date: "2025-03-31"
---

# Loading packages and R-setup
```{r setup}
library(tidyverse)
library(readxl)
library(rstudioapi)
library(lubridate)
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```

# Loading data
```{r data loading} 
Data <- read_xlsx("stress_ratings_anonymized.xlsx")

```

# Data manipulation

## Data cleaning:
```{r data cleaning}
# Remove column 17 (is empty) and incomeplete entries (= remove all rows/entries that contain NAs besides for the last 3 variables -> dOC variables for the lv. 2 task)
CleanData <- Data %>%  
  select(-17) %>% 
  filter(if_all(-c(id_dOC, dOC_cycle_date, dOC_cycle_day), ~ !is.na(.)))
```

## Data aggregation:
```{r data aggregation}
# Give out a tibble containing the mean and standard deviation for each mood and stress item for each participant
CleanData %>% 
  group_by(id) %>% 
  summarise(mean_stressed_24h = mean(stressed_24h),
            SD_stressed_24h = sd(stressed_24h),
            mean_stressed_moment = mean(stressed_moment),
            SD_stressed_moment = sd(stressed_moment),
            mean_depressed = mean(depressed),
            SD_depressed = sd(depressed),
            mean_satisfied = mean(satisfied),
            SD_satisfied = sd(satisfied),
            mean_motivated = mean(motivated),
            SD_motivated = sd(motivated),
            mean_aroused = mean(aroused),
            SD_aroused = sd(aroused))

# Give out a tibble containing the mean and standard deviation for each mood and stress item for each group
CleanData %>% 
  group_by(group) %>% 
  summarise(mean_stressed_24h = mean(stressed_24h),
            SD_stressed_24h = sd(stressed_24h),
            mean_stressed_moment = mean(stressed_moment),
            SD_stressed_moment = sd(stressed_moment),
            mean_depressed = mean(depressed),
            SD_depressed = sd(depressed),
            mean_satisfied = mean(satisfied),
            SD_satisfied = sd(satisfied),
            mean_motivated = mean(motivated),
            SD_motivated = sd(motivated),
            mean_aroused = mean(aroused),
            SD_aroused = sd(aroused))
```

## Data transformation:
```{r data transformation}
# Scatterplots for association between self reportet stress and mood
# Negative mood: x = stressed_24h, y = depressed, confidence interval = 99%
ggplot(CleanData, aes(x= stressed_24h, y= depressed))+
  geom_point(col = "#7699D4", size=0.8)+
  geom_smooth(method= "lm",col = "#480355", se = TRUE, level= 0.99, fill = "#9448BC")+
  labs(
    title= "How stress influences negative mood",
    subtitle= "(Self-reported perception, scores from 0 to 100)",
    x = "How stressed did you feel today? (stressed_24h)",
    y = "How depressed did you feel today? (depressed)"
  )+
  theme(panel.background = element_blank())

#Positive mood: x = stressed_24h, y = satisfied, confidence interval = 99%
ggplot(CleanData, aes(x= stressed_24h, y= satisfied))+
  geom_point(col = "#7699D4", size=0.8)+
  geom_smooth(method= "lm",col = "#480355", se = TRUE, level= 0.99, fill = "#9448BC")+
  labs(
    title= "How stress influences positive mood",
    subtitle= "(Self-reported perception, scores from 0 to 100)",
    x = "How stressed did you feel today? (stressed_24h)",
    y = "How satisfied did you feel today? (satisfied)"
  )+
  theme(panel.background = element_blank())
```


# Visualization
```{r visualization}
# Convert timestamps to weekdays with lubridate (I use timeStampStop, since this seems to make the most sense to me considering your definition of the different time stamps)
WeekdayData <- CleanData %>% 
  mutate(
    date = as_datetime(timeStampStop),
    days_since_start = as.integer(difftime(date, min(date), units = "days")),
    weekday = factor((days_since_start%%7)+1,
                     levels = 1:7,
                     labels= c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"))
  )

# Create 4 plots with ggplot, one for each mood. In each plot there are 3 line graphs, one for each medication group(as shown on the legend)
# Plot for depressed: (confidence interval =75%)
ggplot(WeekdayData,aes(x=weekday, y=depressed, col = group, group = group))+
  geom_smooth(method = "loess", se = TRUE, level =0.75)+
  labs(
    title="Average level of depression per Weekday in dependency of the medication group",
    x="Weekday",
    y="Average depression (self-reported on a scale from 1 to 100)")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for satisfied: (confidence interval =75%)
ggplot(WeekdayData,aes(x=weekday, y=satisfied, col = group, group = group))+
  geom_smooth(method = "loess", se = TRUE, level =0.75)+
  labs(
    title="Average level of satisfaction per Weekday in dependency of the medication group",
    x="Weekday",
    y="Average satisfaction (self-reported on a scale from 1 to 100)")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for motivated: (confidence interval =75%)
ggplot(WeekdayData,aes(x=weekday, y=motivated, col = group, group = group))+
  geom_smooth(method = "loess", se = TRUE, level =0.75)+
  labs(
    title="Average level of motivation per Weekday in dependency of the medication group",
    x="Weekday",
    y="Average motivation (self-reported on a scale from 1 to 100)")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot for aroused: (confidence interval =75%)
ggplot(WeekdayData,aes(x=weekday, y=aroused, col = group, group = group))+
  geom_smooth(method = "loess", se = TRUE, level =0.75)+
  labs(
    title="Average level of arousation per Weekday in dependency of the medication group",
    x="Weekday",
    y="Average arousation (self-reported on a scale from 1 to 100)")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Level 2
```{r visualization level 2}
#Make 5 graphs, one for the stress level and one for each mood. If I find the time I will maybe later also try to add the cycle phases like in your example graph.
#Graph for stress: (confidence interval=95%)
CleanData %>% 
  filter(group == "dOC") %>% 
  ggplot(aes(x=dOC_cycle_day, y=stressed_24h))+
  geom_smooth(method="loess", se=TRUE)+
  labs(
    title = "Mean stress rating by cycle phase (day)",
    x="Cycle day",
    y="Mean stress level (self-reported on a scale from 1 to 100)"
  )+
  theme_minimal()

#Graph for depressed: (confidence interval=95%)
CleanData %>% 
  filter(group == "dOC") %>% 
  ggplot(aes(x=dOC_cycle_day, y=depressed))+
  geom_smooth(method="loess", se=TRUE)+
  labs(
    title = "Mean depression rating by cycle phase (day)",
    x="Cycle day",
    y="Mean depression level (self-reported on a scale from 1 to 100)"
  )+
  theme_minimal()

#Graph for satisfied: (confidence interval=95%)
CleanData %>% 
  filter(group == "dOC") %>% 
  ggplot(aes(x=dOC_cycle_day, y=satisfied))+
  geom_smooth(method="loess", se=TRUE)+
  labs(
    title = "Mean satisfaction rating by cycle phase (day)",
    x="Cycle day",
    y="Mean satisfaction level (self-reported on a scale from 1 to 100)"
  )+
  theme_minimal()

#Graph for motivated: (confidence interval=95%)
CleanData %>% 
  filter(group == "dOC") %>% 
  ggplot(aes(x=dOC_cycle_day, y=motivated))+
  geom_smooth(method="loess", se=TRUE)+
  labs(
    title = "Mean motivation rating by cycle phase (day)",
    x="Cycle day",
    y="Mean motivation level (self-reported on a scale from 1 to 100)"
  )+
  theme_minimal()

#Graph for aroused: (confidence interval=95%)
CleanData %>% 
  filter(group == "dOC") %>% 
  ggplot(aes(x=dOC_cycle_day, y=aroused))+
  geom_smooth(method="loess", se=TRUE)+
  labs(
    title = "Mean arousation rating by cycle phase (day)",
    x="Cycle day",
    y="Mean arousation level (self-reported on a scale from 1 to 100)"
  )+
  theme_minimal()
```
## Kommentar:
- I made the line plots with geom_smooth, beacuse it is easy to generate confidence intervals with this. I hope this is ok, if I should calculate the mean first and just plot the mean as a geom_line, just tell me.
- The way I understand it i should always do the plots for each mood seperateley, or did you imagine a solution where all 4 moods are in some way calculated together for a general mood rating
- Because of that I also have a lot of redundant code I will try to reduce this if find some solutions, maybe you have an idea how?
- It was a really cool project and I had a lot of fun doing this, thanks! :)
