---
title: "Bene-Klein"
output: md_document
---


```{r setup,include=FALSE}
library(tidyverse)
library(readxl)
library(rstudioapi)
library(lubridate)
library(patchwork)
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# data manipulation
After the necessary packages an the dataset were loaded the dataset was cleaned up by removing all Na's and datasets that contain Na's.

After that the data will be summarized by the 'id' of the patient with mean and deviation of 'stressed_24h','stressed_moment', 'depressed', 'satisfied', 'motivated' and 'aroused'.
Then the same is made by grouping the data by 'group'.
```{r data cleaning,echo=FALSE}
data <- read_xlsx("stress_ratings_anonymized.xlsx") %>%  
  select(-17) %>% 
  drop_na(-c(id_dOC, dOC_cycle_date, dOC_cycle_day))
```
```{r data aggregation,include=FALSE}
# Give out a tibble containing the mean and standard deviation for each mood and stress item for each participant
data %>% 
  group_by(id) %>% 
  summarise(
    across(
      c(stressed_24h, stressed_moment, depressed, satisfied, motivated, aroused),
      list(mean = mean, SD = sd),
      .names = "{.fn}_{.col}"
    )
  )

# Give out a tibble containing the mean and standard deviation for each mood and stress item for each group
data %>% 
  group_by(group) %>% 
  summarise(
    across(
      c(stressed_24h, stressed_moment, depressed, satisfied, motivated, aroused),
      list(mean = mean, SD = sd),
      .names = "{.fn}_{.col}"
    )
  )
```
Then the  correlation between a stress moment and the a positive and negative mood is created with the stress moment beeing 'stressed_moment' and the positive/negative mood beeing 'motivated'/'tense'.
```{r data transformation plots,echo=FALSE}
# Define custom colors for the datapoints of each group
custom_cols <- c("sOC" = "red", "dOC" = "darkgreen", "OC" = "lightblue")

plot_function <- function(X, Y) {
  ggplot(data, aes(x = .data[[X]], y = .data[[Y]], col = group)) +
    geom_count(size = 0.8) +
    scale_color_manual(values = custom_cols) +
    geom_smooth(method = "lm", col = "#480355", se = TRUE, level = 0.99, fill = "#9448BC") +
    expand_limits(x = c(0, max(data[[X]], na.rm = TRUE)), y = c(0, max(data[[X]], na.rm = TRUE))) +
    labs(
      title = "How stress influences negative mood",
      subtitle = "(Self-reported perception, scores from 0 to 100)",
      x = paste0("How stressed did you feel today? (", X, ")"),
      y = paste0("How was your mood today? (", Y, ")")
    ) +
    theme(
      panel.background = element_blank(),
      axis.line = element_line(color = "black")
    )
}
plot_function("stressed_moment","tense")

plot_function("stressed_moment","motivated")
```

# Visualization
The visualization of the participants mood and stress are shown during weekdays.

The result shows that the majority of all medication starter **sOC** is feeling most depressed at the middle of the week (Th). 

For the **dOC** group the mood varies much more during the week.
For the **OC** group the mood varies even less and is even is better in comparison to **doc**. 

```{r visualization, echo=FALSE}
# Convert timestamps to weekdays 
Weekdaydata <- data %>% 
  mutate(
    date = as_datetime(timeStampStop),
    days_since_start = as.integer(difftime(date, min(date), units = "days")),
    weekday = factor((days_since_start%%7)+1,
                     levels = 1:7,
                     labels= c("Mo","Tu","We","Th","Fr","Sa","Su"))
  )

long_data <- Weekdaydata %>%
  pivot_longer(cols = c(depressed, satisfied, motivated, aroused),
               names_to = "mood",
               values_to = "value")

# Combined plot function
combined_plot <- ggplot(long_data, aes(x = weekday, y = value, color = group, group = group)) +
  geom_smooth(method = "loess", se = TRUE, level = 0.75) +
  scale_color_manual(values = custom_cols) +
  facet_wrap(~ mood, ncol = 2, scales = "free_y") +
  labs(
    title = "Mood per Weekday by Medication Group",
    x = "Weekday",
    y = "Mood score (1 to 100)"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.background = element_blank(),
    axis.line = element_line(color = "black")
  )

# Show the plot
combined_plot
```