---
title: 'First project attempt - SunKyoungs project'
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(ggplot2)
library(gridExtra)
```
## General project topic

Companies often discover unique patterns when organizing and visualizing consumer data. This project aims to visualize shopping trends of American consumers by age group and gender, focusing on their Purchase Amount and Frequency of Purchases. You need to create four separate graphs:

  1. Female consumers’ purchase amount by age group,
  2. Male consumers’ purchase amount by age group,
  3. Female consumers’ frequency of purchases by age group,
  4. Male consumers’ frequency of purchases by age group.

## The Data Set (processed)
```{r import data}
# Define the URL of the raw csv file
url <- "https://raw.githubusercontent.com/OscarL7/Shopping-Trends-Analysis/main/shopping_trends_updated.csv"

# Import the csv file 
data_raw <- read.csv(url)

# Data processing
data_pro <- data_raw %>%
  select(all_of(c("Customer.ID", "Age", "Gender", "Purchase.Amount..USD.", "Frequency.of.Purchases"))) %>%
  mutate(`Frequency [1/year]` = case_when(
    Frequency.of.Purchases == "Annually" ~ 1,
    Frequency.of.Purchases == "Bi-Weekly" ~ 52/2,
    Frequency.of.Purchases == "Every 3 Months" ~ 4,
    Frequency.of.Purchases == "Fortnightly" ~ 52/2,
    Frequency.of.Purchases == "Monthly" ~ 12,
    Frequency.of.Purchases == "Quarterly" ~ 4,
    Frequency.of.Purchases == "Weekly" ~ 52,
    TRUE ~ 10000
  ))

# Preview the data
head(data_pro)
```

## Visualisation

Purchase amount by age for two genders:

```{r visualisation I}
LabsAge <- c("18-22", "23-27", "28-32", "33-37", "38-42",
             "43-47", "48-52", "53-57", "58-62", "63-67",
             "68-70")
  
# Grouping ages into 5-year intervals and calculating total purchase amount by age group and gender
p1 <- data_pro %>%
  mutate(AgeGroup = cut(Age, breaks = seq(18, 72, by = 5), right = FALSE),
         Gender = factor(Gender, levels = c("Male", "Female"))) %>%
  group_by(AgeGroup, Gender) %>%
  summarise(TotalPurchase = sum(Purchase.Amount..USD.), .groups = 'drop') %>%
  ggplot(aes(x = AgeGroup, y = TotalPurchase, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Consumers Purchase Amount by Age and Gender",
       x = "Age",
       y = "Total\n Purchase\n Amount\n [USD]") +
  scale_x_discrete(labels = LabsAge) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        axis.title.y = element_text(angle = 0, vjust = 1)) 

p1
```

Purchase frequency (per year) by age for two genders (Numeric frequency was calculated based on the Frequency of Purchases):
```{r visualisation II}
p2 <- data_pro %>%
  mutate(AgeGroup = cut(Age, breaks = seq(18, 72, by = 5), right = FALSE),
         Gender = factor(Gender, levels = c("Male", "Female"))) %>%
  group_by(AgeGroup, Gender) %>%
  summarise(FreqPurchase = sum(`Frequency [1/year]`), .groups = 'drop') %>%
  ggplot(aes(x = AgeGroup, y = FreqPurchase, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Consumers Frequency of Purchase by Age and Gender",
       x = "Age",
       y = "Frequency\n of\n Purchase\n [1/year]") +
  scale_x_discrete(labels = LabsAge) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"),
        axis.title.y = element_text(angle = 0, vjust = 1)) 

p2


```