---
title: "Bye Bye Birdie by Christian"
output: md_document
date: "2023-06-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Project Solution of Jasmin's 'Bye Bye Birdie' Project
First we need to load the data set and take the variables of interest into a working data set. Since the Data Set is organized in Losses, a higher positive number means that that species experienced higher losses and vice versa.To make it more clear, I flipped the values, so that positive values now mean a gain in population instead of a loss and rounded the values to full values. Now, mean, SD and median values represent the gains/losses of that specific species across Europe. 

```{r echo=FALSE}
library(tidyverse)
library(readxl)
df <- read_csv("https://zenodo.org/record/5544548/files/EU%20birds%20decline%20overall%20in%20line%20with%20global%20patterns_species_results.csv?download=1", show_col_types = FALSE)

# Some Data Documentation

# species: species accepted latin name in HBW and BirdLife International (2019)

# common_name: species common name in English

# Loss_mean: Mean loss of individuals (1980-2017)
# Loss_sd: Standard deviation of mean loss of individuals 
# Loss_med: Median loss of individuals (1980-2017)

# total_proportional_change: Proportional change 1980:2017
# annual rate of change:		Annual rate of change 1980:2017

# population_trend_long:		a categorical assessment of the long-term trend in national species abundance: (I) Increasing, (D) Decreasing, (F) Fluctuating around zero, (U) Uncertain, (UNK) Unknown, or (S) Stable.
df$`annual rate of change`
df_reduced <- subset(df, select = c(species, common_name, Loss_mean, Loss_sd, Loss_med, total_proportional_change, `annual rate of change`))
list <- c("Loss_mean", "Loss_sd", "Loss_med")
for (i in list) {
  df_reduced[i] <- round(df_reduced[i],1)
  
}
list <- c("total_proportional_change", "annual rate of change")
for (i in list) {
  df_reduced[i] <- round(df_reduced[i],3)
  
}
# Since the dataframe is orientated in Losses, it is somewhat confusing, since negative values mean population growth, while positive values equal a decline --> reframe the dataset
bird <- df_reduced

bird<-bird|>mutate(across(c(Loss_mean, Loss_sd, Loss_med),function(x)(x*-1) ))
knitr::kable(head(bird[1:20,]))
```
Now with the working Data Set, the next step would be to implement the "other" category. In the graph you provided, the cutoff was at these two species:
  - Cyanistes caeruleus 
  - Passer montanus 

We want to summarize every species, that has a lower/higher mean value than these two species, into a "other" category. To accomplish this, we have to create a new variable, that copies the name of the species if it is outside the range specified, or changes its name to "other". 

```{r echo=FALSE}
# match("Cyanistes caeruleus",bird$species)
# match("Passer montanus", bird$species)     #only needed to find index position: 370 & 353 
x1<-print(bird$Loss_mean[370])
x2<-print(bird$Loss_mean[353])

x=0
for (i in bird$Loss_mean) {
  x=x+1
  if (x1>i & x2<i){
    bird$category_name[x]<-"other"
    bird$category_species[x]<-"other"
  }  else {
    bird$category_name[x]<-bird$common_name[x]
    bird$category_species[x]<-bird$species[x]
    }
}



test <- bird
# filter nach Bedingungen
test2<-test|> 
  filter(category_name=="other" & Loss_mean<0)|>
  mutate(Loss_mean=sum(Loss_mean),Loss_med=sum(Loss_med))
'mean(test2$Loss_mean)
mean(test2$Loss_med)
nrow(test2)'
  # Mean: -280815724, Median: -281267170, Name: other species; N = 163

test3<-test|> 
  filter(category_name=="other" & Loss_mean>0)|>
  mutate(Loss_mean=sum(Loss_mean),Loss_med=sum(Loss_med))
'mean(test3$Loss_mean)
mean(test3$Loss_med)
nrow(test3)'
  # Mean: 118702281, Median: 114432915, Name: other species; N = 198


# adding the values to the dataframe and removing those, that are not necessary for analysis (all the duplicate other entries)
bird<-bird|>
  mutate(Loss_mean = ifelse(category_name=="other" & Loss_mean>0,mean(test3$Loss_mean),
                            ifelse(category_name=="other" & Loss_mean<0,mean(test2$Loss_mean),Loss_mean)),
         Loss_med = ifelse(category_name=="other" & Loss_med>0,mean(test3$Loss_med),
                            ifelse(category_name=="other" & Loss_med<0,mean(test2$Loss_med),Loss_med)),
         category_name = ifelse(category_name=="other" & Loss_mean>0,"other species; N = 198",
                            ifelse(category_name=="other" & Loss_mean<0,"other species; N = 163",category_name)))

bird<-bird|>select(c(Loss_mean,Loss_med,category_name))|>distinct()
bird<-bird[c(-1,-4),]
# Notiz: Das war wahrscheinlich ein extrem "roundabout way" das Problem zu lösen aber, hey irgendwie hat es doch noch geklappt...


bird<-bird|>
  mutate(across(c(Loss_mean, Loss_med), function(x)(x/1000000)))

list <- c("Loss_mean", "Loss_med")
for (i in list) {
  bird[i]<-round(bird[i],1)
}  

knitr::kable(bird)
```

With the data frame assembled, now we can tackle the visualization. 

```{r echo=FALSE}

bird|>
  ggplot(aes(x="",Loss_mean, fill=Loss_mean))+ 
  geom_col(width = 0.3)+
  geom_text(aes(label = category_name), position = position_stack(vjust = 0.5))+
  scale_y_continuous()


```

final markdown notes:
no code, only important graphs, tables (use knitr::kable(df[], caption="xy")), focus on visualizations