---
title: "initialsolutioncars_Svenja"
author: "SvenjaPoerstel"
date: "2025-06-03"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse) 
library(readr) 
library(dplyr)
library(ggplot2)
```

```{r}
# read csv
car <- read_csv("Auto_data.csv", show_col_types = FALSE)
```

## Data preparation

```{r}

## 1: Ensure each variable has its own column and each observation its own row
#check structure and summary of the data set 

str(car)
summary(car)
glimpse(car)

#check if values are missing
colSums(is.na(car))

## 2: Convert categorical variables into factors with descriptive labels
car <- car %>%
  mutate(origin = factor(origin,
                         levels = c(1, 2, 3),
                         labels = c("USA", "Europe", "Japan")))
#transfer horsepower to numeric


## 3: Identify and report any duplicate rows in the dataset
duplicates <- car[duplicated(car), ]
n_duplicates <- nrow(duplicates)
n_duplicates
duplicates

## 4: Discuss whether duplicates are meaningful or should be removed.
## 5: Identify columns with missing values.
colSums(is.na(car))

## 6: Summarize the number and percentage of missing values.
missing_summary <- car%>%
  summarise_all(~sum(is.na(.))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "missing_count") %>%
  mutate(percent_missing = round((missing_count / nrow(car)) * 100, 2))

missing_summary

## 7: Decide how to handle the missing values (e.g., remove, impute, or analyze separately).
## 8: Create a new variable for the **power-to-weight ratio** and store it in a new column.

car <- car %>%
  mutate(power_to_weight = horsepower / weight)

```

##Data visualization 
#Scatterplot: relationship between mpg and horsepower

```{r}

p1 <- ggplot(car, aes(x = horsepower, y = mpg)) +
  geom_point(alpha = 0.6, color = "blue") +
  geom_smooth(method = "lm", se = TRUE, color = "darkblue") +
  labs(title = "Relationship between MPG and Horsepower",
       x = "Horsepower (PS)",
       y = "Miles per Gallon (mpg)") +
  theme_minimal()
print(p1)

```

#Scatterplot: relationship between weight and acceleration

```{r}
p2 <- ggplot(car, aes(x = weight, y = acceleration)) +
  geom_point(alpha = 0.6, color = "red") +
  geom_smooth(method = "lm", se = TRUE, color = "darkred") +
  labs(title = "Relationship between Weight and Acceleration",
       x = "Weight (lbs)",
       y = "Acceleration (0-60 mph in seconds)") +
  theme_minimal()
print(p2)
```

#trend analysis

```{r}

car_yearly <- car %>%
  group_by(year) %>%
  summarise(
    avg_mpg = mean(mpg, na.rm = TRUE),
    avg_hp = mean(horsepower, na.rm = TRUE),
    avg_acc = mean(acceleration, na.rm = TRUE)
  )

p3 <- ggplot(car_yearly, aes(x = year)) +
  geom_line(aes(y = avg_mpg, color = "mpg"), size = 1) +
  geom_smooth(aes(y = avg_mpg, color = "mpg"), method = "loess", se = FALSE, linetype = "dashed") +
  geom_line(aes(y = avg_hp, color = "horsepower"), linewidth = 1) +
  geom_smooth(aes(y = avg_hp, color = "horsepower"), method = "loess", se = FALSE, linetype = "dashed") +
  geom_line(aes(y = avg_acc, color = "acceleration"), size = 1) +
  geom_smooth(aes(y = avg_acc, color = "acceleration"), method = "loess", se = FALSE, linetype = "dashed") +
  scale_color_manual(
    name = "Variable",
    values = c("mpg" = "blue", "horsepower" = "red", "acceleration" = "green")
  ) +
  labs(title = "Trend Analysis of Car Characteristics over Years",
       x = "Year",
       y = "Average Value") +
  theme_minimal()

print(p3)

```
## Results 

-Negative Relationship between MPG and Horsepower and weight and acceleration, but both data are fluctuating
-Acceleration is mostly stable over the time, there is a small increase in mpg over time. Horsepower decreased a lot 

