---
title: "Solution for Ziyi"
date: "2026-03-25"
output: md_document
---

## Project Solution for: Surprisal-Based Comparison of Human and LLM Language Generation

By Jakob Raatschen


#### Loading the data set and checking the first few rows
```{r, echo=FALSE, warning = FALSE, message = FALSE}
library(knitr)
library(tidyverse)
library(patchwork)

# set current working directory to the directory of this Rmd file
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
# load data
data<-read_csv("./data/Processed final_data/final.csv")
kable(head(data))

```

```{r, echo = FALSE, message = FALSE}
# Data cleaning and aggregation

# remove NAs
data <- data %>%
  drop_na()

# Removing items 13 and 28 for data set used for comparison with human participants
data_human <- data %>%
  filter(!(item %in% c(13, 28)))

# Aggregating surprisal values (by generator and condition) - comparison with HUMAN
data_agg_human <- data_human %>%
  group_by(generator, context, condition) %>%
  summarise(mean_surprisal = mean(surprisal), .groups = "drop")

data_agg<- data %>%
  group_by(generator, context, condition) %>%
  summarise(mean_surprisal = mean(surprisal), .groups = "drop")

```

## Data Visualizations

As dependent variable we use the mean surprisal value, which is aggregated by generator and context/condition. To compare surprisal patterns across generators, the results will be visualized using line plots.

For the visualizations including human participants (I & II), Items 13 & 28 have been removed, for the visualization comparing only LLMs (II), all items are included.

### Visualization I

```{r, echo = FALSE, message = FALSE}
# add colour-palette vector
colours <- c("#A01818", "#9DCCEF", "#00767B")

p1<- data_human %>%
  mutate(condition = factor(condition, levels = c("RA", "EW", "OH"))) %>%
  group_by(generator, context, condition) %>%
  summarise(se = sd(surprisal, na.rm = TRUE) / sqrt(n()),
            mean_surprisal = mean(surprisal), .groups = "drop") %>%
  ggplot(aes(x = context, y = mean_surprisal, color = fct_reorder2(generator, context, mean_surprisal), group = generator)) +
  geom_line(linewidth = .8) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_surprisal - se, ymax = mean_surprisal + se), width = 0.2) +
  # add colours
  scale_color_manual(values = c("gpt-3.5-turbo-16k" = colours[1], 
                                "gpt-5.2-chat-latest" = colours[2], "human" = colours[3])) +
  labs(title = "Mean Surprisal by Context and Generator",
       x = "Context",
       y = "Mean Surprisal",
       color = "Generator") +
  theme_minimal() +
  theme(
    legend.position = "right",
    panel.grid.minor = element_blank()
  ) +
  facet_wrap(~condition)
p2 <- p1 +
  labs(title = "Mean Surprisal with Y-Axis starting at 0") +
  expand_limits(y = 0)

p1
p2
```


### Visualization II

```{r, echo = FALSE, message = FALSE}
p3<- data_human %>%
  group_by(generator, context) %>%
  summarise(sd = sd(surprisal, na.rm = T),
            se = sd / sqrt(n()),
            mean_surprisal = mean(surprisal)) %>%
  ggplot(aes(x = context, y = mean_surprisal, color = fct_reorder2(generator, context, mean_surprisal), group = generator)) +
  geom_line(linewidth = .8) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_surprisal - se, ymax = mean_surprisal + se), width = 0.2) +
  # add colours
  scale_color_manual(values = c("gpt-3.5-turbo-16k" = colours[1], 
                                "gpt-5.2-chat-latest" = colours[2], "human" = colours[3])) +
  labs(title = "Mean Surprisal by Context and Generator",
       x = "Context",
       y = "Mean Surprisal",
       color = "Generator") +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    # remove legend for this plot, since it will be shown in the second plot
    legend.position = "none"
  )

p4<- p3 +
  labs(title = "") +
  theme(
    legend.position = "right"
  ) +
  expand_limits(y = 0)

p3 | p4
```


**Interpretation:** 

In a highly predictable preceeding context the mean surprisal value of the two GPT generators is closest and both have a considerable difference to the human participants, who in general have lower surprisal values. In a medium predictable preceeding context, the GPT 3.5 still is closer to the GPT 5.2, but its surprisal value approaches more that of the human participants.

However, the distances are very small, which can be seen when looking at the plot with the y-axis starting at 0.

### Visualization III

```{r, echo = FALSE, message = FALSE}
p5<- data %>%
  mutate(condition = factor(condition, levels = c("RA", "EW", "OH"))) %>%
  filter(generator != "human") %>%
  group_by(generator, condition) %>%
  summarise(sd_val = sd(surprisal, na.rm = T),
            se = sd_val / sqrt(n()),
            mean_surprisal = mean(surprisal)) %>%
  ggplot(aes(x = condition, y = mean_surprisal, color = fct_reorder2(generator, condition, mean_surprisal), group = generator)) +
  geom_line(linewidth = .8) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_surprisal - se, ymax = mean_surprisal + se), width = 0.2) +
  # add colours
  scale_color_manual(values = c("gpt-3.5-turbo-16k" = colours[1], 
                                "gpt-5.2-chat-latest" = colours[2])) +
  labs(title = "Mean Surprisal by Condition and Generator",
       x = "Condition",
       y = "Mean Surprisal",
       color = "Generator") +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank()
  )

p6 <- p5 +
  labs(title = "") +
  expand_limits(y = 0) +
  theme(
    legend.position = "right"
  )

p5 | p6
```

**Interpretation:**

The difference between both GPT generators is highest in the OH condition (when no specific particles are present). I would conclude from this that when no specific particles are present, the two models are less specificly trained and therefore less restricted in their generation leading to the largest differences in output.

However, the distances are very small, which can be seen when looking at the plot with the y-axis starting at 0.


## Additional Visualization (Correlation between surprisal values of  generators)

```{r, echo = FALSE, message = FALSE, warning = FALSE}

# Scatterplots comparing surprisal values of different generators for each item, colored by context
generator_combinations <- combn(unique(data$generator), 2, simplify = FALSE)
map(generator_combinations, ~ {
  df_plot <- data %>%
    filter(generator %in% .x) %>%
    group_by(item, context, condition, generator) %>%
    summarise(item_surprisal = mean(surprisal, na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(
      names_from = generator, 
      values_from = item_surprisal
    )
  r_value <- round(cor(df_plot[[.x[1]]], df_plot[[.x[2]]], 
                       method = "pearson", use = "complete.obs"), 2)
  ggplot(df_plot, aes(x = .data[[.x[1]]], 
                      y = .data[[.x[2]]], 
                      color = context)) +
    geom_point(size = 1.5, alpha = 0.5) +
    geom_smooth(method = "lm", se = FALSE, color = "black") +
    annotate("text", x = Inf, y = Inf, 
             label = paste("r =", r_value),
             hjust = 9.1, vjust = 2.1, size = 3, fontface = "bold") +
    labs(title = paste("Correlation:", .x[1], "vs", .x[2]),
         x = paste("Surprisal (", .x[1], ")", sep=""),
         y = paste("Surprisal (", .x[2], ")", sep="")) +
    theme_minimal() +
    theme(legend.position = "right", panel.grid.minor = element_blank())
})
```

**Interpretation:**

While the correlation across all generators is quite high, interestingly the smallest correlation is between the two GPT generators. This might be explained based on the parameters used for both models.

A systematic difference based on the context is not really apparent.