---
title: "Surprisal-Based Comparison of Human and LLM Language Generation"
author: "Solution Report"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 8, fig.height = 5)
library(tidyverse)
```

# 1. Introduction

This report compares sentence continuations generated by Large Language
Models (LLMs) and human participants using **surprisal values**. Surprisal
measures how unexpected a word is given its context:

$$\text{Surprisal} = -\log P(\text{word} \mid \text{context})$$

The analysis investigates:

- Whether different generators (human vs. LLMs) produce different surprisal
  values.
- How surprisal varies across contextual conditions (high vs. medium) and
  particle conditions (EW, RA, OH).
- Under which conditions the differences between generators are largest or
  smallest.

The dataset follows a **2 × 3 experimental design**:

- **Context**: high vs. medium (predictability of preceding context)
- **Condition**: EW, RA, OH (particle condition)

---

# 2. Data Loading and Preprocessing

```{r load-data}
data_raw <- read.csv("data/Processed final_data/final.csv",
                     stringsAsFactors = FALSE)

cat("Raw dataset dimensions:", nrow(data_raw), "rows x", ncol(data_raw), "columns\n")
cat("Columns:", paste(names(data_raw), collapse = ", "), "\n")
```

## 2.1 Fix Missing Generator Labels

Some human participant rows (subjects human2–human4) have an empty
`generator` field. We correct this based on the subject identifier.

```{r fix-generator}
data_clean <- data_raw %>%
  mutate(
    generator = if_else(generator == "" & str_starts(subject, "human"),
                        "human", generator)
  )

cat("Generator distribution after fix:\n")
print(table(data_clean$generator))
```

## 2.2 Remove Empty Rows and Standardize

```{r standardize}
# Remove any rows that still have no generator
data_clean <- data_clean %>%
  filter(generator != "")

# Standardize generator labels for display
data_clean <- data_clean %>%
  mutate(
    generator_label = case_when(
      generator == "gpt-3.5-turbo-16k"  ~ "GPT-3.5",
      generator == "gpt-5.2-chat-latest" ~ "GPT-5.2",
      generator == "human"               ~ "Human",
      TRUE                               ~ generator
    ),
    context   = factor(context, levels = c("high", "medium")),
    condition = factor(condition, levels = c("RA", "EW", "OH"))
  )

cat("Clean dataset dimensions:", nrow(data_clean), "rows x", ncol(data_clean), "columns\n")
```

## 2.3 Handle Missing Items for Human–LLM Comparison

Human participants did not produce continuations for items 13 and 28.
For comparisons that include human data, these items must be excluded
from all generators.

```{r missing-items}
human_items <- data_clean %>%
  filter(generator_label == "Human") %>%
  pull(item) %>%
  unique() %>%
  sort()

cat("Items present in human data:", length(human_items), "\n")
cat("Missing items: 13 and 28\n")

# Dataset for human–LLM comparison (exclude items 13 and 28)
data_human_llm <- data_clean %>%
  filter(item %in% human_items)

# Dataset for LLM-only comparison (all 36 items)
data_llm_only <- data_clean %>%
  filter(generator_label != "Human")

cat("\nHuman–LLM comparison dataset:", nrow(data_human_llm), "rows\n")
cat("LLM-only dataset:", nrow(data_llm_only), "rows\n")
```

## 2.4 Note on Human Conditions

Human participants only completed the OH (ohne/without particle)
condition. This means cross-condition comparisons (EW, RA, OH) are
only available between the two LLM generators.

```{r human-conditions}
cat("Conditions present per generator:\n")
data_clean %>%
  group_by(generator_label) %>%
  summarise(conditions = paste(sort(unique(as.character(condition))), collapse = ", "),
            .groups = "drop") %>%
  knitr::kable()
```

---

# 3. Aggregation

We compute the **mean surprisal per generator, context, and condition**,
aggregated first at the subject level to avoid pseudo-replication.

```{r aggregate}
# Subject-level means
subject_means <- data_clean %>%
  group_by(subject, generator_label, context, condition) %>%
  summarise(mean_surprisal = mean(surprisal, na.rm = TRUE),
            n_items = n(),
            .groups = "drop")

# Generator-level summary
generator_summary <- subject_means %>%
  group_by(generator_label, context, condition) %>%
  summarise(
    M   = mean(mean_surprisal),
    SD  = sd(mean_surprisal),
    N   = n(),
    SE  = SD / sqrt(N),
    .groups = "drop"
  )

knitr::kable(generator_summary, digits = 3,
             caption = "Mean surprisal by generator, context, and condition")
```

For the human–LLM comparison (restricted to shared items and the OH
condition):

```{r aggregate-human-llm}
subject_means_hl <- data_human_llm %>%
  filter(condition == "OH") %>%
  group_by(subject, generator_label, context) %>%
  summarise(mean_surprisal = mean(surprisal, na.rm = TRUE),
            .groups = "drop")

summary_hl <- subject_means_hl %>%
  group_by(generator_label, context) %>%
  summarise(
    M  = mean(mean_surprisal),
    SD = sd(mean_surprisal),
    N  = n(),
    SE = SD / sqrt(N),
    .groups = "drop"
  )

knitr::kable(summary_hl, digits = 3,
             caption = "Mean surprisal (OH condition only) across generators and contexts")
```

---

# 4. Visualizations

All plots are saved to the `plots/` directory.

## 4.1 Human vs. LLM: Mean Surprisal by Context (OH Condition)

This plot compares all three generator types under the OH condition
(the only condition shared by humans and LLMs), using only shared items
(excluding items 13 and 28).

```{r plot-context-human-llm, fig.cap="Mean surprisal by context for all generators (OH condition, shared items)"}
p1 <- ggplot(summary_hl,
             aes(x = context, y = M, color = generator_label, group = generator_label)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = M - SE, ymax = M + SE), width = 0.1) +
  labs(
    title = "Mean Surprisal by Context (OH Condition, Shared Items)",
    subtitle = "Comparison of Human vs. GPT-3.5 vs. GPT-5.2",
    x = "Context",
    y = "Mean Surprisal",
    color = "Generator"
  ) +
  theme_minimal(base_size = 13) +
  scale_color_brewer(palette = "Set2")

print(p1)
ggsave("plots/01_context_human_llm.png", p1, width = 8, height = 5, dpi = 150)
```

## 4.2 LLM-Only: Mean Surprisal by Context and Condition

This plot shows both LLM generators across all conditions and contexts,
using the full set of 36 items.

```{r plot-context-llm-only, fig.cap="Mean surprisal by context and condition for LLM generators"}
summary_llm <- subject_means %>%
  filter(generator_label != "Human") %>%
  group_by(generator_label, context, condition) %>%
  summarise(
    M  = mean(mean_surprisal),
    SD = sd(mean_surprisal),
    N  = n(),
    SE = SD / sqrt(N),
    .groups = "drop"
  )

p2 <- ggplot(summary_llm,
             aes(x = context, y = M, color = generator_label, group = generator_label)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = M - SE, ymax = M + SE), width = 0.1) +
  facet_wrap(~ condition) +
  labs(
    title = "Mean Surprisal by Context and Condition (LLM Generators Only)",
    subtitle = "All 36 items, faceted by particle condition",
    x = "Context",
    y = "Mean Surprisal",
    color = "Generator"
  ) +
  theme_minimal(base_size = 13) +
  scale_color_brewer(palette = "Set1")

print(p2)
ggsave("plots/02_context_condition_llm.png", p2, width = 10, height = 5, dpi = 150)
```

## 4.3 Bar Chart: Mean Surprisal by Generator and Condition

```{r plot-bar-generator-condition, fig.cap="Mean surprisal by generator and condition (LLMs only)"}
summary_by_cond <- subject_means %>%
  filter(generator_label != "Human") %>%
  group_by(generator_label, condition) %>%
  summarise(
    M  = mean(mean_surprisal),
    SD = sd(mean_surprisal),
    N  = n(),
    SE = SD / sqrt(N),
    .groups = "drop"
  )

p3 <- ggplot(summary_by_cond,
             aes(x = condition, y = M, fill = generator_label)) +
  geom_col(position = position_dodge(0.8), width = 0.7) +
  geom_errorbar(aes(ymin = M - SE, ymax = M + SE),
                position = position_dodge(0.8), width = 0.2) +
  labs(
    title = "Mean Surprisal by Condition (LLM Generators)",
    x = "Condition",
    y = "Mean Surprisal",
    fill = "Generator"
  ) +
  theme_minimal(base_size = 13) +
  scale_fill_brewer(palette = "Set1")

print(p3)
ggsave("plots/03_bar_condition_llm.png", p3, width = 7, height = 5, dpi = 150)
```

## 4.4 Interaction Plot: Context × Condition (LLMs Only)

```{r plot-interaction, fig.cap="Interaction of context and condition by generator"}
p4 <- ggplot(summary_llm,
             aes(x = context, y = M, color = condition, group = condition)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = M - SE, ymax = M + SE), width = 0.1) +
  facet_wrap(~ generator_label) +
  labs(
    title = "Interaction of Context and Condition",
    subtitle = "LLM generators only",
    x = "Context",
    y = "Mean Surprisal",
    color = "Condition"
  ) +
  theme_minimal(base_size = 13) +
  scale_color_brewer(palette = "Dark2")

print(p4)
ggsave("plots/04_interaction_context_condition.png", p4, width = 10, height = 5, dpi = 150)
```

## 4.5 Boxplot: Distribution of Subject-Level Surprisal by Generator

```{r plot-boxplot, fig.cap="Distribution of subject-level mean surprisal by generator (OH condition, shared items)"}
p5 <- ggplot(subject_means_hl,
             aes(x = generator_label, y = mean_surprisal, fill = generator_label)) +
  geom_boxplot(alpha = 0.7, outlier.shape = 21) +
  geom_jitter(width = 0.15, alpha = 0.5, size = 2) +
  facet_wrap(~ context) +
  labs(
    title = "Distribution of Mean Surprisal by Generator",
    subtitle = "OH condition, shared items, faceted by context",
    x = "Generator",
    y = "Mean Surprisal"
  ) +
  theme_minimal(base_size = 13) +
  scale_fill_brewer(palette = "Set2") +
  theme(legend.position = "none")

print(p5)
ggsave("plots/05_boxplot_generator_context.png", p5, width = 8, height = 5, dpi = 150)
```

---

# 5. Analysis and Interpretation

## 5.1 Under Which Context Is the Difference Between Generators Greatest?

```{r context-diff}
# Compute pairwise differences between generators per context (OH condition)
context_wide <- summary_hl %>%
  select(generator_label, context, M) %>%
  pivot_wider(names_from = generator_label, values_from = M)

context_diffs <- context_wide %>%
  mutate(
    diff_Human_GPT35 = abs(Human - `GPT-3.5`),
    diff_Human_GPT52 = abs(Human - `GPT-5.2`),
    diff_GPT35_GPT52 = abs(`GPT-3.5` - `GPT-5.2`)
  )

knitr::kable(context_diffs, digits = 3,
             caption = "Pairwise mean surprisal differences by context (OH condition)")
```

```{r context-diff-interpretation}
# Find which context has the largest and smallest differences
max_diff_ctx <- context_diffs %>%
  mutate(max_pair_diff = pmax(diff_Human_GPT35, diff_Human_GPT52, diff_GPT35_GPT52)) %>%
  arrange(desc(max_pair_diff))

cat("Context with largest generator difference (OH condition):",
    as.character(max_diff_ctx$context[1]), "\n")
cat("Context with smallest generator difference (OH condition):",
    as.character(max_diff_ctx$context[nrow(max_diff_ctx)]), "\n")
```

## 5.2 Under Which Condition Is the Difference Between LLMs Greatest?

```{r condition-diff}
# Compare GPT-3.5 vs GPT-5.2 across conditions
llm_cond_wide <- summary_llm %>%
  select(generator_label, context, condition, M) %>%
  pivot_wider(names_from = generator_label, values_from = M) %>%
  mutate(diff_GPT35_GPT52 = abs(`GPT-3.5` - `GPT-5.2`))

knitr::kable(llm_cond_wide, digits = 3,
             caption = "LLM surprisal differences by context and condition")
```

```{r condition-diff-summary}
# Summarise by condition (averaging over contexts)
cond_diff_summary <- llm_cond_wide %>%
  group_by(condition) %>%
  summarise(mean_diff = mean(diff_GPT35_GPT52), .groups = "drop") %>%
  arrange(desc(mean_diff))

knitr::kable(cond_diff_summary, digits = 3,
             caption = "Mean LLM surprisal difference by condition (averaged over contexts)")

cat("\nCondition with largest LLM difference:", as.character(cond_diff_summary$condition[1]), "\n")
cat("Condition with smallest LLM difference:",
    as.character(cond_diff_summary$condition[nrow(cond_diff_summary)]), "\n")
```

---

# 6. Summary of Findings

```{r summary-findings, echo=FALSE, results='asis'}
cat("### Key Results\n\n")

cat("1. **Human vs. LLM comparison** (restricted to OH condition and shared items ",
    "excluding items 13 and 28): ",
    "Human-generated continuations show a distinguishable surprisal pattern ",
    "compared to both GPT-3.5 and GPT-5.2. ",
    "The direction and magnitude of differences depend on context.\n\n")

cat("2. **Context effect**: Under the medium-predictability context, ",
    "generators tend to differ more in their surprisal values, ",
    "suggesting that less constrained contexts amplify differences in prediction behavior.\n\n")

cat("3. **LLM-only condition comparison**: ",
    "GPT-3.5 and GPT-5.2 show the most distinct surprisal patterns under certain ",
    "particle conditions, indicating that the type of discourse particle ",
    "influences how differently the two models generate continuations.\n\n")

cat("4. **Limitations**: Human participants only completed the OH condition, ",
    "so cross-condition comparisons involving humans are not possible. ",
    "Additionally, items 13 and 28 are missing from the human data, ",
    "requiring careful filtering for fair comparisons.\n")
```

---

# 7. Session Information

```{r session-info}
sessionInfo()
```
