---
title: "emily-works"
output: md_document
---

```{r setup, include=FALSE}

library(tidyverse)
library(ggplot2)

```

# Data Manipulation

## Data Import

```{r data-manipulation, include=FALSE}

raw_data <- read_csv("data/laptopData.csv")
head(raw_data)

```


- Remove empty rows with `drop_na()`
- Remove invalid symbols with `str_replace_all()`
- Remove the "GB" text extension from the Ram column and convert it to integer using `as.integer()`
- Remove the "kg" text extension from the Weight column and convert it to numeric using `as.double()`
- Extract the continuous numerical variable representing processor clock speed in GHz from the Cpu column using `str_extract()` and convert it to numeric using `as.double()`
- Create new columns (`is_IPS` and `is_Retina`) based on the presence of specific keywords in the ScreenResolution column using `str_detect()`
- Extract the pure pixel dimensions (Width and Height) from the ScreenResolution column using `str_extract()` and converts them to integer using `as.integer()` and stores them in separate columns
- Create a new column (`OpSys`) that groups categories into broader groups using `case_when()` and `str_detect()`

```{r data-cleaning, include=FALSE}

clean_data <- raw_data %>%
  drop_na() %>%  
  mutate(
    across(everything(), ~ str_replace_all(., "\\?", NA_character_)),
    Ram = as.integer(str_replace_all(Ram, "GB", "")),
    Weight = as.double(str_replace_all(Weight, "kg", "")),
    Cpu_clock = as.double(str_extract(Cpu, "\\d+\\.?\\d*(?=GHz)"))
    ) %>%
  mutate(
    is_IPS = str_detect(ScreenResolution, "IPS"),
    is_Retina = str_detect(ScreenResolution, "Retina"),
    ScreenResolution = str_extract(ScreenResolution, "\\d+x\\d+"),
    Width = as.integer(str_extract(ScreenResolution, "^\\d+")),
    Height = as.integer(str_extract(ScreenResolution, "(?<=x)\\d+")),
    OpSys = case_when(
      str_detect(OpSys, "Windows") ~ "Windows",
      str_detect(OpSys, "Mac OS X|macOS") ~ "macOS",
      TRUE ~ "Other")
      )
  
```

``` {r clean data check, echo = FALSE, results = 'asis'}

library(knitr)
kable(clean_data [1:5, ], caption = "Data after Manipulation")

```

# Visualization

I'll first convert the price to USD for better interpretability, assuming an exchange rate of 1 USD = 83 INR.
```{r price conversion, include = FALSE}

clean_data_USD <- clean_data %>%
  mutate(Price = as.numeric(Price), 
         Price_USD = Price / 83)
```

## 1. Price vs. Numeric Features

Creating three scatter plots with density overlays to explore relationships between:

- RAM vs. Price: how memory capacity correlates with cost
- CPU Clock Speed vs. Price: processor speed's impact on pricing
- Weight vs. Price: physical weight's relationship to price

```{r plots, include=FALSE}

plot_price_relationship <- function(data, x_var, x_label, plot_title) {
  data %>%
    ggplot(aes(x = .data[[x_var]], 
               y = Price_USD)) +
    geom_point(alpha = 0.3, 
               width = 0.2, 
               height = 0) +
    stat_density_2d(aes(fill = after_stat(level)), 
                    geom = "polygon", 
                    alpha = 0.2) +
    scale_fill_viridis_c() +
    theme_minimal() +
    labs(title = plot_title, 
         x = x_label, 
         y = "Price (USD)",
         fill = "Density")
}

```

### RAM & Price
```{r plot RAM, echo=FALSE, warning=FALSE}
plot_price_relationship(clean_data_USD, "Ram", "RAM (GB)", "Price vs. RAM with Market Density")
```

### CPU & Price
```{r plot CPU, echo=FALSE, warning=FALSE}
plot_price_relationship(clean_data_USD, "Cpu_clock", "CPU Clock Speed (GHz)", "Price vs. CPU with Market Density")
```

### Weight & Price
```{r plot Weight, echo=FALSE, warning=FALSE}
plot_price_relationship(clean_data_USD, "Weight", "Weight (kg)", "Price vs. Weight with Market Density")
```

## 2. Price vs. Categorical Features (Distribution Plot)

### By Company
```{r plot Company, echo=FALSE, height=1000}

clean_data_USD %>%
  ggplot(aes(x = Company, 
             y = Price_USD)) +
  geom_violin(fill = "blue", 
              alpha = 0.7) +
  geom_boxplot(width = 0.1, 
               alpha = 0.5) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Price Distribution by Company",
       x = "Brand",
       y = "Price (USD)")

```

### By Laptop Type
```{r plot LaptopType, echo=FALSE, height= 800}

clean_data_USD %>%
  ggplot(aes(x = TypeName, 
             y = Price_USD)) +
  geom_violin(fill = "blue") +
  geom_boxplot(width = 0.1) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Price Distribution by Laptop Type",
       x = "Brand",
       y = "Price (USD)")

```

## 3. Price Distribution Overlap

```{r ridgeline plot, echo=FALSE}

library(ggridges)

clean_data_USD %>%
  filter(TypeName %in% c("Notebook", 
                         "Ultrabook", 
                         "Gaming")) %>%
  ggplot(aes(x = Price_USD,
             fill = TypeName)) +
  geom_density(alpha = 0.4) +
  scale_fill_viridis_d() +
  labs(title = "Price Distribution Overlap: Notebook vs Ultrabook vs Gaming",
       x = "Price in USD",
       y = "Laptop Type",
       fill = "Type") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "bottom")


```





