---
title: "Project-Solution"
output: github_document
always_allow_html: true
---


```{r setup, include=FALSE}

library(tidyverse)
library(jsonlite)
library(ggrepel)
library(scales)
library(gridExtra)
library(rnaturalearth)
library(rnaturalearthdata)
library(stringdist)
library(purrr)
library(grid)
library(ggplot2)
library(tidytext)
library(knitr) 

# transfer speaker counts in millions/billions
format_million_billion <- label_number(
  scale_cut = cut_si(""),
  accuracy = 0.1
)

#figure size
knitr::opts_chunk$set(
  fig.width = 10,
  fig.height = 7,
  dpi = 150,
  out.width = "100%"
)


```

```{r load_data,echo=FALSE}

raw <- jsonlite::fromJSON("world_languages_integrated.json", simplifyVector = FALSE)

```

# Fun with languages

There are over 7000 languages spoken in the world. While some regions are home to a rich diversity of languages, individual languages may also be spoken across multiple countries. The provided dataset collects information about over 7000 languages (including sign languages), such as geographical information, information about religion and about language families. I am particularly interested in the regions with the greatest language diversity and in the (geographical) distribution of languages. I also want to take a look at the corresponding language families.

## Questions and Tasks
1. Where is the place with the most diversity (regarding languages)?
2. How are these languages distributed?
3. What are the languages and the language families with the most speakers? What languages have the most speakers in their family?


## Clean Data

Two things have to be solved before any plot can be drawn:

1. **Flattening.** `extract_language_entry()` pulls exactly four fields out of
   each entry and returns a one-row tibble; `map_dfr()` stacks them into a
   rectangular data frame. The `%||%` operator substitutes `NA` whenever a
   field is absent, so a missing value never silently drops a whole record.
  
2. **Country names.** The country strings in the JSON do not always match the
   `admin` names used by `rnaturalearth` (e.g. *"USA"* vs *"United States of
   America"*). I therefore fuzzy-match every country against the map's names.

```{r clean data}
# Load world map for country matching
world_basemap <- ne_countries(scale = "medium", returnclass = "sf") 
map_names <- world_basemap$admin

# Fuzzy‑match country names to map data
match_country <- function(x) {
  if (is.na(x)) return(NA)
  distances <- stringdist(x, map_names, method = "jw")
  map_names[which.min(distances)][1]
}

# Extract relevant fields from each language entry
extract_language_entry <- function(entry) {
  tibble(
    language = entry$name %||% NA_character_,
    country  = {
      ctry <- entry$speaker_count$metadata$countries
      if (is.null(ctry) || length(ctry) == 0) NA_character_
      else as.character(unlist(ctry))   
    },
    speakers = entry$speaker_count$count %||% NA_real_,
    family   = {
      path <- entry$language_history$family_tree$path
      if (!is.null(path) && length(path) >= 3) as.character(path[[3]])
      else NA_character_
    }
  )
}

# Build cleaned language dataset
language_data <- raw %>%
  map_dfr(extract_language_entry) %>%
  filter(!is.na(language)) %>%
  mutate(country_map_sf = sapply(country, match_country))

# data summary
n_rows <- nrow(language_data)
n_country_na <- sum(is.na(language_data$country))
n_speakers_na <- sum(is.na(language_data$speakers))
n_family_na <- sum(is.na(language_data$family))
```

The cleaned dataset contains `r n_rows` rows in total.

*Missing values summary:*

- Missing country information: `r n_country_na`
- Missing speaker counts: `r n_speakers_na`
- Missing language family information: `r n_family_na`


## Visualization

### 1. Where is the place with the most diversity (regarding languages)?

I read "diversity" as *the number of distinct languages spoken in a country*.
Because the long format contains one row per language–country pair, a language
spoken in ten countries would otherwise be counted ten times in the same
country if the source lists it repeatedly; `distinct()` before `count()`
guarantees that every language is counted **once per country**.

```{r q1-count-languages}
# Find the 20 countries with the most languages
country_diversity <- language_data %>%
  drop_na(country_map_sf) %>%
  distinct(language, country_map_sf) %>%
  count(country_map_sf, name = "n_languages")
 
top20_country_diversity <- country_diversity %>%
  arrange(desc(n_languages)) %>%
  slice_head(n = 20)
 
```

**Top 20 countries by number of languages**
```{r q1-table, echo=FALSE}
kable(
  top20_country_diversity,
  col.names = c("Country", "Number of languages"),
  align = "lc"
)
```

On a linear colour scale everything except the top few countries would collapse into a single shade, so I map the
fill to `log10(n_languages)` and re-label the colour bar with the original
counts (1, 3, 10, … 1000). Countries with no data keep a neutral grey, which
distinguishes "no information" from "few languages".

```{r q1-diversity-map}
# Log‑transform for color scale
country_diversity_log <- country_diversity %>%
  mutate(
    log_lang = log10(n_languages),
    log_lang = ifelse(is.infinite(log_lang), NA, log_lang)
  )


plot_diversity_map <- world_basemap %>%
  left_join(country_diversity_log, by = c("admin" = "country_map_sf")) %>%
  ggplot() +
  geom_sf(aes(fill = log_lang), color = "white", size = 0.1) +
  scale_fill_gradientn(
    colours = c("#5BC0EB", "#9DD9D2", "#9e3dff", "#F7A8B8", "#FF4D6D"),
    na.value = "grey90",
    name = "Languages",
    breaks = log10(c(1, 3, 10, 30, 100, 300, 600, 1000)),
    labels = c("1", "3", "10", "30", "100", "300", "600", "1000"),
    guide = guide_colorbar(
      barheight = 12,   
      barwidth  = 0.8,  
      title.position = "top"
    )
  ) +
  labs(title = "Language Diversity") +
  theme_void() +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5),
    legend.position = "right",   
    legend.title = element_text(size = 10),
    legend.text  = element_text(size = 8)
  )

plot_diversity_map
```

The map shows that linguistic diversity is not evenly spread: it concentrates
in a tropical belt — Papua New Guinea, Indonesia, Nigeria, India — whereas
large, historically centralised states (Russia, most of Europe, the Americas)
are comparatively homogeneous.

### 2. How are these languages distributed?

Two different notions of "big" are worth separating: a language can have many
**speakers** (Mandarin) or be present in many **countries** (English). I
therefore summarise each language once, with both measures. `first(speakers)`
is used rather than `sum()` because the speaker count is a *language-level*
attribute that is repeated on every country row — summing it would multiply
the count by the number of countries.
```{r q2-summarise}

# Summarize languages by total speakers and number of countries
lang_summary <- language_data %>%
  filter(!is.na(country), !is.na(speakers)) %>%
  group_by(language) %>%
  summarise(
    total_speakers = first(speakers),
    n_countries    = n_distinct(country),
    .groups        = "drop"
  )

top5_by_speakers <- lang_summary %>%
  arrange(desc(total_speakers)) %>%
  slice_head(n = 5)

```

**Top 5 languages by number of speakers**
```{r q2-table, echo=FALSE}
kable(
  top5_by_speakers,
  col.names = c("Language", "Total speakers", "Countries"),
  align = "lcc"
)
```


```{r}
top5 <- top5_by_speakers$language

spoken_pairs <- language_data %>%
  filter(language %in% top5) %>%
  distinct(language, country_map_sf) %>%
  drop_na(country_map_sf) %>%
  mutate(spoken = TRUE)

map_facets <- crossing(
    language = top5,
    admin    = world_basemap$admin
  ) %>%
  left_join(spoken_pairs, by = c("language", "admin" = "country_map_sf")) %>%
  mutate(spoken = !is.na(spoken))

world_basemap %>%
  filter(admin != "Antarctica") %>%          
  left_join(map_facets, by = "admin") %>%   
  ggplot() +
  geom_sf(aes(fill = spoken), color = "white", linewidth = 0.05) +
  scale_fill_manual(
    values = c("TRUE" = "#F7A8B8", "FALSE" = "grey92"),
    guide  = "none"
  ) +
  coord_sf(ylim = c(-56, 84), expand = FALSE) +   # 裁掉上下空白
  facet_wrap(~ language, ncol = 3) +
  labs(title = "Where the five largest languages are spoken") +
  theme_void(base_size = 10) +
  theme(
    plot.title    = element_text(face = "bold", size = 13, hjust = 0.5, margin = margin(t = 5, b = 14) ),
    strip.text    = element_text(face = "bold", size = 9, margin = margin(t = 4, b = 4)),
    panel.spacing = unit(0.8, "lines")
  )

```


```{r}
# Remove NA or 0 country counts
lang_summary_clean <- lang_summary %>%
  filter(
    !is.na(total_speakers),
    !is.na(n_countries),
    n_countries > 0
  )

# 10 most spoken
top10_speakers <- lang_summary_clean %>%
  arrange(desc(total_speakers)) %>%
  slice(1:10)

# 10 most wide-spread
top10_countries <- lang_summary_clean %>%
  arrange(desc(n_countries)) %>%
  slice(1:10)

selected_langs <- top10_speakers %>%
  bind_rows(top10_countries) %>%
  distinct(language, .keep_all = TRUE) %>%
  mutate(label_id = row_number())

plot_q2_2 <- ggplot(selected_langs, aes(
  x = total_speakers,
  y = n_countries
)) +
  geom_point(size = 5, color = "#1D3557") +

  # number labels
  geom_label_repel(
    aes(label = label_id),
    size = 2,
    color = "grey50",
   
    fontface = "bold",
    label.size = 0.3,
    label.r = unit(0.25, "lines"),
    box.padding = 0.2,       
    point.padding = 0.1
  ) +

  # language labels
  geom_text_repel(
    aes(label = language),
    size = 2,
    color = "black",
    box.padding = 0.6,       
    point.padding = 0.4,
    segment.color = "grey50"
  ) +

  scale_x_continuous(labels = function(x) format_million_billion(x)) +
  labs(
    title = "Top Languages by Speakers and Country Coverage",
    x = "Number of Speakers (M/B)",
    y = "Number of Countries"
  ) +
  theme_minimal(base_size = 10)

plot_q2_2
```
### 3. Languages and language families with the most speakers

A language family aggregates its languages, so its speaker total is the sum
over its members.
```{r q3-family-totals}

# Keep one row per language to avoid double‑counting speakers
lang_by_family <- language_data %>%
  filter(!is.na(speakers), !is.na(family)) %>%
  distinct(language, speakers, family)

# Compute total speakers per language family
family_summary <- lang_by_family %>%
  group_by(family) %>%
  summarise(total_speakers = sum(speakers), .groups = "drop") %>%
  arrange(desc(total_speakers))

top5_families <- family_summary %>% slice_head(n = 5) %>% pull(family)

```

**Top 5 language families by number of speakers**
```{r q3-family-table, echo=FALSE}
kable(
  family_summary %>% slice_head(n = 5),
  col.names = c("Language family", "Total speakers"),
  align = "lc"
)
```

For each of these five families we draw a stacked bar of its ten largest
languages. The stacking shows two things at once: the **height** is the family
total, the **segments** show how that total is composed — whether the family is
dominated by one giant language or spread over several mid-sized ones.

```{r q3-family-plots, fig.height=8}
# Plot top‑10 languages within a family using a unified blue gradient palette
plot_family_bar <- function(fam_name) {
  lang_by_family %>%
    filter(family == fam_name) %>%
    slice_max(speakers, n = 10) %>%
    mutate(language = fct_reorder(language, speakers)) %>%
    
    ggplot(aes(x = fam_name, y = speakers, fill = language)) +
    geom_col(width = 0.5, color = "white", linewidth = 0.3) +
    scale_y_continuous(labels = format_million_billion) +
    scale_fill_brewer(palette = "Paired") +
    labs(
      title = fam_name,
      x = NULL,
      y = "Total Speakers"
    ) +
    theme_minimal(base_size = 10) +
    theme(
      plot.title = element_text(face = "bold", size = 11, hjust = 0.5,
                                 margin = margin(b = 5)),
      legend.key.size = unit(0.3, "cm"),
      legend.text = element_text(size = 6),
      legend.title = element_text(size = 7)
    )
}


family_plots <- purrr::map(top5_families, plot_family_bar)

big_family_plot <- grid.arrange(
  grobs = family_plots,
  ncol = 3
)

```

# Data Source & License

This project uses the [World Languages dataset](https://huggingface.co/datasets/lukeslp/world-languages) by lukeslp, licensed under the MIT License. See `license.txt` in this repository for the full license text.




