---
title: "Wenjing_solution"
output: md_document
---

```{r import_data,echo=FALSE,message=FALSE,include=FALSE}
#install packages for data cleaning
library(tidyverse)
library(dplyr)

#install packages for importing json. file data 
#install.packages("jsonlite",repos = "https://cran.rstudio.com/")

#install packages for world map visualization 
#install.packages("rnaturalearth",repos = "https://cran.rstudio.com/")
#install.packages("rnaturalearthdata",repos = "https://cran.rstudio.com/")

#data import
library(jsonlite)
data <- fromJSON("https://huggingface.co/datasets/lukeslp/world-languages/raw/main/world_languages_integrated.json")


```


```{r reform_dataframe ,echo=FALSE,message=FALSE}
#done: reform data with column "language_name","speaking_country","speaker_count"
# fill missing value with 'NA' in tibble

#optimize codes by using pipeline

re_df <- tibble(
  language_name = data$name,
  speaker_count = data$speaker_count$count) %>% mutate(speaking_country={
    raw_countries=data$speaker_count$metadata$countries
    map_chr(raw_countries,~{
      if(is.null(.x)||all(is.na(.x))){
        NA_character_
      }else{
        paste(.x, collapse = ", ")
      }
    })
    
  }) 
 
```


Calculating the number of languages spoken in each country:
```{r data_cleaning_Q1, echo=FALSE}

#calculating the number of languages spoken in each country
language_num_in_countries <- re_df %>% 
  drop_na(speaking_country) %>%separate_rows(speaking_country,sep = ",\\s*") %>% group_by(speaking_country) %>% summarise(language_count=n_distinct(language_name)) %>% arrange(desc(language_count)) 
language_num_in_countries
```

Plot1: showing linguistic diversity per countries
```{r data_visualization_Q1,echo=FALSE,message=FALSE,warning=FALSE,results='hide',fig.width=12,fig.height=8,out.width="100%"}

library(rnaturalearth)
# world_map data is also prepared for the second plot
world_map <- ne_countries(scale = "small",returnclass = "sf")

map_data <- world_map %>% left_join(language_num_in_countries, by=c("name"="speaking_country") )

#plot the number of languages spoken per country
ggplot(data = map_data) +
    geom_sf(aes(fill = language_count), color = "black", size = 0.1) +
    scale_fill_viridis_c(name = "Number of Languages", 
                         option = "viridis", 
                         na.value = "gray",
                         direction = -1) + 
    labs(title = "Linguistic Diversity by Country",
         subtitle = "Number of languages spoken per country")+
    theme_minimal() +
    theme(legend.position = "bottom",
          legend.key.width = unit(2, "cm"))

```

```{r data_cleaning_Q2.1,echo=FALSE}
# Find the five languages with the largest number of speakers(as highlighted data)
top5_speakers <- re_df %>% arrange(desc(re_df$speaker_count)) %>% head(5) %>%  select(language_name, speaking_country) %>% separate_rows(speaking_country,sep = ",\\s*")

highlight_data <- world_map %>% right_join(top5_speakers,by=c("name"="speaking_country")) %>% select(language_name,geometry)


```

```{r data_cleaning_Q2.2,echo=FALSE}
lang_stats <- re_df %>%
  # separate countries and remove missing values）
  separate_rows(speaking_country, sep = ",\\s*") %>%
 drop_na(speaking_country) %>%
  group_by(language_name) %>%
  summarise(
    speaker_count = first(speaker_count),              # speaker_count
    country_count = n_distinct(speaking_country)       # country_count
  ) %>%
  ungroup()

top_widespread <- lang_stats %>%
  slice_max(country_count, n = 10, na_rm = TRUE) %>%
  pull(language_name)

top_spoken <- lang_stats %>%
  slice_max(speaker_count, n = 10, na_rm = TRUE) %>%
  pull(language_name)

plot3_data <- bind_rows(
  lang_stats %>% filter(language_name %in% top_spoken),
  lang_stats %>% filter(language_name %in% top_widespread)
) %>%
  distinct(language_name, .keep_all = TRUE) %>%
  mutate(
    category = case_when(
      language_name %in% top_spoken & language_name %in% top_widespread ~ "Both Top 10s",
      language_name %in% top_spoken ~ "Only Spoken Top 10",
      language_name %in% top_widespread ~ "Only Widespread Top 10",
      TRUE ~ "Other"
    )
  ) %>%
  arrange(desc(speaker_count))


```


Plot2.1 showing geographic distribution of five largest languages
```{r data_visualization_Q2.1,echo=FALSE}
ggplot()+geom_sf(data=world_map,fill="grey",color="white",size=0.1)+geom_sf(data=highlight_data,fill="red",color="black",size=0.1)+facet_wrap(~language_name,ncol=3)+theme_minimal()+labs(title = "Geographic Distribution of five largest languages",subtitle = "Red: countries where the language is spoken ")

```

```{r data_visualization_Q2.2, echo=FALSE}

# !!for plot 2.2 :I do not follow the requirement "labeling the points with numbers enclosed in circles and providing a corresponding legend." because I think directly enclosing language name in a circle is also clear.

ggplot(plot3_data, aes(x = speaker_count, y = country_count, fill = category)) +
  geom_point(shape = 21, color = "grey", size = 8, stroke = 1.2) +
  geom_text(aes(label = language_name), size = 4, color = "black", fontface = "bold") +
  scale_fill_manual(
    values = c(
      "Both Top 10s" = "red",          
      "Only Spoken Top 10" = "yellow",    
      "Only Widespread Top 10" = "green" 
    ),
    name = "Language Category"
  ) +scale_x_log10(label=scales::comma)+scale_y_log10(label=scales::comma)+ #make data distribution more linear
  labs(
    title = "Top 10 Most Spoken vs. Most Widespread Languages",
    subtitle = "Red = both Top 10s | Blue = only speakers | Green = only countries",
    x = "Number of Speakers (log scale)",
    y = "Number of Countries (log scale)",
  ) +
  theme_minimal() +theme(legend.position = 'bottom')
```

Top-5 language_families:
```{r data_clean_Q3,echo=FALSE}

lang_path <- tibble(language=data$language_history$name,path=data$language_history$family_tree$path) %>% filter(!sapply(path,is.null)) %>% mutate(family=map_chr(path, ~ .x[2])) %>% select(language,family)


lang_speaker_num <- re_df %>% filter(language_name %in% lang_path$language) %>% select(language_name,speaker_count)

top5_families <- lang_path %>% left_join(lang_speaker_num,by=c("language"="language_name")) %>% drop_na() %>% group_by(family) %>% summarise(count=sum(speaker_count)) %>% arrange(desc(count)) %>% head(5)
top5_families

top10_speakers_per_country <- lang_path %>% left_join(lang_speaker_num,by=c("language"="language_name")) %>% drop_na() %>% group_by(family) %>% slice_max(speaker_count,n=10,with_ties = F) %>% filter(family %in% top5_families$family)
```
Top 5 Language Families and Their Most Spoken Languages:
```{r language_fam_visualization,echo=FALSE,fig.width=12,fig.height=8,out.width="100%"}
  
top10_speakers_per_country %>%
  mutate(Family = factor(family, levels = top5_families)) %>%
  ggplot(aes(x = reorder(language, speaker_count), 
             y = speaker_count, 
             fill = language)) +
  geom_col() +
  coord_flip() +
  facet_wrap(~family, scales = "free_y", nrow = 2, ncol = 3) +
  labs(
    title = "Top 5 Language Families and Their Most Spoken Languages",
    subtitle = "Each bar represents a language within the family",
    x = "",
    y = "Number of Speakers"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5, color = "gray50"),
    axis.text.y = element_text(size = 8),
    legend.position = "none",
    strip.text = element_text(size = 14, face = "bold", color = "white"),
    strip.background = element_rect(fill = "steelblue", color = NA),
    panel.spacing = unit(0.5, "lines")
  ) +
  scale_fill_viridis_d()

```



Data attribution:
This dataset is `world_languages_integrated.json` by Luke Steuber (MIT License)
Data sources: 
1. Glottolog 5.0 (Max Planck Institute for Evolutionary Anthropology) https://glottolog.org/
Geographic coordinates, language family classification, glottocodes, and macroareas (License: CC-BY-4.0) <https://creativecommons.org/licenses/by/4.0/>
2. Joshua Project <https://joshuaproject.net/>
Speaker population aggregates, Bible translation status, and religion data
Full license information <https://huggingface.co/datasets/lukeslp/world-languages/blob/main/LICENSE.md>
This project is not for commercial purposes but for educational purposes only.

