---
title: "Niclas38"
output: md_document
date: "2025-04-02"
---

# Solution for "Film genre and popularity analysis-Project" by @yhykelly

In this File I propose my solution for the "Film genre and popularity analysis-Project" by yhykelly.
It is about a movie data set and the task is to further analyse the data to find trends in the popularity and sucsess of a movie based on statistics and measuarable variables.

#Loading packages and R-setup
```{r, echo=FALSE}
library(tidyverse)
library(rstudioapi)
library(jsonlite)
library(lubridate)
library(ggridges)
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```
#Loading data
```{r, echo=FALSE}
Data <- read_csv("movies_metadata.csv")
```
# Data manipulation

## Data tidying:
- Select only needed columns
- Extract information from JSON Format
- Filter for duplicates & invalid values
- extracting the year from the release-date
```{r, echo=FALSE}
CleanData <- Data %>%
  select(id, title,release_date, production_countries,original_language,genres,vote_average,popularity,budget,revenue,runtime) %>% # select columns
  filter(runtime > 0,
         revenue > 0,
         budget > 0) %>% # filter for invalid values
  distinct() %>% #filter for duplicates
  mutate(genres = map(genres, ~fromJSON(str_replace_all(.x,"'", "\""))),
         genres = map(genres, ~pluck(.,"name"))) %>% 
  mutate(production_countries = map(production_countries, ~fromJSON(str_replace_all(.x,"'", "\""))),
         production_countries = map(production_countries, ~pluck(.,"name"))) %>% 
  unnest(genres) %>%
  unnest(production_countries) %>% #convert JSON formats in list formats and unnest the lists
  mutate(release_date = year(release_date)) # extract year from release date


```

## Data wrangling
- group by genres, year and language
- calculate mean budget, revenue, profit (= revenue - budget), vote_average and popularity for each
```{r, echo=FALSE}
#grouped by genre:
CleanData %>% 
  group_by(genres) %>% 
  summarise(Amount_Of_Movies = n_distinct(title),
            AvgBudget = mean(budget),
            AvgRevenue = mean(revenue),
            AvgProfit = mean(revenue - budget),
            AvgVote = mean(vote_average),
            AvgPopularity = mean(popularity)) %>% 
  knitr::kable()

#grouped by year:
CleanData %>% 
  group_by(release_date) %>% 
  summarise(Amount_Of_Movies = n_distinct(title),
            AvgBudget = mean(budget),
            AvgRevenue = mean(revenue),
            AvgProfit = mean(revenue - budget),
            AvgVote = mean(vote_average),
            AvgPopularity = mean(popularity)) %>% 
  knitr::kable()


#grouped by language:
CleanData %>% 
  group_by(original_language) %>% 
  summarise(Amount_Of_Movies = n_distinct(title),
            AvgBudget = mean(budget),
            AvgRevenue = mean(revenue),
            AvgProfit = mean(revenue - budget),
            AvgVote = mean(vote_average),
            AvgPopularity = mean(popularity)) %>% 
  knitr::kable()

```


# Visualization

## Line chart:
To make it less messy I only took the top 10 countries with the overall most movies and compared their amount of movies over the time.
```{r, echo=FALSE}
#Select top 10 countries
Top10Countries <- CleanData %>% 
  group_by(production_countries) %>% 
  summarise(amount = n_distinct(title)) %>% 
  arrange(desc(amount)) %>% 
  slice_head(n=10) %>% 
  ungroup()

#group data by year and genre and generate a lineplot for each (of the top 10) genre
CleanData %>%
  group_by(release_date,production_countries) %>% 
  summarise(amount = n_distinct(title)) %>%
  filter(production_countries %in% Top10Countries) %>% 
  ggplot(aes(x=release_date,y=amount,colour = production_countries))+
  geom_line()+
  labs(
    title = "Number of films by yer and country",
    x = "Year",
    y = "Amount of movies"
  )+
  theme_minimal()

```

## Ridgeline plot
- I used the "ggridges" package and geom_density_ridges to create the ridgeline plot
- I did the same again only for the top 10 genres to make it cleaner but it didn't change a lot 
```{r, echo=FALSE}
# Plot the amount of movies of each genre against the time in a ridgelineplot
CleanData %>%
  ggplot(aes(x = release_date, y=genres, fill = genres))+
  geom_density_ridges(scale=1, bandwidth = 7.5,na.rm=T)+
  labs(
    title = "number of genres through diffrent year",
    x = "Year",
    y = "Genre"
  )+
  theme_minimal()

# Trying the same for the top 10 genres -> maybe this looks a bit better?
# Define Top 10 Genres
Top10Genres <- CleanData %>% 
  group_by(genres) %>% 
  summarise(amount = n_distinct(title)) %>% 
  arrange(desc(amount)) %>% 
  slice_head(n=10) %>% 
  ungroup()

#Plot it
CleanData %>%
  filter(genres %in% Top10Genres$genres) %>% 
  ggplot(aes(x = release_date, y=genres, fill = genres))+
  geom_density_ridges(scale=1, bandwidth = 7.5,na.rm=T)+
  labs(
    title = "number of genres through diffrent year",
    x = "Year",
    y = "Genre"
  )+
  theme_minimal()
```

## Scatter plot
- Created a "WinLoss" column that shows if the movie was proftiable or lost money.
- Created a Scatterplot of budget and revenue and colorize the datapoints depending if the movie made profit or lost money
- Added a trendline -> All movies above it are more profitable than the average movie, all below it are less profitable
- Added the same plot again with same axis scaling -> shows the real ratio but looks a bit ugly (I was curious how it would look :))
```{r, echo =FALSE}
CleanData %>% 
  mutate(WinLoss = ifelse(revenue>budget, "Movie was profitable", "Movie lost money")) %>% #Create WinLoss column
  ggplot(aes(x=budget,y=revenue, colour = WinLoss))+ #plot scatterplot
  geom_point(size=0.75)+
  geom_smooth(method = "lm",col= "#480355")+ #add trendline
  labs(
    title="Correlation between revenue and budget",
    x = "Budget",
    y = "Revenue"
  )+
  theme_minimal()

#second plot with even axis scaling -> shows the real ratio between budget and revenue but is ugly -> just because i was curious :)

CleanData %>% 
  mutate(WinLoss = ifelse(revenue>budget, "Movie was profitable", "Movie lost money")) %>% #Create WinLoss column
  ggplot(aes(x=budget,y=revenue, colour = WinLoss))+ #plot scatterplot
  geom_point(size=0.75)+
  geom_smooth(method = "lm",col= "#480355")+ #add trendline
  labs(
    title="Correlation between revenue and budget",
    x = "Budget",
    y = "Revenue"
  )+
  coord_fixed()+
  theme_minimal()

```

