---
title: "Project Bernhard"
output:
  html_document:
    df_print: paged
---


# Tasks


```{r chunk options, include=FALSE}

knitr::opts_chunk$set (echo=FALSE, warning=FALSE, message=FALSE, dpi = 300, root.dir = "./Projects/FerSoe/Project2" )


```


Make a bar plot that shows the following:

*   Y-axis: the mean of Nmin mith the given unit
*   X-axis: the grouped numleg-values as shown in the exampel below
  *   the numleg-groups shall contain sub groups as shown in the exampel below
  *   the subgroups shall represent the grouped sowndiv-values
  *   the subgroups must be easily distinguishable (of different colors or symbology)
  *   each bar needs to have an error bar representing the standard error
  *   the symbology/colors of sowndiv shall be described in a legend box



# Methods

```{r include=FALSE}
library(tidyverse)

```


```{r}

read.csv("https://dr-eberle-zentrum.github.io/Advanced-data-processing-with-R/Projects/bernhard-ebersbach/nmin_jena_2022.csv",sep = ";") -> data_or

# data_or |> 

head(data_or,10)
```





# Solutions

## Select and rename columns

```{r}


data <- data_or |> 
  rename (Nmin = Nmin..mg.kg.) |> 
  select(plotcode, sowndiv, Nmin, numleg) 


head(data)
```
```{r}

data |> 
  count(sowndiv)

data |> 
  count(numleg)

```

## Grouping of variables

```{r}

# as factor
data$sowndiv <- as.factor(data$sowndiv)
data$numleg <- as.factor(data$numleg)

# recode sowndiv
levels(data$sowndiv)
data$sowndiv_cat <- data$sowndiv
levels(data$sowndiv_cat) <- c("0-1", "0-1", "2-4", "2-4", "8-16", "8-16", "60")
levels(data$sowndiv_cat)

# recode numleg
levels(data$numleg)
data$numleg_cat <- data$numleg
levels(data$numleg_cat) <- c("0-1", "0-1", "2-3", "2-3", "4-5", "4-5", "6-8", "6-8", "12")
levels(data$numleg_cat)

```
"0-1" = 1,
         "2-3" = 2,
         "2-3" = 3

+
  geom_errorbar(aes ())
  
  
```{r}
RColorBrewer::display.brewer.all()
```
  
  
```{r}

data |> 
  ggplot(aes(numleg_cat, Nmin, fill = sowndiv)) +
  geom_bar(stat = "identity", position = position_dodge())+
  scale_fill_brewer ("Diversity Level \n (snowdiv)") +
  theme(panel.background = element_blank(), panel.grid.major = element_line("grey"), panel.grid.major.x = element_blank())+
  labs(y = "Nmin [mg/kg]", x = "Number of legume species (numleg)") 

```

## Errorbars

```{r}
# Statistics




data_sd$sd <- apply(data,1, sd, na.rm=T)


```


```{r}

data_sd <- data

data_sd<-data_sd |> 
  group_by(numleg_cat) |> 
  summarise(sd = sd (Nmin))

df1  = transform(data, mean=rowMeans(data$Nmin), sd=apply(data$Nmin,1, sd))

df <- df %>%
  mutate(sd = transform(sd_by_level, factor_column, mean))

sd_by_level <- data %>%
  group_by(numleg_cat) %>%
  summarize(sd = sd(Nmin))

# Use transform to add the sd to the original dataframe
data_sd <- data %>%
  mutate(sd = transform(sd_by_level, numleg_cat, sd))


data_sd <- merge(data, sd_by_level, by = "numleg_cat")
```


  
```{r}



data |> 
  ggplot(aes(numleg_cat, Nmin, fill = sowndiv)) +
  geom_bar(stat = "identity", position = position_dodge())+
  scale_fill_brewer ("Diversity Level \n (snowdiv)") +
  theme(panel.background = element_blank(), panel.grid.major = element_line("grey"), panel.grid.major.x = element_blank())+
  labs(y = "Nmin [mg/kg]", x = "Number of legume species (numleg)") +
    geom_errorbar( aes(x=numleg_cat, ymin=Nmin-sd, ymax=Nmin+sd))


```

Merge sd of the groupings (numleg) with the main dataset.
