---
title: "Chaoicci"
output: md_document
date: "2026-06-05"
---

```{r libraries, include = FALSE}
library(readxl)
library(tidyverse)
library(ggplot2)
library(lme4)
library(lmerTest)
library(stdmod)
```


```{r data-import, include = FALSE}
# Task 1: Import data and prepare for analysis
data1 <- "data.xlsx" |> read_excel(1)
data2 <- "data.xlsx" |> read_excel(2)
```


```{r quality-index, include = FALSE}
# Task 2: Sum score for quality index before and after intervention
data_qis <- data1 |> 
  # there is one na value: ! we cannot sum with this!
  drop_na() |> 
  rowwise() |> 
  mutate(qi_pre = i1t0 + i2t0 + i3t0 + i4t0 + i5t0,
         qi_post = i1 + i2 + i3 + i4 + i5) |> 
  ungroup()
```

```{r sweets-mean, include = FALSE}
# Task 3: Calculate mean for sweets and merge
data_mean_sweets <- data2 |> 
  rowwise() |> 
  mutate(child_id = `id child`,
         age = `id age`,
         sweets_mean = mean(sweet1, sweet2, sweet3, sweet4)) |> 
  ungroup() |>  
  select(child_id, age, sweets_mean)

data_full <- full_join(data_qis, data_mean_sweets, by = "child_id") |> 
  mutate(Intervention = factor(Intervention, 
                               labels = c("Reality Check", "Sugar Shock")),
         child_id = factor(child_id))
```
# Task 1-3
After importing, calculating the quality index before and after the intervention,
calculating the mean of sweet 1-4 and merging the two tables, the dataset looks
like this:
```{r table, echo = FALSE}
knitr::kable(
  data_full[1:5, ],
  caption = "Dataset"
)
```


# Task 4: Comparison of the two interventions
```{r comparison-of-interventions, echo = FALSE}
# Task 4: Compare the two interventions in terms of their effect on the increase 
# in quality visually
data_full |> 
  group_by(Intervention, qi_pre, qi_post) |> 
  mutate(freq_linetype = n()) |> 
  ungroup() |> 
  ggplot(aes(color = freq_linetype, linewidth = freq_linetype))  + 
  geom_point(aes(x = 1, y = qi_pre)) +
  geom_point(aes(x = 2, y = qi_post)) +
  geom_segment(aes(x = 1, xend = 2, y = qi_pre, yend = qi_post)) +
  facet_wrap(~ Intervention) + 
  scale_color_gradient(high = "orange", low = "blue") +
  theme_minimal() +
  theme(legend.position = "none", 
        axis.ticks = element_blank(),
        panel.spacing.x = unit(2, "lines"),
        plot.margin = margin(10, 35, 10, 20)) +
  labs(x = "Intervention type",
       y = "Behavioral Indicator") +
  scale_x_continuous(breaks = c(1, 2),
                     labels = c("QIpre", "QIpost")) 
```

The individual changes before and after the intervention are plotted. Size and 
color of the lines indicate the frequency of the line.

# Task 5: Test whether motivation moderates the relationship between the intervention and the QIpost
Note: I added `child_id` as a random effect (as: `(1 | ID)`), since a random 
effect is needed for the `lmer()` function. 
```{r relationship-intervention-qi_post, warning = FALSE, echo = FALSE, message = FALSE}
# Task 5: Test whether motivation moderates relationship between intervention
# and QIpost
model <- lmer(qi_post ~  Intervention * Motivation + qi_pre + (1 | child_id), 
            data_full)
model |>  summary()
## Ich hoffe ich habe das richtig angepasst, ich weiß nicht was Moderation ist
## oder wie man das mit R macht

# test whether relationship between Intervention and QIpost is moderated by 
# motivation and control for initial Qpre
# outcome: qipost
# predictor: intervention
# moderator: motivation
# control for: qipre

```
There seems to be no significant interaction between the intervention and the
motivation (InterventionSugar Shock:Motivation : Est = 0.33717, SE = 0.35259, 
t = 0.956, Pr(>|t|) = 0.34096 > .05) and therefore no significant moderation 
effect of motivation on intervention.


# Task 6: Plot the moderation
```{r moderation-plot, echo = FALSE}
# Task 6: Plot Task 5
model |> 
  plotmod(x = Motivation, w = Intervention) +
  labs(y = "QIpost",
       title = "Moderation effect of motivation") +
  theme_classic() 

```

