---
title: "Chaoicci"
output: md_document
date: "2026-06-05"
---

```{r libraries, include = FALSE}
library(readxl)
library(tidyverse)
library(ggplot2)
library(lme4)
library(lmerTest)
library(stdmod)
```


```{r quality-index, include = FALSE}
# Task 2: Sum score for quality index before and after intervention
data_qis <- "data.xlsx" |> 
  read_excel(1) |> 
  # there is one na value: ! we cannot sum with this!
  drop_na() |> 
  rowwise() |> 
  mutate(qi_pre = matches("^i.t0$") |> 
           c_across() |> 
           sum(),
         qi_post = matches("^i.$") |> 
           c_across() |> 
           sum()) |>  
  ungroup()
```

```{r sweets-mean, include = FALSE}
# Task 3: Calculate mean for sweets and merge
data_mean_sweets <- "data.xlsx" |> 
  read_excel(2) |> 
  rowwise() |> 
  mutate(child_id = `id child`,
         age = `id age`,
         sweets_mean = matches("^sweet.$") |> c_across() |> mean()) |> 
  ungroup() |>  
  select(child_id, age, sweets_mean)

data_full <- full_join(data_qis, data_mean_sweets, by = "child_id") |> 
  mutate(intervention = factor(Intervention, 
                               labels = c("Reality Check", "Sugar Shock")),
         motivation = Motivation,
         helpfulness = Helpfulness,
         child_id = factor(child_id)) |> 
  select(-c(Intervention, Motivation, Helpfulness))
```
# Task 1-3
After importing, calculating the quality index before and after the intervention,
calculating the mean of sweet 1-4 and merging the two tables, the dataset looks
like this:
```{r table, echo = FALSE}
knitr::kable(
  data_full[1:5, ],
  caption = "Dataset"
)
```


# Task 4: Comparison of the two interventions
```{r comparison-of-interventions, echo = FALSE}
# Task 4: Compare the two interventions in terms of their effect on the increase 
# in quality visually
data_full |> 
  group_by(intervention, qi_pre, qi_post) |> 
  mutate(freq_linetype = n(),
         qi_incr_or_decr = (qi_post - qi_pre) |> 
           map_chr(\(x) if(x > 0) "incr" else if(x == 0) "same" else "decr") |>
           as.factor()) |> 
  ungroup() |> 
  ggplot(aes(linewidth = freq_linetype, col = qi_incr_or_decr))  + 
  geom_point(aes(x = 1, y = qi_pre)) +
  geom_point(aes(x = 2, y = qi_post)) +
  geom_segment(aes(x = 1, xend = 2, y = qi_pre, yend = qi_post)) +
  facet_wrap(~ intervention) + 
  theme_minimal() +
  theme(legend.position = "none", 
        axis.ticks = element_blank(),
        panel.spacing.x = unit(2, "lines"),
        plot.margin = margin(10, 35, 10, 20),
        plot.caption = element_text(hjust = 0.5)) +
  labs(x = "Intervention type",
       y = "Behavioral Indicator",
       caption = "The individual changes before and after the intervention are plotted.
       Size of the lines indicates the frequency of the line, color indicates whether
       the line is increasing (green), stable (red) or decreasing (blue).") +
  scale_x_continuous(breaks = c(1, 2),
                     labels = c("QIpre", "QIpost")) 
  
```




```{r change-qi_pre-qi_post, echo = FALSE}
# Martin's idea
data_full |> 
  mutate(qi_diff = qi_post - qi_pre) |> 
  ggplot(aes(x = qi_diff)) + 
  geom_bar() +
  facet_wrap(~ intervention) + 
  labs(title = "Change between QIpre and QIpost", 
       subtitle = str_c("with \U03BC = ", (data_full$qi_post - data_full$qi_pre) |>
                          mean() |> 
                          round(2), 
                        " and \U03C3 = ", (data_full$qi_post - data_full$qi_pre) |> 
                          sd() |> 
                          round(2)),
       x = "Difference between QIpost and QIpre",
       y = "Count") +
  theme_bw()
```


# Task 5: Test whether motivation moderates the relationship between the intervention and the QIpost
Note: I added `child_id` as a random effect (as: `(1 | ID)`), since a random 
effect is needed for the `lmer()` function. 

The summary of the model looks like this:
```{r relationship-intervention-qi_post, warning = FALSE, echo = FALSE, message = FALSE}
# Task 5: Test whether motivation moderates relationship between intervention
# and QIpost
model <- lmer(qi_post ~  intervention * motivation + qi_pre + (1 | child_id), 
            data_full)
model |>  summary()
## Ich hoffe ich habe das richtig angepasst, ich weiß nicht was Moderation ist
## oder wie man das mit R macht

# test whether relationship between Intervention and QIpost is moderated by 
# motivation and control for initial Qpre
# outcome: qipost
# predictor: intervention
# moderator: motivation
# control for: qipre

```
There seems to be no significant interaction between the intervention and the
motivation (InterventionSugar Shock:Motivation : Est = 0.33717, SE = 0.35259, 
t = 0.956, Pr(>|t|) = 0.34096 > .05) and therefore no significant moderation 
effect of motivation on intervention.


# Task 6: Plot the moderation
```{r moderation-plot, echo = FALSE}
# Task 6: Plot Task 5
model |> 
  plotmod(x = motivation, w = intervention) +
  labs(y = "QIpost",
       title = "Moderation effect of motivation") +
  theme_classic() 

```

