---
title: "Wenjing_solution"
author: "Wenjing"
date: "2026-06-30"
output: md_document
---
# Topic: The impact of various deterrents on children’s consumption of sweets-from Lea

```{r setup, echo=FALSE}
knitr::opts_chunk$set(include = TRUE, warning = FALSE, message = FALSE)

#install.packages("readr")
#install.packages("lme4")
#install.packages("stdmod")
library(readxl)
library(tidyverse)
library(dplyr)
library(lme4)
library(stdmod)
library(knitr)

```

## Load the file (btw. remove the missing values)

### quick overview of the data

1.The file consists of two sheets: "Intervention data" and "questionaire". 

2.Each child was assigned to one of **two interventions**: "Reality Check" or "Sugar Shock".(in dataset we use "1" for "Reality Check" and "2" for "Sugar Shock"). 

3.Each type experiment was conducted in **three rounds**, with a pre- and post-intervention measurement of the child's consumption of sweets. The pre-intervention measurement is represented by the columns "i1t0" to "i5t0", while the post-intervention measurement is represented by the columns "i1" to "i5".

```{r Data_import}
file.exists("data.xlsx")


Intervention_data <- read_excel("data.xlsx",  sheet = "Intervention data") %>% drop_na()
#revise: "glimpse()" is better to preview the data.
glimpse(Intervention_data)


Q_data <- read_excel("data.xlsx", sheet = "questionaire") %>% drop_na()
glimpse(Q_data)

```

## Data cleaning
- Calculating the total value before intervention `QIpre`
- Calculating the total value before intervention `QIpost`
 - using `across()` and `c()` to select the columns in a more "tidyverse" style
```{r QIpre_and_QIpost }

#Intervention_data$QIpre <- rowSums(Intervention_data[,c("i1t0","i2t0","i3t0","i4t0","i5t0")])
#Intervention_data$QIpost <- rowSums(Intervention_data[,c("i1","i2","i3","i4","i5")])
#revise: in a more "tidyverse" style

Intervention_data <- Intervention_data %>% mutate(QIpre=rowSums(across(c(i1t0:i5t0))),QIpost=rowSums(across(c(i1:i5))))
glimpse(Intervention_data)

```

- Calculating mean of sweet to get children's attitude toward sweet consumption, and then merge it with Intervention_data by "child_id" column.
```{r data_calculation}
# Calculating mean of sweet
df_withmean <- Q_data %>% 
  mutate(sweet_mean=pmap_dbl(select(.,sweet1,sweet2,sweet3,sweet4),~mean(c(...)))) %>%
  select("id child","id age",sweet_mean) %>%
  rename("child_id"="id child") %>%
  right_join(Intervention_data,by="child_id")
df_withmean

```

## Data visualization
1. To compare the effect of two interventions, we make the plot where each line represents one participant, connecting their pre score to post score. We need two datasets: 

- the individual QIpre & QIpost scores

```{r plot1_data_t}

plot_data <-df_withmean  %>% select(child_id, Intervention, QIpre,QIpost) %>%group_by(child_id, Intervention) %>%
mutate(trial_id = row_number()) %>%
ungroup() %>%
mutate(pair_id = paste(child_id, Intervention, trial_id, sep = "_")) %>% pivot_longer(cols = c(QIpre,QIpost),names_to = "time",values_to = "score") %>% mutate(
    time = factor(time, 
                  levels = c("QIpre","QIpost"))
  )  
glimpse(plot_data)

# here I keep plot_data, because i need to used it to create "summary_data". and they will be used separately in the next step.
```

- the mean score of all participants in each intervention group to present tendency
```{r plot1_data_summarise}
 
summary_data <- plot_data %>% group_by(child_id,Intervention,time) %>% summarise(mean=mean(score,na.rm=TRUE), .groups = "drop") 
glimpse(summary_data)
```
  
From plot, we can conclude that there is no obvious difference between two interventions, but the "Suger shock" intervention seems to have a slightly better effect on reducing children's consumption of sweets. 
```{r Data_Visualization}
ggplot() +
  geom_line(data=plot_data,alpha = 0.2, linewidth = 1.8,aes(x=time,y=score,group = pair_id)) +
  geom_point(alpha = 0.4, size = 1.2) +geom_line(data=summary_data,aes(x=time,y=mean,group=child_id,colour=factor(child_id)),linewidth=1.5)+geom_point(data=summary_data,aes(x=time,y=mean),color="red",size=2)+facet_wrap(~ Intervention,
    ncol = 2,
    labeller = labeller(Intervention = c(`1` = "Reality Check",
                                         `2` = "Sugar Shock"))
  ) +
  scale_y_continuous(breaks = 0:5, limits = c(0, 5)) +scale_color_viridis_d()+
  labs(x = "Intervention type", y = "Behavioral Indicator",legend="Child Id") +
  theme_minimal(base_size = 13) +
  theme(
    strip.text = element_text(hjust = 0.5, face = "bold"),
    legend.position = "bottom"
  )
```
2. We need to make a plot to explore whether **motivation** moderates the relationship between the **intervention** and the QIpost in quality (for each type) and control for the initial QIpre.

We should use `model2 <- lmer(QIpost ~ intervention_type * motivaton + QIpre, data = data)`. Because

`lmer() function` is used to fit linear mixed-effects models, which are useful when you have data that is grouped or clustered, such as **repeated measurements on the same subjects**. In this case, we have repeated measurements of QIpost for each child, and we want to account for the variability between children.

The random intercept variance was near 0 which indicates minimal between participants variablity after controlling baseline quality. 
```{r plot_data_sencond_plot }

model_lmer <- lmer(QIpost ~ Intervention * Motivation + QIpre + (1 | child_id), data = df_withmean)
summary(model_lmer)


model <- lm(QIpost ~ Intervention * Motivation + QIpre, data = df_withmean) 
summary(model)


#use a linear plot with the motivation on the x-Axis, QIpost quality on the y-axis and one line (colour) for each Intervention #type. It should look like an interaction plot in the end


model %>% plotmod(x=Motivation,w=Intervention)+labs(y="QIpost quality",title="Motivation effect")





```
Conclusion: The interaction between intervention type and motivation was not significant ( p = 0.341), suggesting that motivation did not significantly moderate the relationship between intervention and QIpost.


