---
title: "Project_description"
author: "Raseswari Mondal Supti"
date: "2025-05-13"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Comparison of salaries of global remote data scientists!

Remote working became a norm during covid-19. And with remote work came more concern about salary. Let's see using the [data](https://github.com/Dr-Eberle-Zentrum/Data-projects-with-R-and-GitHub/blob/main/Projects/Supti7/Salaries_data_scientists.csv) from [kaggle](https://www.kaggle.com/datasets/abhinavshaw09/data-science-job-salaries-2024?resource=download) whether there is any difference in salary of remote data scientists in different countries.

Here is a first look at the data:

```{r, data}
data <- read.csv("Salaries_data_scientists.csv")
knitr::kable(head(data))
```


A short description of some of the variables are given below:

- __work_year__ : The year the salary was paid.
- __experience_level__ : The experience level in the job during the year: EN -  Entry-level / MI - Mid-level / SE - Senior-level / EX - Executive-level / Director.
- __job_title__ : The role worked in during the year.
- __salary__ : The total gross salary amount paid.
- __remote_ratio__ : The overall amount of work done remotely, where the values are as follows: 0 no remote work (less than 20%),
50 partially remote, 100 fully remote (more than 80%).
- __company_location__ : The country of the employer's main office [as an ISO 3166 country code](https://www.iso.org/iso-3166-country-codes.html)

## Data Peperation tips

- Use the data for 2024 only.
- For the missing data, whether it is regarding the inclusion of countries for salary or remote work ratio, you can choose to remove the rows. Or you can choose to fill the missing values with the mean or median of the respective column. Trust your judgement.
- If there is any redundancy for the job title, meaning same job with different recording style (e.g. "Cybersecurity Engineer" and "Cyber Security Engineer"), please unify them.


## Visualization Goal of the project

The goal of this project is to summarize the visual information primarily.The purpose is to see the regional differences in remote ratio and salaries.Use __choropleth__ maps to visualize the data.

-   Show first the percentage of the remote work ratio. Choose one type e.g. 100% or 50% remote work ratio and generate a graph that shows the the total number of jobs among different countries globally using that ratio. It might be helpful to use the percentage for which you have the most available data.
- For this first plot, use all types of job title as the focus is on the remote work ratio.

-   The second visualization task you can choose either of the two tasks:
    - Show the average salary of top 5 jobs with entry level and/or mid level.
    - Generate two more choropleth or violin plots for North America (NA) and Europe (EU) with the percentage you have chosen for the first graph. If you have confusion regarding countries like Russia, please exclude them.


## Ideas for further work

- Do a hypothesis test to compare the average salary difference between USA and Non-USA region. In order to do that, first you need to sample jobs with same title but different salary from this two regions.Here, you can choose the jobs that have a significant number of jobs in both regions, e.g. Data Scientist, Data/ML Engineer, etc.
- The statistical test to do so, I will leave it to you.

## Quesitons of interest with this project

- With the first task, we want to see, globally what is the number of jobs among different countries that have a certain remote work ratio. This will help us understand the __global trend of remote work__.

- The second task will help us understand the __salary differences among different job titles and experience levels__.

- The plots for __NA__ and __EU__ will help us understand the __regional differences in remote work ratio__ as the number of jobs is quite large in these regions.

- The hypothesis test will help us understand if there is a __significant difference in salary between USA and Non-USA regions__ for the same job title.

Have fun and looking forward to seeing your work!

Kind regards, Supti
