---
title: "Svenjasinitialprojectsolution.rmd"
author: "SvenjaPoerstel"
date: "2025-06-30"
output: md_document
---

```{r}
library(tidyverse) 
library(ggplot2) 
library(readr) 
library(dplyr) 
```

```{r}
# read csv
ms <- read_csv2("data_MS.csv", show_col_types = FALSE)
```

##Data manipulation

```{r}

# (1) MS duration in months (+++), created a new column but didn't delete the old one
ms$MSDuration_months <- case_when(
  grepl("Years", ms$`MS Duration`) ~ as.integer(gsub("Years", "", ms$`MS Duration`)) * 12,
  grepl("Year", ms$`MS Duration`) ~ as.integer(gsub("Year", "", ms$`MS Duration`)) * 12,
  grepl("Months?", ms$`MS Duration`) ~ as.integer(gsub("Months", "", ms$`MS Duration`)),
  TRUE ~ NA_integer_
)

# (2) new column related to diseases 

ms$Other_Diseases_Count <- ifelse(
  ms$`Other Diseases` == "-" | is.na(ms$`Other Diseases`) | trimws(ms$`Other Diseases`) == "",
  0,
  sapply(ms$`Other Diseases`, function(x) {
    # Anzahl Kommas zählen
    n_commas <- lengths(regmatches(x, gregexpr(",", x)))
    # Anzahl Krankheiten = Kommas + 1
    n_commas + 1
  })
)

# (3) new columns (treated with Corticosteroid)
ms$Corticosteroid <- ifelse(str_detect(ms$Medication, regex("corticosteroid", ignore_case = TRUE)), "Yes", "No")

# (4) new columns (treated with B-cell depleting therapy (+)
ms$BCellDepletion <- ifelse(str_detect(ms$Medication, regex("rituximab", ignore_case = TRUE)), "Yes", "No")

# (5) use coded names according to the schema instead of clear names and delete column names
ms <- ms %>% 
  mutate(
    PatientCode = sprintf("MS_XS_%03d", as.integer(as.character(`No.`)))
  ) %>% 
  select(-`Patient Name`)
print(ms)
```

## Visualization

# Scatterplot

```{r}
ggplot(ms, aes(x = `Cortisol Test Result`, y = `BAI Score`, color = Sex, shape = Corticosteroid)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = FALSE,
              aes(color = Sex, linetype = Corticosteroid, group = interaction(Sex, Corticosteroid))) +
  labs(title = "BAI Score vs Cortisol Level",
       x = "Cortisol (ng/mL)",
       y = "BAI Score") +
  theme_minimal()
```  
  
```{r}
# correlation value for four groups f+, f-, m+, m-
cor_groups <- ms %>%
  group_by(Sex, Corticosteroid) %>%
  summarise(Correlation = cor(`Cortisol Test Result`, `BAI Score`, use = "complete.obs"),
            .groups = "drop")
print(cor_groups)
```

# chart showing

```{r}
# Group MS duration into 6-month intervals
ms$MSDurationGroup <- cut(ms$MSDuration_months, 
                          breaks = seq(0, max(ms$MSDuration_months, na.rm = TRUE) + 6, by = 6), 
                          right = FALSE)

```

# Calculate mean and standard deviation of NO test values for each MS duration group
ms %>%
  group_by(MSDurationGroup) %>%
  summarise(mean_NO = mean(`NO test result`, na.rm = TRUE),
            sd_NO = sd(`NO test result`, na.rm = TRUE)) %>%
  ggplot(aes(x = `MS Duration`, y = mean_NO)) +
  geom_point(size = 3, color = "darkblue") +
  geom_errorbar(aes(ymin = mean_NO - sd_NO, ymax = mean_NO + sd_NO), width = 0.2) +
  labs(title = "NO Test Results by MS Duration (6-Month Groups)",
       x = "MS Duration Group",
       y = "Average NO Test Value") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
all_groups <- c("0–6", "6–12", "12–18", "18–24", "24–30", "30–36", "36–42", "42–48", "48–54", "54–60")

# 2. Mittelwerte und SD berechnen
summary_data <- ms %>%
  group_by(MSDurationGroup) %>%
  summarise(
    mean_NO = mean(`NO test result`, na.rm = TRUE),
    sd_NO = sd(`NO test result`, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  # 3. Fehlende Gruppen auffüllen
  complete(MSDurationGroup = all_groups)

# 4. Plot
ggplot(summary_data, aes(x = MSDurationGroup, y = mean_NO)) +
  geom_point(size = 3, color = "darkblue", na.rm = TRUE) +
  geom_errorbar(aes(ymin = mean_NO - sd_NO, ymax = mean_NO + sd_NO),
                width = 0.2, color = "darkblue", na.rm = TRUE) +
  labs(title = "NO Test Results by MS Duration (6-Month Groups)",
       x = "MS Duration Group",
       y = "Average NO Test Value") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```