---
title: "MS Patient Data Analysis"
author: "Hamed"
date: "`r Sys.Date()`"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(stringr)
```

```{r load-data}
data <- read.csv2("~/Data-projects-with-R-and-Github/Projects/flomaili/data_MS.csv")
```

### Clean and Prepare Data

```{r clean-data}
data <- data %>%
  mutate(
    ms_duration_clean = str_to_lower(str_replace_all(`MS.Duration`, " ", "")),
    ms_duration_value = as.numeric(str_extract(ms_duration_clean, "\\d+")),
    ms_duration_unit = case_when(
      str_detect(ms_duration_clean, "year") ~ "year",
      str_detect(ms_duration_clean, "month") ~ "month",
      TRUE ~ NA_character_
    ),
    `MS Duration (months)` = case_when(
      ms_duration_unit == "year" ~ ms_duration_value * 12,
      ms_duration_unit == "month" ~ ms_duration_value,
      TRUE ~ NA_real_
    ) %>% round()
  )

data <- data %>%
  mutate(
    other_disease_clean = na_if(trimws(`Other.Diseases`), "-"),
    Num_Other_Diseases = case_when(
      is.na(other_disease_clean) ~ 0L,
      TRUE ~ str_count(other_disease_clean, ",") + 1L
    )
  )

data <- data %>%
  mutate(
    medication_clean = str_to_lower(trimws(Medication)),
    Corticosteroid_Treatment = str_detect(medication_clean, "corticosteroid"),
    Bcell_Treatment = str_detect(medication_clean, "rituximab")
  )

data <- data %>%
  mutate(Patient_ID = sprintf("MS_XS_%03d", row_number())) %>%
  select(-"Patient.Name", everything())
```

### Cortisol vs. BAI Score

```{r plot-cortisol-bai}
ggplot(data, aes(x = `Cortisol.Test.Result`, y = `BAI.Score`, color = Sex, shape = Corticosteroid_Treatment)) +
  geom_point(size = 3) +
  theme_minimal() +
  labs(title = "BAI Score vs Cortisol Test Result", x = "Cortisol", y = "BAI Score")
```

### Correlation by Sex and Corticosteroid Treatment

```{r correlation-plot}
data <- data %>%
  filter(!is.na(Sex)) %>%
  mutate(Sex_Cort = paste0(str_trim(Sex), "_", ifelse(Corticosteroid_Treatment, "+", "-")))

ggplot(data, aes(x = `Cortisol.Test.Result`, y = `BAI.Score`, color = Sex_Cort)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~Sex_Cort) +
  theme_minimal() +
  labs(title = "Correlation per Sex & Corticosteroid Group")

correlations <- data %>%
  group_by(Sex_Cort) %>%
  summarise(correlation = cor(`Cortisol.Test.Result`, `BAI.Score`, use = "complete.obs"))
correlations
```

### NO Test Results by MS Duration Group

```{r no-ms-duration}
data <- data %>%
  mutate(MS_Group = cut(`MS Duration (months)`, breaks = seq(0, max(`MS Duration (months)`, na.rm = TRUE) + 6, 6), right = FALSE))

summary_ms <- data %>%
  group_by(MS_Group) %>%
  summarise(mean_NO = mean(`NO.test.result`, na.rm = TRUE),
            sd_NO = sd(`NO.test.result`, na.rm = TRUE))

ggplot(summary_ms, aes(x = MS_Group, y = mean_NO)) +
  geom_point() +
  geom_errorbar(aes(ymin = mean_NO - sd_NO, ymax = mean_NO + sd_NO), width = 0.2) +
  theme_minimal() +
  labs(title = "Mean NO Test Result by MS Duration Group", x = "MS Duration Group (months)", y = "NO Test Result")
```

### Median Number of Diseases by Age Group

```{r diseases-by-age}
data <- data %>%
  mutate(Total_Diseases = Num_Other_Diseases + 1,
         Age_Group = cut(Age, breaks = seq(0, 100, by = 15), right = FALSE))

summary_age <- data %>%
  group_by(Age_Group) %>%
  summarise(median_diseases = median(Total_Diseases, na.rm = TRUE))

ggplot(summary_age, aes(x = Age_Group, y = median_diseases)) +
  geom_col(fill = "skyblue") +
  theme_minimal() +
  labs(title = "Median Number of Diseases by Age Group", x = "Age Group", y = "Median Disease Count")

