---
title: "Danas attempt for Mihai's project"
author: "Dana Jabari"
date: "23.01.2023"
output: html_document
---

# Hey Mihai,

my last project will be yours :)

```{r, include = FALSE}

library(dplyr)
library(ggplot2)

```

1. First task: Bar plots illustrating the CO2 uptake means (over the conc condition) 
for type and treatment groups, together with their respective standard errors.

  * My struggle: How to make the error bar in the middle?


```{r, include = FALSE}

mean_uptake <- CO2 |> 
  group_by(Type, Treatment, conc) |> 
  summarize(means = mean(uptake), SE = sd(uptake) / sqrt(length(uptake)), .groups = "drop") |>
  ungroup()

```


```{r, echo = FALSE, results = "hide" }

mean_uptake |> 
  ggplot() +
  geom_bar(mapping = aes(x = conc, y = means, fill = Treatment), 
           stat = "identity", position = "dodge") +
  geom_errorbar(aes(x = conc,  ymin = means - SE, ymax = means + SE),
               width=0.4, colour="black", alpha=0.9) +
  facet_wrap(~ Type)

```

2. Line plots for each plant's CO2 uptake over the conc condition (x-Axis) for all plants from Quebec.
  * This kind of worked? At first the left axis the numbers were overlaying and I could not
    figure out, how to fix this. So I removed them. If you have a tip, I would love to know!

```{r, echo = FALSE, results = "hide"}
CO2 |> 
  filter(Type == "Quebec") |> 
  ggplot(CO2, mapping = aes(x = conc, y = uptake)) +
  geom_line() +
  facet_grid(Plant ~ Type) + 
  labs(title = "Plants of Quebec") +
  scale_y_discrete(guide=guide_axis(n.dodge=3))
  theme(axis.text.y.left =element_text(vjust = 0.5,
                             hjust = 0.5),
        axis.text = element_text(angle = 90)
        )
             
```

3. Formulating Hypothesises.
    * H0: Concentration level has no influence on the CO2 uptake of the plants.
    * H1: Concentration leven has an influence on the CO2 uptake of the plants.

```{r, echo = FALSE}

summary(aov(uptake ~ conc, CO2))

```

With p < .001 H0 will be abolished. 
This means, that concentration has a significant 
impact on the C02 uptake of the plants.


Thank you for your project!

Best, 
Dana
