---
 title: "Solution for Miguel"
 author: "Martin"
 output: md_document
---

```{r setup, echo=FALSE, show=FALSE}

#load without warnings
suppressMessages(library(tidyverse))

# set working directory to source file location
library(this.path)
setwd(this.path::here())
```


I loaded the data from the GitHub repository in one go and all data files into one table with respective animal annotations.

```{r import}

data <-
  # generate URLs to download
  str_c("https://github.com/Dr-Eberle-Zentrum/Data-projects-with-R-and-GitHub/raw/main/Projects/MiguelDLM/",
        "smooth_stress_tensor ("
        ,c("bear","hyaena","lion","wolf"),
        ").csv") |>
  # deal with blanks etc. in URLs
  URLencode() |>
  # read data
  # read_csv("https://github.com/Dr-Eberle-Zentrum/Data-projects-with-R-and-GitHub/raw/main/Projects/MiguelDLM/smooth_stress_tensor.csv") |>
  read_csv(id="animal") |>
  # simplify stress column name
  rename( stress = `Von Misses Stress`) |>
  # extract animal from url
  mutate( animal = str_extract(animal, "(?<=\\()[a-z]+(?=\\))") )
```

Given that, I rescaled removed the top 2% of stress values and rescaled the coordinates to a range from 0 to 1. I then segmented the data into 10 segments per dimension via rounding and dropping the maximal value per dimension (eventually extending the dropped 2% by one more), to ensure 10 segments.

```{r segment}
dataSegmented <-
  data |>
  group_by(animal) |>
  # remove 2% highest stress values
  filter( stress <= quantile(data$stress, 0.98)) |>
  # compute relative coordinate of X,Y,Z in range 0 to 1
  mutate( across(c(X,Y,Z), scales::rescale) ) |>
  # round to 1 decimal to get 10 segments
  # (but produces 11 segments due to rescale [0,1])
  mutate( across(c(X,Y,Z), ~floor(.x*10)/10) ) |>
  # drop value 1 to ensure 10 segments
  filter(X<1, Y<1, Z<1) |>
  ungroup()
```

Finally, I aggregated the mean stress data for each dimension and segment in one data frame and plotted the results.

```{r plot}
# aggregate mean stress data for each dimension in one data frame
dataSegmented |>
  # store the stress level per dimension and coordinate
  pivot_longer(cols=c(X,Y,Z), names_to="dimension", values_to="coord") |>
  # aggregate respective mean values
  group_by(animal, dimension, coord) |>
  summarize( stress = mean(stress), .groups="drop") |>
  # plotting
  ggplot(aes(col=animal)) +
  # per coordinate dimension
  facet_wrap(~dimension) +
  geom_point(aes(x=coord, y=stress), size=2) +
  geom_line(aes(x=coord, y=stress)) +
  labs(title="Mean stress in 10 segments of the model",
       x="Relative coordinate",
       y="Mean stress")
```

Now we should start beautifying the plot...



