---
title: "MiguelsProject_byMeri"
author: "Meri Nehlsen"
date: "2024-06-10"
output: md_document
---
# Solution to Miguel's Project by Meri
```{r setup}
library(tidyverse)
library(ggplot2)
library(gridExtra)
```

Importing the datasets:
```{r DataImport}
smoothstress_bear <- read_csv("https://github.com/Dr-Eberle-Zentrum/Data-projects-with-R-and-GitHub/raw/main/Projects/MiguelDLM/smooth_stress_tensor%20(bear).csv")
smoothstress_lion <- read_csv("https://github.com/Dr-Eberle-Zentrum/Data-projects-with-R-and-GitHub/raw/main/Projects/MiguelDLM/smooth_stress_tensor%20(lion).csv")
smoothstress_wolf <- read_csv("https://github.com/Dr-Eberle-Zentrum/Data-projects-with-R-and-GitHub/raw/main/Projects/MiguelDLM/smooth_stress_tensor%20(wolf).csv")
smoothstress_hyaena <- read_csv("https://github.com/Dr-Eberle-Zentrum/Data-projects-with-R-and-GitHub/raw/main/Projects/MiguelDLM/smooth_stress_tensor%20(hyaena).csv")
```

Next I created the relative intervals across the length of the respective axises. (For the data wrangling steps I should probably write functions (which I am not super comfortable with) so for now, this also works).

```{r IntervalCreation}
bear_xbreaks <- c(seq(min(smoothstress_bear$X), max(smoothstress_bear$X), length.out = 11))
bear_ybreaks <- c(seq(min(smoothstress_bear$Y), max(smoothstress_bear$Y), length.out = 11))
bear_zbreaks <- c(seq(min(smoothstress_bear$Z), max(smoothstress_bear$Z), length.out = 11))

lion_xbreaks <- c(seq(min(smoothstress_lion$X), max(smoothstress_lion$X), length.out = 11))
lion_ybreaks <- c(seq(min(smoothstress_lion$Y), max(smoothstress_lion$Y), length.out = 11))
lion_zbreaks <- c(seq(min(smoothstress_lion$Z), max(smoothstress_lion$Z), length.out = 11))

wolf_xbreaks <- c(seq(min(smoothstress_wolf$X), max(smoothstress_wolf$X), length.out = 11))
wolf_ybreaks <- c(seq(min(smoothstress_wolf$Y), max(smoothstress_wolf$Y), length.out = 11))
wolf_zbreaks <- c(seq(min(smoothstress_wolf$Z), max(smoothstress_wolf$Z), length.out = 11))

hyaena_xbreaks <- c(seq(min(smoothstress_hyaena$X), max(smoothstress_hyaena$X), length.out = 11))
hyaena_ybreaks <- c(seq(min(smoothstress_hyaena$Y), max(smoothstress_hyaena$Y), length.out = 11))
hyaena_zbreaks <- c(seq(min(smoothstress_hyaena$Z), max(smoothstress_hyaena$Z), length.out = 11))
```

Next I am calculating the mean Stress across the intervalls.
```{r MeanStress}
bear_x <- smoothstress_bear %>% 
  mutate(breaks = cut(X, bear_xbreaks)) %>% 
  group_by(breaks) %>% 
  summarise(MeanStress = mean(`Von Misses Stress`))
bear_y <- smoothstress_bear %>% 
  mutate(breaks = cut(Y, bear_ybreaks)) %>% 
  group_by(breaks) %>% 
  summarise(MeanStress = mean(`Von Misses Stress`))
bear_z <- smoothstress_bear %>% 
  mutate(breaks = cut(Z, bear_zbreaks)) %>% 
  group_by(breaks) %>% 
  summarise(MeanStress = mean(`Von Misses Stress`))

lion_x <- smoothstress_lion %>% 
  mutate(breaks = cut(X, lion_xbreaks)) %>% 
  group_by(breaks) %>% 
  summarise(MeanStress = mean(`Von Misses Stress`))
lion_y <- smoothstress_lion %>% 
  mutate(breaks = cut(Y, lion_ybreaks)) %>% 
  group_by(breaks) %>% 
  summarise(MeanStress = mean(`Von Misses Stress`))
lion_z <- smoothstress_lion %>% 
  mutate(breaks = cut(Z, lion_zbreaks)) %>% 
  group_by(breaks) %>% 
  summarise(MeanStress = mean(`Von Misses Stress`))

wolf_x <- smoothstress_wolf %>% 
  mutate(breaks = cut(X, wolf_xbreaks)) %>% 
  group_by(breaks) %>% 
  summarise(MeanStress = mean(`Von Misses Stress`))
wolf_y <- smoothstress_wolf %>% 
  mutate(breaks = cut(Y, wolf_ybreaks)) %>% 
  group_by(breaks) %>% 
  summarise(MeanStress = mean(`Von Misses Stress`))
wolf_z <- smoothstress_wolf %>% 
  mutate(breaks = cut(Z, wolf_zbreaks)) %>% 
  group_by(breaks) %>% 
  summarise(MeanStress = mean(`Von Misses Stress`))

hyaena_x <- smoothstress_hyaena %>% 
  mutate(breaks = cut(X, hyaena_xbreaks)) %>% 
  group_by(breaks) %>% 
  summarise(MeanStress = mean(`Von Misses Stress`))
hyaena_y <- smoothstress_hyaena %>% 
  mutate(breaks = cut(Y, hyaena_ybreaks)) %>% 
  group_by(breaks) %>% 
  summarise(MeanStress = mean(`Von Misses Stress`))
hyaena_z <- smoothstress_hyaena %>% 
  mutate(breaks = cut(Z, hyaena_zbreaks)) %>% 
  group_by(breaks) %>% 
  summarise(MeanStress = mean(`Von Misses Stress`))
```

Here I am combining the data of the different animals per dimension into one dataframe.
```{r DimensionDataframes}
x_stress <- data.frame(RelativeCoordinate = c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1),
                       stress_bear = bear_x$MeanStress,
                       stress_lion = lion_x$MeanStress,
                       stress_wolf = wolf_x$MeanStress,
                       stress_hyaena = hyaena_x$MeanStress)
x_stress_long <- x_stress %>% 
  pivot_longer(cols = starts_with("stress_"),
               names_to = "animal",
               values_to = "MeanStress") %>% 
  mutate(animal = str_replace(animal, "stress_",""))

y_stress <- data.frame(RelativeCoordinate = c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1),
                       stress_bear = bear_y$MeanStress,
                       stress_lion = lion_y$MeanStress,
                       stress_wolf = wolf_y$MeanStress,
                       stress_hyaena = hyaena_y$MeanStress)
y_stress_long <- y_stress %>% 
  pivot_longer(cols = starts_with("stress_"),
               names_to = "animal",
               values_to = "MeanStress") %>% 
  mutate(animal = str_replace(animal, "stress_",""))

z_stress <- data.frame(RelativeCoordinate = c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1),
                       stress_bear = bear_z$MeanStress,
                       stress_lion = lion_z$MeanStress,
                       stress_wolf = wolf_z$MeanStress,
                       stress_hyaena = hyaena_z$MeanStress)
z_stress_long <- z_stress %>% 
  pivot_longer(cols = starts_with("stress_"),
               names_to = "animal",
               values_to = "MeanStress") %>% 
  mutate(animal = str_replace(animal, "stress_",""))
```

This was my first attempt at plotting, but the facet_grid works much better.

```{r GridPlotting}
px <- ggplot(x_stress_long, aes(x=RelativeCoordinate, y = MeanStress, color = animal))+
  geom_point()+
  geom_line()+
  theme_classic()+
  xlab("Relative Coordinate")+
  ylab("Mean van Misses Stress")+
  scale_color_manual(values = c(
    "bear" = "brown",
    "hyaena" = "#E15634",
    "lion" ="#FFBF46",
    "wolf" = "grey"),
  guide ="none")

py <- ggplot(x_stress_long, aes(x=RelativeCoordinate, y = MeanStress, color = animal))+
  geom_point()+
  geom_line()+
  theme_classic()+
  xlab("Relative Coordinate")+
  ylab("Mean van Misses Stress")+
  scale_color_manual(values = c(
    "bear" = "brown",
    "hyaena" = "#E15634",
    "lion" ="#FFBF46",
    "wolf" = "grey"),
  guide ="none")

pz <- ggplot(x_stress_long, aes(x=RelativeCoordinate, y = MeanStress, color = animal))+
  geom_point()+
  geom_line()+
  theme_classic()+
  xlab("Relative Coordinate")+
  ylab("Mean van Misses Stress")+
  scale_color_manual(values = c(
    "bear" = "brown",
    "hyaena" = "#E15634",
    "lion" ="#FFBF46",
    "wolf" = "grey"))

grid.arrange(px, py, pz, nrow = 1)
  
```

To plot using `facet_grid` I combine all dimensions into one dataframe.

```{r FacetPlotting}
x_stress_long$Dimension <- "X"
y_stress_long$Dimension <- "Y"
z_stress_long$Dimension <- "Z"
df_list <- list(x_stress_long, y_stress_long, z_stress_long)

stress_long <- Reduce(function(x,y) merge(x,y, all = TRUE), df_list, accumulate = FALSE)

p <- ggplot(stress_long, aes(x=RelativeCoordinate, y = MeanStress, color = animal))+
  geom_point()+
  geom_line()+
  theme_classic()+
  xlab("Relative Coordinate")+
  ylab("Mean van Misses Stress")+
  scale_color_manual(values = c(
    "bear" = "brown",
    "hyaena" = "#E15634",
    "lion" ="#FFBF46",
    "wolf" = "grey"))

p+facet_grid(cols = vars(Dimension))
```

