---
title: "Solution_byArlene"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("readxl")
library("dplyr")
```

So far, I was only able to combine the different data into one dataframe. I will try to create the plots in the next week!

## Sheet 1 & 2: Survival & Developmental_Time

At first I looked at the structure of the different excel sheets. As sheet 1 (survival) and sheet 2 (developmental time) had the same structure, I only needed two functions for both of them: First I made a function to rename the columns to mean and sd plus numbers.
 
```{r}
# Function to rename columns to mean and sd
rename_columns <- function(df, start, prefix) {
  df %>%
    rename_with(~ paste0(prefix, seq_along(seq(start, ncol(df), by = 2))), 
                .cols = seq(start, ncol(df), by = 2))
}

```

I then implemented a function to reorganize the sheets so there are no empty cells:
```{r}
reorganize_sheets <- function(data, sheetname, rowname1, rowname2, rowname3){
  # load data and rename columns
  data <- data %>% 
    rename_columns(2, paste0(sheetname, "_mean_")) %>%
    rename_columns(3, paste0(sheetname, "_sd_"))
  # change first column name to index
  colnames(data)[1] <- "index"
  # change "index" for rows 4-6
  new_index_names <- c(paste0(sheetname, "_" ,rowname1,"_mean"), paste0(sheetname, "_" ,rowname2,"_mean"), paste0(sheetname, "_" ,rowname3,"_mean"))
  data[c(4:6),1] <- new_index_names
  
  # one matrix for mean
  mean_cols <- data %>% 
    select(contains("mean")) %>% 
    as.matrix()

  # one matrix for sd
  sd_cols <- data %>% 
    select(contains("sd")) %>% 
    # remove first 3 rows of sd_cols as they are empty
    slice(-c(1:3)) %>% 
    as.matrix()
  
  # combine data
  combined_data <- rbind(mean_cols, sd_cols) %>% 
    as.data.frame()
  
  # change the index names
  new_index_names_sd <- c(paste0(sheetname, "_" ,rowname1,"_sd"), paste0(sheetname, "_" ,rowname2,"_sd"), paste0(sheetname, "_" ,rowname3,"_sd"))
  combined_data$index <- c(data$index, new_index_names_sd)
  
  # change structure: pivot
  reshaped_data <- combined_data %>%
    pivot_longer(cols = contains("mean"), names_to = "mean", values_to = "value") %>%
    pivot_wider(names_from = index, values_from = value) %>%
    select(-"mean")
  
  return(reshaped_data)
}

```

Now let's change the structure of the survival and development data.
But first, here is an example how the sheet for survival looked before reshaping it:

```{r}
head(read_excel("flydata.xlsx", sheet = "Survival", col_names = F))
```


```{r}
data_survival <- read_excel("flydata.xlsx", sheet = "Survival", col_names = F) %>% 
  reorganize_sheets("Survival", "Females", "Males", "Total")

data_developm <- read_excel("flydata.xlsx", sheet = "Survival", col_names = F) %>% 
  reorganize_sheets("Developmental_Time", "L-P", "P-E", "L-E")
```
It now looks like this, which is already way cleaner!

```{r}
head(data_survival)
```
But now I also want to have a new column with the sex in the survival dataframe. I wrote a function for that.

```{r}
transform_data <- function(data, exclude_columns, new_sex, mean_column, sd_column) {
  data %>%
    # only select one sex
    select(-all_of(exclude_columns)) %>%
    # add new column with sex
    mutate(sex = new_sex) %>%
    # rename columns so sex not in column name anymore
    rename(survival_mean = !!sym(mean_column),
           survival_sd = !!sym(sd_column))
}
```

Now let's use this function on the data. In the following, I already combined the new dataframes with the development dataframe.

```{r}
data_survival_developm <-
  bind_rows(
    transform_data(
      data_survival,
      exclude_columns = c("Survival_Males_mean", "Survival_Males_sd", "Survival_Total_mean", "Survival_Total_sd"),
      new_sex = "F",
      mean_column = "Survival_Females_mean",
      sd_column = "Survival_Females_sd"),
    transform_data(
      data_survival,
      exclude_columns = c("Survival_Females_mean", "Survival_Females_sd", "Survival_Total_mean", "Survival_Total_sd"),
      new_sex = "M",
      mean_column = "Survival_Males_mean",
      sd_column = "Survival_Males_sd"),
    transform_data(
      data_survival,
      exclude_columns = c("Survival_Females_mean", "Survival_Females_sd", "Survival_Males_mean", "Survival_Males_sd"),
      new_sex = "Total",
      mean_column = "Survival_Total_mean",
      sd_column = "Survival_Total_sd")) %>% 
  merge(data_developm, by = c("n", "Line", "Diet"), all = F)
```

So the data for survival and development now look like this:
```{r}
head(data_survival_developm)
```


Now that the development and survival data are combined, I will look at metabolic pools.

## Sheet 3: Metabolic Pools

First I wrote a function to read and process the metabolic pool data. The dataframe looked like this at first:

```{r}
read_excel("flydata.xlsx", sheet = "Metabolic_Pools", col_names = F)
```


```{r}

process_metabolic_pool <- function(sheetname, rows) {
  # read data
  read_excel("flydata.xlsx", sheet = "Metabolic_Pools", col_names = F) %>% 
    # delete irrelevant notes at the bottom of excel sheet
    slice(-c(18:58)) %>% 
    # only get relevant data (drymass, proteins or driglycerides)
    slice(rows) %>%
    # reorganize sheets with function above
    reorganize_sheets(sheetname, "HPS", "EPS", "LPS") %>%
    # now convert it more meaningful format
    pivot_longer(cols = starts_with(sheetname), names_to = "temp", values_to = sheetname) %>%
    separate(temp, into = c("metric", "diet", "type"), sep = "_") %>%
    select(-metric) %>% 
    pivot_wider(names_from = type, values_from = sheetname)
}
```

Now I use this function for all three different types of metabolic pools (drymass, protein and triglycerides) and also renamed the columns.

```{r}
data_drymass <- process_metabolic_pool("Drymass", c(1:3, 5:7)) %>% 
  rename(mean_drymass = mean, sd_drymass = sd)

data_protein <- process_metabolic_pool("Protein", c(1:3, 10:12)) %>% 
  rename(mean_protein = mean, sd_protein = sd)

data_triglycerides <- process_metabolic_pool("Triglycerides", -c(4:14)) %>% 
  rename(mean_triglycerides = mean, sd_triglycerides = sd)
```

This is how drymass looks like now as an example:

```{r}
head(data_drymass)
```

## Combine into one dataframe

Now I finally combined the processed data into one final dataframe:
```{r}
final_data <- merge(data_drymass, data_protein, by=c("line", "n", "sex", "diet")) %>% 
  merge(data_triglycerides, by=c("line", "n", "sex","diet")) %>% 
  # rename so dataframes have same column names
  rename(Line = line,
         Diet = diet) %>% 
  left_join(data_survival_developm, data_metabolic, by = c("Line", "n", "Diet", "sex"))
```

This is how it looks like now:

```{r}
head(final_data)
```

## Plot 

I will plot the data in the next week.
