---
title: "Solution_byArlene"
output: md_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warnings = F, message=FALSE}

suppressMessages({
  suppressWarnings({
    library("readxl")
    library("dplyr")
    library(tidyr)
    library("gridExtra")
    library(ggsignif)
    library(ggrepel)
  })
})
```


## Sheet 1 & 2: Survival & Developmental_Time

At first I looked at the structure of the different excel sheets. As sheet 1 (survival) and sheet 2 (developmental time) had the same structure, I only needed two functions for both of them: First I made a function to rename the columns to mean and sd plus numbers.
 
```{r}
# Function to rename columns to mean and sd
rename_columns <- function(df, start, prefix) {
  df %>%
    rename_with(~ paste0(prefix, seq_along(seq(start, ncol(df), by = 2))), 
                .cols = seq(start, ncol(df), by = 2))
}

```

I then implemented a function to reorganize the sheets so there are no empty cells:
```{r}
reorganize_sheets <- function(data, sheetname, rowname1, rowname2, rowname3){
  # load data and rename columns
  data <- data %>% 
    rename_columns(2, paste0(sheetname, "_mean_")) %>%
    rename_columns(3, paste0(sheetname, "_sd_"))
  # change first column name to index
  colnames(data)[1] <- "index"
  # change "index" for rows 4-6
  new_index_names <- c(paste0(sheetname, "_" ,rowname1,"_mean"), paste0(sheetname, "_" ,rowname2,"_mean"), paste0(sheetname, "_" ,rowname3,"_mean"))
  data[c(4:6),1] <- new_index_names
  
  # one matrix for mean
  mean_cols <- data %>% 
    select(contains("mean")) %>% 
    as.matrix()

  # one matrix for sd
  sd_cols <- data %>% 
    select(contains("sd")) %>% 
    # remove first 3 rows of sd_cols as they are empty
    slice(-c(1:3)) %>% 
    as.matrix()
  
  # combine data
  combined_data <- rbind(mean_cols, sd_cols) %>% 
    as.data.frame()
  
  # change the index names
  new_index_names_sd <- c(paste0(sheetname, "_" ,rowname1,"_sd"), paste0(sheetname, "_" ,rowname2,"_sd"), paste0(sheetname, "_" ,rowname3,"_sd"))
  combined_data$index <- c(data$index, new_index_names_sd)
  
  # change structure: pivot
  reshaped_data <- combined_data %>%
    pivot_longer(cols = contains("mean"), names_to = "mean", values_to = "value") %>%
    pivot_wider(names_from = index, values_from = value) %>%
    select(-"mean")
  
  return(reshaped_data)
}

```

Now let's change the structure of the survival and development data.
But first, here is an example how the sheet for survival looked before reshaping it:

```{r, message=FALSE}
head(read_excel("flydata.xlsx", sheet = "Survival", col_names = F))
```


```{r, warnings = F, message=FALSE}
data_survival <- read_excel("flydata.xlsx", sheet = "Survival", col_names = F) %>% 
  reorganize_sheets("Survival", "Females", "Males", "Total")

data_developm <- read_excel("flydata.xlsx", sheet = "Developmental_Time", col_names = F) %>% 
  reorganize_sheets("Developmental_Time", "L-P", "P-E", "L-E")
```
It now looks like this, which is already way cleaner!

```{r}
head(data_survival)
```
But now I also want to have a new column with the sex in the survival dataframe. I wrote a function for that.

```{r}
transform_data_surviv <- function(data, exclude_columns, new_sex, mean_column, sd_column) {
  data %>%
    # only select one sex
    select(-all_of(exclude_columns)) %>%
    # add new column with sex
    mutate(sex = new_sex) %>%
    # rename columns so sex not in column name anymore
    rename(survival_mean = !!sym(mean_column),
           survival_sd = !!sym(sd_column))
}
```

Now let's use this function on the survival data.

```{r}
data_survival <-
  bind_rows(
    transform_data_surviv(
      data_survival,
      exclude_columns = c("Survival_Males_mean", "Survival_Males_sd", "Survival_Total_mean", "Survival_Total_sd"),
      new_sex = "F",
      mean_column = "Survival_Females_mean",
      sd_column = "Survival_Females_sd"),
    transform_data_surviv(
      data_survival,
      exclude_columns = c("Survival_Females_mean", "Survival_Females_sd", "Survival_Total_mean", "Survival_Total_sd"),
      new_sex = "M",
      mean_column = "Survival_Males_mean",
      sd_column = "Survival_Males_sd"),
    transform_data_surviv(
      data_survival,
      exclude_columns = c("Survival_Females_mean", "Survival_Females_sd", "Survival_Males_mean", "Survival_Males_sd"),
      new_sex = "Total",
      mean_column = "Survival_Total_mean",
      sd_column = "Survival_Total_sd"))
```

Let's also do this for development.

```{r}
transform_data_developm <- function(data, exclude_columns, new_developm, mean_column, sd_column) {
  data %>%
    # only select one sex
    select(-all_of(exclude_columns)) %>%
    # add new column with sex
    mutate(developmental_time = new_developm) %>%
    # rename columns so sex not in column name anymore
    rename(developmental_mean = !!sym(mean_column),
           developmental_sd = !!sym(sd_column))
  }
```


```{r}

data_developm <-
  bind_rows(
    transform_data_developm(
      data_developm,
      exclude_columns = c("Developmental_Time_P-E_mean", "Developmental_Time_P-E_sd", "Developmental_Time_L-E_mean", "Developmental_Time_L-E_sd"),
      new_developm = "L-P",
      mean_column = "Developmental_Time_L-P_mean",
      sd_column = "Developmental_Time_L-P_sd"),
    transform_data_developm(
      data_developm,
      exclude_columns = c("Developmental_Time_P-E_mean", "Developmental_Time_P-E_sd", "Developmental_Time_L-P_mean", "Developmental_Time_L-P_sd"),
      new_developm = "L-E",
      mean_column = "Developmental_Time_L-E_mean",
      sd_column = "Developmental_Time_L-E_sd"),
    transform_data_developm(
      data_developm,
      exclude_columns = c("Developmental_Time_L-E_mean", "Developmental_Time_L-E_sd", "Developmental_Time_L-P_mean", "Developmental_Time_L-P_sd"),
      new_developm = "P-E",
      mean_column = "Developmental_Time_P-E_mean",
      sd_column = "Developmental_Time_P-E_sd"))
```
And now I combine the two:

```{r}
data_survival_developm <- data_developm %>% 
  rename(Line = line,
         Diet = diet,
         n_developm = n) %>% 
  merge(data_survival, by = c("Line", "Diet"), all = F) %>% 
  rename(n_survival = n)
```


So the data for survival and development now look like this:
```{r}
head(data_survival_developm)
```


Now that the development and survival data are combined, I will look at metabolic pools.

## Sheet 3: Metabolic Pools

First I wrote a function to read and process the metabolic pool data. The dataframe looked like this at first:

```{r, results='hide', message=FALSE}

read_excel("flydata.xlsx", sheet = "Metabolic_Pools", col_names = F)
```


```{r}

process_metabolic_pool <- function(sheetname, rows) {
  # read data
  read_excel("flydata.xlsx", sheet = "Metabolic_Pools", col_names = F, n_max=17) %>% 
    # only get relevant data (drymass, proteins or driglycerides)
    slice(rows) %>%
    # reorganize sheets with function above
    reorganize_sheets(sheetname, "HPS", "EPS", "LPS") %>%
    # now convert it more meaningful format
    pivot_longer(cols = starts_with(sheetname), names_to = "temp", values_to = sheetname) %>%
    separate(temp, into = c("metric", "diet", "type"), sep = "_") %>%
    select(-metric) %>% 
    pivot_wider(names_from = type, values_from = sheetname)
}
```

Now I use this function for all three different types of metabolic pools (drymass, protein and triglycerides) and also renamed the columns.

```{r, warnings = F, message=FALSE}
data_drymass <- process_metabolic_pool("Drymass", c(1:3, 5:7)) %>% 
  rename(mean_drymass = mean, sd_drymass = sd)

data_protein <- process_metabolic_pool("Protein", c(1:3, 10:12)) %>% 
  rename(mean_protein = mean, sd_protein = sd)

data_triglycerides <- process_metabolic_pool("Triglycerides", -c(4:14)) %>% 
  rename(mean_triglycerides = mean, sd_triglycerides = sd)
```

This is how drymass looks like now as an example:

```{r}
head(data_drymass)
```

## Combine into one dataframe

Now I finally combined the processed data into one final dataframe:

```{r}
final_data <- merge(data_drymass, data_protein, by=c("line", "n", "sex", "diet")) %>% 
  merge(data_triglycerides, by=c("line", "n", "sex","diet")) %>% 
  # rename so dataframes have same column names
  rename(Line = line,
         Diet = diet,
         n_metabolic = n) %>% 
  left_join(data_survival_developm, data_metabolic, by = c("Line", "Diet", "sex"))
```

This is how it looks like now:

```{r}
head(final_data)
```
As a final step, lets convert the columns containing numerical data to numeric.

```{r, warnings = F, message = F}
# check if column only contains numeric values

is_numeric_column <- function(x) {
  all(!is.na(as.numeric(x)))
}

# change them to numeric
final_data[] <- lapply(final_data, function(x) if (is_numeric_column(x)) as.numeric(x) else x)
```


## Plot 

Now let's create the plots.

### Plot for survival

```{r, warning=F}

# Plot absolute values
plot_absolute_survival <- final_data %>% 
  ggplot( aes(x = Diet, y = survival_mean, fill = Diet)) +
  geom_violin(trim = FALSE, color = NA) +
  scale_fill_manual(values = c("HPS" = "yellow", "EPS" = "lightgreen", "LPS" = "lightblue")) +
  geom_boxplot(width = 0.2, fill = NA, color = "white") +
  theme_minimal() +
  theme(legend.position = "none") +
  ylab("Absolute variance survival") +
  xlab("Diet")+
  geom_signif(comparisons = list(c("EPS", "HPS") ,c("EPS","LPS"), c("HPS","LPS")), 
              test = t.test,
              map_signif_level=c("***"=0.001, "**"=0.01, "*"=0.05),
              # hide.ns does not work?
              hide.ns = T)

# Plot variance
plot_variance_survival <- final_data %>%
  ggplot(aes(x = Diet, y = survival_sd, fill = Diet)) +
  geom_violin(trim = FALSE, color = NA) +
  scale_fill_manual(values = c("HPS" = "yellow", "EPS" = "lightgreen", "LPS" = "lightblue")) +
  geom_boxplot(width = 0.2, fill = NA, color = "white") +
  theme_minimal() +
  theme(legend.position = "none") +
  ylab("variance (SD)") +
  xlab("Diet")

# show both in one 
grid.arrange(plot_absolute_survival, plot_variance_survival, ncol = 1)
```


### Plot for development


```{r, warning=F}
# Plot absolute values
plot_absolute_developm <- final_data %>% 
  ggplot( aes(x = Diet, y = developmental_mean, fill = Diet)) +
  geom_violin(trim = FALSE, color = NA) +
  scale_fill_manual(values = c("HPS" = "yellow", "EPS" = "lightgreen", "LPS" = "lightblue")) +
  geom_boxplot(width = 0.2, fill = NA, color = "white") +
  theme_minimal() +
  theme(legend.position = "none") +
  ylab("Absolute variance developmental time") +
  xlab("Diet")+
  geom_signif(comparisons = list(c("EPS", "HPS") ,c("EPS","LPS"), c("HPS","LPS")), 
              test = t.test,
              map_signif_level=c("***"=0.001, "**"=0.01, "*"=0.05),
              # hide.ns does not work?
              hide.ns = TRUE)

# Plot variance
plot_variance_developm <- final_data %>%
  ggplot(aes(x = Diet, y = developmental_sd, fill = Diet)) +
  geom_violin(trim = FALSE, color = NA) +
  scale_fill_manual(values = c("HPS" = "yellow", "EPS" = "lightgreen", "LPS" = "lightblue")) +
  geom_boxplot(width = 0.2, fill = NA, color = "white") +
  theme_minimal() +
  theme(legend.position = "none") +
  ylab("variance (SD)") +
  xlab("Diet")

# show both in one 
grid.arrange(plot_absolute_developm, plot_variance_developm, ncol = 1)
```
## Plot for metabolic pools
### Proteins

```{r, warning=F}
# Plot absolute values
plot_absolute_protein <- final_data %>% 
  ggplot( aes(x = Diet, y = mean_protein, fill = Diet)) +
  geom_violin(trim = FALSE, color = NA) +
  scale_fill_manual(values = c("HPS" = "yellow", "EPS" = "lightgreen", "LPS" = "lightblue")) +
  geom_boxplot(width = 0.2, fill = NA, color = "white") +
  theme_minimal() +
  theme(legend.position = "none") +
  ylab("Absolute variance protein in metabolic pool") +
  xlab("Diet")+
  geom_signif(comparisons = list(c("EPS", "HPS") ,c("EPS","LPS"), c("HPS","LPS")), 
              test = t.test,
              map_signif_level=c("***"=0.001, "**"=0.01, "*"=0.05),
              # hide.ns does not work?
              hide.ns = T)

# Plot variance
plot_variance_protein <- final_data %>%
  ggplot(aes(x = Diet, y = sd_protein, fill = Diet)) +
  geom_violin(trim = FALSE, color = NA) +
  scale_fill_manual(values = c("HPS" = "yellow", "EPS" = "lightgreen", "LPS" = "lightblue")) +
  geom_boxplot(width = 0.2, fill = NA, color = "white") +
  theme_minimal() +
  theme(legend.position = "none") +
  ylab("variance (SD)") +
  xlab("Diet")

# show both in one 
grid.arrange(plot_absolute_protein, plot_variance_protein, ncol = 1)
```
### Triglycerides

```{r, warning=F}
# Plot absolute values
plot_absolute_triglycerides <- final_data %>% 
  ggplot( aes(x = Diet, y = mean_triglycerides, fill = Diet)) +
  geom_violin(trim = FALSE, color = NA) +
  scale_fill_manual(values = c("HPS" = "yellow", "EPS" = "lightgreen", "LPS" = "lightblue")) +
  geom_boxplot(width = 0.2, fill = NA, color = "white") +
  theme_minimal() +
  theme(legend.position = "none") +
  ylab("Absolute variance triglycerides in metabolic pool") +
  xlab("Diet")+
  geom_signif(comparisons = list(c("EPS", "HPS") ,c("EPS","LPS"), c("HPS","LPS")), 
              test = t.test,
              map_signif_level=c("***"=0.001, "**"=0.01, "*"=0.05),
              # hide.ns does not work?
              hide.ns = T)

# Plot variance
plot_variance_triglycerides <- final_data %>%
  ggplot(aes(x = Diet, y = sd_triglycerides, fill = Diet)) +
  geom_violin(trim = FALSE, color = NA) +
  scale_fill_manual(values = c("HPS" = "yellow", "EPS" = "lightgreen", "LPS" = "lightblue")) +
  geom_boxplot(width = 0.2, fill = NA, color = "white") +
  theme_minimal() +
  theme(legend.position = "none") +
  ylab("variance (SD)") +
  xlab("Diet")

# show both in one 
grid.arrange(plot_absolute_triglycerides, plot_variance_triglycerides, ncol = 1)
```
### Drymass

```{r, warning=F}
# Plot absolute values
plot_absolute_drymass <- final_data %>% 
  ggplot( aes(x = Diet, y = mean_drymass, fill = Diet)) +
  geom_violin(trim = FALSE, color = NA) +
  scale_fill_manual(values = c("HPS" = "yellow", "EPS" = "lightgreen", "LPS" = "lightblue")) +
  geom_boxplot(width = 0.2, fill = NA, color = "white") +
  theme_minimal() +
  theme(legend.position = "none") +
  ylab("Absolute variance drymass in metabolic pool") +
  xlab("Diet")+
  geom_signif(comparisons = list(c("EPS", "HPS") ,c("EPS","LPS"), c("HPS","LPS")), 
              test = t.test,
              map_signif_level=c("***"=0.001, "**"=0.01, "*"=0.05),
              # hide.ns does not work?
              hide.ns = T)

# Plot variance
plot_variance_drymass <- final_data %>%
  ggplot(aes(x = Diet, y = sd_drymass, fill = Diet)) +
  geom_violin(trim = FALSE, color = NA) +
  scale_fill_manual(values = c("HPS" = "yellow", "EPS" = "lightgreen", "LPS" = "lightblue")) +
  geom_boxplot(width = 0.2, fill = NA, color = "white") +
  theme_minimal() +
  theme(legend.position = "none") +
  ylab("variance (SD)") +
  xlab("Diet")

# show both in one 
grid.arrange(plot_absolute_drymass, plot_variance_drymass, ncol = 1)
```

### Relationship triglycerides and protein

```{r}
# Lineare Regression berechnen
fit <- lm(mean_triglycerides ~ mean_protein, data = final_data)

# R^2 und p-Wert berechnen
r_squared <- summary(fit)$r.squared
p_value <- summary(fit)$coefficients[2, 4]

# Plot erstellen
plot_protein_triglycerides <- ggplot(final_data, aes(x = mean_protein, y = mean_triglycerides, color = Diet)) +
  geom_point(size = 3, shape = 16, alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  annotate("text", x = Inf, y = Inf, label = paste("R² =", round(r_squared, 2), "\n", "p =", format(p_value, digits = 2)),
           hjust = 1.1, vjust = 1.1, size = 5) +
  labs(title = "Relationship between protein- und triglycerides",
       x = "Protein (Mean)",
       y = "Triglycerides (Mean)") +
  scale_color_manual(values = c("HPS" = "yellow", "EPS" = "lightgreen", "LPS" = "lightblue"))+
  theme_minimal() +
  theme(legend.position = "right")

# Potenzielle Ausreißer markieren
outliers <- final_data %>% 
  filter(mean_protein > mean(mean_protein) + 2 * sd(mean_protein) | mean_triglycerides > mean(mean_triglycerides) + 2 * sd(mean_triglycerides))

plot_protein_triglycerides <- plot_protein_triglycerides +
  geom_text_repel(data = outliers, aes(label = Line), size = 3, arrow = arrow(length = unit(0.02, "npc")))

# Plot anzeigen
print(plot_protein_triglycerides)
```




