---
title: "Bike Counts Project (R2)"
author: "Mohamed Elsherif"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: md_document
---

```{r setup, include = F}
options(scipen=999)
```

***

### Bike count data

<div class="alert alert-block alert-info">
(@) Load all csv files into data frames. For each, create a variable and name the variable according to the respective filename. Familiarize yourself with the data.
</div>

```{r libraries}
library(readr)
```

```{r data-import}
# Define file paths
data_path <- "C:/Users/Admin/Downloads/DS400 Data Science Project Management WS23-24/Assignments/Assignment2/data/"

# Load datasets
bike_counts <- read_csv(paste0(data_path, "bike_counts.csv"))
counter_sites <- read_csv(paste0(data_path, "counter_sites.csv"))
holidays <- read_csv(paste0(data_path, "holidays.csv"))
weather <- read_csv(paste0(data_path, "weather.csv"))

# Display the first few rows of each dataset
head(bike_counts)
head(counter_sites)
head(holidays)
head(weather)
```

<div class="alert alert-block alert-info">
(@) Convert all date or datetime columns to a date/datetime datatype. In R, this should have been done automatically when importing the csv.
</div>

```{r date-time-check}

```

<div class="alert alert-block alert-info">
(@) Aggregate the `bike_counts` dataset by `counter_site` such that the number of bikes per hour is now the sum of the `bike_counts` across all channels. Call the resulting object `bike_counts_agg`.
</div>

```{r}
# Load necessary libraries
library(readr)
library(dplyr)
# Aggregate bike counts by counter_site_id and time
bike_counts_agg <- bike_counts %>%
  group_by(counter_site_id, time) %>%
  summarise(bike_count = sum(bike_count), .groups = 'drop')

# Display the aggregated dataset
head(bike_counts_agg)
```

***

### Merging the data

<div class="alert alert-block alert-info">
(@) Merge all datasets by the appropriate key. Use a left-join for all datasets except the `weather` data. The first 'left' dataset should be the `bike_counts_agg`. For the `weather` data use an inner-join. Call the result `merged`.
</div>

<div class="alert alert-block alert-warning">
As you might have noticed, for the `holidays` dataset there is no column in the `bike_counts` data that can be directly used for the merge. Therefore, you first have to create a new column which contains just the date of the respective timestamp. In R it is possible with the `as.Date()` function. After the merge, drop the `date` column as it is not needed anymore.
</div>

```{r}
# Merge bike_counts_agg with counter_sites using left join
merged <- bike_counts_agg %>%
  left_join(counter_sites, by = "counter_site_id")

# Add date column for merging with holidays
merged <- merged %>%
  mutate(date = as.Date(time)) %>%
  left_join(holidays, by = "date") %>%
  select(-date)  # Drop the date column after the merge

# Merge the resulting data with weather using inner join
merged <- merged %>%
  inner_join(weather, by = c("time" = "time"))

# Display the merged dataset
head(merged)
```

***

### Missing values

<div class="alert alert-block alert-info">
(@) For the columns contained in bike_counts and counter_sites there should be no missing values. For the holidays columns, however, you will encounter a lot of missing values after the join. Briefly describe the reason for this. Can you come up with a reasonable approach to fill these missing values? Implement your approach.
</div>

```{r}
library(tidyr) 
# Function to check for missing values and return column names with missing values
check_missing_values <- function(df) {
  missing_cols <- sapply(df, function(x) sum(is.na(x)) > 0)
  colnames(df)[missing_cols]
}

# Check for missing values in each table
bike_counts_missing <- check_missing_values(bike_counts_agg)
counter_sites_missing <- check_missing_values(counter_sites)
holidays_missing <- check_missing_values(holidays)
weather_missing <- check_missing_values(weather)
merged_missing <- check_missing_values(merged)

# Display columns with missing values
print("Bike Counts missing columns:")
print(bike_counts_missing)

print("Counter Sites missing columns:")
print(counter_sites_missing)

print("Holidays missing columns:")
print(holidays_missing)

print("Weather missing columns:")
print(weather_missing)

print("Merged missing columns:")
print(merged_missing)

# Fill missing values in holidays columns with 0
merged <- merged %>%
  mutate(across(starts_with("semester_break"), ~replace_na(., 0))) %>%
  mutate(across(starts_with("school_holiday"), ~replace_na(., 0))) %>%
  mutate(across(starts_with("public_holiday"), ~replace_na(., 0)))

# Check for remaining missing values
merged_missing_after_fill <- check_missing_values(merged)
print("Merged missing columns after fill:")
print(merged_missing_after_fill)
```

<div class="alert alert-block alert-info">
(@) Since we used an inner-join for the `weather` data, there should be no missing values as well. Can you confirm this?
</div>
<div class="alert alert-block alert-info">
The missing values in the holidays columns arise because not all timestamps in the bike_counts_agg dataset have corresponding dates in the holidays dataset. Specifically, the holidays dataset only marks certain days (e.g., public holidays, school holidays), and if there is no entry for a specific date, it results in a missing value when joined.
</div>

```{r}
holidays_missing <- check_missing_values(holidays)
holidays_missing
```

***

### Visual data exploration

Before estimating a model on the data, we try to get a feeling for possible patterns in the data. That helps to feed the data to the model in a meaningful way.

#### A: Bike counts per week
The first thing we want to have a look at is just the time series of aggregated bike counts.

<div class="alert alert-block alert-info">
(@) Aggregate the `bike_counts` (per hour) to bike counts per week. Hint: For the grouping part you might want to have a look at the API documentation of  [{lubridate}](https://lubridate.tidyverse.org/reference/round_date.html) (R). 
</div>

```{r}
library(lubridate)
# Aggregate bike counts to weekly counts
bike_counts_weekly <- bike_counts %>%
  mutate(week = floor_date(time, unit = "week")) %>%
  group_by(counter_site_id, week) %>%
  summarise(bike_count = sum(bike_count), .groups = 'drop')

# Display the aggregated dataset
head(bike_counts_weekly)

```

<div class="alert alert-block alert-info">
(@) Create a faceted lineplot that shows the number of bikes per week. For each counter site there should be a facet with the respective counter site name as label, arranged such that the resulting figure has 3 rows and 1 column. The x and y axis should be shared so that the subplots are comparable.
</div>

```{r}
library(ggplot2)
# Merge with counter_sites to get the counter site names
bike_counts_weekly <- bike_counts_weekly %>%
  left_join(counter_sites, by = "counter_site_id")

# Create faceted line plot
ggplot(bike_counts_weekly, aes(x = week, y = bike_count)) +
  geom_line() +
  facet_wrap(~ counter_site, scales = "free_y", nrow = 3) +
  labs(title = "Weekly Bike Counts by Counter Site", x = "Week", y = "Number of Bikes") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

***

#### B: Bike counts per hour and weekday

Next, we want to look for timetable-ish patterns.

<div class="alert alert-block alert-info">
(@) Aggregate the `bike_counts` dataset to get the average number of bikes per hour separated by weekday (across all counter sites).
</div>

```{r}
# Extract weekday from the time column
bike_counts <- bike_counts %>%
  mutate(weekday = wday(time, label = TRUE, abbr = FALSE))

# Aggregate to get the average number of bikes per hour by weekday
avg_bikes_per_weekday <- bike_counts %>%
  group_by(weekday) %>%
  summarise(avg_bike_count = mean(bike_count, na.rm = TRUE))

# Display the aggregated dataset
print(avg_bikes_per_weekday)
```

<div class="alert alert-block alert-info">
(@) Plot the resulting average bike count in a heatmap such that it looks like a timetable (`weekday` on the x-axis and `hour` on the y-axis).
</div>

```{r}
# Extract weekday and hour from the time column
bike_counts <- bike_counts %>%
  mutate(weekday = wday(time, label = TRUE, abbr = FALSE),
         hour = hour(time))

# Aggregate to get the average number of bikes per hour by weekday and hour
avg_bikes_per_weekday_hour <- bike_counts %>%
  group_by(weekday, hour) %>%
  summarise(avg_bike_count = mean(bike_count, na.rm = TRUE), .groups = 'drop')

# Create heatmap
ggplot(avg_bikes_per_weekday_hour, aes(x = weekday, y = hour, fill = avg_bike_count)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "red") +
  labs(title = "Average Bike Count by Weekday and Hour",
       x = "Weekday", y = "Hour", fill = "Avg Bike Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


***

### Fixed effects regression

Finally, we want to regress the number of bikes per hour (`bike_count`) on the other variables.

<div class="alert alert-block alert-info">
(@) Which variables should be included as fixed effects? What about the `time` column? Since we are not estimating a time series model, it does not make sense to include it just as it is. However, there is certainly relevant information contained in the timestamp. By referencing to the results of the last section, state which information of the timestamp could be relevant/predictive. Extract this information by creating new variables in the data frame.    
    *(Hint: You can use `month(time)` from the {lubridate} package (R) to extract for example the month of a timestamp)*    
    Would you include these variables as numeric variables or fixed effects?
</div>

```{r}
# Extract relevant information from the timestamp
merged <- merged %>%
  mutate(hour = hour(time),
         day_of_week = wday(time, label = TRUE, abbr = FALSE),
         month = month(time, label = TRUE, abbr = FALSE))

# View the new variables
head(merged)

# Load necessary libraries
library(lmtest)

# Run the regression model
model <- lm(bike_count ~ hour + day_of_week + month + temperature + windspeed + humidity + heaven + rain + snow + thunderstorms + fog + semester_break + school_holiday + public_holiday + counter_site_id, data = merged)

# Summary of the model
summary(model)

```

<div class="alert alert-block alert-info">
(@) Make a train-test-split. In this case, just choose a single week and counter site of your choice as test set and use the rest for the estimation. Set a seed for reproducibility. Estimate the model on the training sample. To make sure the model outputs only positive values, we apply a trick: Instead of using the bike_count in levels as dependent variable, use the square root.
</div>

```{r}
set.seed(123)
# Select a specific week and counter site for the test set
test_start_date <- as.Date("2021-07-01")
test_end_date <- as.Date("2021-07-07")
test_counter_site_id <- 100003358

test_set <- merged %>%
  filter(date(time) >= test_start_date & date(time) <= test_end_date & counter_site_id == test_counter_site_id)

# Create the training set
train_set <- merged %>%
  filter(!(date(time) >= test_start_date & date(time) <= test_end_date & counter_site_id == test_counter_site_id))

# Fit the regression model on the training set using the square root of bike_count as the dependent variable
model <- lm(sqrt(bike_count) ~ hour + day_of_week + month + temperature + windspeed + humidity + heaven + rain + snow + thunderstorms + fog + semester_break + school_holiday + public_holiday + counter_site_id, data = train_set)

# Summary of the model
summary(model)

```


<div class="alert alert-block alert-info">
(@) Make predictions on the test set. Remember, that we fitted the model to the square root of the `bike_counts`. Compute the mean absolute error of your predictions.
</div>

```{r}
# Make predictions on the test set
test_set$predicted_bike_count <- predict(model, newdata = test_set)

# Transform the predicted values back to the original scale
test_set$predicted_bike_count <- test_set$predicted_bike_count^2

# Compute the mean absolute error
mae <- mean(abs(test_set$predicted_bike_count - test_set$bike_count))
print(paste("Mean Absolute Error:", mae))
```

<div class="alert alert-block alert-info">
(@) Plot the prediction on your test set as a time series. Also add the 95% confidence interval of your prediction as filled area as well as the actual values as points.
</div>

```{r}
# Calculate confidence intervals
test_set$lower_ci <- predict(model, newdata = test_set, interval = "confidence")[, "lwr"]
test_set$upper_ci <- predict(model, newdata = test_set, interval = "confidence")[, "upr"]


# Plotting
ggplot(data = test_set, aes(x = time)) +
  geom_line(aes(y = predicted_bike_count), color = "blue", linetype = "solid", size = 1) +
  geom_point(aes(y = bike_count), color = "red", size = 2) +
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci), fill = "lightblue", alpha = 0.5) +
  labs(title = "Prediction of bike counts per hour",
       subtitle = "Counter Site: Fuß- & Radtunnel Südportal - Derendinger Allee",
       x = "Date",
       y = "Bike Count") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 14, face = "italic"))

```