---
title: "A Visual Exploration of the Artistic Landscape in Colombia (1940-2018)"
author: "Julia Quach"
date: "2023-06-22"
output: md_document
always_allow_html: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(rio)
library(kableExtra)
library(cowplot)
df <- import_list("PG_BD_INVITACIONES_SM05-20-2021.xlsx", rbind = TRUE)

```

-------------------------------------------------------------------------------

# Motivation

-------------------------------------------------------------------------------


In this project, we would like to get an insight into the art scene in Colombia in the second half of the 20th century. For this, we observe and visualise the transformation of the art scene of Bogot치. In particular, we take a closer look at the relations between the artists and galleries at that time as well as the art critics and their media outlets.

-------------------------------------------------------------------------------

# Aim

-------------------------------------------------------------------------------


To visualise the relations between artists, exposition and gallery, we opt for a Sankey diagram.


-------------------------------------------------------------------------------

# Data Wrangling

-------------------------------------------------------------------------------


The available data set is an xlsx.-file with eight sheets, each sheet representing one decade between 1940 and 2018. Each row represents one gallery and for the galleries, the columns represent information about the name of the gallery, the title, the inauguration data, the artists and many more. In this project, we will focus on the variables "Nombre, Galeria, Espacio", "Artistas", and "A침o de fecha de inauguracion", i.e., galleries, artists and years. If information about the gallery, artist or year is not available, we drop the whole row representing that gallery.

```{r data wrangling, echo=FALSE}
# Reducing the df to the relevant columns

data <- df %>% 
  subset( select = c("Nombre / Galeria / Espacio", "Artistas" ,"A침o de fecha de inauguracion (aaaa)")) %>% # Reducing the df to the relevant columns
  rename( c(galleries = "Nombre / Galeria / Espacio", artists = "Artistas", year = "A침o de fecha de inauguracion (aaaa)" ) ) %>%  # Renaming columns
  separate_rows( artists, sep = ", ") %>% # Splitting strings of artists and saving as new variable
  mutate( across( c( artists, galleries ), function(x){ iconv(x, to='ASCII//TRANSLIT') } ) ) %>%
  mutate( year = as.integer( str_extract( year, regex("^[0-9]{4}") ) ) ) %>%  # Cast year format
  na.omit() %>% # remove all rows with NA entries 
  mutate( decade = year%/%10*10) # add decade information


knitr::kable( data[1:5,1:3], col.names = c("Galleries", "Artists", "Year"), caption = "Table 1: First five rows of our data set after reduction of our initial data set.") %>%
    kable_styling(c("bordered", "striped"), full_width = T)


```

We classify the galleries by their decade and group the data set by the artists and galleries. This allows to count how many times a gallery invited an artist.


```{r grouping, echo=FALSE}

data <- data %>% 
  group_by(artists, galleries,decade) %>% 
  count(artists) %>% 
  ungroup()


knitr::kable( data[1:5,], col.names = c("Galleries", "Artists", "Decade", "Number of Invitations"), caption = "Table 2: First five rows of our data set after computing the decade and counting of invitations") %>%
    kable_styling(c("bordered", "striped"), full_width = T)

```

-------------------------------------------------------------------------------

# Visualisation

-------------------------------------------------------------------------------

To avoid a cluttered visualisation, we either choose a minimum number of invitations for the artists to be included in the Sankey diagram or a certain decade which we want to visualise.

```{r visualization, echo=FALSE}

# creating a df that only permits cases with invitations over the threshold 'minNumInvitations'

minNumInvitations <- 3
df1 <- data[ ( data$n > minNumInvitations ), ]

# creating a df that only permits cases from one decade
chosenDecade <- "1990"
df2 <- data[ ( data$decade == chosenDecade ), ]

### WIP #########################################################################
# Choose data frame which you want to plot
# x <- df1
# 
# # Plotting
# 
# links <- data.frame(
#   source=x$galleries,
#   target=x$artists,
#   value=x$n
# )
# 
# nodes <- data.frame(
#   name=c(as.character(links$source),
#          as.character(links$target))|>unique()
# )
# 
# links$IDsource <- match(links$source, nodes$name)-1
# links$IDtarget <- match(links$target, nodes$name)-1
# 
# #Network
# sankey <- sankeyNetwork(Links = links, Nodes = nodes, 
#                         Source = "IDsource", Target = "IDtarget",
#                         Value = "value", NodeID = "name",
#                         sinksRight = FALSE)
# sankey

```

-------------------------------------------------------------------------------

# Findings

-------------------------------------------------------------------------------

