---
title: "Solution for Jakob Raatschen's Project by Armin Burkhardt"
output: md_document
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(this.path)
library(khroma)
library(scales)

setwd(this.path::this.dir())

massoni <- read_csv("Massoni_2017/data_Massoni_2017.csv") # Study 1
palser <- read_csv("Palser_2018/data_Palser_2018.csv")    # Study 2
hu <- read_csv("Hu_2017/data_Hu_2017.csv")                # Study 3

highcontrast <- color("high contrast")
```

## Cleaning the data

The data for all 3 studies is cleaned, Confidences are rescaled and the data is merged into new dataframes.

```{r, message=FALSE, echo=FALSE}

# note that rescaling has to be done per dataset because the ranges are not the same


# prepare each dataset
massoni <- massoni %>% 
  filter(Subj_idx >= 67, Subj_idx <= 120) %>%           # second study only
  mutate(Study_id = "Massoni&Roux", 
         Title = "Massoni & Roux\n(Construct: Perception)",
         Orig_idx = Subj_idx,
         Confidence = rescale(Confidence, to = c(0, 100)))       # rescale from 0-1 to 0-100

palser <- palser %>%
  drop_na() %>%                                                  # exclude responses with exceeded deadline
  filter(Condition == "Baseline") %>%                            # following the suggestion to make this study more comparable to the others
  rename(Orig_idx = Subj_idx) %>%
  mutate(Study_id = "Palser_et_al", 
         Title = "Palser et al.\n(Construct: Perception)",
         Subj_idx = Orig_idx + max(massoni$Subj_idx),            # in preperation for merging, shift by largest massoni subject index
         Confidence = rescale(Confidence, to = c(0, 100)),       # rescale
         Accuracy = as.double(Response == Stimulus))             # add accuracy column

hu <- hu %>%
  rename(Orig_idx = Subj_idx) %>%
  mutate(Study_id = "Hu_et_al", 
         Title = "Hu et al.\n(Construct: Memory)",
         Subj_idx = Orig_idx + max(palser$Subj_idx),             # in preperation for merging, shift by largest palser subject index
         Confidence = rescale(Confidence, to = c(0, 100)))       # rescale from 1-6 to 0-100


# merge 
confidence <- bind_rows(
  massoni,
  palser,
  hu
) %>%
  select(Subj_idx, Orig_idx, Study_id, Title, Accuracy, Confidence)
  
performance <- confidence %>%
  group_by(Subj_idx, Orig_idx, Study_id, Title) %>%
  summarise(average_performance = mean(Accuracy, na.rm = TRUE)) %>%
  ungroup()


```


Resulting in these two final dataframes:
```{r, message=FALSE, echo=FALSE}
library(knitr)
kable(head(confidence %>% select(-Title)))
kable(head(performance %>% select(-Title)))
```

## Data visualization 1: Confidence ratings and performance accuracy across studies

Firstly, filter out all subjects that have only correct or only incorrect responses.

```{r, message=FALSE}
confidence_filtered <- confidence %>%
  group_by(Subj_idx) %>%
  filter(sum(Accuracy == 1) > 0, sum(Accuracy == 0) > 0) %>%
  ungroup()
```

Now the means per subject and study can be calculated:

```{r, message=FALSE}

# in preperation for plotting, create txt labels
confidence_filtered <- confidence_filtered %>%
  mutate(Label = ifelse(Accuracy == 1, "Correct", "Incorrect"))


confidence_means_individual <- confidence_filtered %>%
  group_by(Subj_idx, Study_id, Label, Title) %>%
  summarise(mean_conf = mean(Confidence, na.rm = TRUE))


confidence_means_group <- confidence_filtered %>%
  group_by(Study_id, Label, Title) %>%
  summarise(mean_conf = mean(Confidence, na.rm = TRUE))

```


Final plot:

```{r viz-1}  
ggplot() +
  # individual lines
  geom_line(data = confidence_means_individual, 
            aes(x = Label, y = mean_conf, group = Subj_idx, color = Title), 
            alpha = 0.4) +  # slightly transparent to make group lines stand out more
  # individual points
  geom_point(data = confidence_means_individual, 
             aes(x = Label, y = mean_conf, color = Title), 
             alpha = 0.4,
             size = 1) +
  # group lines
  geom_line(data = confidence_means_group, 
            aes(x = Label, y = mean_conf, group = Study_id, color = Title), 
            linewidth = 2) +
  # group points
  geom_point(data = confidence_means_group, 
             aes(x = Label, y = mean_conf, color = Title), 
             size = 3) +
  facet_wrap(~ Title) +  # panel for each study
  theme_minimal() +
  labs(
    title = "Confidence ratings by decision Accuracy across Studies",
    x = "",
    y = "Avg. Confidence",
  ) +
  theme(legend.position = "none") + 
  scale_color_manual(values = highcontrast(3))

```



## Data visualization 2: Performance accuracy across studies

This was very straightforward thanks to the amazing drawing provided. The following plot shows where most participants fall in terms of their average performance accuracy across the three studies (violin: actual distribution of the data, boxplot: median and range, jitter: individual data points spread out)

```{r viz-2}

ggplot(performance, aes(x = Title, y = average_performance, fill = Study_id)) +
  geom_violin(alpha = 0.5, trim = FALSE) +
  # boxplot with outliers hidden as jitter also shows them
  geom_boxplot(width = 0.1, fill = "white", outlier.shape = NA) +
  # each individual participants performance score (jittered as in the drawing)
  geom_jitter(width = 0.1, alpha = 1, size = 2, aes(fill = Study_id), color = "black", shape = 21) +
  theme_minimal() +
  labs(
    title = "Average performance accuracy across studies",
    x = "Study",
    y = "Performance Accuracy"
  ) +
  theme(legend.position = "none") +
  scale_fill_manual(values = highcontrast(3))

```


## Closing remarks

The data very much confirms that correct decisions usually have higher confidence ratings than incorrect ones, as you stated in your visualization 1 task, and its also interesting to see how well the participants performed across the three studies!

Thank you for this interesting project and especially for the pretty drawings, they helped a lot!


