---
title: "Solution for Jakob Raatschen's Project by Armin Burkhardt"
output: md_document
---

```{r, echo=FALSE}
library(tidyverse)
library(this.path)

setwd(this.path::this.dir())

massoni <- read_csv("Massoni_2017/data_Massoni_2017.csv") # Study 1
palser <- read_csv("Palser_2018/data_Palser_2018.csv")    # Study 2
hu <- read_csv("Hu_2017/data_Hu_2017.csv")                # Study 3
```


## Cleaning the data

### Study 1

```{r}
massoni <- massoni %>%
  filter(Subj_idx >= 67, Subj_idx <= 120) %>%                    # second study only
  mutate(Study_id = "Massoni&Roux")                              # add study_id column

# calculate average performance per subject
massoni_subj_perf <- massoni %>%
  group_by(Subj_idx, Study_id) %>%                               # group by subject and keep study_id
  summarise(average_performance = mean(Accuracy, na.rm = TRUE))  # calc average performance
```

### Study 2

```{r}
palser <- palser %>%
  drop_na() %>%                                                  # exclude responses with exceeded deadline
  filter(Condition == "Baseline") %>%                            # following the suggestion to make this study more comparable to the others
  mutate(Study_id = "Palser_et_al",                              # add study_id column
         Accuracy = ifelse(Response == Stimulus, 1, 0))          # add accuracy column

# again calculate average performance per subject, same as above
palser_subj_perf <- palser %>%
  group_by(Subj_idx, Study_id) %>%
  summarise(average_performance = mean(Accuracy, na.rm = TRUE))
```


### Study 3

```{r}
# following a similar pattern as above:
hu <- hu %>%
  mutate(Study_id = "Hu_et_al")

hu_subj_perf <- hu %>%
  group_by(Subj_idx, Study_id) %>%
  summarise(average_performance = mean(Accuracy, na.rm = TRUE))

```



### All studies

Rescaling of the confidence ratings to 0 to 100:
```{r}
massoni <- massoni %>%
  mutate(Confidence = Confidence * 100)

# palser is already scaled well, except for the bounds 0 and 99, but Ill just ignore that

hu <- hu %>%
  mutate(Confidence = (Confidence - 1) / 5 * 100) # rescale from 1-6 to 0-100
```

### Merging the datasets to aggregated dataframes:

Firstly, shift subject indices to avoid overlap:
```{r}  
palser <- palser %>%  
  mutate(Subj_idx = Subj_idx + max(massoni$Subj_idx)) # shift by largest massoni subject index

hu <- hu %>%
  mutate(Subj_idx = Subj_idx + max(palser$Subj_idx))  # shift by largest palser subject index

```

Now merge the datasets into two new dataframes:

```{r}
select_relevant <- function(df) {
  df %>%
    select(Subj_idx, Study_id, Accuracy, Confidence)
}

confidence <- bind_rows(
  select_relevant(massoni),
  select_relevant(palser),
  select_relevant(hu)
)

performance <- bind_rows(
  massoni_subj_perf,
  palser_subj_perf,
  hu_subj_perf
)
```

resulting in these two final dataframes:
```{r}
head(confidence)
head(performance)
```

## Data visualization 1: Confidence ratings and performance accuracy across studies

Firstly, filter out all subjects that have only correct or only incorrect responses.

```{r}
subjects_to_include <- confidence %>%
  group_by(Subj_idx) %>%
  summarise(n_correct   = sum(Accuracy == 1),
            n_incorrect = sum(Accuracy == 0)) %>%
  filter(n_correct > 0, n_incorrect > 0) %>%
  pull(Subj_idx)

confidence_filtered <- confidence %>%
  filter(Subj_idx %in% subjects_to_include)

```

Now the means per subject and study can be calculated:

```{r}

# in preperation for plotting, create txt labels
confidence_filtered <- confidence_filtered %>%
  mutate(Label = ifelse(Accuracy == 1, "Correct", "Incorrect"))


confidence_means_individual <- confidence_filtered %>%
  group_by(Subj_idx, Study_id, Label) %>%
  summarise(mean_conf = mean(Confidence, na.rm = TRUE))


confidence_means_group <- confidence_filtered %>%
  group_by(Study_id, Label) %>%
  summarise(mean_conf = mean(Confidence, na.rm = TRUE))

```


Final plot:

```{r}  
ggplot() +
  # individual lines
  geom_line(data = confidence_means_individual, 
            aes(x = Label, y = mean_conf, group = Subj_idx, color = Study_id), 
            alpha = 0.6) +  # slightly transparent to make group lines stand out more
  # individual points
  geom_point(data = confidence_means_individual, 
             aes(x = Label, y = mean_conf, color = Study_id), 
             alpha = 0.6) +
  # group lines
  geom_line(data = confidence_means_group, 
            aes(x = Label, y = mean_conf, group = Study_id, color = Study_id), 
            linewidth = 1.5) +
  # group points
  geom_point(data = confidence_means_group, 
             aes(x = Label, y = mean_conf, color = Study_id), 
             size = 3) +
  facet_wrap(~ Study_id) +  # panel for each study
  theme_minimal() +
  labs(
    title = "Confidence ratings by decision Accuracy across Studies",
    x = "",
    y = "Avg. Confidence",
    color = "Study"
  ) +
  theme(legend.position = "right")

```







