---
title: "Solution by Joshua"
output: md_document
---


```{r, include = FALSE}
library(tidyverse)
knitr::opts_knit$set(root.dir = here::here())
```

Study 1
```{r, include = FALSE}
study_id <- factor(c("Masoni", "Palser", "Hu"), levels = c("Masoni", "Palser", "Hu"), ordered = TRUE)

masoni <- read_csv("Projects/JakobRaatschen/Massoni_2017/data_Massoni_2017.csv") %>%
    filter(Study == 2) %>%
    mutate(Study_id = study_id[1]) %>%
    drop_na() %>%
    mutate(Subj_idx = (3 * Subj_idx)) %>%
    select(Study_id, Subj_idx, Stimulus, Response, Confidence, Accuracy)
masoni
masoni_avg_pp <- masoni %>%
    summarise(.by = Subj_idx, Avg_Accuracy = mean(Accuracy))

masoni_avg_pp
```

Study 2
```{r, include = FALSE}
palser <- read_csv("Projects/JakobRaatschen/Palser_2018/data_Palser_2018.csv") %>%
    filter(Condition == "Baseline") %>%
    mutate(Study_id = study_id[2]) %>%
    drop_na() %>%
    mutate(Accuracy = as.numeric(Stimulus == Response)) %>% # I know the compare is bool but the mean can use it anyway
    mutate(Confidence = ifelse((Confidence > 99), 99, Confidence)) %>% # This step actually belongs to the "All Studies part" we normalize the Confidence to 1->99
    mutate(Confidence = ((Confidence - 1) / 98)) %>% # Normalizing the confidence vals to 0->100%
    mutate(Subj_idx = ((3 * Subj_idx) + 1)) %>%
    select(Study_id, Subj_idx, Stimulus, Response, Confidence, Accuracy)
palser
palser_avg_pp <- palser %>%
    summarise(.by = Subj_idx, Avg_Accuracy = mean(Accuracy))
palser_avg_pp
```

Study 3
```{r, include = FALSE}
hu <- read_csv("Projects/JakobRaatschen/Hu_2017/data_Hu_2017.csv") %>%
    mutate(Study_id = study_id[3]) %>%
    drop_na() %>%
    mutate(Confidence = ((Confidence - 1) / 5)) %>% # Normalizing the 1->6 scale
    mutate(Subj_idx = ((3 * Subj_idx) + 2)) %>%
    select(Study_id, Subj_idx, Stimulus, Response, Confidence, Accuracy)
hu
hu_avg_pp <- hu %>%
    summarise(.by = Subj_idx, Avg_Accuracy = mean(Accuracy))

hu_avg_pp
```

# All Studies
For palser the txt says that the scale is 1->99 but that the program was not restricted, which means I would either need to filter out everything above 99 or set it to 99 as the users where just overconfident. Probably the correct thing would be throwing it out, so I've set them to 99 :D

For the participant ID I would rather like a operation that is easily reversible with knowing only little about the data itself. Since we have 3 lists we can simply use modulo or bitshift for that, I will use modulo for it as it is simpler to understand without computer science knowledge.
We already created a factor for the studies, so we use the factor as numeric input for:
$$
i = factornumber - 1 \\\\
y = 3x+i
$$
That way we can easily
$$ 
x = \frac{(y - i)}{3}
$$
i can be calculated simply by doing 
$$ 
i = y\ \%\ 3 
$$

```{r, include = FALSE}
confidence <- full_join(full_join(masoni, palser), hu)
performance <- full_join(full_join(masoni_avg_pp, palser_avg_pp), hu_avg_pp)
```

Data visualization
Spaghetti plot
```{r}
confidence %>%
    filter_out(.by = Subj_idx, ifelse(mean(Accuracy) == 1 || mean(Accuracy) == 0, TRUE, FALSE)) %>%
    mutate(Study_id = factor(Study_id)) %>%
    mutate(Accuracy = factor(Accuracy, levels = c(1, 0), labels = c("Correct", "Incorrect"))) %>%
    group_by(Study_id, Subj_idx, Accuracy) %>%
    summarise(Confidence = mean(Confidence)) %>%
    ggplot(aes(Accuracy, Confidence, color = Study_id)) +
    geom_line(aes(group = Subj_idx), alpha = 0.3) +
    geom_point(aes(group = Subj_idx), alpha = 0.3) +
    stat_summary(aes(group = Study_id), fun = mean, geom = "line", linewidth = 1.5) +
    stat_summary(aes(group = Study_id), fun = mean, geom = "point", size = 3) +
    facet_wrap(~Study_id) +
    theme_minimal()
```
