---
title: "Solution for Samandar's Project by Emily"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(readr)
```

# Data Import
```{r data-import, message = FALSE}

raw_data  <- read_csv("dependencies/SpeciesByKingdomAndClass.csv")

```

## Add Kingdom Data

- lists kingdoms in the order they appear in the data set
- creates column that says TRUE when a change appears (new kingdom)
- replaces NA in 1st observation with TRUE
- fills the NAs with the value of the last non-NA observation
- removes the ClassChange & AllKingdom column 

```{r add-kingdom}

kingdoms <- c("Animalia", "Chromista", "Fungi", "Plantae") 

kingdom_data  <- raw_data %>%
  mutate(ClassChange = Name < lag(Name),     
         ClassChange = replace_na(ClassChange, TRUE),
         AllKingdom = ifelse(ClassChange == TRUE, kingdoms[cumsum(ClassChange)], NA)) %>%
  fill(AllKingdom, .direction = "down") %>%
  mutate(Kingdom = if_else(Name == "Total", NA, AllKingdom)) %>%
  select(!c("ClassChange","AllKingdom"))      

```    


# Data Manipulation

- removes non-IUCN data
- creates new column for NearThreatened species by combining the two columns that contain this information
- renames columns for clarity
- reorders columns

```{r data-manipulation}

manipulated_data <- kingdom_data %>%
  select(!("CR(PE)":"Subtotal (EX+EW+ CR(PE)+CR(PEW))")) %>%
  rowwise() %>%
  mutate(NearThreatened = sum(c_across("LR/cd":"NT or LR/nt"))) %>%
  select(!("LR/cd":"NT or LR/nt")) %>%
  rename("Extinct" = "EX",
         "ExtinctWild" = "EW",
         "CriticallyEndangered" = "CR",
         "Endangered" = "EN",
         "Vulnerable" = "VU",
         "LowRisk" = "LC or LR/lc",
         "DataDeficient" = "DD") %>%
  select(`Name`,`Kingdom`,`Extinct`:`Subtotal (threatened spp.)`, `NearThreatened`, `LowRisk`:`Total`)
```

# Data Visualization

- creates new table
- filters only for classes with at least 1000 recorded species
- creates relative amount table

```{r data-visualization}

data_table_filtered <- manipulated_data %>%
  filter (Total > 1000) %>%
  transmute(`Name`,
            `Kingdom`,
            across(c(`Extinct`:`Total`), ~round ( .x / Total * 100, 2), .names = "{.col} [%]"))

```

## Barplot 1

- creates new column for unaffected species by combining the two columns that contain this information
- groups the data by kingdom and calculates the mean percentage for each category
- creates barplot

```{r barplot-1}

data_table_filtered %>%
  rowwise() %>%
  mutate(`Unaffected [%]` = sum(c_across(`NearThreatened [%]`:`DataDeficient [%]`))) %>%
  group_by(Kingdom) %>%
  summarise(Extinct = mean(`Subtotal (EX+EW) [%]`),
            Affected = mean(`Subtotal (threatened spp.) [%]`),
            Unaffected = mean(`Unaffected [%]`)) %>%
  pivot_longer(cols = c(Extinct, Affected, Unaffected),
               names_to = "Category",
               values_to = "Percentage") %>%
  filter(!is.na(Kingdom)) %>%
  ggplot() +
  geom_bar(aes( x = Kingdom, y = Percentage, fill = Category), stat = "identity", position = "stack") +
  ggtitle("Distribution of Impact Among Kingdoms") +
  xlab("Kingdom") +
  ylab("Average Percentage within Species") +
  scale_fill_manual(values = c("Extinct" = "red", "Affected" = "orange", "Unaffected" = "green"))

```

## Barplot 2

- creates the same table as above, but w/o the 1000 species filter (+ adds a new column for the total affected species)
- creates a new table with the top 10 and bottom 24 classes based on the total affected species
- adds a gap in the middle of the table to separate the two groups


```{r barplot-2-data-preparation}

library(dplyr)

data_table_unfiltered <- manipulated_data %>%
  transmute(`Name`,
            `Kingdom`,
            across(c(`Extinct`:`Total`), ~round ( .x / Total * 100, 2), .names = "{.col} [%]")) %>%
  rowwise() %>%
  mutate(`Affected [%]` = sum(c_across(`Extinct [%]`:`CriticallyEndangered [%]`))) 

data_top_ten <- data_table_unfiltered %>%
  ungroup() %>%
  arrange(desc(`Affected [%]`)) %>% 
  slice_head(n = 10)

data_bottom_twentyfour <- data_table_unfiltered %>%
  ungroup() %>%
  arrange(desc(`Affected [%]`)) %>% 
  slice_tail(n = 24)

gap_plot <- tibble(
  Name = c("gap_one", "gap_two", "gap_three"),
  Kingdom = NA_character_,
  `Affected [%]` = NA_real_)
  
ordered_data <- bind_rows(
  data_top_ten %>% 
    select(Name, Kingdom, `Affected [%]`), 
  gap_plot,
  data_bottom_twentyfour %>% 
    select(Name, Kingdom, `Affected [%]`))

```

---

- "locks" the order of the classes in the barplot by converting the `Name` column to a factor with levels set to the unique values of Name in the order they appear in `ordered_data`
- creates a barplot with the top 10 and bottom 24 classes based on the total affected species
- adds the gap in the middle to separate the two groups (`geom_point()`)

```{r barplot-2-mapping, warning = FALSE, fig.height = 7, fig.width = 10}

ordered_data %>%
  mutate(Name = factor(Name, levels = unique(Name))) %>%
  ggplot() +
  geom_col(aes(x = Name, y = `Affected [%]`, fill = Kingdom)) +
  geom_point(data = gap_plot, aes(x = Name, y = 50), size = 2) +
  scale_x_discrete(labels = function(x) ifelse(grepl("gap_", x), "", x)) +  # Hide gap labels
  scale_fill_grey(na.translate = FALSE) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  ggtitle("Most and Least Affected Classes") +
  xlab("Class")


```

