---
title: "emily-works"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(readr)
```

# Data Import
```{r data-import}
raw_data  <- read_csv("dependencies/SpeciesByKingdomAndClass.csv")
```
On the website you will see that they also have data on which Kingdom the various Classes belong to, we want to add this data semi-automatically.

Manually note the different Kingdoms and in what order they appear in the online table.

Use lag() and/or fill() alongside the fact that the table is ordered lexicographically by Class-name within each Kingdom to automatically find the boundaries between the different Kingdoms.

Add the Kingdom information in a new column named Kingdom

## Add Kingdom Data
```{r add-kingdom}
kingdoms <- c("Animalia", "Chromista", "Fungi", "Plantae") # list kingdoms in the order they appear in

kingdom_data  <- raw_data %>%
  mutate(
    ClassChange = Name < lag(Name), # create column that says TRUE when a change appears (new kingdom)
    ClassChange = replace_na(ClassChange, TRUE) # replaces NA in 1st observation with TRUE
    ) %>%
  mutate(
    Kingdom = ifelse(ClassChange == TRUE, kingdoms, NA)  
    # creates a new column, where TRUE there's a kingdom name
        ) %>%
  fill(Kingdom, .direction ="down") %>%
  # fills the NAs with the value
  select(!("ClassChange"))
  # removes the ClassChange column again


kingdom_data$Kingdom[75] <- NA 
# removes kingdom name from last row bc. its the total & therefor has no kingdom
```

# Data Manipulation
```{r data-manipulation}

manipulated_data <- kingdom_data %>%
  select(!("CR(PE)":"Subtotal (EX+EW+ CR(PE)+CR(PEW))")) %>%
  # removes columns
  rowwise() %>%
  mutate( 
    NearThreatened = sum(across("LR/cd":"NT or LR/nt"))
    # combines columns
        ) %>%
  select(!("LR/cd":"NT or LR/nt")
         # removes columns
         ) %>%
  rename(
    "Extinct" = "EX",
    "Extinct in the Wild" = "EW",
    "Critically Endangered" = "CR",
    "Endangered" = "EN",
    "Vulnerable" = "VU",
    "Low Risk" = "LC or LR/lc",
    "Data Deficient" = "DD"
    )
```

# Data Visualization














