---
title: "emily-works"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(readr)
```

# Data Import
```{r data-import, message = FALSE}
raw_data  <- read_csv("dependencies/SpeciesByKingdomAndClass.csv")
```

## Add Kingdom Data
```{r add-kingdom}

kingdoms <- c("Animalia", "Chromista", "Fungi", "Plantae") # list kingdoms in the order they appear in

kingdom_data  <- raw_data %>%
  mutate(
    ClassChange = Name < lag(Name),     # create column that says TRUE when a change appears (new kingdom)
    ClassChange = replace_na(ClassChange, TRUE),       # replaces NA in 1st observation with TRUE
    AllKingdom = ifelse(ClassChange == TRUE, kingdoms[cumsum(ClassChange)], NA)
  ) %>%
  fill(AllKingdom, .direction = "down") %>%      # fills the NAs with the value
  mutate(
    Kingdom = if_else(Name == "Total", NA, AllKingdom)
    ) %>%
  select(!c("ClassChange","AllKingdom"))     # removes the ClassChange & AllKingdom column 

```    


# Data Manipulation
```{r data-manipulation}

manipulated_data <- kingdom_data %>%
  select(!("CR(PE)":"Subtotal (EX+EW+ CR(PE)+CR(PEW))")) %>%  # removes columns
  rowwise() %>%
  mutate( 
    NearThreatened = sum(c_across("LR/cd":"NT or LR/nt"))  # combines columns
        ) %>%
  select(!("LR/cd":"NT or LR/nt")   # removes columns
         ) %>%
  rename(
    "Extinct" = "EX",
    "ExtinctWild" = "EW",
    "CriticallyEndangered" = "CR",
    "Endangered" = "EN",
    "Vulnerable" = "VU",
    "LowRisk" = "LC or LR/lc",
    "DataDeficient" = "DD"
    ) %>%
  select(`Name`,`Kingdom`,`Extinct`:`Subtotal (threatened spp.)`, `NearThreatened`, `LowRisk`:`Total`)     # changes column order

```

# Data Visualization

Unfortunately I couldn't figure out how the plots are supposed to work :(

First step: create a new seperate table
```{r data-visualization}

data_table <- manipulated_data %>%
  transmute(
    `Name`,
    `Kingdom`,
    across(c(`Extinct`:`Total`), ~ .x / Total * 100, .names = "{.col} [%]")
    )
```

Create barplot 1:


manipulated_data %>%
  mutate(
    Unaffected = sum(across(`LowRisk`:`NearThreatened`)) 
  ) %>%
  filter ( Total > 1000 ) %>%
  group_by(Kingdom) %>%
  summarise(
    Extinct = sum(`Subtotal (EX+EW)`),
    Affected = sum(`Subtotal (threatened spp.)`),
    Unaffected = sum(Unaffected)
  ) %>%
  ggplot() +
   geom_bar(aes( x = Kingdom, y = Total, fill = "Extinct"), stat = "identity", position = "stack")


```



