
---
title: "IUCN Red List Data Analysis – Species Conservation Status"
author: "Michael"
date: "`r Sys.Date()`"
output: 
  md_document:
    variant: markdown_github
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.align = "center")
library(dplyr)
library(tidyr)
library(ggplot2)
library(forcats)
library(knitr)
library(readr)   # <-- ADDED to fix read_csv() error
```

# Introduction

This analysis examines the **conservation status of species** across different taxonomic classes and kingdoms, based on data from the **IUCN Red List**. The IUCN Red List is the world's most comprehensive inventory of the global conservation status of biological species.

**What we will do:**

1. **Import and enrich the data** – add Kingdom information based on taxonomic classification.
2. **Clean and restructure** – remove non‑official categories, combine similar statuses, and rename columns for readability.
3. **Create a percentage table** – express each category as a percentage of the total species count per class.
4. **Visualise differences between Kingdoms** – compare the proportion of extinct, affected, and unaffected species across major kingdoms.
5. **Identify most and least affected Classes** – highlight the classes with the highest and lowest percentages of Critically Endangered species.

---

# 1. Data Import and Wrangling

## 1a) Import CSV into R

The dataset is loaded from the project's `dependencies` folder (portable path).

```{r import}
# Load from dependencies folder (portable)
redlist <- read_csv("~/Downloads/SpeciesByKingdomAndClass.csv")
```

**First rows of the dataset:**

```{r import_head}
head(redlist, 6) %>% kable(caption = "First 6 rows of the raw dataset")
```

**Explanation:** The table shows for each class the number of species in each IUCN category: `EX` (Extinct), `EW` (Extinct in the Wild), `CR` (Critically Endangered), `EN` (Endangered), `VU` (Vulnerable), `NT` (Near Threatened), `LC` (Least Concern), and `DD` (Data Deficient). The `Total` column gives the total number of assessed species for that class.

**Note:** The columns `CR(PE)` and `CR(PEW)` represent "Possibly Extinct" and "Possibly Extinct in the Wild" – these are *not* official IUCN categories and will be removed later.

---

## 1c) Add Kingdom Data

The original dataset does not contain a `Kingdom` column. Instead of using a static lookup table (which breaks when the dataset updates), we use a **dynamic, order‑based method**:

- The dataset is sorted alphabetically within each Kingdom.
- We identify Kingdom boundaries by detecting when a class name is **alphabetically "smaller"** than the previous class name (using `lag()`).
- We assign a Kingdom ID using `cumsum()`, then map IDs to Kingdom names.

```{r add_kingdom}
# Detect Kingdom boundaries based on alphabetical order
redlist <- redlist %>%
  mutate(
    # TRUE if the class name is alphabetically smaller than the previous row
    # (this marks the start of a new Kingdom)
    is_boundary = ifelse(
      lag(Name, default = "ZZZZ") > Name,
      1, 0
    ),
    # Cumulative sum gives a unique ID per Kingdom block
    Kingdom_id = cumsum(is_boundary)
  )

# Map Kingdom IDs to actual Kingdom names (in the order they appear in the data)
# The order in the IUCN table is: ANIMALIA, CHROMISTA, FUNGI, PLANTAE
kingdom_names <- c("ANIMALIA", "CHROMISTA", "FUNGI", "PLANTAE")

redlist <- redlist %>%
  mutate(Kingdom = kingdom_names[Kingdom_id + 1]) %>%  # +1 because IDs start at 0
  select(-is_boundary, -Kingdom_id)  # remove helper columns
```

**First rows after adding Kingdom:**

```{r kingdom_head}
head(redlist, 6) %>% kable(caption = "First 6 rows with Kingdom column added")
```

**Explanation:** The `Kingdom` column now tells us which kingdom each class belongs to. This allows us to group and compare species across the major kingdoms.

---

# 2. Data Manipulation

## 2a) Remove Unwanted Columns

The columns `CR(PE)` (Possibly Extinct) and `CR(PEW)` (Possibly Extinct in the Wild) are *not* official IUCN categories and are removed. The subtotal that includes them is also removed. The remaining subtotals are renamed for clarity.

```{r remove_cols}
redlist_clean <- redlist %>%
  select(-c("CR(PE)", "CR(PEW)", "Subtotal (EX+EW+ CR(PE)+CR(PEW))")) %>%
  rename(
    Extinct_Subtotal = "Subtotal (EX+EW)",
    Threatened_Subtotal = "Subtotal (threatened spp.)"
  )

head(redlist_clean, 6) %>% kable(caption = "After removing non‑official IUCN columns")
```

---

## 2c) Simplify Columns

`LR/cd` (Lower Risk – conservation dependent) and `NT or LR/nt` (Near Threatened) are combined into a single column called **Near Threatened**.

```{r combine_nt}
redlist_clean <- redlist_clean %>%
  mutate(
    `Near Threatened` = `LR/cd` + `NT or LR/nt`
  ) %>%
  select(-c("LR/cd", "NT or LR/nt"))

head(redlist_clean, 6) %>% kable(caption = "After combining Near Threatened categories")
```

---

## 2d) Make Column Names Readable

All column names are replaced with their full, human‑readable names.

```{r rename_cols}
redlist_clean <- redlist_clean %>%
  rename(
    Extinct = EX,
    Extinct_Wild = EW,
    Critically_Endangered = CR,
    Endangered = EN,
    Vulnerable = VU,
    Least_Concern = `LC or LR/lc`,
    Data_Deficient = DD
  )

head(redlist_clean, 6) %>% kable(caption = "After renaming columns to readable names")
```

**Resulting columns:** `Name`, `Kingdom`, `Extinct`, `Extinct_Wild`, `Critically_Endangered`, `Endangered`, `Vulnerable`, `Near Threatened`, `Least_Concern`, `Data_Deficient`, `Extinct_Subtotal`, `Threatened_Subtotal`, `Total`.

---

# 3. Data Visualisation

## 3a) Create Relative Amount Table

We create a new table `redlist_pct` where every numeric column (except `Total`) is expressed as a percentage of the row's `Total`. The `Total` column is set to 100.

```{r pct_table}
# Identify numeric columns (excluding Name, Kingdom, Total)
numeric_cols <- redlist_clean %>%
  select(where(is.numeric), -Total) %>%
  names()

# Create percentage table and rename columns in one pipe
redlist_pct <- redlist_clean %>%
  mutate(
    across(all_of(numeric_cols), ~ (. / Total) * 100),
    Total = 100
  ) %>%
  rename_with(~ paste0(.x, " [%]"), all_of(numeric_cols))

head(redlist_pct, 6) %>% kable(caption = "Relative amounts (percentages of Total)")
```

**Explanation:** In this table, each value shows the percentage of species in that category relative to the total number of species for that class. The `Total` column is always 100. This makes it easy to compare proportions across classes.

---

## 3b) Visualise Difference Between Kingdoms

To compare the conservation status across major Kingdoms, we create a **stacked bar plot**. Only Kingdoms with at least 1,000 species are included.

**Categories:**
- **Extinct:** `Extinct + Extinct_Wild` (EX + EW) – shown in **red**
- **Affected:** `Critically_Endangered + Endangered + Vulnerable` (CR + EN + VU) – shown in **orange**
- **Unaffected / No data:** `Least_Concern + Near Threatened + Data_Deficient` – shown in **blue**

```{r kingdom_plot}
# Prepare data: keep only Kingdoms with >= 1000 species
kingdom_plot_data <- redlist_clean %>%
  group_by(Kingdom) %>%
  filter(sum(Total, na.rm = TRUE) >= 1000) %>%
  summarise(
    Total_species = sum(Total, na.rm = TRUE),
    Extinct = sum(Extinct + Extinct_Wild, na.rm = TRUE),
    Affected = sum(Critically_Endangered + Endangered + Vulnerable, na.rm = TRUE),
    Unaffected = sum(Least_Concern + `Near Threatened` + Data_Deficient, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = c(Extinct, Affected, Unaffected),
               names_to = "Category",
               values_to = "Count") %>%
  group_by(Kingdom) %>%
  mutate(Percentage = Count / sum(Count) * 100)

# Create stacked bar plot with total species labels above bars
p_kingdom <- ggplot(kingdom_plot_data, aes(x = Kingdom, y = Percentage, fill = Category)) +
  geom_bar(stat = "identity", position = "stack") +
  # Add total species count above each bar
  geom_text(
    data = kingdom_plot_data %>% distinct(Kingdom, Total_species),
    aes(x = Kingdom, y = 105, label = paste0("n = ", Total_species)),
    inherit.aes = FALSE,
    size = 3.5
  ) +
  scale_fill_manual(
    values = c("Extinct" = "red", "Affected" = "orange", "Unaffected" = "blue"),
    labels = c("Extinct (EX + EW)", "Affected (CR + EN + VU)", "Unaffected / No data")
  ) +
  labs(
    title = "Conservation status by Kingdom",
    x = "Kingdom",
    y = "Percentage of species",
    fill = "Category"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )

print(p_kingdom)
```

**Interpretation:** The plot shows the proportion of species in each status category for each Kingdom.

- **Animalia** and **Plantae** have relatively high proportions of "Affected" species.
- **Fungi** and **Chromista** have very few species in the extinct or affected categories – largely because fewer species have been assessed.
- The total number of assessed species (`n = ...`) is shown above each bar.

---

## 3c) Visualise Most and Least Affected Classes

We define **"affected"** as the percentage of **Critically Endangered** species within a class. This plot shows the **10 most affected** and the **10 least affected** classes, colour‑coded by Kingdom.

```{r class_plot}
# Calculate Critically Endangered percentage per class
class_cr_pct <- redlist_clean %>%
  mutate(CR_percent = (Critically_Endangered / Total) * 100) %>%
  arrange(desc(CR_percent))

# Top 10 most affected
most_affected <- class_cr_pct %>%
  head(10) %>%
  mutate(Group = "Most Affected")

# Bottom 10 least affected (excluding classes with 0% CR)
least_affected <- class_cr_pct %>%
  filter(CR_percent > 0) %>%
  tail(10) %>%
  mutate(Group = "Least Affected")

# Combine both groups for plotting
plot_data <- bind_rows(most_affected, least_affected) %>%
  mutate(Class = fct_inorder(Name))  # Preserve the order

# Create faceted bar plot
p_classes <- ggplot(plot_data, aes(x = Class, y = CR_percent, fill = Kingdom)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ Group, scales = "free_x") +
  labs(
    title = "Most and least affected Classes (by % Critically Endangered)",
    x = "Class",
    y = "Critically Endangered species (%)",
    fill = "Kingdom"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )

print(p_classes)
```

**Interpretation:**

- **Most affected classes:** **AMPHIBIA** (Amphibians) and **ACTINOPTERYGII** (Ray‑finned fishes) have the highest percentages of Critically Endangered species. Many of the most affected classes belong to **Animalia**.
- **Least affected classes:** The least affected classes (with small but non‑zero CR percentages) include **CHAROPHYCEAE** (Charophytes) and **POLYTRICHOPSIDA** (a moss class), both belonging to **Plantae**. However, the low numbers often reflect limited assessments rather than genuinely low threat levels.

---

# Summary

- **Data import** and **Kingdom enrichment** were successful using a dynamic, order‑based method that is robust to future updates.
- **Cleaning** removed non‑official categories and consolidated the Near Threatened classification.
- The **percentage table** (`redlist_pct`) allows direct comparison of relative conservation statuses across classes.
- **Plot 1 (Kingdom comparison)** shows that the majority of assessed species are in the "Unaffected / No data" category, but **Animalia** and **Plantae** have notable affected proportions.
- **Plot 2 (Class comparison)** highlights **Amphibia** and **ACTINOPTERYGII** as the most affected classes.

---

## ⚠️ Important Notes

1. **File path:** The CSV file is loaded from `dependencies/SpeciesByKingdomAndClass.csv` – this is portable across systems.
2. **Data limitations:** The IUCN Red List data is incomplete for many groups. Fungi and Chromista have very few assessed species, so the percentages may not reflect true global patterns.
3. **Plots:** The plots are rendered as PNG images in the output folder.

```{r session_info, echo=FALSE}
sessionInfo()
```