---
title: "paula0013"
output: md_document
---

# Project 1: crop vs climate 
## solution by paula0013 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readxl)
library(readr)
library(magrittr)
library(tidyverse)
```

Climate change is a huge challenge for farmers worldwide. It is causing extreme weather events, such as droughts and floods, which can have devastating effects on crops and livestock. This project focuses on *crop production in the region of Tübingen and how it has been affected by climate change.*

# Read and transform the data set of the weather data from 1976 to 2021
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Read and transform the crop data for Tübingen
crop_data <- read.csv("https://www.openagrar.de/servlets/MCRFileNodeServlet/openagrar_derivate_00056476/Final_data.csv") %>% 
  #mutate the species 'ww' to 'winter wheat', 'sb' to 'spring barley' and 'grain_maize' to 'grain maize'
  mutate(
    var = case_when(
      var == "ww" ~ "winter wheat",
      var == "sb" ~ "spring barley",
      var == "grain_maize" ~ "grain maize",
      TRUE ~ var
    ))%>%
  # Filter the columns for 'Tübingen', the species 'winter wheat', 'spring barley' and 'grain maize' and the measure in 'yield'
  filter(
    district_no == '8416' & 
    (var == 'winter wheat' | var == 'spring barley' | var == 'grain maize') & 
    (measure == 'yield'))%>%
  select(2,4,5,7)%>%
   # Rename the column measure into measure in yield
  rename('measure in yield' = 'value')%>%
  # Rename the column var into species 
  rename('species' = 'var')

# Read and transform the precipitation data for regions in Germany
precipitation_BW <- read.table(
  "https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/annual/precipitation/regional_averages_rr_year.txt",
  header = TRUE,
  sep = ";",
  skip = 1,
  check.names = FALSE
)%>%
  # Select the column 1 with the years and column 5 with the BW data 
  select(1,5)%>%
  # Rename the column Jahr into year 
  rename('year'= 'Jahr')%>%
  # Rename the column Baden-Württemberg into Precipitation in mm
  rename('Precipitation in mm'= 'Baden-Wuerttemberg')%>%
  # Filter the Year from 1979 to 2021, becouse I have only for this years crop data
  filter(year >= '1979', year <= '2021')

# Read and transform the temperature data for regions in Germany
temperature_BW <- read.table(
  "https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/annual/air_temperature_mean/regional_averages_tm_year.txt",
  header = TRUE,
  sep = ";",
  skip = 1,
  check.names = FALSE
)%>%
  # Select the column 1 with the years and column 5 with the BW data 
  select(1,5)%>%
  # Rename the column Jahr into year 
  rename('year'= 'Jahr')%>%
  # Rename the column Baden-Württemberg into air temperature in °C
  rename('air temperature in °C'= 'Baden-Wuerttemberg')%>%
  # Filter the Year from 1979 to 2021, because I have only for this years crop data
  filter(year >= '1979', year <= '2021')

# Read and transform the frost days data for regions in Germany
frost_days_BW <- read.table(
  "https://opendata.dwd.de/climate_environment/CDC/regional_averages_DE/annual/ice_days/regional_averages_txcs_year.txt", 
  header = TRUE,
  sep = ";",
  skip = 1,
  check.names = FALSE
)%>%
  # Select the column 1 with the years and column 5 with the BW data 
  select(1,5)%>%
  # Rename the column Jahr into year 
  rename('year'= 'Jahr')%>%
  # Rename the column Baden-Württemberg into frost days in number of days 
  rename('frost days in number of days'= 'Baden-Wuerttemberg')%>%
  # Filter the Year from 1979 to 2021, because I have only for this years crop data
  filter(year >= '1979', year <= '2021')

# 
weather_data <- reduce(
  list(precipitation_BW, temperature_BW, frost_days_BW),
  left_join,
  by = "year"
)

data <- left_join(crop_data, weather_data, by = 'year')%>%
  as_tibble()

knitr::kable(data, caption = "Wetter- und Ertragsdaten Tübingen (1976–2021)")
```