---
title: "timow98"
author: "Timo Walcher"
date: "`r Sys.Date()`"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "https://cloud.r-project.org"))
if (!requireNamespace("pacman", quietly = TRUE)) {
  install.packages("pacman")} 
library(pacman)
p_load("tidyverse", "pdftools", "rio", "readxl", "utils")
```

## Data Wrangling
First of all I wanted to make the data more uniformly - I guess it is not really needed as you can just make a list by file type - still might be handy. There is probably a more efficient way
```{r, results='hide'}
# Create new directory
ifelse(!dir.exists("./Data/renamed"), dir.create("./Data/renamed"), FALSE)

# Here is a list of files from the pdf directory
pdflist <- list.files("./Data/PDF/",pattern="^FY\\d{2}\\D*\\.pdf$", full.names = TRUE)
pdflongyear <- list.files("./Data/PDF/",pattern="^FY\\d{4}.*\\.pdf$", full.names = TRUE)

# find years
years <- str_extract(as.numeric(str_extract(pdflongyear, pattern ="\\d{4}")),pattern ="\\d{2}$")

# rename the ones that have 4 year digits
file.copy(from = pdflongyear, 
            to = paste0("./Data/renamed/","FY",years,"_Table_III.pdf"))

crop <- str_extract(pdflist, pattern = "FY.*")
file.copy(from = pdflist, 
            to = paste0("./Data/renamed/", crop))

numbers <- str_extract(str_extract(list.files("./Data/renamed/", "\\.pdf$", full.names = TRUE),
                       str_extract(list.files("./Data/renamed/", "\\.pdf$", 
                                                          full.names = TRUE),
                                   "\\d{2}[^\\d]*")),"\\d{2}")

renamedlist <- list.files("./Data/renamed/",pattern="\\.pdf$", full.names = TRUE)
newname <- paste0("./Data/renamed/","FY", numbers, "_Table_III.pdf")

for (i in seq_along(renamedlist)) {
  file.rename(from = renamedlist[i], to = newname[i])
}
```
Then I wondered how to import the files. I failed at importing them from pdf files, so excel it is. The following code imports all sheets from all excel files.
```{r}
# Get list of filepaths
xlsx_list <- list.files("./Data/Excel", pattern = "\\.xlsx$")
xlsx_list <- paste0("./Data/Excel/", xlsx_list)
xlsx_list <- unlist(xlsx_list)

read_all_sheets <- function(files) {
  # iterate over files and get sheet names
  for (i in seq_along(files)) {
    sheet_names <- excel_sheets(files[i])
    
    # then interate over sheet names and read into df
    for (j in seq_along(sheet_names)) {
      df <- read_excel(files[i], sheet = sheet_names[j])
      # assign name (df+(year-1)+sheetnumber)
      assign(paste0("df",i-1, "_", j), df, envir = .GlobalEnv)
    }
  }
}

# call function to import all sheets from all excel files
read_all_sheets(xlsx_list)
```
Not really sure if that is what I would need to go on, but I still learnt a lot :D
Next we could try to iterate over the dataframes to get specific columns. Instead of storing the data in separate dfs, I could probably also make a list.