---
title: "week4toothgrowth"
output: html_document
---

####################################################################
#' Let's practice our ggplot plotting skills. To this end, we use the
#' "ToothGrowth" data set directly available within R (check help).
#' We want to plot for the different dose levels (X) the respective
#' tooth length. The dose levels should be represented as follows:
#' - 0.5 = non-filled circle in a dark green
#' - 1 = orange filled triangles
#' - 2 = filled circles in a light blue
#' Use jittering to avoid overplotting of data points.
#' We don't want to have any background decorations etc. and useful
#' labels for axes and legend.
#' Ensure there is only ONE legend (showing both color and shape).
####################################################################

* jittering ist wenn man Datenpunkte ein wenig verschiebt wenn nur eine Achse skaliert

```{r tooth}
ggplot(ToothGrowth, aes(x = dose, y = len, shape = factor(dose), color = factor(dose)))+
  geom_point(size = 2)+
  scale_shape_manual(values = c(1, 17, 16))+
  scale_color_manual(values = c("darkgreen", "orange", "lightblue"))+
  geom_jitter(width= 0.2)+
  labs(x= "Dose levels",
        y = "Tooth length")+
  theme_dark()
```

* dose ist als dbl gespeichert nicht als Faktor, ggplot macht je nach Datentyp
verschiedene Sachen
* numerische Werte produzieren Skalen 
* bei Faktorisierung wird aus nem Wert ein Text -> keine Skala mehr sondern eine
Kategorie
* Kategorien kann man beliebig anordnen
* man kann die values auch im shape vektor einer kategorie zuweisen
* Trick für gleiche Legende: wenn dinge den gleichen legenden namen haben, 
dann werden Legenden zusammengeschoben
* Jittering explizit benennen




```{r martinlösung}
ToothGrowth |> 
  mutate(dose = factor(dose)) |> 
  
  ggplot( aes(x = dose, y = len, col=dose, shape=dose ) ) +
  
  geom_jitter(width = 0.2) +
  
  scale_color_manual(
    # name = "Dose (mg/day)",
    values = c(
                "0.5" = "darkgreen", # Outline color
                "1"   = "orange",
                "2"   = "lightblue"
              )
  ) +
  
  labs(
    x = "Dose (mg/day)",
    col = "Dose (mg/day)",
    shape = "Dose (mg/day)",
    y = "Tooth length"
  ) +
  
  theme_minimal()
```


