---
title: "Solution for Jan's project"
author: "Sarah Löber"
date: "2024-07-01"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Solution for Jan's project

This is my attempt for a solution for Jan's project. I will use the dataset that Jan provided.

### Libraries
```{r libraries}
library(tidyverse)
```

### Read in data
```{r data, warning=FALSE}
# read in dataset
food_data <- read.csv("/Users/sarahlober/Data Projects with R and GitHub/Projects/jan-thiele7/data_openfood.csv")
head(food_data)
```

I did not quite grasp the categories in the categorization variables as the head does not tell me much about them, so I'm taking a look at them first.

```{r variables, warning=FALSE}
table(food_data$pnns_groups_1)
table(food_data$pnns_groups_2)
```

### Cleaning the data

I'm getting the data ready for plotting first, then create the actual plot. I'm removing the unknown / not-applicable categories, drop the NAs (if there are any) and change the groups to factor. 

```{r plot data, warning=FALSE}

df <- food_data %>%
  drop_na(pnns_groups_1, nutriscore_grade, ecoscore_grade) %>%
  filter(
    pnns_groups_1 != "unknown" & 
    nutriscore_grade != "unknown" & 
    ecoscore_grade != "unknown" & 
    nutriscore_grade != "not-applicable" & 
    ecoscore_grade != "not-applicable") %>%
  mutate(pnns_groups_2 = factor(pnns_groups_2)
  )
```

Then I had to ask ChatGPT for advice on how to tackle this plot since I was clueless about the shape I needed to put the data in in order to achieve a plot like in the example. 
```{r visualize, warning=FALSE}

```


```{r plot nutriscore, warning=FALSE}

```
