---
title: "Solution for Jan's project"
author: "Sarah Löber"
date: "2024-07-01"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Solution for Jan's project

This is my attempt for a solution for Jan's project. I will use the dataset that Jan provided.

### Libraries
```{r libraries}
library(tidyverse)
library(ggrepel)
```

### Read in data
```{r data, warning=FALSE}
# read in dataset
food_data <- read.csv("./data_openfood.csv")
head(food_data)
```
```

### Cleaning the data

I'm getting the data ready for plotting first, then create the actual plot. I'm removing the unknown / not-applicable categories, drop the NAs (if there are any) and change the groups to factor. I am also summarising groups together so I have less groups in the plot later (which makes it easier to read compared to my first version.)

```{r plot data, warning=FALSE}

df <- food_data %>%
  drop_na(pnns_groups_1, nutriscore_grade, ecoscore_grade) %>%
  filter(
    pnns_groups_1 != "unknown" & 
    nutriscore_grade != "unknown" & 
    ecoscore_grade != "unknown" & 
    nutriscore_grade != "not-applicable" & 
    ecoscore_grade != "not-applicable"
  ) %>%
  mutate(
    pnns_groups_2 = factor(pnns_groups_2),
    pnns_groups_2 = case_when(
      pnns_groups_2 %in% c("Offals", "Meat") ~ "Meat",
      pnns_groups_2 %in% c("Fruit juices", "Fruit nectars") ~ "Fruit juices",
      pnns_groups_2 %in% c("Cereals", "Breakfast cereals") ~ "Cereals",
      pnns_groups_2 %in% c("Teas and herbal teas and coffees", "Unsweetened beverages") ~ "Unsweetened beverages",
      pnns_groups_2 %in% c("Plant-based milk substitutes", "Milk and yogurt") ~ "Milk and yogurt",
      pnns_groups_2 %in% c("Ice cream", "Dairy desserts") ~ "Dairy desserts",
      pnns_groups_2 %in% c("Biscuits and cakes", "Pastries") ~ "Pastries",
      pnns_groups_2 %in% c("Pizza pies and quiches") ~ "Pizza",
      pnns_groups_2 %in% c("Artificially sweetened beverages") ~ "Sweetened beverages",
      pnns_groups_2 %in% c("Dried fruit", "Nuts") ~ "Nuts and dried fruit",
      pnns_groups_2 %in% c("Sandwiches", "Soups") ~ "Sandwiches and soups",
      pnns_groups_2 %in% c("One-dish meals", "Salty and fatty products") ~ "Salty and fatty products",
      TRUE ~ as.character(pnns_groups_2)
    )
  )

```

For this next part, I did not really know how to approach it. I took your idea from the description and started off with a heatmap, however, that did not really work out as I wanted. The rownames have to be unique, therefore cannot be grouped and a matrix takes numeric variables only. It looks cool but did not help me much. 

```{r visualize, warning=FALSE}
# convert data to matrix 
df1 <- df %>%
  select(-c(X, countries_en, product_name, pnns_groups_1))

df1$nutriscore <- as.numeric(as.factor(df1$nutriscore_grade))
df1$ecoscore <- as.numeric(as.factor(df1$ecoscore_grade)) 
df1$pnns_group <- as.numeric(df1$pnns_groups_2) 

rn <- df1[,3]
rn <- make.unique(as.character(rn))
df1 <- df1[,-3]

df1 <- df1 %>%
  select(-c(nutriscore_grade, ecoscore_grade))
rownames(df1) <- rn

matrix <- as.matrix(df1)
heatmap(matrix, scale="column")
```
So I had ChatGPT do it in the end...
```{r data visualization}
nutriscore_levels <- c("a" = 1, "b" = 2, "c" = 3, "d" = 4, "e" = 5)
ecoscore_levels <- c("a" = 1, "b" = 2, "c" = 3, "d" = 4, "e" = 5)

# Compute the centers and sizes of the bubbles
plot_data <- df %>%
  group_by(pnns_groups_2, nutriscore_grade, ecoscore_grade) %>%
  summarise(count = n()) %>%
  ungroup() %>%
  mutate(nutriscore_value = nutriscore_levels[nutriscore_grade],
         ecoscore_value = ecoscore_levels[ecoscore_grade])

# Compute the size of bubbles based on the number of unique combinations within each group
bubble_size <- df %>%
  group_by(pnns_groups_2) %>%
  summarise(size = n_distinct(nutriscore_grade, ecoscore_grade))

# Merge bubble size with plot data
plot_data <- plot_data %>%
  left_join(bubble_size, by = "pnns_groups_2") %>%
  group_by(pnns_groups_2) %>%
  slice_max(order_by = count, n = 1) %>%
  ungroup()
```

And this is the plot that came out of it. The labels are a bit tricky since they overlap and that does not look great in the plot. I changed the size of the circle around the circle so the colours are clearer.
```{r plot nutriscore, warning=FALSE}
nutriscore_colors <- c("a" = "#009e73", "b" = "#d6db32", "c" = "#f7c239", "d" = "#e06666", "e" = "#e21a1a")
ecoscore_colors <- c("a" = "#009e73", "b" = "#d6db32", "c" = "#f7c239", "d" = "#e06666", "e" = "#e21a1a")

# Create the plot
ggplot(plot_data, aes(x = ecoscore_value, y = nutriscore_value, size = size, label = pnns_groups_2)) +
  geom_point(aes(color = nutriscore_grade, fill = ecoscore_grade), alpha = 0.6, shape = 21, stroke = 2, 
             position = position_jitter(width = 0.3, height = 0.3)) +
  scale_size_area(max_size = 14, guide = 'none') +
  scale_color_manual(values = nutriscore_colors) +
  scale_fill_manual(values = ecoscore_colors) +
  scale_x_continuous(breaks = 1:5, labels = names(ecoscore_levels)) +
  scale_y_continuous(breaks = 1:5, labels = names(nutriscore_levels)) +
  theme_minimal() +
  labs(
    title = "Relationship between Nutri- and Ecoscore",
    x = "Ecoscore",
    y = "Nutriscore"
  ) +
  geom_text_repel(size = 2, box.padding = 0.3, point.padding = 0.2)

```
