---
title: "Chaoicci"
output: md_document
---
```{r prerequisites, include = FALSE}
library(tidyverse)
library(readr)
library(ggplot2)
```


```{r data_import-and-cleanup, include = FALSE}
data <- read_csv2("sciFiMovies_dataset.csv")

data_clean <- data |> 
  mutate(`Release Year` = Release_Year,
         `Movie Rating` = Movie_Rating,
         `Gross collection` = Gross_collection |> as.double() |> round(),
         `Gross equivalent` = Gross_equivalent |> as.double() |>  round(),
         Votes = Votes |> str_remove_all(",") |> as.double() |>  round(),
         Genre = Genre |> str_split(",")) |> 
    select(-c(`...1`, 
              Release_Year, 
              Movie_Rating, 
              Gross_collection, 
              Gross_equivalent))
  

```

# Data manipultion Tasks
Note: I calculated the rate of inflation with this formula: `Rate of Inflation = (Gross equivalent / Gross collection) - 1`
```{r data-manipulation, include = FALSE}
data_full <- data_clean |> 
  group_by(`Release Year`) |> 
  mutate(`Overall Score` = `Movie Rating` + Metascore,
         `Rate of Inflation` = ((`Gross equivalent` |> sum()) / 
           (`Gross collection` |> sum())) - 1) |> 
  ungroup()
```
The new information in the table looks somewhat like this:
```{r table, echo = FALSE}
knitr::kable(
  data_full |>
    select(c(Title, `Movie Rating`, Metascore, `Overall Score`, `Release Year`, `Rate of Inflation`)) |> 
    head())
```


# Data visualization Tasks:
## The relationship between movie rating and metascore
Note: I put `Movie Rating` on the x, since there is no variable called 
`Movie_Score` and context suggests that `Movie Rating` was intended.

```{r correlation-plot, echo = FALSE, warning = FALSE, error = FALSE}
data_full |> 
  filter(!is.na(Metascore)) |> 
  ggplot() +
  geom_point(aes(x = `Movie Rating`, y = Metascore)) +
  geom_segment(x = 25, y = 0, xend = 90, yend = 100, color = "red", linewidth = 1) +
  theme_classic()
```

There seems to be a positive correlation. This would suggest that Movies with
a high Movie Rating also tend to have a high Metascore.

The relationship of `Movie Rating` and `Metascore` by `Genre`:
```{r correlation-by-genre, echo = FALSE}
data_full |> 
  filter(!is.na(Metascore)) |> 
  unnest(Genre) |> 
  ggplot() +
  geom_point(aes(x = `Movie Rating`, y = Metascore)) +
  facet_wrap(~ Genre)
```

## How do different directors score?
Note: I only included the directors with the most movies (The 35 directors with the most movies).
```{r directors_plot, echo = FALSE}
data_full |> 
  mutate(Director = Director |> str_remove("Director.*:")) |> 
  drop_na(`Overall Score`) |> 
  group_by(Director) |> 
  summarize(Films = n(),
         `Average Score` = mean(`Overall Score`),
         `SD Score` = sd(`Overall Score`)) |> 
  slice_max(Films, n = 35) |> 
  ggplot(aes(x = Director, y = `Average Score`)) +
  geom_col() + 
  geom_errorbar(aes(ymin = `Average Score` - `SD Score`, 
                    ymax = `Average Score` + `SD Score`),
                color = "red") +
  geom_text(aes(label = str_c("Films: ", Films)), hjust = -2) +
  labs(y = "Combined Score") +
  theme_classic() +
  ylim(c(0, 250)) +
  coord_flip()
```

