---
title: "Data Projects with R and GitHub: Movies and Inflation"
author: "copilot"
date: "2026-07-22"
output: md_document
---

# Movies and Inflation

Movies are a big and important part of our culture. They are a source of
entertainment, a way to tell stories, and a reflection of society. But they are
also a big business, with billions of dollars in revenue generated each year. In
this project, we explore the relationship between movies and inflation using a
dataset that contains information about movies released between 1970 and 2020,
including their ratings, box office performance, and inflation-adjusted gross
collections.

---

## Setup

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
```

---

## Data Loading

The dataset is loaded directly from Zenodo without downloading it locally.

```{r load-data}
movies_raw <- read.csv(
  "https://zenodo.org/records/5653247/files/sciFiMovies_dataset.csv",
  sep       = ";",
  header    = TRUE,
  stringsAsFactors = FALSE,
  check.names = FALSE
)

# Quick overview
glimpse(movies_raw)
```

---

## Data Cleanup

### Step 1 – First column: check uniqueness and remove or assign ID

```{r cleanup-first-col}
# The first column is an unnamed index column (values 0, 1, 2, …).
# Check whether rows are unique without it.
col1_name <- names(movies_raw)[1]   # empty string or "X"
movies_no_first <- movies_raw[, -1]

if (nrow(distinct(movies_no_first)) == nrow(movies_raw)) {
  message("Rows are unique without the first column → removing it.")
  movies <- movies_no_first
} else {
  message("Rows are NOT unique without the first column → keeping it as row ID.")
  movies <- movies_raw %>%
    rename(ID = all_of(col1_name))
}
```

### Step 2 – Remove special characters from variable names

```{r cleanup-varnames}
# Standardise column names: replace spaces / special characters with underscores
names(movies) <- gsub("[^A-Za-z0-9_]", "_", names(movies))
names(movies) <- gsub("__+", "_", names(movies))   # collapse repeated underscores
names(movies) <- gsub("_$", "", names(movies))      # remove trailing underscores
glimpse(movies)
```

### Step 3 – Round numeric values (eliminate decimals)

```{r cleanup-round}
# The Votes column stores numbers with comma thousands-separators ("888,938")
# → remove commas and coerce to numeric first.
movies <- movies %>%
  mutate(
    Votes            = as.numeric(gsub(",", "", Votes)),
    Gross_collection = round(as.numeric(Gross_collection)),
    Gross_equivalent = round(as.numeric(Gross_equivalent))
  )
```

### Step 4 – Ensure Movie_Rating is numeric (not a date)

```{r cleanup-rating-type}
cat("Current class of Movie_Rating:", class(movies$Movie_Rating), "\n")

if (!is.numeric(movies$Movie_Rating)) {
  movies <- movies %>%
    mutate(Movie_Rating = as.numeric(Movie_Rating))
  cat("Movie_Rating converted to numeric.\n")
} else {
  cat("Movie_Rating is already numeric – no change needed.\n")
}
```

### Step 5 – Disassemble Genre into a list of individual genres

```{r cleanup-genre-nest}
# Split the comma-separated genre string into a character vector per row,
# then store as a list-column via nest().
movies_nested <- movies %>%
  mutate(Genre = strsplit(Genre, ",\\s*")) %>%   # split into a character vector per row
  group_by(across(-Genre)) %>%
  nest(genres = Genre) %>%
  ungroup()

# Inspect the first few entries
head(movies_nested$genres, 3)
```

---

## Data Manipulation

### Best movie: highest combined rating score

```{r manip-combined-score}
# Create a new variable as the sum of Movie_Rating and Metascore.
# Note: the two scales differ (0–10 vs. 0–100), so we interpret the task
# as a direct sum that can be used for relative ranking.
movies <- movies %>%
  mutate(
    Metascore    = as.numeric(Metascore),
    Movie_Rating = as.numeric(Movie_Rating),
    Combined_Score = Movie_Rating + Metascore
  )

best_movie <- movies %>%
  slice_max(Combined_Score, n = 1) %>%
  select(Title, Release_Year, Movie_Rating, Metascore, Combined_Score)

print(best_movie)
```

### Inflation rate per year

```{r manip-inflation}
# For each year: rate of inflation based on the difference between the
# inflation-adjusted gross (Gross_equivalent) and the nominal gross
# (Gross_collection).
# Formula: inflation_rate = (sum_equivalent - sum_gross) / sum_gross * 100

inflation_by_year <- movies %>%
  group_by(Release_Year) %>%
  summarise(
    sum_gross      = sum(Gross_collection, na.rm = TRUE),
    sum_equivalent = sum(Gross_equivalent, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    Inflation_Rate = (sum_equivalent - sum_gross) / sum_gross * 100
  ) %>%
  arrange(Release_Year)

print(inflation_by_year)
```

The inflation rates are highest for the oldest years (1970s) because the dollar
has lost significant purchasing power over the decades. As we approach more recent
years, the difference between nominal and inflation-adjusted gross collections
shrinks, and the implied inflation rate decreases accordingly.

---

## Data Visualization

### 1. Scatter plot: Movie_Rating vs. Metascore

```{r viz-scatter, fig.width=7, fig.height=5}
ggplot(movies, aes(x = Movie_Rating, y = Metascore)) +
  geom_point(alpha = 0.4, color = "steelblue") +
  geom_smooth(method = "lm", se = TRUE, color = "firebrick", linewidth = 0.8) +
  labs(
    title   = "Correlation between IMDB Rating and Metascore",
    x       = "IMDB Movie Rating",
    y       = "Metascore",
    caption = "Each point represents one movie; red line shows the linear trend."
  ) +
  theme_minimal()
```

**Interpretation:** The scatter plot shows a moderate positive correlation between
the IMDB Movie Rating and the Metascore. Movies with higher IMDB ratings tend to
also receive higher Metascores, indicating that both rating systems largely agree
on movie quality, although there is considerable spread around the trend line.

### 2. Scatter plot faceted by genre

```{r viz-scatter-facets, fig.width=12, fig.height=10, warning=FALSE}
# Unnest the genre list-column so every genre–movie combination gets its own row
movies_unnested <- movies %>%
  mutate(Genre = strsplit(Genre, ",\\s*")) %>%
  unnest(Genre) %>%
  mutate(Genre = str_trim(Genre))

ggplot(movies_unnested, aes(x = Movie_Rating, y = Metascore)) +
  geom_point(alpha = 0.3, color = "steelblue", size = 0.8) +
  geom_smooth(method = "lm", se = FALSE, color = "firebrick", linewidth = 0.6) +
  facet_wrap(~Genre) +
  labs(
    title = "IMDB Rating vs. Metascore by Genre",
    x     = "IMDB Movie Rating",
    y     = "Metascore"
  ) +
  theme_minimal() +
  theme(strip.text = element_text(size = 7))
```

### 3. Horizontal bar plot: average combined score by director

```{r viz-director-bars, fig.width=10, fig.height=14}
# Clean up the Director column (raw format: "Director:Name" or "Directors:Name1, Name2")
# For simplicity, extract the first director's name.
director_stats <- movies %>%
  mutate(
    Director = str_remove(Director, "^Directors?:\\s*"),  # remove prefix
    Director = str_trim(str_extract(Director, "^[^,]+"))  # first director only
  ) %>%
  group_by(Director) %>%
  summarise(
    mean_score = mean(Combined_Score, na.rm = TRUE),
    sd_score   = sd(Combined_Score,   na.rm = TRUE),
    n_movies   = n(),
    .groups    = "drop"
  ) %>%
  arrange(mean_score) %>%
  mutate(Director = factor(Director, levels = Director))  # preserve order in plot

ggplot(director_stats,
       aes(x = mean_score, y = Director)) +
  geom_col(fill = "steelblue", width = 0.7) +
  geom_errorbar(
    aes(xmin = mean_score - sd_score,
        xmax = mean_score + sd_score),
    width = 0.3, color = "gray30"
  ) +
  geom_text(
    aes(label = paste0("n=", n_movies)),
    hjust = -0.2, size = 2.8, color = "gray20"
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Average Combined Score by Director",
    x     = "Average Combined Score (IMDB + Metascore)",
    y     = NULL
  ) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 7))
```

---

## Optional Tasks

### Highlight a favourite actor's movies in the scatter plot

```{r viz-actor-highlight, fig.width=7, fig.height=5}
# Example: highlight all movies featuring Leonardo DiCaprio
favourite_actor <- "Leonardo DiCaprio"

movies_actor <- movies %>%
  mutate(
    Featured = str_detect(Star, fixed(favourite_actor))
  )

ggplot(movies_actor, aes(x = Movie_Rating, y = Metascore, color = Featured)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(
    values = c("FALSE" = "gray70", "TRUE" = "tomato"),
    labels = c("Other movies", paste0(favourite_actor, "'s movies"))
  ) +
  geom_smooth(data = filter(movies_actor, !Featured),
              method = "lm", se = FALSE, color = "gray40", linewidth = 0.6) +
  labs(
    title  = paste("Movies featuring", favourite_actor, "highlighted"),
    x      = "IMDB Movie Rating",
    y      = "Metascore",
    color  = NULL
  ) +
  theme_minimal()
```

### Does a good rating correlate with higher earnings?

```{r viz-score-vs-gross, fig.width=7, fig.height=5}
ggplot(movies, aes(x = Gross_equivalent, y = Combined_Score)) +
  geom_point(alpha = 0.3, color = "steelblue") +
  geom_smooth(method = "lm", se = TRUE, color = "firebrick", linewidth = 0.8) +
  scale_x_continuous(labels = scales::label_number(scale = 1e-6, suffix = "M")) +
  labs(
    title   = "Do highly-rated films make more money?",
    x       = "Inflation-adjusted Gross (millions USD)",
    y       = "Combined Score (IMDB + Metascore)",
    caption = "Positive slope would suggest higher-rated films earn more."
  ) +
  theme_minimal()
```

**Interpretation:** This scatter plot reveals whether critically and publicly
acclaimed movies also tend to be bigger box-office successes. A weak or moderate
positive correlation is expected, but many blockbusters achieve enormous revenues
despite moderate ratings.
