---
title: "Solution by Joshua"
output: md_document
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(lubridate)
library(tidyquant)

visual_dir <- "joshuaschlucke-visualisation"
if (!dir.exists(visual_dir)) dir.create(visual_dir)

knitr::opts_chunk$set(
    fig.path = file.path(visual_dir, ""),
    fig.width = 11,
    fig.height = 7,
    dpi = 300
)
```

# 1. Data reading & cleaning

In its current state, the data is impossible to work with.

- **Reading the data**: The file should be read twice, once to generate sensible column names from the multi-row header and once to read the data itself (ignoring the headers. For this, `read_csv` features the argument `skip`).

  Make sure all data types are as expected (eg. `Datetime` should be a datetime object, `Open` should be a number, ...)

- **Formatting**: Pivot this dataframe to get a dataframe with the following columns:

  - `Datetime`
  - `Symbol`
  - `Type` (Factor with two levels: `Stock` and `Index`)
  - `Open`
  - `High`
  - `Low`
  - `Close`
  - `Volume`

  This should result in one very tidy table.

- **Aggregation**: To make plotting the data simpler, create a new dataframe where the data is aggregated from hourly to daily data.

  > the new `Open` should be the `Open` of the first hour of the day
  >
  > the new `Close` should be the `Close` of the last hour of the day
  >
  > the new `High` should be the maximum `High` of all hours of the day
  >
  > the new `Low` should be the minimum `Low` of all hours of the day
  >
  > the new `Volume` should be the sum of all hourly volumes of the day

```{r read-and-clean, warning=FALSE, include=FALSE}
# Read the two header rows and build column names in one step
header_rows <- readr::read_csv(
    "stock_data.csv",
    col_names = FALSE,
    n_max = 2,
    show_col_types = FALSE
) %>%
  t() %>%
  {str_c(.[,1],.[,2], sep = "_")} %>%
  str_replace(.,"Price_Ticker","Datetime")

# Read the actual data, skipping the header rows
raw_data <- readr::read_csv(
    "stock_data.csv",
    skip = 3,
    col_names = header_rows,
    col_types = cols(
        Datetime = col_datetime(),
        .default = col_double()
    )
)

# Pivot to tidy format
indices <- c("^GSPC", "^IXIC")

long_data <- raw_data %>%
    pivot_longer(
        cols = -Datetime,
        names_to = c("Metric", "Symbol"),
        names_sep = "_",
        values_to = "Value"
    ) %>%
    pivot_wider(
        names_from = Metric,
        values_from = Value
    ) %>%
    drop_na() %>%
    mutate(
        Type = if_else(Symbol %in% indices, "Index", "Stock"),
        Type = factor(Type, levels = c("Stock", "Index"))
    ) %>%
    select(Datetime, Symbol, Type, Open, High, Low, Close, Volume)

# Aggregate hourly data to daily data

daily_data <- long_data %>%
    arrange(Symbol, Datetime) %>%
    group_by(Symbol, Type, Date = as_date(Datetime)) %>%
    summarise(
        Open = first(Open),
        Close = last(Close, na_rm = TRUE),
        High = max(High, na.rm = TRUE),
        Low = min(Low, na.rm = TRUE),
        Volume = sum(Volume, na.rm = TRUE),
        .groups = "drop"
    )

```

# 2. Visualization

- **Price**: Use `ggplot2`s facetting and `geom_line` to plot the daily price of all stocks and indices over time. You can use `Open` or `Close` for this.

- **Volume**: Either in the same plots as above or in a new plot, plot the trading volume of all stocks and indices over time as barplots (`geom_bar`).

- **Optional: Candlesticks**: Instead of using a line chart to plot the price, use candlesticks (eg. using `geom_boxplot` or `geom_segment` and `geom_rect`) to represent the daily price data for all stocks.

  > The top of the body is the open price, the bottom of the body is the close price, and the wicks are the high and low prices.
  >
  > A candlestick is red if the price decreased during the day, and green if it increased.

## Daily price (Close)

```{r plot-price, warning=FALSE}
price_plot <- ggplot(daily_data, aes(x = Date, y = Close, color = Type)) +
    geom_line(linewidth = 0.6, alpha = 0.9) +
    facet_wrap(~Symbol, scales = "free_y") +
    labs(
        title = "Daily Close Price by Symbol",
        x = "Date",
        y = "Close Price",
        color = "Type"
    ) +
    theme_minimal()

price_plot
```

## Daily volume

```{r plot-volume, warning=FALSE}
volume_plot <- ggplot(daily_data, aes(x = Date, y = Volume, fill = Type)) +
    geom_col(alpha = 0.75) +
    facet_wrap(~Symbol, scales = "free_y", ncol = 1) +
    labs(
        title = "Daily Trading Volume by Symbol",
        x = "Date",
        y = "Volume (Millions)",
        fill = "Type"
    ) +
    scale_y_continuous(labels = scales::label_number(suffix = "M", scale = 1e-6)) +
    theme_minimal()

volume_plot
```

## Optional: Candlesticks (stocks only)
Candlestick plot with tidyquant, but still without whiskers (no matter the scaling) because they are only shown on short time frames or really zoomed in.
```{r plot-candlesticks, warning=FALSE, fig.asp=1.5}
candlestick_plot <- daily_data %>%
    filter(Type == "Stock") %>%
    ggplot(aes(x = Date, open = Open, high = High, low = Low, close = Close)) +
    geom_candlestick(
        colour_up = "#2E8B57", colour_down = "#B22222",
        fill_up   = "#2E8B57", fill_down  = "#B22222", size = 0.8
    ) +
    facet_wrap(~Symbol, scales = "free_y", ncol = 1) +
    labs(
        title = "Daily Candlesticks (Stocks Only)",
        x = "Date",
        y = "Price"
    ) +
    theme_minimal()

candlestick_plot
```

# 3. Pattern analysis & correlation (hourly)

For this part we will go back to hourly data, as we want to analyze patterns throughout the day.

- Compute the percentage change in price (between `Open` or `Close`) for each hour (compared to the previous hour) for all stocks and indices. This will be used in the following tasks.

- **Price Movements/Volatility**: At which hour of the day do prices across all stocks & indices move/change the most? (In this case: When is the absolute percentage change between the `Open` price and the `Close` price the highest on average?)

  > Visualize the average price movement per hour of the day here (eg. using a bar plot).

- **Correlation**: Take a look at how closely stock and index prices follow each other. Are they correlated?

  - **Visualization**: Plot all stock prices against the S&P 500 price to visually inspect their correlation (line or point plot). Its up to you whether you use individual facets per stock or plot all in one plot. Do they (stocks and the index) seem to move together?
  - **Optional: Quantification**: Quantify the correlation between stock and index price changes (percentage changes from above) by calculating the correlation coefficient. R has a built-in function for this, called `cor` (check `?cor`, usage hint: `cor(df$stock_change, df$index_change)`). You can do this for each stock/S&P 500 pair and plot the results to see which stock correlates the most with the S&P 500.

```{r hourly-analysis, warning=FALSE}
# Hourly percentage change in Close vs previous hour
# Calculate changes within a trading day to avoid overnight gaps.
hourly_changes <- long_data %>%
    mutate(
        Datetime_ny = with_tz(Datetime, "America/New_York"),
        Date_ny = as_date(Datetime_ny)
    ) %>%
    arrange(Symbol, Datetime) %>%
    group_by(Symbol, Type, Date_ny) %>%
    mutate(
        pct_change = (Close - lag(Close)) / lag(Close),
        hour = hour(Datetime_ny),
        minute = minute(Datetime_ny),
        intrahour_move = abs((Close - Open) / Open)
    ) %>%
    ungroup()

# Filter to regular US trading hours (09:30–16:00 New York time)
hourly_trading <- hourly_changes %>%
    filter(
        (hour > 9 & hour < 16) |
            (hour == 9 & minute == 30) |
            (hour == 16 & minute == 0)
    )

# Average absolute intrahour movement by hour across all symbols
hourly_volatility <- hourly_trading %>%
    group_by(hour) %>%
    summarise(
        avg_abs_move = mean(intrahour_move, na.rm = TRUE),
        .groups = "drop"
    )
```

## Price movements/volatility by hour

```{r plot-hourly-volatility, warning=FALSE}
vol_plot <- ggplot(hourly_volatility, aes(x = factor(hour), y = avg_abs_move)) +
    geom_col(fill = "#4C72B0") +
    labs(
        title = "Average Absolute Intrahour Price Movement by Hour (US/Eastern)",
        x = "Hour of Day (US/Eastern)",
        y = "Average |(Close - Open) / Open|"
    ) +
    theme_minimal()

vol_plot
```

## Correlation with S&P 500

```{r correlation-analysis, warning=FALSE}
# Visual inspection: stock prices vs S&P 500 (hourly close)
sp500_hourly <- long_data %>%
    filter(Symbol == "^GSPC") %>%
    select(Datetime, SP500_Close = Close)

stock_vs_sp500 <- long_data %>%
    filter(Type == "Stock") %>%
    select(Datetime, Symbol, Stock_Close = Close) %>%
    left_join(sp500_hourly, by = "Datetime")

scatter_plot <- ggplot(stock_vs_sp500, aes(x = SP500_Close, y = Stock_Close)) +
    geom_point(alpha = 0.4, size = 0.6) +
    facet_wrap(~Symbol, scales = "free_y") +
    labs(
        title = "Stock Prices vs S&P 500 (Hourly Close)",
        x = "S&P 500 Close",
        y = "Stock Close"
    ) +
    theme_minimal()

scatter_plot

# Quantify correlation of hourly percentage changes
sp500_changes <- hourly_trading %>%
    filter(Symbol == "^GSPC") %>%
    select(Datetime, sp500_pct = pct_change)

stock_changes <- hourly_trading %>%
    filter(Type == "Stock") %>%
    select(Datetime, Symbol, stock_pct = pct_change)

correlation_df <- stock_changes %>%
    left_join(sp500_changes, by = "Datetime") %>%
    group_by(Symbol) %>%
    summarise(
        correlation = cor(stock_pct, sp500_pct, use = "complete.obs"),
        .groups = "drop"
    )

cor_plot <- ggplot(correlation_df, aes(x = reorder(Symbol, correlation), y = correlation)) +
    geom_col(fill = "#55A868") +
    coord_flip() +
    labs(
        title = "Correlation of Hourly % Changes with S&P 500",
        x = "Stock",
        y = "Pearson Correlation"
    ) +
    theme_minimal()

cor_plot

```

# 4. Interpretation

Summary of key findings from the computed results (peak-movement hour and correlation strengths).

Based on the computed hourly movements and correlations:

- The hour with the highest average intrahour movement (US/Eastern) is: `r hourly_volatility %>% slice_max(avg_abs_move, n = 1, with_ties = FALSE) %>% pull(hour)` am (avg abs move: `r scales::percent(hourly_volatility %>% slice_max(avg_abs_move, n = 1, with_ties = FALSE) %>% pull(avg_abs_move), accuracy = 0.01)`).

- The strongest positive correlation with S&P 500 hourly returns is: `r correlation_df %>% slice_max(correlation, n = 1, with_ties = FALSE) %>% pull(Symbol)` (correlation: `r round(correlation_df %>% slice_max(correlation, n = 1, with_ties = FALSE) %>% pull(correlation), 3)`).

- The weakest correlation with S&P 500 hourly returns is: `r correlation_df %>% slice_min(correlation, n = 1, with_ties = FALSE) %>% pull(Symbol)` (correlation: `r round(correlation_df %>% slice_min(correlation, n = 1, with_ties = FALSE) %>% pull(correlation), 3)`).

Overall, the intrahour volatility pattern tends to spike around market open and taper off later in the session, which is consistent with information-dense opening auctions and early-session price discovery. The stock–S&P 500 correlations are positive but vary in strength, which reflects differences in sector exposure (e.g., tech-heavy names tend to track the index more closely than defense-oriented names).
