---
title: "Solution for Historical_Stock_Data_Analysis"
output: md_document
---

# Stock and Index data
Financial markets are highly complex systems that are influenced by many factors. Both stocks (pieces of ownership for individual companies) and ETFs/Indices (a sort of aggregation of multiple stocks/companies) are represented and traded on these markets. Nowadays, some combination of stocks and ETFs (which pretty much track indices) make up the long term investments of many people. In this project, you will take a closer look at the historical data of a few stocks and some indices in which they are included, analyzing their performance and the relationship between them. How closely do you think stock and index prices correlate?


# Get tidy data

```{r setup, echo=FALSE, message=FALSE,warning=FALSE}
library(readr)
library(dplyr)
library(stringr)
library(tidyverse)
library(tidyr)
library(forcats)
library(ggplot2)
library(scales)
library(knitr)
library(lubridate)
library(tidyquant)
```

## Data cleaning

- read data and combine the two header rows into meaningful column names.
```{r rename, echo=FALSE, message=FALSE, warning=FALSE}
setwd("~/Data-projects-with-R-and-GitHub/Projects/ArminBurkhardt")

col_names <- read_csv("stock_data.csv", n_max = 2, col_names = FALSE) %>%
  mutate(row = row_number()) %>%
  pivot_longer(
    cols = -row,
    names_to = "col_id",
    values_to = "value"
  ) %>%
  pivot_wider(
    names_from = row,
    values_from = value,
    names_prefix = "row_"
  ) %>%
  mutate(
    col_name = if_else(
      col_id == "X1",
      "Datetime",
      paste(row_2, row_1, sep = "_")
    )
  ) %>%
  pull(col_name)
```

- Skip the header rows and apply the new column names.
```{r actual_data, echo=FALSE, message=FALSE}
#read the dataset again, skip the first two rows
data_raw <- read_csv("stock_data.csv", skip = 2,col_names = col_names, show_col_types = FALSE)
```

- Convert columns to appropriate data types.
```{r fix_data_types, echo=FALSE, message=FALSE}
data_raw <- data_raw %>%
  mutate(
    # Convert datetime to proper datetime format
    Datetime = as.POSIXct(Datetime,format = "%Y-%m-%d %H:%M:%S"),
    # Convert all other columns to numeric
    across(-Datetime, as.numeric)
  ) %>% 
  filter(!is.na(Datetime))
```

- Check the cleaned dataset.
```{r fig.width= 2, fig.height= 3, out.width= "100%", check_data, echo =FALSE, message=FALSE}
data_raw %>%
  slice_head(n = 5)%>%
  kable()
```
## Data formatting
- Pivot all price related columns into a long format and separate the symbol from the price type.

- Transform the measurement variable (`Measure`) into individual columns.

- Add a `Type` Column indicating whether each symbol is a stock or an index.
```{r, echo=FALSE, message=FALSE}
# Reshape the dataset into long format
data_tidy <- data_raw %>%
  pivot_longer(
    cols = -Datetime,
    names_to = c("Symbol", "Measure"),
    names_sep = "_",
    values_to = "value"
  )%>%
#convert measurement types into separate columns
  pivot_wider(
    names_from = Measure,
    values_from = value
  ) %>%
# classify each symbol as stock or index
   mutate(
    Type = if_else(
      Symbol %in% c("^GSPC", "^IXIC"),
      "Index",
      "Stock"
    ),
    Type = factor(Type, levels = c("Stock", "Index"))
  ) %>%
  select(Datetime, Symbol, Type, Open, High, Low, Close, Volume)
```

- Check tidy dataset
```{r, fig.width= 4, fig.height= 6, out.width= "100%", echo =FALSE, message=FALSE}
data_tidy %>% slice_head(n = 7)%>%
  kable()
```

## Aggregation
- Extract the date from the datetime column

- Summarise the data for each day and symbol according to the required rules.
```{r, echo=FALSE, message=FALSE}
data_daily <- data_tidy %>%
  mutate(Date = as.Date(Datetime))%>%
  arrange(Datetime) %>%
  group_by(Date, Symbol, Type) %>%
  summarise(
    Open = first(Open),
    Close = last(Close),
    High = max(High, na.rm = TRUE),
    Low = min(Low, na.rm = TRUE),
    Volume = sum(Volume, na.rm = TRUE),
    .groups = "drop"
  )
```

- Check aggregated data
```{r, fig.width= 4, fig.height= 6, out.width= "100%",echo =FALSE, message=FALSE}
data_daily %>% slice_head(n = 5)%>%
  kable()
```

# Visualization

## Price
Plot the daily closing prices for all stocks and indices using `facet`.
```{r, fig.width=4, fig.height=10, out.width="100%", echo=FALSE, message=FALSE}
data_label <- data_daily %>%
  mutate(
    Symbol_label = if_else(
      Type == "Index",
      paste("index", str_remove(Symbol, "^\\^")),
      paste("stock", Symbol)
    )
  )
ggplot(data_label, 
       aes(x = Date, y = Close, color = Type)) +
  geom_line() +
  facet_wrap(~ Symbol_label, scales = "free_y", ncol = 1) +
  labs(
    title = "Daily Closing Prices of Stocks and Indices",
    x = "Date",
    y = "Close Price",
    color = "Type"
  ) +
  theme_minimal()+
  theme(strip.text = element_text(face = "bold"))
```

## Volume

- Visualize the daily trading volume for all stocks and indices using bar plots with `facet`.
```{r,fig.width=4, fig.height=10, out.width="100%",echo=FALSE, message=FALSE}
ggplot(data_label, 
       aes(x = Date, y = Volume, fill = Type)) +
  geom_col(alpha = 0.8) +
  facet_wrap(~ Symbol_label, scales = "free_y", ncol = 1) +
  scale_y_continuous(labels = label_number(
    scale_cut = cut_si(""),
    accuracy = 0.1
  ))+
  labs(
    title = "Daily Trading Volume of Stocks and Indices",
    x = "Date",
    y = "Volume",
    fill = "Type"
  ) +
  theme_minimal() +
  theme(
    strip.text = element_text( face = "bold")
  )
```

## Candlesticks

- Represent daily price data using candlesticks based on open, high, low, and close values.
```{r,fig.width=4, fig.height=10, out.width="100%", echo=FALSE, message=FALSE}
ggplot(data_label,
       aes(x = Date, y = Close))+
  geom_candlestick(
    aes(
      open = Open,
      high = High,
      low = Low,
      close = Close
    ),
    colour_up = "green",
    colour_down = "red",
    fill_up = "green",
    fill_down = "red"
  )+
  facet_wrap(~Symbol_label, scales = "free_y", ncol = 1)+
  labs(
    title = "candlestick chart of daily prices",
    x = "Date",
    y = "Price"
  )+
  theme_dark()+
  theme(strip.text = element_text(face = "bold"))

```

(I wanted to include a bar chart showing the daily volume in the same graph, but as the y-axes couldn’t be merged and required scaling, I found it too difficult to achieve, so I gave up.)

# Pattern analysis & correlation
## Percentage computing
- Calculate the percentage change in closing price compared to the previous hour for each stock and index.
```{r, echo=FALSE, message=FALSE}
hourly_change <- data_tidy %>%
  arrange(Symbol, Datetime) %>%
  group_by(Symbol) %>%
  mutate(
    price_change = (Close - lag(Close))/lag(Close)*100
  ) %>%
  ungroup()
```

- check the computed percentage changes
```{r, echo =FALSE, message=FALSE}
hourly_change %>%
  select(Datetime, Symbol, Close, price_change) %>% 
  mutate(
    price_change = paste0(round(price_change, 2), "%")
    )%>%
  slice_head(n = 7)%>%
  kable()
```
## Volatility movement (bar-plot)
-  Calculate the absolute percentage change between the hourly open and close prices.
```{r, echo = FALSE, message=FALSE}
hourly_volatility <- data_tidy %>%
  mutate(
    hour = format(Datetime, "%H"),
    movement_pct = abs((Close - Open) / Open * 100)
  )
```

- Aggregate the average absolute price movement by hour across all stocks and indices.
```{r, echo = FALSE,message=FALSE}
movement_summary <- hourly_volatility %>%
  group_by(hour) %>%
  summarise(
    avg_movement = mean(movement_pct, na.rm = TRUE),
    .groups = "drop"
  )
```

- Visualize the average price movement per hour of the day using a bar plot.
```{r, fig.width=7, fig.height=3.5, out.width="100%", echo=FALSE, message=FALSE}
volatility_by_hour_ny <- data_tidy %>%
  mutate(
    Datetime_ny = with_tz(as.POSIXct(Datetime, tz = "UTC"), "America/New_York"),
    Hour_ny = hour(Datetime_ny)
  ) %>%
  arrange(Symbol, Datetime_ny) %>%
  group_by(Symbol) %>%
  mutate(abs_pct_change = abs(Close / lag(Close) - 1)) %>%
  ungroup() %>%
  group_by(Hour_ny) %>%
  summarise(
    avg_abs_change = mean(abs_pct_change, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(volatility_by_hour_ny, aes(x = Hour_ny, y = avg_abs_change)) +
  geom_col(fill = "steelblue", width = 0.6) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    title = "Average Absolute Price Movement by Hour of Day (New York Time)",
    x = "Hour of Day (New York Time)",
    y = "Average Absolute Percentage Change"
  ) +
  theme_minimal()
```
- Answer: Prices change the most at 7.00 (NewYork Time) on average.

## Correlation

### line-plot

- As the data on the y-axis varied too widely (If draw a plot with the y-axis showing price, the difference between the index and the stocks is too big and hard to compare), making it difficult to visualise clearly, so there are two way to show the correlation between index and the stocks

  - use a percentage change trend chart for comparison (two lines)
```{r,fig.width= 4, fig.height= 10, out.width= "100%", echo = FALSE,message=FALSE}
# Extract S&P500
sp500 <- data_daily %>%
  filter(Symbol == "^GSPC") %>%
  select(Date, sp500 = Close)
# Keep only stocks
stocks <- data_daily %>%
  filter(Symbol %in% c("AAPL", "NVDA", "AMZN", "LMT", "V"))
# Merge
data_compare <- stocks %>%
  left_join(sp500, by = "Date")
# long format
data_compare_long <- data_compare %>%
  pivot_longer(
    cols = c(Close, sp500),
    names_to = "Series",
    values_to = "Price"
  )
# percentage growth
data_compare_pct <- data_compare_long %>%
  group_by(Symbol, Series) %>%
  arrange(Date, .by_group = TRUE) %>%
  mutate(
    pct_growth = (Price / first(Price) - 1)
  ) %>%
  ungroup()
ggplot(data_compare_pct, 
       aes(x = Date, y = pct_growth, color = Series)) +
  geom_line() +
  facet_wrap(~ Symbol, scales = "free_y", ncol = 1) +
  scale_y_continuous(labels = percent_format(scale = 1)) +
  labs(
    title = "Percentage Growth of Stocks vs S&P 500",
    x = "Date",
    y = "Percentage Growth",
    color = "Series"
  ) +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold"))
```
  - or plot a scatter plot with the index value (growth percentage) in x Axis and stock value (growth percentage) in y axis:
```{r, fig.width= 4, fig.height= 10, out.width= "100%", echo = FALSE,message=FALSE}
data_compare_scatter <- data_compare_pct %>%
  select(Date, Symbol, Series, pct_growth)%>%
  pivot_wider(
    names_from = Series,
    values_from = pct_growth) %>%
  rename(stock_pct_growth = Close, index_pct_growth = sp500
  )

ggplot(data_compare_scatter, 
       aes(x = stock_pct_growth, y = index_pct_growth)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "red")+
  facet_wrap(~ Symbol, scales = "free", ncol = 1) +
  scale_x_continuous(labels = percent_format()) +
  scale_y_continuous(labels = percent_format())+
  labs(
    title = "Percentage Growth of Stocks vs S&P 500",
    x = "S&P 500 Percentage Growth",
    y = "Stock Percentage Growth"
  ) +
  theme_minimal() +
  theme(strip.text = element_text(face = "bold"))
```
- Interpretation：This plot illustrates the relationship between the percentage growth of individual stocks and the index (S&P 500). By adding a fitted linear trend line using `geom_smooth`, the correlation between the two can be observed more intuitively: the more closely the data points cluster around the fitted line, the stronger the linear relationship between the stock and the index; conversely, greater dispersion indicates a weaker correlation. Based on the plot, AMZN and NVDA appear to have data points that lie closer to the fitted line, suggesting a stronger alignment with the overall market trend.
  
### correlation coefficient
- calculate:

```{r, fig.width=6, fig.height=2, echo=FALSE, message=FALSE}
# Extract the hourly percentage changes of the S&P 500 as the reference series.
sp500_change <- hourly_change %>%
  filter(Symbol == "^GSPC") %>%
  select(Datetime, sp500_change = price_change) 
# Keep the hourly percentage changes for the five stocks.
  stock_change <- hourly_change %>%
  filter(Symbol %in% c("AAPL", "NVDA", "AMZN", "LMT", "V")) %>%
  select(Datetime, Symbol, stock_change = price_change)
# Match each stock’s hourly percentage changes with the corresponding S&P 500 changes 
  cor_data <- stock_change %>%
  left_join(sp500_change, by = "Datetime") 
# calculate
  cor_results <- cor_data %>%
  group_by(Symbol) %>%
  summarise(
    correlation = cor(stock_change, sp500_change, use = "complete.obs"),
    .groups = "drop"
  ) %>%
  arrange(desc(correlation))
kable(cor_results)
```
  
- plot
```{r,fig.width= 5, fig.height= 3, out.width= "100%", echo =FALSE, message=FALSE}
ggplot(cor_results, 
       aes(x = reorder(Symbol, correlation), y = correlation)) +
  geom_col(fill = "steelblue", width = 0.5) +
  labs(
    title = "Correlation of Stock Price Changes with S&P 500",
    x = "Stock",
    y = "Correlation Coefficient"
  ) +
  theme_minimal()
```

- Now the question could be answered, stock prices are generally positively correlated with the S&P 500, although the strength of the correlation varies across companies. Both the growth percentage scatter plot and the correlation coefficient illustrate this point, AMZN and NVDA shows the higher correlation with the S&P 500 and is therefore the more similar in terms of price movements.
