---
output: md_document
---

# Tim's take on Yunji's project

Installs pacman ("packagemanager") if needed

```{r}
if (!require("pacman")) install.packages("pacman")
```

Use pacman to load add-on packages as desired

```{r}
pacman::p_load(pacman,
               readr,
               maps,
               ggplot2,
               dplyr,
               viridis,
               eurostat,
               ggrepel,
               tidyverse)
```

## Data

#### First, load the datasets

```{r}


Poverty_rate <- read_csv("At-risk-of-poverty rate.gz", show_col_types = FALSE, 
                         col_select = c("freq", "unit", "geo", "TIME_PERIOD", "OBS_VALUE"))

Household_Income <- read_csv("Income of households.gz", show_col_types = FALSE, 
                             col_select = c("freq", "unit", "direct", "na_item", "geo", "TIME_PERIOD", "OBS_VALUE"))

Crime_data_1 <- read_csv("Crime_data_for_Task1.gz", show_col_types = FALSE, 
                         col_select = c("freq", "unit", "iccs", "geo", "TIME_PERIOD", "OBS_VALUE"))

Crime_data_2 <- read_csv("Crime_data_for_Task2.gz", show_col_types = FALSE, 
                         col_select = c("freq", "unit", "iccs", "geo", "TIME_PERIOD", "OBS_VALUE"))
```

## Task 1

```{r}

#Only keep the data for 2019

Crime_data_1.2 <- filter(Crime_data_1, TIME_PERIOD == "2019" )

```

#### Plotting

still under construction 🤔

```{r eval=FALSE, include=FALSE}

# load spacial data of whole europe

df60 <- get_eurostat_geospatial(output_class = "df", resolution = 60)

df60 %>%
  # reduce to germany
  filter(str_starts(NUTS_ID,"DE")) |> 
  # plot shapes for each NUTS level (LEVL_CODE)
  ggplot(aes(x=long,y=lat,group=group)) +
  geom_polygon(fill=NA,color="black") + 
  facet_wrap(~LEVL_CODE)
```

```{r eval=FALSE, include=FALSE}

Ger <- map_data("world") |>  
  filter(region =="Germany")

#ggplot() +
#  geom_polygon(data = Ger, aes(x=long, y = lat, #group = group), fill="grey", alpha=0.3) +
#  geom_point( data= Crime_data_1.2, aes(x=long, #y=lat, size= OBS_VALUE, color= OBS_VALUE)) +
#  scale_size_continuous(range=c(1,12)) +
#  scale_color_viridis(trans="log") +
#  theme_void() + 
#  ylim(50,59) + 
#  coord_map() 
```

## Task2

```{r include=FALSE}

#lets only look at data from 2008

Crime_data_2.1 <- filter(Crime_data_2, TIME_PERIOD == "2008")

Household_Income2 <-filter(Household_Income, TIME_PERIOD == "2008")

Poverty_rate2 <- filter(Poverty_rate, TIME_PERIOD == "2008")
```

```{r include=FALSE}

#merge the datasets together by countries

Household_Income2 <-
  Household_Income2 |> 
  rename(Household_Income_EUR_HAB = OBS_VALUE) |> 
  select( - c(freq, unit, direct, na_item, TIME_PERIOD))

Poverty_rate2 <-
  Poverty_rate2 |>
  rename(Poverty_rate_PC = OBS_VALUE) |> 
  select(- c(freq, unit, TIME_PERIOD))

Crime <- Crime_data_2.1 |> 
 left_join(Household_Income2, by="geo") |> 
 left_join(Poverty_rate2, by="geo") |>
  na.omit()

```

Splitting the unit variable by NR and P_HTHAB. Creating a new data set with the observed values from NR and P_HTHAB in the same row, so we can use for y the P_HTHAB observed values and for the size the NR observed value. (before I created this new data set, I tried using a ifelse function for "y" and the "size" in the aes function, but it wasn't working. So, there is prob a more elegant way to do it :)

```{r include=FALSE}

Crime_NR <- subset(Crime, unit == "NR", select = -unit) |> 
  rename(OBS_VALUE_NR = OBS_VALUE)
Crime_P_HTHAB <- subset(Crime, unit == "P_HTHAB", select = -unit) |> 
  rename(OBS_VALUE_P_HTHAB = OBS_VALUE)

Crime_merged <- merge(Crime_NR, Crime_P_HTHAB, by = c("geo", "freq", "iccs", "TIME_PERIOD", 
                                                      "Household_Income_EUR_HAB", "Poverty_rate_PC"))
```

#### The Plot

I only did it for 2008

```{r}

ggplot(Crime_merged, 
       aes(x = Household_Income_EUR_HAB, 
           y = OBS_VALUE_P_HTHAB, 
           size = OBS_VALUE_NR, 
           color = Poverty_rate_PC)) +
  geom_point(alpha = 0.7) +
  geom_text_repel(aes(label = geo), vjust = 1.5, size = 2, color = "black") +
  scale_size_continuous(range = c(3, 10)) +
  scale_color_viridis_c() +
  facet_wrap(~iccs, scales = "free", ncol = 4, 
             labeller = labeller(iccs = c(ICCS0101 = "Intentional Homicide", ICCS02011 = "Assault",
                                          ICCS0401 = "Robbery", ICCS0501 = "Burglary", 
                                          ICCS05012 = "Burglary of private residential premises",
                                          ICCS0502 = "Theft", 
                                          ICCS050211 = "Theft of a motorized land vehicle"))) +
  labs(x = "Income of Households (Euro per inhabitant)",
       y = "Police-recorded Offence (Per hundred thousand inhabitants)",
       size = "Police-recorded Offence (Number)",
       color = "At Risk of Poverty Rate") +
  theme_minimal() +
  theme(legend.position = "bottom", 
        strip.text = element_text(size = 6.3, margin = margin(0, 0, 5, 0)),
        #axis.text.x = element_text(angle = 45, hjust = 1))
        text = element_text(size = 7))
       
 
```
