---
title: "project-description"
output: md_document
date: "2024-11-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

introduce shortly the general topic of your task (to put it in context)
introduce the data set(s), it’s origin and important information of what-is-what, etc.
define data manipulation goals if needed (also fine if this is the main focus of the project!)
define the visualization goal(s) and possible intermediates


## Volleyball Passing- USA and TU


## Topic of task

The topic of the task is to analyze the volleyball passing data of USA and TU and see how the number of passes lead to a point win, or point loss . The data set is collected from the volleyball matches

## Data Set
https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/27560
Coaching Action Research Excellence explores the use of quantitative research data to improve the focus of coaching practice from a tactical and technical perspective. Software development designed to assist meaningful Sport Data Collection that is in line with principles of Game Sense and Athlete Centered Coaching will be featured in this Book that is scheduled for Publication in 2015. (2014-10-09)

```{r cars}
```

## Data Manipulation Goals
There are data of games for the USA Sitting Volleyball team and the University of Tulsa Volleyball team. The dataset contains the average number of passes made by the teams if they won the point or lost the point verses other volleyball teams of another country (China, Russia, Brazil, etc).

The data manipulation goals are to find the correlation between the number of passes and the number of points won or lost.


## Visualization Goals
I want a correlation graph between passing score and point win or loss.
