---
title: "timow98"
author: "Timo Walcher"
date: "`r Sys.Date()`"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(repos = c(CRAN = "https://cloud.r-project.org"))
if (!requireNamespace("pacman", quietly = TRUE)) {
  install.packages("pacman")} 
library(pacman)
p_load(tidyverse, cowplot)
```

## Data Manipulation Goals
1. Converting the ‘Date’ column to a proper date format to facilitate time-based analysis. 
  
Read the CSV file and display the head of the data. We can format the date column when importing the file.
```{r}
NVDA <- read_csv("NVDA_messedUp_final.csv", 
    col_types = cols(Date = col_date(format = "%Y-%m-%d"))) # Date format
head(NVDA)
```
2. Calculating weekly, monthly and yearly closing values from the dataset to identify trends and patterns in the stock performance over time.

Here we can use dplyr to get a new dataframe with the last Close value of every week, month and year.
```{r}
NVDA_monthly <- NVDA %>%
  mutate(Date_YM = format(Date, "%Y-%m")) %>%
  group_by(Date_YM) %>%
  summarise(monthly_close = last(Close))

NVDA_weekly <- NVDA %>%
  mutate(Date_YM = format(Date, "%Y-%W"),
         calendarweek = format(Date, "%W")) %>%
  group_by(Date_YM) %>%
  summarise(weekly_close = last(Close),
            calendarweek = first(calendarweek))

NVDA_yearly <- NVDA %>%
  mutate(year = format(Date, "%Y")) %>%
  group_by(year) %>%
  summarise(yearly_close = last(Close))

```
Now we have 3 new Dataframes.  
For the second Plot we need to extend the weekly dataframe.
```{r}
# Transform Date_YM column into datatype date 
NVDA_weekly <- NVDA_weekly %>%
  mutate(Date_YMD = as.Date(paste0(Date_YM,"-01"), format = "%Y-%W-%d"),
         Date_YM = NULL)
# which somehow gives back wrong week numbers

# Also add a column in NVDA_weekly with year so that we can join the yearly dataframe
NVDA_weekly <- NVDA_weekly %>%
  mutate(year = lubridate::year(Date_YMD))

# Merge both dataframes based on year column
#NVDA_merged <- merge(NVDA_weekly, NVDA_yearly, by = "year", all.x = TRUE, suffixes = c("_weekly", "_yearly"))

# Left join NVDA_weekly with NVDA_yearly
NVDA_yearly$year <- as.double(NVDA_yearly$year) 
NVDA_merged <- left_join(NVDA_weekly, NVDA_yearly, by = "year")

# make new column "percentage" showing the deviation of a week to the yearly average
NVDA_merged <- NVDA_merged %>%
  mutate(deviation = weekly_close - yearly_close,
         percentage = (deviation / yearly_close) * 100)

```

## Visualisation Plot 1
The closing values are now used to produce two charts
*A plot with a logarithmic scale to highlight percentage changes over time.

First of all, the new dataframe with the column month is of type character. I need to change it back to the date format.
```{r}
NVDA_monthly$Date_YM <- as.Date(paste0(NVDA_monthly$Date_YM, "-01"))
```

When using logarithmic scales, NA values would create an error message, so first of all we need to do something about that. It would be possible to add just a tiny constant to all the NA values or just delete them.
```{r}
NVDA_monthly <- NVDA_monthly %>% drop_na()
```
Adding a small constant would look like this:
NVDA_monthly$mean_close_p <- NVDA_monthly$monthly_close + 1e-6
But it would not be useful because there is only one NA value which is basically the recent month.

Now we can plot it.
```{r}
# This is for the labels on the x-axis
break_years <- as.Date(c("2000-01-01", "2010-01-01", "2020-01-01"))

p1 <- ggplot(NVDA_monthly, aes(x = Date_YM, y = monthly_close)) +
  geom_line(group = 1, color = "blue", linewidth = 1) +
  scale_y_log10(breaks = c(1, 10, 100, 1000), 
                label = \(tickposition) scales::comma(tickposition))+
  scale_x_date(breaks = break_years, labels = format(break_years, "%Y"), date_labels = "%Y") +
  labs(
    x = "Date",
    y = "Monthly values",
    title = "NVIDIA STOCK CHART - Logarithmic scale") +
  theme(panel.grid = element_line(colour = "lightgrey"),
        panel.background = element_rect(fill = "white"))
  
```

*A plot with a linear scale to provide a clear representation of absolute stock values.
```{r}
# This is for the labels on the x-axis
break_years <- as.Date(c("2000-01-01", "2010-01-01", "2020-01-01"))

p2 <- ggplot(NVDA_monthly, aes(x = Date_YM, y = monthly_close)) +
  geom_line(group = 1, color = "black", linewidth = 1) +
  scale_y_continuous(breaks = c(0,200,400,600,800), limits = c(0,800))+
  scale_x_date(breaks = break_years, labels = format(break_years, "%Y"), date_labels = "%Y") +
  labs(
    x = "Date",
    y = "Monthly values",
    title = "NVIDIA STOCK CHART - Linear scale") +
  theme(panel.grid = element_line(colour = "lightgrey"),
        panel.background = element_rect(fill = "white"))
```
Now both plots next to each other
```{r plot_grid_output, fig.width=10, fig.height=5}
plot_grid(p1, p2, ncol = 2)
```

## Visualisation Plot 2
```{r}

```

