---
title: "Solution for Karim's Project by Armin"
output: md_document
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(this.path)
library(khroma)
library(scales)
library(knitr)

setwd(this.path::this.dir())

BLData <- read_csv("https://raw.githubusercontent.com/barrolee/BarroLeeDataSet/refs/heads/master/BLData/BL2013_MF_v2.2.csv") 
```


## Data cleaning & manipulation

The validation check shows that the percentage sums far exceed 100%. More precisely, the range of the percentage sums are as follows:

```{r data_val, echo=FALSE}
check_sums <- function(df) {
  df %>%
    mutate(perc_sum = lu + lp + lpc + ls + lsc + lh + lhc) %>%
    summarise(`Minimum sum` = min(perc_sum), `Maximum sum` = max(perc_sum)) %>%
    kable()
}

BLData %>%
  check_sums()
```

This is especially interesting, because education levels are mutually exclusive, so the percentage sums should always add up to 100%.
However if one was to consider that people with `completed` education are also counted in the more generalized education level (`some`), the following ranges are calculated:

```{r data_val1, echo=FALSE}

DFData_adjusted <- BLData %>%
  mutate(
    lp = lp - lpc,
    ls = ls - lsc,
    lh = lh - lhc
  )

check_sums(DFData_adjusted)
```

This makes much more sense than the previous results, thus the adjusted data will be used from now on.

Testing for implausible values and missing values:

```{r, echo=FALSE, message=FALSE}
# temporary variable to make the visible code prettier
data <- DFData_adjusted %>%
  select(lu:lhc)
```

```{r data_val2}
min(data)
max(data)
sum(is.na(data))
```

There are no values below 0 or above 100, and there are no missing values. 

Are all expected country × year combinations present?

```{r data_val3, echo=FALSE}
expected_combinations <- expand.grid(
  country = unique(BLData$country),
  year = unique(BLData$year)
)

missing_combinations <- anti_join(expected_combinations, BLData, by = c("country", "year"))
nrow(missing_combinations) == 0
```
All expected combinations are present, so there are no missing country × year combinations.

**Renaming cryptic names**

```{r, message=FALSE}
DFData_adjusted <- DFData_adjusted %>%
  rename(
    `No education` = lu,
    `Some primary` = lp,
    `Completed primary` = lpc,
    `Some secondary` = ls,
    `Completed secondary` = lsc,
    `Some tertiary` = lh,
    `Completed tertiary` = lhc,
    `Years of schooling` = yr_sch,
    Region = region_code
  )
```

