---
title: "Global Education Attainment Analysis"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 10,
  fig.height = 6
)
```

```{r load-packages}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(scales)
library(forcats)
```

## Data

The analysis uses the **Barro-Lee Educational Attainment Dataset (v2.2, 2013)**.  
Data source: <http://barrolee.com> or directly via  
<https://raw.githubusercontent.com/barrolee/BarroLeeDataSet/refs/heads/master/BLData/BL2013_MF_v2.2.csv>

```{r load-data}
data_path <- "BL2013_MF_v2.2.csv"
data_url  <- "https://raw.githubusercontent.com/barrolee/BarroLeeDataSet/master/BLData/BL2013_MF_v2.2.csv"

bl_raw <- tryCatch(
  read_csv(data_url, show_col_types = FALSE),
  error = function(e) read_csv(data_path, show_col_types = FALSE)
)
```

### First rows of the dataset

```{r preview-table}
kable(
  head(bl_raw, 5),
  caption = "First 5 rows of the Barro-Lee dataset",
  align   = "c",
  digits  = 2
)
```

**Key columns**

| Column | Description |
|--------|-------------|
| `lu`   | % with no education |
| `lp`   | % with some primary education |
| `lpc`  | % with completed primary education |
| `ls`   | % with some secondary education |
| `lsc`  | % with completed secondary education |
| `lh`   | % with some higher/tertiary education |
| `lhc`  | % with completed higher/tertiary education |
| `yr_sch` | Average years of schooling |
| `WBcode` | World Bank country code |
| `region_code` | World region |

The education-level variables are mutually exclusive: each person is counted only under
their **highest** attained education level.

---

## Data Cleaning & Validation

```{r data-cleaning}
# This dataset is already filtered to combined sex (MF only)
# Retain only standard 5-year age groups (exclude aggregate rows: ageto==999)
bl <- bl_raw |>
  filter(ageto != 999)
```

### Check 1 — Education level groups sum to 100

The dataset uses three mutually exclusive **level groups**: primary (`lp`), secondary
(`ls`), and higher (`lh`). Within each group, the `*c` variant (`lpc`, `lsc`, `lhc`)
counts those who **completed** that level. Accordingly, `lu + lp + ls + lh ≈ 100` for
each row, while `lu + lp + lpc + ls + lsc + lh + lhc` will naturally exceed 100 (as
`lpc`, `lsc`, `lhc` are nested subsets of their parent level).

```{r check-sum}
share_check <- bl |>
  mutate(total_share = lu + lp + ls + lh) |>
  summarise(
    n_rows            = n(),
    mean_total        = round(mean(total_share, na.rm = TRUE), 2),
    min_total         = round(min(total_share,  na.rm = TRUE), 2),
    max_total         = round(max(total_share,  na.rm = TRUE), 2),
    n_off_by_gt0.5pct = sum(abs(total_share - 100) > 0.5, na.rm = TRUE)
  )

kable(share_check, caption = "Share sum validation: lu + lp + ls + lh should be ~100")
```

### Check 2 — Missing and negative values

```{r check-missing}
edu_cols <- c("lu", "lp", "lpc", "ls", "lsc", "lh", "lhc", "yr_sch", "pop")

missing_check <- bl |>
  summarise(across(all_of(edu_cols), ~sum(is.na(.)))) |>
  pivot_longer(everything(), names_to = "column", values_to = "n_missing")

negative_check <- bl |>
  summarise(across(all_of(edu_cols), ~sum(. < 0, na.rm = TRUE))) |>
  pivot_longer(everything(), names_to = "column", values_to = "n_negative")

checks <- left_join(missing_check, negative_check, by = "column")
kable(checks, caption = "Missing and negative value counts per column")
```

### Check 3 — Country × year coverage

```{r check-coverage}
expected_years <- seq(1950, 2010, by = 5)

incomplete <- bl |>
  group_by(country) |>
  summarise(n_years = n_distinct(year), .groups = "drop") |>
  filter(n_years < length(expected_years))

if (nrow(incomplete) == 0) {
  cat("All countries have data for all", length(expected_years), "expected years.\n")
} else {
  cat(nrow(incomplete), "countries have incomplete year coverage:\n")
  kable(incomplete, caption = "Countries with fewer than 13 year observations")
}
```

---

## Visualization 1 — Regional Trends in Average Years of Schooling

*Addresses Q1 (Has the education gap between regions narrowed?) and Q2 (Are there
countries where schooling stagnated?)*

Population-weighted average years of schooling per region per year:

```{r viz1}
regional_trends <- bl |>
  filter(!is.na(yr_sch), !is.na(pop), pop > 0, !is.na(region_code)) |>
  group_by(region_code, year) |>
  summarise(
    yr_sch_wtd = sum(yr_sch * pop) / sum(pop),
    .groups = "drop"
  )

ggplot(regional_trends, aes(x = year, y = yr_sch_wtd, color = region_code)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = seq(1950, 2010, 10)) +
  scale_color_brewer(palette = "Dark2") +
  labs(
    title    = "Regional Trends in Average Years of Schooling (1950–2010)",
    subtitle = "Population-weighted regional averages across 5-year age groups",
    x        = "Year",
    y        = "Average Years of Schooling",
    color    = "Region"
  ) +
  theme_bw() +
  theme(
    legend.position  = "bottom",
    legend.title     = element_text(face = "bold"),
    plot.title       = element_text(face = "bold")
  ) +
  guides(color = guide_legend(nrow = 3))
```

---

## Visualization 2 — Education Distribution by Age Group (Population Pyramid Style)

*Addresses Q4 (Which age group benefits most?) and Q1*

Germany (left) vs. India (right) for the years 1960 and 2010, with stacked bars
showing the share in each education level per 5-year age group.

```{r viz2, fig.width=12, fig.height=7}
# Correct mutually exclusive education shares:
# lu, lp-lpc, lpc, ls-lsc, lsc, lh-lhc, lhc  (sum = lu + lp + ls + lh ≈ 100)

pyr_data <- bl |>
  filter(
    country %in% c("Germany", "India"),
    year    %in% c(1960, 2010),
    agefrom >= 15, ageto <= 64
  ) |>
  mutate(
    `No Education`       = lu,
    `Some Primary`       = lp  - lpc,
    `Complete Primary`   = lpc,
    `Some Secondary`     = ls  - lsc,
    `Complete Secondary` = lsc,
    `Some Tertiary`      = lh  - lhc,
    `Complete Tertiary`  = lhc
  ) |>
  select(country, year, agefrom, ageto,
         `No Education`, `Some Primary`, `Complete Primary`,
         `Some Secondary`, `Complete Secondary`,
         `Some Tertiary`, `Complete Tertiary`) |>
  pivot_longer(
    cols      = `No Education`:`Complete Tertiary`,
    names_to  = "edu_level",
    values_to = "share"
  ) |>
  mutate(
    age_group = factor(
      paste0(agefrom, "–", ageto),
      levels = paste0(seq(15, 60, 5), "–", seq(19, 64, 5))
    ),
    edu_level = factor(edu_level,
                       levels = c("No Education", "Some Primary", "Complete Primary",
                                  "Some Secondary", "Complete Secondary",
                                  "Some Tertiary", "Complete Tertiary")),
    # Germany bars extend left (negative), India bars extend right (positive)
    share_directed = ifelse(country == "Germany", -share, share)
  )

ggplot(pyr_data,
       aes(x = share_directed, y = age_group, fill = edu_level)) +
  geom_bar(stat = "identity") +
  facet_wrap(~year, labeller = label_both) +
  geom_vline(xintercept = 0, color = "black", linewidth = 0.6) +
  scale_x_continuous(
    labels = function(x) paste0(abs(x), "%"),
    limits = c(-100, 100),
    breaks = seq(-100, 100, 25)
  ) +
  scale_fill_brewer(palette = "RdYlGn") +
  labs(
    title    = "Education Level Distribution by Age Group",
    subtitle = "← Germany (left)  |  India (right) →",
    x        = "Share of Age Group (%)",
    y        = "Age Group",
    fill     = "Education Level"
  ) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    plot.title      = element_text(face = "bold"),
    plot.subtitle   = element_text(face = "bold", size = 11)
  ) +
  guides(fill = guide_legend(nrow = 2))
```

---

## Visualization 3 — No-Education Share vs. Average Years of Schooling

*Addresses Q3 (Do countries with more uneducated population have lower average
schooling?) — shown here for the year 2010*

Each point is one country; colour encodes world region. Values are
population-weighted averages across all 5-year age groups.

```{r viz3}
scatter_year <- 2010

viz3_data <- bl |>
  filter(year == scatter_year, !is.na(lu), !is.na(yr_sch),
         !is.na(pop), pop > 0, !is.na(region_code)) |>
  group_by(country, region_code) |>
  summarise(
    lu_wtd     = sum(lu     * pop) / sum(pop),
    yr_sch_wtd = sum(yr_sch * pop) / sum(pop),
    .groups    = "drop"
  )

ggplot(viz3_data,
       aes(x = yr_sch_wtd, y = lu_wtd, color = region_code)) +
  geom_point(size = 2.5, alpha = 0.8) +
  scale_color_brewer(palette = "Dark2") +
  labs(
    title    = paste0("No-Education Share vs. Average Years of Schooling (",
                      scatter_year, ")"),
    subtitle = "Each point = one country; population-weighted values across age groups",
    x        = "Average Years of Schooling",
    y        = "Share with No Education (%)",
    color    = "Region"
  ) +
  theme_bw() +
  theme(
    legend.position = "right",
    plot.title      = element_text(face = "bold")
  ) +
  guides(color = guide_legend(ncol = 1))
```

---

## Visualization 4 — Choropleth Map: Gain in Average Years of Schooling (1950–2010)

*Addresses Q1 and Q2*

Countries coloured by the absolute increase in average years of schooling between
1950 and 2010.  **Red** = negative or zero gain; **Green** = highest gains;
**Light grey** = no data.

```{r viz4, fig.width=12, fig.height=7}
# Population-weighted country average for 1950 and 2010
gain_data <- bl |>
  filter(year %in% c(1950, 2010), !is.na(yr_sch), !is.na(pop), pop > 0) |>
  group_by(country, WBcode, year) |>
  summarise(yr_sch_avg = sum(yr_sch * pop) / sum(pop), .groups = "drop") |>
  pivot_wider(names_from = year, values_from = yr_sch_avg,
              names_prefix = "yr_sch_") |>
  mutate(gain = yr_sch_2010 - yr_sch_1950) |>
  filter(!is.na(gain))

# Map BL country names → ggplot2 map_data region names
name_fixes <- c(
  "Brunei Darussalam"                                    = "Brunei",
  "Congo"                                                = "Republic of Congo",
  "Cote dIvoire"                                         = "Ivory Coast",
  "Dominican Rep."                                       = "Dominican Republic",
  "Iran (Islamic Republic of)"                           = "Iran",
  "Lao People's Democratic Republic"                     = "Laos",
  "Libyan Arab Jamahiriya"                               = "Libya",
  "Republic of Korea"                                    = "South Korea",
  "Republic of Moldova"                                  = "Moldova",
  "Russian Federation"                                   = "Russia",
  "Syrian Arab Republic"                                 = "Syria",
  "Trinidad and Tobago"                                  = "Trinidad",
  "United Kingdom"                                       = "UK",
  "United Republic of Tanzania"                          = "Tanzania",
  "Viet Nam"                                             = "Vietnam",
  "China, Hong Kong Special Administrative Region"       = "Hong Kong",
  "China, Macao Special Administrative Region"           = "Macao"
)

gain_map <- gain_data |>
  mutate(map_region = dplyr::recode(country, !!!name_fixes))

# Build world map and join
world_map  <- map_data("world")
world_edu  <- world_map |>
  left_join(gain_map, by = c("region" = "map_region"))

# Compute diverging colour scale centred at 0
min_gain  <- min(gain_data$gain, na.rm = TRUE)
max_gain  <- max(gain_data$gain, na.rm = TRUE)
zero_frac <- (0 - min_gain) / (max_gain - min_gain)   # position of 0 in [0, 1]

ggplot(world_edu, aes(x = long, y = lat, group = group, fill = gain)) +
  geom_polygon(color = "white", linewidth = 0.08) +
  scale_fill_gradientn(
    colors = c("darkred", "red", "lightyellow", "lightgreen", "darkgreen"),
    values = c(0, zero_frac * 0.98, zero_frac, zero_frac + 0.02, 1),
    limits = c(min_gain, max_gain),
    na.value = "lightgray",
    name = "Gain\n(years)"
  ) +
  coord_fixed(ratio = 1.3) +
  labs(
    title    = "Gain in Average Years of Schooling per Country (1950–2010)",
    subtitle = "Red = negative / zero gain  |  Green = highest gains  |  Light grey = no data"
  ) +
  theme_void() +
  theme(
    legend.position  = "bottom",
    legend.direction = "horizontal",
    legend.key.width = unit(2, "cm"),
    plot.title       = element_text(face = "bold", size = 13),
    plot.subtitle    = element_text(size = 10)
  )
```
