---
title: "Solution for Karim"
output: md_document
---

Setup
```{r, echo = TRUE, message = FALSE}
library(readr)
library(knitr)
library(khroma)
library(tidyverse)
data <- read_csv("https://raw.githubusercontent.com/barrolee/BarroLeeDataSet/master/BLData/BL2013_MF_v2.2.csv")
```

## Validation Checks


#### Sum of education level percentages

```{r, echo = TRUE, message = FALSE}
# Sum up to 100?
edu_lvl <- subset(data, select = c("lu", "lp", "lpc", "ls", "lsc", "lh", "lhc"))
min<-which.min(rowSums(edu_lvl))
max<-which.max(rowSums(edu_lvl)) # Max Value almost at 200, something must be wrong here with the data
# What to do? Should from a threshold on entries be removed?
sum(edu_lvl[min,])
sum(edu_lvl[max,])
```

The sum of education levels should be at ~100, the max. value however, is at almost 200. This is because the completed education level (e.g., lpc) is a subset of that education level in general (e.g., lp). Therefore, for the general education levels the completed subset needs to be subtracted:

```{r, echo = TRUE, message = FALSE}  
data <- data %>%
  mutate(
    lp = lp - lpc,
    ls = ls - lsc,
    lh = lh - lhc
  )
```

#### Missing values?

```{r, echo = TRUE, message = FALSE}  
sum(is.na(data)) # No
```
There are NO missing values

#### Implausible values?

```{r, echo = TRUE, message = FALSE}  
# Are there any implausible values (negative percentage)
min(edu_lvl) # no values below 0
max(edu_lvl) # no values above 100
```
The max value at any single education level is 100, the min value 0

#### Check if for every country there are all year entries

```{r, echo = TRUE, message = FALSE}  
# 1. Create a grid of all possible combinations
all_combinations <- data %>%
  expand(country, year)
# 2. Find which combinations are NOT in your original data
missing_entries <- all_combinations %>%
  anti_join(data, by = c("country", "year"))
print(missing_entries)
```
There are NO missing combinations of country and year

BUT: for age groups from 15, 25 & 75 entries are always double (the second time with ageto == 999). This needs to be removed!
```{r, echo = TRUE, message = FALSE}
data <- data %>%
  filter(!(agefrom %in% c(15, 25, 75) & ageto == 999))
```


## Renaming columns for improved readability
```{r, echo = TRUE, message = FALSE}
data <- data %>%
  rename(No_education = "lu",
         Primary_education  = "lp",
         Primary_education_completed = "lpc",
         Secondary_education = "ls",
         Secondary_education_completed = "lsc",
         Higher_education = "lh",
         Higher_education_completed = "lhc")

kable(head(subset(data, select = c("No_education", "Primary_education", "Primary_education_completed", "Secondary_education", "Secondary_education_completed", "Higher_education", "Higher_education_completed")), 1))
```

# Data Visualizations

## Visualization 1

```{r, echo = TRUE, message = FALSE}
## Compute population-weighted regional average years of schooling (`yr_sch`) for each region and year and plot their evolution over time (1950-2010) using a line plot with one line per region.

# Group by Region and Year and calculate the population-weighted average years of schooling
regional_avg <- data %>%
  group_by(region_code, year) %>%
  summarise(
    total_population = sum(pop, na.rm = TRUE),
    weighted_yr_sch = weighted.mean(yr_sch, w = pop, na.rm = TRUE),
    .groups = "drop"
  )

# Plot the evolution of population-weighted average years of schooling over time for each region
ggplot(regional_avg, aes(x = year, y = weighted_yr_sch, color = as.factor(region_code))) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = seq(1950,2010, by = 10)) +
  scale_color_nightfall(discrete = TRUE) +
  labs(title = "Population-Weighted Average Years of Schooling Over Time",
       subtitle = "Analysis by Region (1950-2010)",
       x = "Year",
       y = "Population-Weighted Average Years of Schooling",
       color = "Region") +
  theme_minimal() +
  theme(
    legend.position = "right",
    panel.grid.minor = element_blank()
  )
```

## Visualization 2

```{r, echo = TRUE, message = FALSE}  
# 1. Prepare data with Netherlands/Fiji and "Incomplete" math
pyramid_data <- data %>%
  filter(country %in% c("Netherlands", "Fiji"), 
         year %in% c(1960, 2010)) %>%
  mutate(
    age_range = paste0(agefrom, "-", ageto),
    # Ensure the axis stays in numerical order (15-19 before 20-24)
    age_range = reorder(age_range, agefrom)
  ) %>%
  pivot_longer(
    cols = c("No_education", 
             "Primary_education", "Primary_education_completed", 
             "Secondary_education", "Secondary_education_completed", 
             "Higher_education", "Higher_education_completed"),
    names_to = "education_level",
    values_to = "percentage"
  ) %>%
  mutate(
    # Netherlands on the left (negative)
    adj_percentage = ifelse(country == "Netherlands", -percentage, percentage),
    
    # OPPOSITE ORDER: Higher education will now be at the center line
    education_level = factor(education_level, levels = rev(c(
      "No_education", 
      "Primary_education", "Primary_education_completed", 
      "Secondary_education", "Secondary_education_completed", 
      "Higher_education", "Higher_education_completed"
    )))
  )

# 2. Plotting
ggplot(pyramid_data, aes(x = adj_percentage, 
                         y = age_range, 
                         fill = education_level)) +
  geom_col(width = 0.8) +
  geom_vline(xintercept = 0, color = "black", linewidth = 0.6) +
  facet_wrap(~year) +
  # Limits set to 115 to keep bars away from the edges
  scale_x_continuous(labels = abs) + 
  scale_fill_nightfall(
    discrete = TRUE,
    reverse = TRUE,
    labels = c("Higher (Comp)", "Higher (Inc)", "Secondary (Comp)", 
               "Secondary (Inc)", "Primary (Comp)", "Primary (Inc)", "No Education")
  ) +
  labs(
    title = "Education Comparison: Netherlands (Left) vs. Fiji (Right)",
    subtitle = "Years 1960 and 2010 | Advanced degrees at the center",
    x = "Percentage of Population",
    y = "Age Groups",
    fill = "Education Level"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 7),
    panel.grid.major.y = element_blank()
  )

```
