---
title: "Tackling_Laura_Project"
author: "Nicolai Schwarz"
date: "2024-01-08"
output: html_document
---

```{r setup, include=FALSE}
# get this packages
library(tidyverse)
library(dplyr)
library(stringr)
library(emo)
library(readr)
knitr::opts_chunk$set(echo = TRUE)
 rm(list = ls())
```

Okay let's have a look, first we need to change columns and values from team roster.

Therefor we could recycle the `auto_import_csv`-function from my other Project Solution.

***

```{r folder_content_roster, echo=FALSE}
folder_path_roster <- "C:\\Users\\Pirate\\Documents\\Data-projects-with-R-and-GitHub\\Projects\\laura-burk\\volleyball-project\\team-rosters"

 files <- list.files(folder_path_roster)

 # need to change working directon
files
```

***

```{r all_paths, echo=TRUE}
# Ain't working here due to wrong working direction

folders <- list.files(getwd(), full.names = TRUE) %>%
  .[!grepl("\\.", .)]

folders

```

***

```{r import, echo=TRUE }

auto_import_csv <- function(file_path){
  
  file_list <- list.files(path = file_path, pattern = "\\.csv$", full.names = TRUE)        #-> here we could change the auto import to other data type's 
  data_frames <- list()
  
  for (file in file_list) {
    df_name <- tools::file_path_sans_ext(basename(file))  # uses filename as dataframe
    
    
    data_frames[[df_name]] <- read.csv2(file, fileEncoding = "cp852")
  }
  
  return(data_frames)
}
all_rosters <- auto_import_csv(folder_path_roster)
str(all_rosters)
#?row.names
```

***

Okay god damn it, it took me fucking long time to notice, i just had to use the `read.csv2` function instead.

should we supercharge the process now by using the `package(air)`?

***


```{r airup, echo=TRUE, eval=FALSE}
install.packages("devtools") 
devtools::install_github("soumyaray/air")
```
***

Okay instead of going through the whole API stuff. let's do some minor adjustements by hand.

Let's go for height first:

***
`#df <- mutate(df, Height = as.numeric(Height)/100)
  #df$Height <- suppressWarnings(as.numeric(df$Height) / 100)`

```{r adjust height in cm, echo=TRUE}
all_rosters2 <- lapply(all_rosters, function(df) {
  df %>%
     mutate(Height = as.numeric(Height),
           Height = ifelse(Height == 0 | Height > 3, NA, Height/100))

    #mutate(Height = ifelse(Height == 0, NA,
     #                      ifelse(as.numeric(substr(Height, 1, 1)) >= 3, as.numeric(paste0("1", substring(Height, 2))),
      #                            as.numeric(Height))))
})


str(all_rosters2)
all_rosters2
```

***

And again did some useless extra work, due to value `999` equals `NA`and in the end its just an doc....

So might be better, first mutate gender, get rid of unnecessary Positions and than finally fix the height into cm.

**For the nationality might be easiest to implement an already existing country code list.**

***

```{r adjust gender, echo=TRUE}
all_rosters3 <- lapply(all_rosters, function(df) {
 df %>%
    mutate(Gender = ifelse(str_detect(Gender, "^[fF]|2"), "female", "male"))
})


#str(all_rosters3)
#all_rosters3
```

***

Okay that was easy, probably should wrap it in a function to use it later on, on the other csv's aswell.

Or should i punch all csv's in a list first ? `r emo::ji("thinking")`

***

Let's find some nice RegEx to fix the position Issues, or lets take a look in wiki first, might we could use something from there.

-> Therefor we take a look what's the unique entry's in the "Position" column's

***

```{r unique_positions, echo=TRUE}

unique_positions <- unique(unlist(lapply(all_rosters3, function(df) df$Position)))

print(unique_positions)
```
***

Okay lets make it more recyclable.

***

```{r unique_values, echo=TRUE}

get_unique_values <- function(data, column){
  unique_values <- unique(unlist(lapply(data, function(df) df[[column]])))
  return(unique_values)
}
get_unique_values(all_rosters3, "Position")

```

***

Okay lets go...

***

```{r adjust position, echo=TRUE}
all_rosters4 <- lapply(all_rosters, function(df) {
 df %>%
    mutate(Position = str_replace_all(Position, "^[zZ].*|^[sS]e.*", "setter"),
           Position = str_replace(Position, "^[mM].*", "middle blocker"))
})


#all_rosters4
```

***

Yep this semse to be working, we could go on like that, but anyway there's still a lot of repeating, maybe we could just replace it with a **" adjustment map"**


***
```{r adjust_map position, echo=TRUE}

adjust_map <- c("^[zZ].*|^[sS]e.*" = "setter",
                "^[mM].*" = "middle blocker",
                "^[aA]u.*|^[oO]u.*" = "outside hitter",
                "^[dD].*|^[oO]p.*" = "opposite hitter",
                "^[uU].*|^[lL].*" = "libero",
                "^[tT]r.*|^[cC]h.*|.*[hH]ead.*" = "head coach")

all_rosters5 <- lapply(all_rosters, function(df) {
  df %>%
    mutate(Position = str_replace_all(Position, adjust_map))
})

#str(all_rosters5)

#all_rosters5
```

***

Let's check what we got left, with our `get_unique_values()` function

***

```{r check_whats_left, echo=TRUE}
get_unique_values(all_rosters5, "Position")
```
***

looks pretty nice so far, how can we kick out, all unnecessary positions without calling them single again?

`r emo::ji("thinking")` maybe just use an `ifelse` function before the `str_detect()`, nah this is getting to long and ugly

instead we could just use the `filter`function, but im still not confident it works proper, i should check it.
Aslong i'll kepp it in an extra chunk.

```{r adjust_map position_kick_rest, echo=TRUE}

all_rosters6 <- lapply(all_rosters, function(df) {
  df %>%
    mutate(Position = str_replace_all(Position, adjust_map))%>%
    filter(Position %in% adjust_map)

})

#str(all_rosters6)

#all_rosters6
```

***
Let's have a look again whats left.
```{r check_whats_left_filterd, echo=TRUE}
get_unique_values(all_rosters6, "Position")
```

*** 

seems like it's working, even if im still not sure, how R pulls out the right words (position names) from the whole `adjust_map` instead of looking for the pattern insight 
&#x1F937;

***

Anyway next we need to fix the country code, would love to use a predefined country-code-list. 
Due to its capability to recycle, but sounds like im missing the point again.
So first take a look what values we got.
Might be some simple Regex fixes the issues faster.

***
```{r check_country, echo=TRUE}
get_unique_values(all_rosters, "Nationality")
```

***

Alright, this looks to me like, we should do a semi automatic way, luckily there's already packages for `library(countrycode)`
`r emo::ji("thinking")`

Let's fix this German issues by hand and with some simple Regex and let the package do the rest.

***

```{r adjust_nation, echo=TRUE}

all_rosters7 <- lapply(all_rosters6, function(df) {
 df %>%
    mutate(Nationality = str_replace_all(Nationality, "^[dD].*|^[gG].*", "Deutschland"),
           Nationality = str_replace(Nationality, ".*wenia$", "Slovenia"),
            Nationality = str_replace(Nationality, ".*lie$","Australien"))
})
```

***

```{r, IOC, echo=TRUE}
#install.packages("countrycode")

library(countrycode)

apply_countrycode_de <- function(df) {
  df$Nationality <- countrycode(df$Nationality, origin = "country.name.de", destination = "ioc", nomatch = NULL)
  return(df)
}

all_rosters8 <- lapply(all_rosters7, apply_countrycode_de)

apply_countrycode_en <- function(df) {
  df$Nationality <- countrycode(df$Nationality, origin = "country.name.en", destination = "ioc", nomatch = NULL)
  return(df)
}

all_rosters9 <- lapply(all_rosters8, apply_countrycode_en)

#all_rosters9
#?`countrycode-package`
#?countrycode
```

***
Let's check
***
```{r check_final_country, echo=TRUE}
get_unique_values(all_rosters9, "Nationality")
```

***
Okay we could shorten this last fucker by using second variable in the function and get the origin part in ones.

And than put all the single parts together

***

okay this is working lets go for the topscorer `tsv` files
this should be working simple by...

***

Alright, got no glue what to do.
Let's start to import the `topscorer`files
recycling all our previous stuff.

***
```{r folder_content_topscore, echo=FALSE}
folder_path_topscorers <- "C:\\Users\\Pirate\\Documents\\Data-projects-with-R-and-GitHub\\Projects\\laura-burk\\volleyball-project\\topscorers"

 files2 <- list.files(folder_path_topscorers)

files2
```

***

Okay first had to make the `auto_import_csv`function in a more comon usable `auto_import`function to make it capable for more datatyp's

***
```{r import, echo=TRUE }

auto_import <- function(file_path, datatyp){
  
  file_list <- list.files(path = file_path, pattern = paste("\\.", datatyp, "$", sep = ""), full.names = TRUE)        #-> here we could change the auto import to other data type's 
  data_frames <- list()
  
  for (file in file_list) {
    df_name <- tools::file_path_sans_ext(basename(file))  # uses filename as dataframe
    
    read_call <- paste0("read.", datatyp, "2", sep = "")
    data_frames[[df_name]] <- do.call(read_call, list(file, fileEncoding = 'cp852', row.names = NULL))
  }
  
  return(data_frames)
}
all_rosters <- auto_import(folder_path_roster, datatyp="csv")
#str(all_rosters)
#all_rosters
#?row.names
```

***

-> God damn it, missread, the `topscorer`dat's are `.txt` instead of `.tsv` format
need to change the endings, to get the `auto_import`function working 

***

**I want to use my `auto_import`function !!!**

So let's change the dat endings of the topscorers files.

***
```{r txt_to_tsv, echo=TRUE, eval=FALSE}
txt_files <- list.files(path = folder_path_topscorers, pattern = "\\.txt$", full.names = TRUE)

for (file in txt_files) {
  tsv_file <- sub("\\.txt$", ".tsv", file)
  file.copy(file, tsv_file)
}
```

***

Let's do it with a modified suffix version

***

```{r import_all, echo=TRUE }

auto_import <- function(file_path, datatyp, suffix = ""){
  
  file_list <- list.files(path = file_path, pattern = paste("\\.", datatyp, "$", sep = ""), full.names = TRUE)        #-> here we could change the auto import to other data type's  -> put that part in a seperate variable inside the function
  data_frames <- list()
  
  for (file in file_list) {
    df_name <- tools::file_path_sans_ext(basename(file))  # uses filename as dataframe
    
    read_call <- paste0("read_", datatyp, suffix, sep = "") # need to paste the "2" here to read (german) encodeing <- solved it with suffix
    data_frames[[df_name]] <- do.call(read_call, list(file, locale = locale(encoding = "cp852")))
  }
  
  return(data_frames)
}

```

```{r import_topscore, echo=TRUE}
all_topscores <- auto_import(folder_path_topscorers, datatyp = "tsv")
str(all_topscores)
```

finally solved it

***

***

## # need to change working directon for:

** not working in Solution, due to we changed df from list of df into tibble**

***

```{r check_games, echo=TRUE, eval=FALSE}
get_unique_values(all_games_clean, "Geschlecht")
na_count <- sapply(all_games, function(df) sum(is.na(df$Geschlecht)))

na_count


```

***

```{r adjust_games, echo=TRUE}
all_games_clean <- lapply(all_games, function(df) {
    df %>%
    setNames(gsub(" ", "_", colnames(.))) %>%
  mutate(Mannschaft_1 = gsub("Ř", "ü", Mannschaft_1),
         Mannschaft_2 = gsub("Ř", "ü", Mannschaft_2),
         Gastgeber = gsub("Ř", "ü", Gastgeber),
         Austragungsort = gsub("Ř", "ü", Austragungsort),
         Geschlecht = ifelse(str_detect(Geschlecht, "^[fF]|2"), "female", "male"))
})
  
all_games_clean
```

***

Alright,so far, so good!

Next mutating all these values requiered for the barplot.

Let's get a list of all Teams first
***
```{r all_games_teams,echo=TRUE}
all_games_teams <- get_unique_values(all_games_clean, "Mannschaft_1")
all_games_teams
```
***

now let's check how often each of them played at home and away

***

```{r home_games,echo=TRUE}
num_home_games <-lapply(all_games_teams, function(team){
  lapply(all_games_clean, function(df){
    df%>%
    filter(Mannschaft_1 == team)%>%
    nrow()
  })
})

num_away_games <-lapply(all_games_teams, function(team){
  lapply(all_games_clean, function(df){
    df%>%
    filter(Mannschaft_2 == team)%>%
    nrow()
  })
})

num_total_games <-lapply(all_games_teams, function(team){
  lapply(all_games_clean, function(df){
    df%>%
    filter(Mannschaft_1 == team|Mannschaft_2 == team)%>%
    nrow()
  })
})

```



***

This already looks like, as we can put it together.

***

```{r try}
count_team_games <- function(team, df) {
  home_games <- df %>%
    filter(Mannschaft_1 == team) %>%
    nrow()
  
  away_games <- df %>%
    filter(Mannschaft_2 == team) %>%
    nrow()
  
  total_games <- df %>%
    filter(Mannschaft_1 == team | Mannschaft_2 == team) %>%
    nrow()
  
  return(c(Home = home_games, Away = away_games, Total = total_games))
}

num_games_in_once <- lapply(all_games_teams, function(team) {
  lapply(all_games_clean, count_team_games, team = team)
})
num_games_in_once
```

***

Next go for win and losses

***

```{r try victory, echo=TRUE}
num_victory <-lapply(all_games_teams, function(team){
  lapply(all_games_clean, function(df){
    df%>%
    filter((Mannschaft_1 == team & Satzpunkte_1 == 3) | (Mannschaft_2 == team & Satzpunkte_2 == 3))%>%
    nrow()
  })
})

num_losses <-lapply(all_games_teams, function(team){
  lapply(all_games_clean, function(df){
    df%>%
    filter((Mannschaft_1 == team & Satzpunkte_1 != 3) | (Mannschaft_2 == team & Satzpunkte_2 != 3))%>%
    nrow()
  })
})
num_losses
```

***
For all the average stuff im not sure if it aint  more easy to combine all games into one df to get total counts more easy for each team.

Anyway we can count total of sets now ans than let's see.

***

```{r total_sets_played_per_game, echo=TRUE}

num_sets_per_game_played <-lapply(all_games_teams, function(team){
  lapply(all_games_clean, function(df){
    df%>%
    filter(Mannschaft_1 == team|Mannschaft_2 == team)%>%
    summarise(
      Sum_Satzpunkte = sum(Satzpunkte_1 + Satzpunkte_2),
      Avg_Satzpunkte = mean(Satzpunkte_1 + Satzpunkte_2)) # actually could put the whole sum function in the mean function
  })
})

```
***

**Note:** * This is the total and average number of setsplayed per game* **not** average sets to win, wait a minute.....

than it should be **right?**

Next Average points per set

***
**Dunno here why its working with backticks but aint with normal ""**

```{r avg_points_per_game, echo=TRUE}

avg_points_per_set<- lapply(all_games_teams, function(team) {
  lapply(all_games_clean, function(df) {
    df %>%
      filter(Mannschaft_1 == team | Mannschaft_2 == team) %>%
      summarise(
        Avg_Ballpunkte_first_set = mean(ifelse(Mannschaft_1 == team, `Satz_1_-_Ballpunkte_1`, `Satz_1_-_Ballpunkte_2`)),
        Avg_Ballpunkte_sec_set = mean(ifelse(Mannschaft_1 == team, `Satz_2_-_Ballpunkte_1`, `Satz_2_-_Ballpunkte_2`)),
        Avg_Ballpunkte_third_set = mean(ifelse(Mannschaft_1 == team, `Satz_3_-_Ballpunkte_1`, `Satz_3_-_Ballpunkte_2`)),
        Avg_Ballpunkte_fourth_set = mean(ifelse(Mannschaft_1 == team, `Satz_4_-_Ballpunkte_1`, `Satz_4_-_Ballpunkte_2`)),
        Avg_Ballpunkte_fifth_set = mean(ifelse(Mannschaft_1 == team, `Satz_5_-_Ballpunkte_1`, `Satz_5_-_Ballpunkte_2`))
        
      )
  })
})

avg_points_per_set
```