---
title: "Dragonflies - Lizzola Project Solution"
output: md_document
date: "`r Sys.Date()`"
---
  
# Dragonflies - to find out under which weather conditions autumn mosaic dragonfly prefers to fly

## Data manipulation

```{r data set, echo=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(tidyverse)
library(kableExtra)
library(knitr)
library(fs)
library(lubridate)
library(ggplot2)


if (!is.null(current_input())) {
  rmd_dir <- path_dir(current_input())
} else {
  rmd_dir <- getwd()
}
weather_data <- read.csv(file.path(rmd_dir, "Weather_Data_Poland.csv"))
```

Form mean values from the extracted and revised weather data for the morning (10 a.m. to 2 p.m.) and midday (2 p.m. to 4 p.m.) periods. Creation of a new column based on the wind direction in degrees (True Dir.) with the wind direction in compass direction (e.g. NW, SSO)
```{r revision, echo=TRUE, message=FALSE, warning=FALSE}
transformed_weather_data <- weather_data %>%
  slice(-1) %>%
  select(Time, Temp, Wind.Speed, True.Dir.) %>%
  rename("Time in yyyy-MM-dd hh:mm:ss" = Time, 
         "Temperature in Celsius" = Temp, 
         "Wind speed in m/s" = Wind.Speed, 
         "Wind direction in degrees" = True.Dir.) %>%
  separate("Time in yyyy-MM-dd hh:mm:ss", into = c("Date", "Time"), sep = " ") %>%
  mutate( 
    period = case_when(
    Time %in% c("10:00:00", "11:00:00", "12:00:00", "13:00:00") ~ "Morning",
    Time %in% c("14:00:00", "15:00:00") ~ "Afternoon",
    TRUE ~ NA_character_)
    ) %>%
  group_by(Date, period) %>%
  filter(period %in% c("Morning", "Afternoon")) %>% # here I changed Midday to Afternoon that both dfs can be merged
  mutate(
    `Mean Temperature in Celsius` = mean(as.numeric(`Temperature in Celsius`), na.rm = TRUE),
    `Mean Wind Speed in m/s` = mean(as.numeric(`Wind speed in m/s`), na.rm = TRUE),
    `Mean Wind Direction in degrees` = mean(as.numeric(`Wind direction in degrees`), na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(
    `Wind direction in compass direction` = case_when(
      `Wind direction in degrees` >= 337.5 | `Wind direction in degrees` < 22.5 ~ "N",
      `Wind direction in degrees` >= 22.5 & `Wind direction in degrees` < 67.5 ~ "NE",
      `Wind direction in degrees` >= 67.5 & `Wind direction in degrees` < 112.5 ~ "E",
      `Wind direction in degrees` >= 112.5 & `Wind direction in degrees` < 157.5 ~ "SE",
      `Wind direction in degrees` >= 157.5 & `Wind direction in degrees` < 202.5 ~ "S",
      `Wind direction in degrees` >= 202.5 & `Wind direction in degrees` < 247.5 ~ "SW",
      `Wind direction in degrees` >= 247.5 & `Wind direction in degrees` < 292.5 ~ "W",
      `Wind direction in degrees` >= 292.5 & `Wind direction in degrees` < 337.5 ~ "NW",
      TRUE ~ NA_character_
    ))
```

Merging the data from the revised weather file with the dragonfly catch data
```{r merge, echo=TRUE, message=FALSE, warning=FALSE} 

if (!is.null(current_input())) {
  rmd_dir <- path_dir(current_input())
} else {
  rmd_dir <- getwd()
}
catch_data <- read.csv(file.path(rmd_dir, "Fangdaten_Poland.csv"), sep = ";")

revised_catch_data <- catch_data %>%
  mutate(across(everything(), ~na_if(.x, ""))) %>%
  mutate(across(everything(), ~na_if(.x, "NA"))) %>%
  filter(if_any(everything(), ~ !is.na(.))) %>%
  mutate(Datum = format(dmy(Datum), "%Y-%m-%d"))

merged_data <- left_join(transformed_weather_data, revised_catch_data, by = c("Date" = "Datum", "period" = "Uhrzeit"))

```

## Data visualization

```{r visualization, echo=TRUE, message=FALSE, warning=FALSE}

```