---
title: "Distribution of Speakers Time in Lord of the Rings"
output: md_document
---

## Distribution of Speakers Time in Lord of the Rings

------------------------------------------------------------------------

## Introduction

Lord of the Rings is a fantasy novel trilogy written by J.R.R. Tolkien. Until today, it is one of the most popular and influential works of fantasy literature, and has been adapted into several films, video games, and other media. The story follows a group of characters as they embark on a quest to destroy a powerful ring that has the potential to enslave the world.

In this analysis, we will explore the distribution of speakers time in Lord of the Rings. For that, not the minutes of speaking in the movies, but the number of words spoken by each character in the three books will be used as a proxy for the time spent speaking.

------------------------------------------------------------------------

## Aim of the Analysis

This analysis will help us understand which characters have the most dialogue and how the speaking time is distributed among the characters in the story. We will also explore any patterns or trends in the distribution of speakers time, and how it may relate to the overall narrative of the novel. This also in light of the gender representation in the novel, as the movies are often criticized for their lack of diverse representation.

------------------------------------------------------------------------

## Data

The data used for this analysis will consist of two datasets.

### 1st Dataset: Words by Character

This dataset is concerned with the number of words spoken contains the number of words spoken by each character in all three volumes of Lord of the Rings. The data is organized in a tabular format, with each row representing a character and the number of words they spoke in a particular chapter of the novel. The columns include the name of the film (book), the chapter, the character's name, race and the number of words spoken by them:

|            Film            |   Chapter    | Character |  Race  | Words |
|:--------------------------:|:------------:|:---------:|:------:|:-----:|
| The Fellowship of the Ring | 01: Prologue |   Bilbo   | Hobbit |   4   |
| The Fellowship of the Ring | 01: Prologue |  Elrond   |  Elf   |   5   |
| The Fellowship of the Ring | 01: Prologue | Galadriel |  Elf   |  460  |

The columns are defined as follows:

-   **Film**: The name of the individual novel (The Fellowship of the Ring, The Two Towers, The Return of the King).
-   **Chapter**: The chapter of the specified novel in which the character speaks.
-   **Character**: The name of the character who speaks.
-   **Race**: The race (species) of the specified character.
-   **Words**: The number of words spoken by the specified character in the specified chapter within the specified book.

### 2nd DataSet: Character Information

The second dataset contains information about the characters in Lord of the Rings, including their name, race, gender and realm. The dataset is available in the project folder under the name `InformationByCharacter.csv`. The creator of this dataset is me, Emily. The data was collected and compiled by me, based on information from the novels and other sources, such as the LOTR-Wiki. The data is organized in a tabular format, with each row representing a character:

| Character |  Race  | Gender |   Realm   |
|:---------:|:------:|:------:|:---------:|
|  Aragorn  |  Men   |  Male  |  Gondor   |
|   Arwen   |  Elf   | Female | Rivendell |
|   Bilbo   | Hobbit |  Male  | The Shire |

The columns are defined as follows:

-   **Name**: The name of the specified character.
-   **Race**: The race (species) of the specified character.
-   **Gender**: The gender of the specified character.
-   **Realm**: The realm (location) of which the specified character is mainly associated with.

------------------------------------------------------------------------

# Tasks

------------------------------------------------------------------------

## 1. Data Import and Manipulation

------------------------------------------------------------------------

```{r setup, echo=FALSE, message=FALSE}

library(tidyverse)
library(ggplot2)
library(ggrepel)
library(patchwork)
library(knitr)
```

### 1.1. Fixing variable names

```{r Realm_name_change,echo=FALSE, message=FALSE}


words_data<-read_csv("WordsByCharacter.csv")

info_data<-read_csv("InformationByCharacter.csv", locale=locale(encoding="Windows-1252"))%>%
  mutate(Realm=case_when(Realm=="Lothl—rien"~"Lothlórien",TRUE~Realm))

info_data%>%distinct(Realm)

info_data %>%
  select(Realm) %>%
  kable(col.names = "Realm")
```

: Realm "Lothelórien" was written as "Lothl-rien", so changed the name for it.

### 1.2. Cleaning up NA

```{r cleaning_NA,echo=FALSE, message=FALSE}

sum(is.na(words_data))
sum(is.na(info_data))

info_data%>%
  filter(if_any(everything(),is.na))

info_data <- info_data %>%
  mutate( Gender = replace_na(Gender, "Unknown"),
          Realm  = replace_na(Realm, "Unknown"))

info_data %>%
  filter(Gender=="Unknown" | Realm=="Unknown")%>%
  select(Gender,Realm)%>%
  kable(col.names = c("Gender", "Realm"))
```

: Found several NA data, and changed them into "Unknown" so that the other information of the column still remains.

### 1.3. Fixing Character Name

```{r Checking_charater_names}

words_names <- unique(words_data$Character)

info_names  <- unique(info_data$Character)

setdiff(unique(words_data$Character), unique(info_data$Character))

info_data$Character[info_data$Character=="Bosun"]<-"Boson"

```

: In Info_data, there was "Bosun" instead of "Boson", so fixed it to unify the names with words_data

### 1.4. Merging dataset

```{r merging_datasets}

lotr_data<- words_data %>%
  left_join(info_data%>%
      select(-Race),by = "Character")
```

: Merged dataset by Character

------------------------------------------------------------------------

## 2. Data Analysis & Visualization

------------------------------------------------------------------------

### 2.1 Speakers Time by Volume

**Total number of spoken words by book**

```{r table_total_words, echo=FALSE, message=FALSE}

words_spoken <- lotr_data %>%
  group_by(Film) %>%
  summarise( total_words = sum(Words),.groups="drop")

words_spoken %>%
  knitr::kable(
    col.names = c("Book", "  Total words"))
```

-   What is the total number of words spoken?

The Fellowship Of The Ring: `r words_spoken$total_words[words_spoken$Film == "The Fellowship Of The Ring"]` The Return Of The King: `r words_spoken$total_words[words_spoken$Film == "The Return Of The King"]` The Two Towers: `r words_spoken$total_words[words_spoken$Film == "The Two Towers"]`

-   Which book has the highest number of words spoken?

Highest: `r words_spoken$Film[which.max(words_spoken$total_words)]`

-   Which book has the lowest number of words spoken?

Lowest: `r words_spoken$Film[which.min(words_spoken$total_words)]`

**Visualization for "Number of Words spoken in each Volume"**

```{r donut_chart, echo=FALSE, message=FALSE}

words_spoken<- lotr_data%>%
  group_by(Film)%>%
  summarise(total_words=sum(Words, na.rm=T), .groups="drop")



words_spoken%>%
  mutate(percentage=total_words/sum(total_words)*100,
         label=paste0(round(percentage,2),"%"))%>%
  ggplot(aes(x=2,y=total_words,fill=Film))+
  geom_col()+
  geom_text(aes(label=label),position=position_stack(vjust=0.5))+
  coord_polar(theta="y")+
  xlim(0,2.5)+
  theme_void()+
  labs(fill="Books")+
  annotate("text",x = 0,y= 0,label = "Number of Words\nby Volume")+  
  scale_fill_grey(start = 0.3, end = 0.6)

```

### 2.2. Speaking Time by Character

#### 2.2.1. Total Speaking Time by Character across all 3 books

**Total number of words spoken by each character across all three books**

```{r table_each_character, echo=FALSE, message=FALSE}

character_total <- lotr_data %>%
  group_by(Character) %>%
  summarise(total_words = sum(Words, na.rm = TRUE),
            .groups = "drop") %>%
  arrange(desc(total_words))

character_total%>%
  slice_head(n=9)%>%
  kable(col.names = c("Character", "Total Words"))
```

**Visualization for "Top 9 Characters with the most speaking time"**

```{r top9_chart_books, echo=FALSE, message=FALSE}

top9_data <- lotr_data %>%
  group_by(Character, Gender) %>%
  summarise(total_words = sum(Words, na.rm=TRUE),.groups="drop") %>%
  mutate(Character=fct_lump(
    Character,
    n=9, 
    w=total_words, 
    other_level="Other"),
  Gender=if_else(
    Character=="Other",
    "Other",
    Gender)) %>%
  group_by(Character, 
           Gender) %>% 
  summarise(total_words=sum(total_words),
            .groups="drop") %>%
  mutate(percentage=total_words/sum(total_words)*100,
         label=paste0(round(percentage,2),"%")) %>%
  arrange(desc(total_words))%>%
  mutate(percentage=total_words/sum(total_words)*100,
         label=paste0(round(percentage,2),"%"),
         ymin=lag(cumsum(percentage),default=0),
         ymax=cumsum(percentage),
         ypos=(ymin+ymax)/2)

top9_data%>%
  ggplot()+
  geom_rect(
    aes(ymin=ymin,ymax=ymax,xmin=0,xmax=1,fill=Gender),
    color="white")+
  geom_segment(aes(x=1, xend=1.3, y=ypos, yend=ypos))+
  geom_text(aes(x=1.5,y=ypos,label=Character),size=2.5)+
  geom_text(aes(x=0.8,y=ypos,label=label),color="white",size=2)+
  coord_polar(theta="y")+
  theme_void()+
  labs(title="Total Speaking Time by Character across all 3 books", fill="Characters")+
  scale_fill_manual(values=c("Male"="blue","Female"="red","Other"="grey"))
```

#### 2.2.2. Visualizaion for "Total Speaking Time by Character for each individual book"

```{r top9_chart_character, fig.width=9, fig.height=18, echo=FALSE, message=FALSE}

top9_by_book<-lotr_data%>%
  mutate(Film=factor(Film,levels=c("The Fellowship Of The Ring","The Two Towers","The Return Of The King"))) %>%
  group_by(Film,Character,Gender)%>%
  summarise(total_words=sum(Words),
          .groups="drop")%>%
  group_by(Film)%>%
  mutate(Character=fct_lump(
    Character,
    n=9,
    w=total_words,
    other_level="Other"))%>%
  mutate(Gender=if_else(
    Character=="Other",
    "Other",
    Gender))%>%
  group_by(Film,Character,Gender)%>%
  summarise(total_words=sum(total_words),.groups="drop")%>%
  group_by(Film)%>%
  arrange(desc(total_words))%>%
  mutate(percentage=total_words/sum(total_words)*100,
         label=paste0(round(percentage,2),"%"),
         ymax=cumsum(percentage),
         ymin=lag(ymax,default=0),
         ypos=(ymin+ymax)/2) %>%
  ungroup()

top9_by_book %>%
  ggplot()+
  geom_rect(aes(ymin=ymin,ymax=ymax,xmin=0,xmax=1,fill=Gender),color="white")+
  geom_segment(aes(x=1, xend=1.2, y=ypos, yend=ypos))+
  geom_text(aes(x=1.3, y=ypos, label=Character),size=2.5) +
  geom_text(aes(x=0.8,y=ypos,label=label),color="white",size=2)+
  facet_wrap(~Film,ncol=1) +
  coord_polar(theta = "y") +
   scale_fill_manual(values=c("Male"="blue","Female"="red","Other"="grey"))+
  theme_void()+
  theme(legend.position="right")
```

### 2.3. Speakers Time of Frodo

Here we loloked at three different aspects related to Frodo's speaking time, and visualized three charts into one scene.

#### 2.3.1. Speaking Time of Frodo across all 3 books

```{r Frodo_3_books,echo=FALSE, message=FALSE}

frodo_data<- lotr_data%>%
  mutate(
    category=if_else(
       Character =="Frodo","Frodo","Other"))%>%
  group_by(category)%>%
  summarise(
    words=sum(Words, narm=TRUE),
    .groups="drop")%>%
  mutate(percent=words/sum(words)*100,
         label=paste0(category,"\n",round(percent,2),"%"))

frodo_data %>%
  select(category,percent)%>%
  mutate(percent = paste0(round(percent, 2), "%")) %>%
  kable(
    col.names = c("Character", "Percent")
  )

```

: Frodo had 7.14% of speaking throughout all three books

#### Visualization for Speaking Time of Frodo across all 3 books

```{r pie_chart_Frodo,echo=FALSE, message=FALSE}

p_frodo_book<-frodo_data%>%ggplot(aes(x="",y=words,fill=category))+
  geom_col()+
  geom_text(aes(label=label),position=position_stack(vjust=0.5),size=4,color="white")+
  coord_polar(theta="y")+
  theme_void()+
  labs(title="Speaking Time of Frodo in all 3 books", fill="Character")+
  scale_fill_manual(values=c("Frodo"="Blue","Other"="Grey"))+
  theme(plot.title = element_text(hjust = 0.5))

```

#### 2.3.2. Percentage of chapters where Frodo speaks

```{r Frodo_chapters,echo=FALSE, message=FALSE}

all_chapters<-lotr_data%>%
distinct(Film,Chapter)%>%
nrow()

chapters_frodo<-lotr_data%>%
filter(Character=="Frodo")%>%
distinct(Film,Chapter)%>%
nrow()

chapters_piechart<-lotr_data%>%
  group_by(Film,Chapter)%>%
  summarise(chapters_frodo=any(Character=="Frodo"),.groups="drop")%>%
  mutate(all_chapters=n(),category=if_else(chapters_frodo,"Frodo", "Other"))%>%
  group_by(category)%>%
  summarise(chapters=n(),all_chapters=first(all_chapters),.groups="drop")%>%
  mutate(percent=chapters/all_chapters *100,
         label=paste0(category,"\n",round(percent,2),"%"))

chapters_piechart %>%
  select(category,percent)%>%
  mutate(percent = paste0(round(percent, 2), "%")) %>%
  kable(
    col.names = c("Character", "Percent")
  )

```

: Frodo spoke in 30.32% of the chapters throughout all three books.

#### Visualization for Percentage of chapters where Frodo speaks

```{r pie_chart2_Frodo,echo=FALSE, message=FALSE}
p_frodo_chapters<-chapters_piechart%>%
  ggplot(aes(x="",y=chapters,fill=category))+
  geom_col()+
  geom_text(aes(label=label),position=position_stack(vjust=0.5),size=4,color="white")+
  coord_polar(theta="y")+
  theme_void()+
  labs(title="Percentage of Chapters Frodo speaks", fill="Character")+
  scale_fill_manual(values=c("Frodo"="Blue","Other"="Grey"))+
  theme(plot.title = element_text(hjust = 0.5))

```

#### 2.3.3. Percentage of Frodo speaking for each chapters in each book

```{r Frodo_percentage_each_chapter,echo=FALSE, message=FALSE}

book_by_order<-c("The Fellowship Of The Ring","The Two Towers","The Return Of The King")

frodo_chapters<-lotr_data%>%
  mutate(Film=factor(Film,levels=book_by_order),row_fixed=row_number())%>%
  group_by(Film,Chapter)%>%
  summarise(chapter_order=min(row_fixed),total_words=sum(Words),frodo_words=sum(if_else(Character=="Frodo",Words,0)), .groups="drop")%>%
  arrange(Film,chapter_order)%>%
  group_by(Film)%>%
  mutate(chapter_for_each_book=row_number())%>%
  ungroup()%>%
  mutate(chapter_label=row_number(),percent_label=frodo_words/total_words*100)

frodo_chapters %>%
  select(Film,Chapter,percent_label)%>%
  mutate(percent_label = paste0(round(percent_label, 2), "%")) %>%
  kable(
    col.names = c("Volume","Chapter", "Percent")
  )
```

#### Visualization for Percentage of Frodo speaking for each chapters in each book

```{r line_graph_Frodo}
book_names<-frodo_chapters%>%
  group_by(Film)%>%
  summarise(start=min(chapter_label),end=max(chapter_label),middle=(start+end)/2,.groups="drop")%>%
  arrange(Film)%>%
  mutate(book_label=c("Fellowship","Two Towers","Return"))
  
book_boundaries<-book_names%>%
  slice_head(n=nrow(book_names)-1)%>%
  mutate(boundary=end+0.5)

p_frodo_lines<-frodo_chapters%>%
  ggplot(aes(x=chapter_label,y=percent_label,))+
  geom_line(linewidth=0.6)+
  geom_point(size=1)+
  geom_text(data=frodo_chapters%>%filter(frodo_words>0),
            aes(label=chapter_for_each_book),vjust=-0.7,size=2)+
  geom_vline(data=book_boundaries,aes(xintercept=boundary),
             linewidth=0.5,color="grey")+
  scale_x_continuous(breaks=book_names$middle, labels=book_names$Film)+
  scale_y_continuous(limits=c(0,100),labels=scales::label_percent(scale=1))+
  labs(title="Frodo's Speaking Percentage by Chapter",x="",y="% of words")+
  theme_minimal()
```

#### 2.3.4. All three charts related to Frodo as a Speaker

```{r final_plot, fig.width=15, fig.height=10, echo=FALSE, message=FALSE}


plot_all<-p_frodo_lines / (p_frodo_chapters | plot_spacer()| p_frodo_book)+plot_layout(widths=c(1,0.2,1))

plot_all

```
