---
title: "solution for timo's project"
author: "Elise Bücklein"
date: "`r Sys.Date()`"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
#install.packages("magrittr")
library(magrittr)
library(reshape2)
library(ggplot2)
library(gridExtra)
library(grid)
```

## Loading and cleaning up the data
```{r}
# # Read lines from the file
# lines <- read_lines("Projects/timow98/SoilData.csv")
# 
# # Remove the second line before reading in the file
# units <- lines[2] # save the line with the units for later
# lines <- lines[-2]
# 
# # Create a temporary file-like object
# temp_file <- tempfile()
# write_lines(lines, temp_file)
# 
# # Read the modified lines into a data frame
# df <- read.delim2(temp_file, sep = ";", 
#                   na = c("99", "99999,99", 99999))
# 
df_col <- read_delim("Projects/timow98/SoilData.csv", delim = ";",
                      n_max = 1,
                      show_col_types = F)

df <- read_delim("Projects/timow98/SoilData.csv", delim = ";",
                     locale = locale(decimal_mark = ","),
                     skip = 2,
                     col_names = F,
                     na = c("99", "99999,99", 99999),
                     show_col_types = F)
```

### Clean-Up
1.1 the column KAKpot2 is not needed    
1.2 the header displays numbers which do not make much sense. I would like to get rid of them by using rename_with() in combination with regular expressions.    
1.3 there is a typo in the column "Expozition". It should actually be "Exposition" 
1.4 integrate the unit into the header
```{r}
# Remove the random numbers in the variable names
names(df_col) <- str_replace(names(df_col), "_.*", "") # 1.2

# Correct spelling mistake 
names(df_col)[9] <- "Exposition" # 1.3

# Adding an underscore before each unit and removing all empty spaces # 1.4 ff.
df_col %<>%
  mutate_all(~ paste0("_", .)) %>%
  mutate_all(~ str_remove_all(., " ")) %>%
  mutate_all(~ str_remove_all(., "_\\[\\]"))

# Removing all white spaces in the names of the variables
names(df_col) <- trimws(names(df_col), "right")

# Adding the thing to each variable name
units <- df_col[1,] %>% unname() %>% unlist()
names(df_col) <- paste0(names(df_col), units)

names(df) <- names(df_col)

# Remove variable "KAKpot 2
df$`KAKpot 2_[cmol/kg]` <- NULL # 1.1
```

1.6 two columns need to be calculated: base_saturation_[%] by dividing "Kationen" by "KAKpot" and SOM_[%] by multiplying Corg with 1,72  
```{r}
df %<>%
  mutate(`base_saturation_[%]` = `Kationen_[mmol/kg]`/`KAKpot_[mmol/kg]`,
         `SOM_[%]` = 1.72 * `Corg_[%]`)
```

1.7 Column S+U+T is not always 100%. Write a new column and put in "more", "less" and "exactly" depending on the value in the column S+U+T. Then, write a query that counts and prints the number of values.
```{r}
df %<>%
  mutate(`S+U+T_descr` = case_when(`S+U+T_[%]` > 100 ~ "more",
                                `S+U+T_[%]` < 100 ~ "less", 
                                `S+U+T_[%]` == 100 ~ "exactly",
                                .default = NA))
df %>%
  count(`S+U+T_descr`)
```

## Visualization
### Stacked Barplot
Create a stacked barplot that shows the values for Ca, Mg and Kationen of the **last** horizon of the profiles 55, 71, 102 and 109.
Use the same colors as in the example plot (viridis)

```{r}
# make a dataset that includes only the desired values
df_barplot <- df %>%
  filter(Profil %in% c(55, 71, 102, 109)) %>% # only these profiles are relevant
  group_by(Profil) %>%
  slice_max(Horizont) %>% # only keep the last horizon of each profile
  ungroup %>%
  select(c("Profil", "Ca_[mmol/kg]", "Mg_[mmol/kg]", "Kationen_[mmol/kg]")) # only keep Ca, Mg and Kationen variables

# shape the dataset into long format
df_barplot <- pivot_longer(df_barplot, cols = -Profil, names_to = "variable", values_to = "value") %>%
  mutate(Profil = factor(Profil))

# create the barplot
# x axis should be the 4 profiles
# y axis should be the three kinds of variables, but stacked 
ggplot(df_barplot, aes(x = Profil, y = value, fill = variable)) +
  geom_bar(stat = "identity", position = "stack") + 
  scale_fill_viridis_d(labels = c("Ca", "Kationen", "Mg")) + # color should come from viridis
  labs(y = "mmol/kg") +
  theme_minimal()
```

### Piechart of grain sizes

3 piecharts of the "Ah" horizons of the profiles 82, 111 and 134 next to each other where I can see the portion of S (Sand), U (silt) and T

```{r}
# make a dataset that includes only the desired values
df_pie <- df %>%
  filter(Profil %in% c(82, 111, 134)) %>% # only these profiles are relevant
  filter(grepl("^Ah\\s*$", Horizontbezeichnung)) %>% # only keep the horizons with "Ah"
  select(c("Profil", "S_[%]", "U_[%]", " T_[%]"))

# melt to long format
df_pie_long <- pivot_longer(df_pie, -Profil, names_to = "variable", values_to = "value")

# create pie charts for each profile
df_pie_long %>%
  ggplot(aes(x = "", y = value, fill = variable)) +
  geom_bar(stat = "identity", width = 1) +
  geom_text(aes(label = paste0(round(value), "%")), position = position_stack(vjust = 0.5)) + 
  coord_polar("y") +
  scale_fill_discrete(labels = c("Sand", "Silt", "Clay")) +  # Change legend labels
  theme_void() +
  facet_wrap(~ Profil) +
  ggtitle("Comparison of grain size composition in Ah horizons") + 
  theme(plot.title = element_text(hjust = 0.5, vjust = 5))
```

