---
title: "solution for timo's project"
author: "Elise Bücklein"
date: "`r Sys.Date()`"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
#install.packages("magrittr")
library(magrittr)
library(reshape2)
library(ggplot2)
library(gridExtra)
library(grid)
```

## Loading and cleaning up the data
```{r}
# Read lines from the file
lines <- read_lines("Projects/timow98/SoilData.csv")

# Remove the second line before reading in the file
units <- lines[2] # save the line with the units for later
lines <- lines[-2]

# Create a temporary file-like object
temp_file <- tempfile()
write_lines(lines, temp_file)

# Read the modified lines into a data frame
df <- read.delim2(temp_file, sep = ";", 
                  na = c("99", "99999,99", 99999))
```

### Clean-Up
1.1 the column KAKpot2 is not needed    
1.2 the header displays numbers which do not make much sense. I would like to get rid of them by using rename_with() in combination with regular expressions.    
1.3 there is a typo in the column "Expozition". It should actually be "Exposition" 
```{r}
# Remove the random numbers in the variable names
names(df) <- str_replace(names(df), "_.*", "") # 1.2

# Correct spelling mistake 
names(df)[9] <- "Exposition" # 1.4

# Change back the two column names that were supposed to have "+" instead of "."
names(df)[c(28, 30)] <- str_replace_all(names(df)[c(28, 30)], "\\.", "+")
```

1.4 integrate the unit into the header

```{r}
# Make a vector out of the string that holds the units
units <- unlist(strsplit(units, ";"))

# Adding an underscore before each unit
units <- paste("_", units, sep = "")
# manually correcting one error
units[8] <- "_[%]"

# Adding the thing to each variable name
names(df) <- paste(names(df), units, sep = "")

# Remove the unit if it is empty
names(df) <- str_replace(names(df), "_\\[\\]$", "")

# Remove variable "KAKpot 2
df$"KAKpot.2_[cmol/kg]" <- NULL # 1.1
```


1.6 two columns need to be calculated: base_saturation_[%] by dividing "Kationen" by "KAKpot" and SOM_[%] by multiplying Corg with 1,72  
```{r}
df %<>%
  mutate("base_saturation_[%]" = !!sym("Kationen_[mmol/kg]")/!!sym("KAKpot_[mmol/kg]"),
         "SOM_[%]" = 1.72 * !!sym("Corg_[%]"))
```

1.7 Column S+U+T is not always 100%. Write a new column and put in "more", "less" and "exactly" depending on the value in the column S+U+T. Then, write a query that counts and prints the number of values.
```{r}
df %<>%
  mutate(S_U_T = (!!sym("S_[%]") + !!sym("U_[%]") + !!sym("T_[%]") )) %>%
  mutate(S_U_T_descr = case_when(S_U_T > 100 ~ "more",
                                S_U_T < 100 ~ "less", 
                                S_U_T == 100 ~ "exactly",
                                .default = NA))
```

## Visualization
### Stacked Barplot
Create a stacked barplot that shows the values for Ca, Mg and Kationen of the **last** horizon of the profiles 55, 71, 102 and 109.
Use the same colors as in the example plot (viridis)

```{r}
# make a dataset that includes only the desired values
df_barplot <- df %>%
  filter(Profil %in% c(55, 71, 102, 109)) %>% # only these profiles are relevant
  group_by(Profil) %>%
  filter(Horizont == max(Horizont)) %>% # only keep the last horizon of each profile
  ungroup %>%
  select(c("Profil", "Ca_[mmol/kg]", "Mg_[mmol/kg]", "Kationen_[mmol/kg]")) # only keep Ca, Mg and Kationen variables

# shape the dataset into long format
df_barplot <- pivot_longer(df_barplot, cols = -Profil, names_to = "variable", values_to = "value") %>%
  mutate(Profil = factor(Profil))

# create the barplot
# x axis should be the 4 profiles
# y axis should be the three kinds of variables, but stacked 
ggplot(df_barplot, aes(x = Profil, y = value, fill = variable)) +
  geom_bar(stat = "identity", position = "stack") + 
  scale_fill_viridis_d(labels = c("Ca", "Kationen", "Mg")) + # color should come from viridis
  labs(y = "mmol/kg") +
  theme_minimal()
```

### Piechart of grain sizes

3 piecharts of the "Ah" horizons of the profiles 82, 111 and 134 next to each other where I can see the portion of S (Sand), U (silt) and T

```{r}
# make a dataset that includes only the desired values
df_pie <- df %>%
  filter(Profil %in% c(82, 111, 134)) %>% # only these profiles are relevant
  filter(grepl("^Ah\\s*$", Horizontbezeichnung)) %>% # only keep the horizons with "Ah" (and also "Ah " - thanks for nothing! ;))
  select(c("Profil", "S_[%]", "U_[%]", "T_[%]"))

# melt to long format
df_pie_long <- pivot_longer(df_pie, -Profil, names_to = "variable", values_to = "value")

# create pie charts for each row
plots <- lapply(unique(df_pie_long$Profil), function(profile) {
  df_subset <- filter(df_pie_long, Profil == profile)
  ggplot(df_subset, aes(x = "", y = value, fill = variable)) +
    geom_bar(stat = "identity", width = 1) +
    geom_text(aes(label = paste0(round(value), "%")), position = position_stack(vjust = 0.5)) + 
    coord_polar("y") +
    labs(caption = paste("Profil", profile)) +
    scale_fill_discrete(labels = c("Sand", "Silt", "Clay")) +  # Change legend labels
    theme_void() +
    theme(plot.caption = element_text(hjust=0.5, size=rel(1.2)))
})

# arrange next to each other
grid.arrange(grobs = plots, ncol = 3,
             top = textGrob("Comparison of grain size composition in Ah horizons")
)
```

