---
title: "project-solution by johannaroever"
author: "Johanna Röver"
date: "`r Sys.Date()`"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("tidyverse")
library(tidyverse)
```

# Solution for Soil Data Project

```{r, include=T}
# get data from .csv
soilData <- read_delim("SoilData.csv", delim = ";", col_types = list(.default = col_character()),
escape_double = FALSE,  trim_ws = TRUE)

```

## 1. Data cleanup

**1.1 the column KAKpot2** is not needed  

```{r, echo=TRUE}
# delete column KAKpot2
  soilData <- subset(soilData, select=-c(`KAKpot 2`))
```

**1.2 the header** displays numbers which do not make much sense. I would like to get rid of them by using rename_with() in combination with regular expressions. 
```{r, echo=TRUE}

# delete numbers at end of column names
soilData <- rename_with(soilData , .fn=str_replace, pattern = "_\\d*", replacement='')

```

**1.3 there is a second header** with the respective units. Leaving this line will interfere with the datatype in R, that's why I want to delete this line and instead integrate the unit into the header such like this: Tiefe\_\[cm\]. For this dataset you will need to work with backticks. If you are not familiar with them, you can read about them [<span style="color:blue">here</span>](https://jhudatascience.org/intro_to_r/resources/quotes_vs_backticks.html). Also, the function scan() could be useful (depending on your aproach) because it enables you to read for example specific data rows directly from the source file.  

```{r, echo=TRUE}

 # add units to header 
 colnames(soilData)   <- paste(colnames(soilData), soilData[1,], sep = "_")

 # clean header columns without units of unnecessary characters
 colnames(soilData) <- gsub(("_\\[\\]"), "", colnames(soilData))
 
 # Replace whitespace in between header parts for "_"
 colnames(soilData) <- str_replace_all(colnames(soilData), '([:alpha:]+)(\\s)([:upper:][:lower:]+)', '\\1_\\3')
 
 # clean header columns of unnecessary whitespaces
 colnames(soilData) <- gsub(("\\s"), "", colnames(soilData))

# delete first row
 soilData <- soilData[-c(1),]
```

**1.4 there is a typo in the column "Expozition"**. It should actually be "Exposition"
```{r, echo=TRUE}
names(soilData)[names(soilData) == "Expozition"] <- "Exposition"
```

**1.5 99999 and 99999,99 are the NA values**, but R does not know this, so all these values need to be set to NA. It should be noted that some of these NA values might be actual numbers, but some might also be of datatype character.  

```{r, echo=TRUE}
# uncertain if this ignores cases depending on the data type but the result looks like it worked
soilData[soilData=="99999" | soilData=="99999,99" ] <- NA
```

**1.6 two columns need to be calculated**: base_saturation_\[\%\] by dividing "Kationen" by "KAKpot" and SOM_\[\%\] by multiplying Corg with 1,72  

```{r, echo=TRUE}
# convert numbers to use "." instead of ","
soilData <- as.data.frame(sapply(soilData, gsub,pattern=",", replacement="."))
soilData <- soilData %>%
  mutate(`base_saturation_[%]` = round(as.numeric(`Kationen_[mmol/kg]`)/as.numeric(`KAKpot_[mmol/kg]`), digits=2)) %>%
  mutate(`SOM_[%]` = round(as.numeric(`Corg_[%]`)*1.72, digits=2))
```

**1.7 Column S+U+T is not always 100%**. I would like to see how often I got more, less and exactly 100%. Write a new column and put in "more", "less" and "exactly" depending on the value in the column S+U+T. Then, write a query that counts and prints the number of values (for example "10x less than 100%, 3x exactly 100% and 24x more than 100%") 

```{r, echo=TRUE}

#TODO

```

## Dataset after Cleanup
```{r, echo=FALSE}
knitr::kable(
  soilData[1:10, ], 
  caption = "Table: SoilData.csv"
)
```

# 2. Visualisation
**2.1 Stacked Barplot**  

I would like to have a stacked barplot that shows the values for Ca, Mg and Kationen of the **last** horizon of the profiles 55, 71, 102 and 109 - for every last horizon of these profiles in the data table one stacked barplot. Here is an example for such an barplot. I would also like to have the same colors as in the example plot (viridis).

```{r, echo=T}
data <- soilData |> 
  filter(Profil == 55 | Profil == 71 | Profil == 102 | Profil == 109) |> 
  group_by(Profil) |> 
  filter(Horizont == max(Horizont)) |> 
  select( Profil, Horizont, Horizontbezeichnung, `Ca_[mmol/kg]`, `Mg_[mmol/kg]`, `Kationen_[mmol/kg]`)

# change data to numeric values
data$`Ca_[mmol/kg]` <- as.numeric(data$`Ca_[mmol/kg]`)
data$`Mg_[mmol/kg]` <- as.numeric(data$`Mg_[mmol/kg]`)
data$`Kationen_[mmol/kg]` <- as.numeric(data$`Kationen_[mmol/kg]`)

# create long table for plotting
dataPlot <- pivot_longer(data, cols = c(`Ca_[mmol/kg]`, `Mg_[mmol/kg]`, `Kationen_[mmol/kg]`), names_to = "group", values_to = "value")

ggplot( dataPlot, aes(x=Profil, y= value, fill=group)) + 
  geom_bar(position = "stack", stat = "identity") +
#  scale_colour_continuous(type=scale_fill_viridis_b()) +
  # TODO: color scale is not used in plot
  scale_color_viridis_d(option="viridis", aesthetics = "fill", labels=c("CA", "Mg", "Kationen"))+
  ggtitle("Last horizon of the profiles 55, 71, 102 and 109") +
  theme_minimal() +
  labs(fill = "") +
  ylab("mmol/kg") +
  xlab("Profilnummer")
  
```


**2.2 Piechart of grain sizes**  
Then I would like to have 3 piecharts of the "Ah" horizons of the profiles 82, 111 and 134 next to each other where I can see the portion of S (Sand), U (silt) and T (clay).
  
  **Die Ausgabe mehrerer Plots nebeneinander funktioniert noch nicht**
```{r, echo=T}
toDraw <- c(82, 111, 134)

piechart <- function(profil){
  data <- soilData |> 
  filter(Profil == profil) |> 
  filter( str_detect("Ah", Horizontbezeichnung) )
  # select only relevant columns
  data <-  data |> 
  select( Profil, Horizont, Horizontbezeichnung, `S_[%]`, `U_[%]`, `T_[%]`)

  # create long table for plotting
  dataPlot <- pivot_longer(data, cols = c(`S_[%]`, `U_[%]`, `T_[%]`), names_to = "group", values_to = "value")

 # Compute the position of labels
  dataPlot <- dataPlot  |>  
   arrange(desc(group)) |> 
   mutate(prop = as.numeric(value) / sum(as.numeric(value)) *100) |> 
   mutate(ypos = cumsum(prop)- 0.5*prop )

  return (
    ggplot(dataPlot, aes(x="", y = c(group),  fill=group)) +
    geom_bar(stat="identity", width=1) +
    coord_polar("y", start=0) +
    #geom_text(aes(y = ypos, label = group), color = "white", size=6) +
    scale_color_viridis_d(option="viridis", aesthetics = "fill",
                          labels=c("sand", "silt", "clay")) +
    theme_void() +
    theme(plot.caption = element_text(hjust = 0.5, face = "bold")  ) +
    labs(fill="", 
         caption = str_c("Profil ", profil))
  )
}


# TODO: Ausgabe nebeneinander
 par(mfrow = c(1, 2))
piechart(82)
piechart(111)

 
 for(p in toDraw){
   piechart(p)
 }
```
