---
output: md_document
---

## Task 1 - Get in Touch with the Material

### 1) Take a look at the dataset in R

-   How many oberservations do we have?

```{r}
library(readxl)
data <- read_excel("data_smaschkon_2023-02-11_09-11.xlsx", na ="-9")
dim(data)
```

Answer: There are **101** observations in the data set

### 2) Examine the table

-   What is the question behind the items "SC01_14", "TS01_01"?

SC01_14: Ich kann Versuchungen gut wiederstehen.

TS01_01: Wie häufig trägst Du dein Smartphone bei Dir?

### 3) Examine the table

-   What are the possible values for the item "PS03_01", "SC01_14", "TS01_01"? PS03_01: 1;2;3;4;-9 SC01_14: 1;2;3;4;5;-9 TS01_01: 1;7;-9

-   What does "-9" mean? And how should "-9" be presented in R?

That the question was not answered. In R, -9 should be represented as NA

### 4) Back to R

-   Check with R for the observed values for "PS03_01," "SC01_14," "TS01_17" (use, for example, the table() function).

```{r}
variable_names <- c ("PS03_01", "SC01_14", "TS01_17")
  
tables <- lapply(variable_names, function(variable) {
  table(data[[variable]])
})
names(tables) <- variable_names
print(tables)

```

(In "TS01_17" there are **two NAs** as the numbers does not add up to 101. However Number of NAs are not displayed directly)

**or** by using this new-learned alternative (hopefully pleasing the R-Gods)

```{r}
library(tidyverse)
data |>  select("PS03_01", "SC01_14", "TS01_17") |> 
  map(table) |> print()
```

## Task 2 - Building Scales

### 1) Some preparation is needed: Item "SC01_09" must be recoded.

-   Examine the question behind item "SC01_09" again. What is unique about this question, and why should this item be "recoded"?
-   Find a way to recode the item in the dataset.

This question should also be asked in "umgepolt". All other questions for SC are asked in a way that a higher number translates to more self-control. However, when it comes to "SC01_09" a higher number refers to less Self Control. Therefore, I would suggest to recode this item that 1 translates to 5, 2 to 4, 3 stays the same, 4 to 2 and 5 to 1.

```{r}
data$SC01_09 <- 6- data$SC01_09
```

### 2) **Find a way to build the scales.** Different methods can be used. As mentioned earlier, we want row-wise means over the items that belong to the specific scale. (You can use either `Base R` or

```         
`tidyverse`.)
```

Create the Scale for Sleep Quality

```{r}
data$Sleep_Quality <- data$PS03_01
```

Scale for Self-Control

```{r}
library(tidyverse)
#SelfControlVariable <- c("SC01_14", "SC01_02", "SC01_03", "SC01_04", "SC01_05",
#    "SC01_06", "SC01_07", "SC01_08", "SC01_09", "SC01_10", "SC01_11",
#    "SC01_12", "SC01_13")
#data$Self_Control <- rowMeans(data[SelfControlVariable])

##Alternative##
SelfControlVariable <- data |> select(starts_with("SC01_"))
data$Self_Control <- rowMeans(SelfControlVariable)
```

Scale for Smartphone Usage

```{r}
#SmartphoneUsageVariable <- c("TS01_01", "TS01_02" , "TS01_03", "TS01_04", "TS01_05", "TS01_06", #"TS01_07", "TS01_08", "TS01_09", "TS01_10", "TS01_11", "TS01_12", "TS01_13", "TS01_14", "TS01_15", #"TS01_16", "TS01_17", "TS01_18", "TS01_19")
#data$Smartphone_Usage <- rowMeans(data[SmartphoneUsageVariable])

##Alternative##
SmartphoneUsageVariable <- data |> select(starts_with("TS01_"))
data$Smartphone_Usage <- rowMeans(SmartphoneUsageVariable)
```

## Task 3 - Plot Something Beautiful for Me

Creating the Plot

```{r}

color_names <- c("1" = "red", "2" = "blue", "3" = "green", "4" = "yellow")
color_labels <- c("1" = "Very Good", "2" = "Good", "3" = "Bad", "4" = "Very Bad")
library(ggplot2)
ggplot(data, aes(x = Self_Control, y = Smartphone_Usage, color = factor(PS03_01))) +
  geom_point() +
   scale_color_manual(values = color_names, breaks = names(color_names), labels = color_labels) +
 geom_smooth(method = "lm", se = FALSE, aes(group = PS03_01)) +
  theme_dark()+
  labs(
    title = "Correlation between Self-Control and Smartphone Usage",
    x = "Self-Control",
    y = "Smartphone Usage",
    fill = "PS03_01"
  ) +
  guides(color = guide_legend(title = "Sleep Quality"))
 

```

## I still need to work on some details, sry for that
