---
title: "Tennis Project"
author: 
output:
  md_document
---

**Getting Top50 players: **

**Step 1:** *Pick up only useful columns for the task, which are "winner_name", "winner_rank", "loser_rank" "tourney_date", "w_1stIn", "w_1stWon", "l_1stIn", "l_1stWon", "surface".*

**Step 2:** *Select only top 50 players in 2024. One player could have perhaps multiple ranks, so i will select the last rank by "tourney_date" of a player, no matter he won or lost, then we will get players, whose rank eventually in top 50 in 2024.*

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(readr)
library(dplyr)
library(ggplot2)

data <- read_csv("atp_matches_qual_chall_2024.csv", show_col_types = FALSE)
  
top50 <-  
  data %>% select(name = winner_name, rank = winner_rank, winner_name, winner_rank, loser_rank, tourney_date, w_1stIn, w_1stWon, l_1stIn, l_1stWon, surface) %>%
  bind_rows(data %>% #creating a single dataset that includes all players (as either winners or losers) along with relevant match information. This allows for analyzing the players regardless of their match outcomes.
      select(name = loser_name, rank = loser_rank, winner_name, winner_rank, loser_rank, tourney_date, w_1stIn, w_1stWon, l_1stIn, l_1stWon, surface)) %>%
  group_by(name) %>%
  arrange(desc(tourney_date)) %>% 
  filter(rank <= 50)  #choosing players within top50 based on rank

knitr::kable(head(top50, 220), format = "markdown")
```




**Visualization of the players by their final rank in 2024: **

**Step 3:** *And then select all matches of these players, no matter if they were alawys top 50 in these matches. Use this formula w_1stWon/w_1stIn to calculate their first won point of these matches as a winner or loser.*

**Step 4:** *Categorize these matches into two groups: hard and clay, so we get the rate attribute on two surface. Then calculate the average first won point rate on two surface.*

**Step 5:** *Rank players by their final rank in 2024 from high to low to visualize the rate distribution.* 


```{r, message=FALSE, warning=FALSE, echo=FALSE}
data %>%  filter(winner_name %in% top50$name | loser_name %in% top50$name) %>%
  mutate(winner_first_point = w_1stWon / w_1stIn,  
    loser_first_point = l_1stWon / l_1stIn) %>% 
    group_by(surface) %>%  #Step4 starts here 
  mutate(avg_winner_first_point = mean(winner_first_point, na.rm = TRUE)) %>% 
 ggplot(aes(x = reorder(winner_name, winner_rank), 
                 y = winner_first_point, 
                 color = surface, 
                 shape = surface, 
                 fill = surface)) +  
  geom_point(size = 5, stroke = 1.2, color="black") +  
  scale_shape_manual(values = c(21, 24)) +  
  scale_fill_manual(values = c("darkgreen", "darkorange")) +  
  labs(title = "First Point Win Rate of Top 50 Players", 
       x = "Player Name", 
       y = "First Point Win Rate") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


