---
title: "Tabea's Project"
author: "Laurens Mauz"
date: "2023-06-21"
output: md_document
---

```{r setup, include=FALSE}
library(tidyverse)

data <- read.csv("Data_Project1.csv", sep=";")

```

I first tried to make a simple scatterplot for the total mean radius of the three measurements of each Taxa.

```{r ex1}
data %>%
  dplyr::filter(Radius_number == 1) %>%
  dplyr::filter(Taxa == "Populus" | Taxa == "Indet" | Taxa == "Salix") %>%
  ggplot(aes(x=Taxa, y=Radius_tot_mean, color=Radius_tot_mean)) +
    geom_point() +
    geom_jitter() +
    scale_color_gradient(low="blue", high="red") +
    labs(title="Mean Radius of all Measurements",
          x="Taxa",
          y="Total mean Radius",
         color="Mean total Radius") +
    theme_bw() + 
    theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
  


```
I remembered that Martin advised Nico to make Violin plots, so I did as well. I think with violin plots the distribution of the data is a little bit easier to see.

```{r ex1 violin}

data %>%
  dplyr::filter(Radius_number == 1) %>%
  dplyr::filter(Taxa == "Populus" | Taxa == "Indet") %>%
  ggplot(aes(x=Taxa, y=Radius_tot_mean, color=Radius_tot_mean)) +
    geom_violin() +
    geom_point() +
    geom_jitter() +
    scale_color_gradient(low="blue", high="red") +
    labs(title="Mean Radius of all Measurements",
          x="Taxa",
          y="Total mean Radius") +
    theme_bw() + 
    theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
    
    

```

Here I make a Barplot for the number of tree rings. I was not sure why there were rows in the data with the same ID but different values for Standard deviation etc.

```{r ex2}
data %>%
  select(Radius_number, Taxa) %>%
  dplyr::filter(Taxa == "Populus" | Taxa == "Indet" | Taxa == "Salix") %>%
  drop_na() %>%
  group_by(Radius_number) %>%
  count() %>%
  mutate(Radius_number = as.factor(Radius_number)) %>%
  ggplot(aes(x=Radius_number, y=n)) +
    geom_bar(stat="identity") +
    labs(title="Number of Tree-rings",
         x="Number of Rings",
         y="Amount of Trees") +
    theme_bw() + 
    theme(panel.border = element_blank(), panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))





```
I made the scatterfplot of mean radius against the standard deviation of the mean radius and added a linear approximation of the data for INdet and Populus.

```{r ex3}
data %>%
  select(Radius_tot_mean, Radius_tot_sd, Radius_number, Taxa) %>%
  dplyr::filter(Radius_number == 1) %>%
  dplyr::filter(Taxa == "Populus" | Taxa == "Indet") %>%
  mutate(Radius_tot_sd = as.numeric(Radius_tot_sd)) %>% 
  drop_na() %>%
  ggplot(aes(x=Radius_tot_mean, y=Radius_tot_sd, color=Taxa)) +
    geom_point() +
    stat_smooth(method="lm") +
    labs(title="Correlation between Mean Radius and Standard Deviation of Mean Radius",
         x="Mean Radius of all Measurements",
         y="Standard Deviation of Radius") +
         theme_bw() + 
         theme(panel.border = element_blank(), panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
```
I wanted to see how good the linear regression fits the data. You can see that R squared for Populus is ca. 0,7 and for Indet it is ca. 0,35. The linear model seems to fit the Populs Taxa a lot better

```{r linear model}
lm_data_pop <- data %>%
  select(Radius_tot_mean, Radius_tot_sd, Radius_number, Taxa) %>%
  dplyr::filter(Radius_number == 1) %>%
  dplyr::filter(Taxa == "Populus") %>%
  mutate(Radius_tot_sd = as.numeric(Radius_tot_sd)) %>% 
  drop_na() 

lm_data_indet <- data %>%
  select(Radius_tot_mean, Radius_tot_sd, Radius_number, Taxa) %>%
  dplyr::filter(Radius_number == 1) %>%
  dplyr::filter(Taxa == "Indet") %>%
  mutate(Radius_tot_sd = as.numeric(Radius_tot_sd)) %>% 
  drop_na() 
  
lm_std_radius_indet <- lm(Radius_tot_mean~Radius_tot_sd, data = lm_data_indet) #Create the linear regression
summary(lm_std_radius_indet)

lm_std_radius_pop <- lm(Radius_tot_mean~Radius_tot_sd, data = lm_data_pop)

summary(lm_std_radius_pop)




```


