---
title: "factordata"
author: "Victor Asua"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
#packages
library(tidyverse)
```

# 1. creating factors

Imagine data containing four of the 12 possible months.`x = c("Dec", "Apr", "Jam", "Mar")`, Sort() does not sort in a useful way `sort(x)`. levels() function solves this problem
```{r cars}
#data
x1 <- c("Dec", "Apr", "Jan", "Mar")

# arrangement levels of months
month_levels <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun",  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

#creating a factor with month levels

y1 <- factor(x1, levels = month_levels)

# now when sort, it follows the factor levels
sort(y1)

```

## Factor Data manipulation
Using gss_cat data for practicing

```{r modifactor data}
# read data from r source
data("gss_cat")
#glipse into the data
head(gss_cat)

```

## Modifying factor data
changing order of factor levels in visualization. Imagine exploring numbers of hrs on tv vs religion

```{r, vis1}
#data summary
relig_summary <- gss_cat %>%
  group_by(relig) %>%
  summarise(
    #age=mean(age, na.rm=TRUE),
    tvhours=mean(tvhours, na.rm=TRUE),
    n=n()
  )

#plotting
ggplot(relig_summary, aes(tvhours, relig))+
  geom_point()
```
No overall pattern, figure difficult to interpret. reordering the levels of relig using fct_reorder(). it take three arguments:
-f, factor whose levels need to be modified
-x, numeric vector for reordering
-optional, fun,

```{r}
ggplot(relig_summary, aes(tvhours, fct_reorder(relig, tvhours))) + geom_point()
```


## Exploring the distribution of rincome(reportedincome). Improving the default bar appearance
```{r}

#default plot
gss_cat %>% ggplot(aes(x=rincome)) + geom_bar()

```

Improved plot
```{r}
#labels are difficult to read, will change the angle of x-axis to 45 degrees

gss_cat %>% ggplot(aes(x=rincome)) + geom_bar() +#default
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +# adjust xaxis labels
  coord_flip()
```

## Further improve the plot by
- removing the “Not applicable” responses,
- renaming “Lt $1000” to “Less than $1000”,
- using color to distinguish non-response categories (“Refused”, 
- “Don’t know”, and “No answer”) from income levels (“Lt $1000”, …),
- adding meaningful y- and x-axis titles, and
- formatting the counts axis labels to use commas.
```{r}

gss_cat %>%
  filter(!rincome %in% c("Not applicable")) %>% #remove "Not applicable"
  mutate(rincome = fct_recode(rincome, "Less than $1000" = "LT $1000")) %>%# change LT 1000 to Less than 1000
  mutate(rincome_na = rincome %in% c("Refused", "Don't know", "No answer")) %>% ggplot(aes(x=rincome, fill=rincome_na)) +
  geom_bar()+
  coord_flip()+
  scale_y_continuous("Number of Respondents", labels = scales::comma)+
  scale_x_discrete("Respondent's Income")+
  scale_fill_manual(values = c("FALSE" = "black", "TRUE" = "gray"))+
  theme(legend.position = "None")

```

