---
title: "Solution to Zoés Task by Dana"
output: html_notebook
---

Hello! 
I am attempting to work on Zoes Task. I've had some struggles to properly
read the data in, but now I'm a huge step forward.


```{r, echo = FALSE}
library(ggplot2)
library(tidyverse)
library(RPMG)
```


```{r}

dat0 <- read_csv2("https://dr-eberle-zentrum.github.io/Advanced-data-processing-with-R/Projects/zoebuerger/Data_mood_stress.csv") 
  
dat0 |> 
rownames_to_column() |> group_by(timepoint, questionnaire)


# mean of negative PANAS items
# mean_negPANAS <- mean(dat0$questionnaire == "PANAS" & item = c("bekuemmert", "beunruhigt", "schuldig", "erschreckt", "feindselig", "reizbar", "beschaemt", "nervoes", "aengstlich", "furchtsam"))


# mean of positive PANAS items
# mean_posPANAS <-  mean(dat0$questionnaire == "PANAS" & item = c("entschlossen", "begeistert", "aufmerksam", "stark", "interessiert", "aktiv", "wachsam", "angeregt", "stolz", "schwungvoll")),

# recoding positive STAI items
# mutate(recode((factor(questionnaire == "STAI"),
#                 "1" = "5",
#                 "2" = "4"
#                 "3" = "3"
#                 "4" = "2"
#                 "5" = "1")))

# sum posSTAI
# sum_postSTAI <- sum(dat0$questionnaire == "STAI" & dat0$item == c("ruhig", "geborgen", "geloest", "ausgeruht", "wohl", "selbstsicher", "entspannt", "zufrieden", "froh", "vergnuegt"))

# sum of negative STAI items
# sum_negSTAI <- sum(dat0$questionnaire == "STAI" & dat0$item == c("angespannt", "bekuemmert", "aufgeregt", "schiefgehen", "beunruhigt", "nervoes", "zappelig", "erkrampft", "besorgt", "ueberreizt"))


# 
# Attempt for ggplot
# ggplot(dat0, aes(x=timepoint, y= mean_posPANAS, fill=birthcontrol)) + 
#    geom_boxplot()
# ggplot(dat0, aes(x=timepoint, y= mean_neg_PANAS, fill=birthcontrol)) + 
#    geom_boxplot()
# ...... :(


```

I notice that my boxplots might not work this way.
Unfortunately, since I did not manage to tidy the data I could not execute and try the code but I wrote
down the steps that would follow and what I thought of.
If someone is able to give me tips or hints regarding the tidying of the data, i would love to try and visualize the data again and finish it! You can read on the details on my struggles regarding the tidying of the data in the issue i opened a while ago.

Thank you this far! I wish i could have presented more than that but I think I need the tidied data
for that in order to find out whether the code i thought of to continue the tasks work out or not.
