---
title: "Exercises"
author: "Adkhamjon"
date: "2025-04-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
```
## exercises 27
### How Diamond Sizes Vary by Cut, Color, and Clarity

27.4.7 Exercises 1. Add a section that explores how diamond sizes vary by cut, colour, and clarity. Assume you’re writing a report for someone who doesn’t know R, and instead of setting echo = FALSE on each chunk, set a global option.

```{r}
library(ggplot2)
library(dplyr)

```

```{r}
diamonds %>% 
  group_by(cut, color, clarity) %>%
  summarise(mean_carat = mean(carat)) %>%
  ungroup() %>%
  ggplot(aes(x = cut, y = mean_carat, fill = color)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~clarity) +
  labs(title = "Mean Carat Size by Cut, Color, and Clarity",
       x = "Cut",
       y = "Mean Carat Size") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set3")
```





### Largest 20 Diamonds


2.Add a section that describes the largest 20 diamonds, including a table that displays their most important attributes.

```{r}
diamonds %>% 
  arrange(desc(carat)) %>%
  head(20) %>%
  select(cut, color, clarity, carat, price) %>%
  knitr::kable(caption = "Largest 20 Diamonds by Carat Size")
```


### Diamonds Larger Than 2.5 Carats

Modify diamonds-sizes.Rmd to use comma() to produce nicely formatted output. Also include the percentage of diamonds that are larger than 2.5 carats.


```{r}
library(dplyr)
large_diamonds <- diamonds %>%
  filter(carat > 2.5)

percentage_large <- nrow(large_diamonds) / nrow(diamonds) * 100
percentage_large
```
```{r}
comma <- function(x) format(x, digits = 2, big.mark = ",")
```


Out of `r comma(nrow(diamonds))` diamonds, there are `r comma(nrow(large_diamonds))` diamonds larger than 2.5 carats. 

That is approximately `r comma(percentage_large)` of all diamonds.



Set up a network of chunks where d depends on c and b, and both b and c depend on a. Have each chunk print lubridate::now(), set cache = TRUE, then verify your understanding of caching.


```{r a,  cache = TRUE}
library(lubridate)
large_diamonds <- diamonds %>%
  filter(carat > 2.5)

```


```{r b, cache = TRUE, depends = "a"}
large <- nrow(large_diamonds) / nrow(diamonds)* 10000000
large
```

```{r c,  cache = TRUE, depends = "a"}
percentage_large <- nrow(large_diamonds) / nrow(diamonds) * 100
```


```{r d,  cache = TRUE, depends = c("b", "c")}
comma <- function(x) format(x, digits = 2, big.mark = ",")
comma(large)
comma(percentage_large)
```


## exercises 28


Create one plot on the fuel economy data with customised title, subtitle, 
caption, x, y, and colour labels.

```{r}
mpg %>%
  ggplot(aes(x = displ, y = cty)) +
  geom_point(aes(colour = class)) +
  geom_smooth(se = FALSE) +
  labs(title = "Fuel Economy by Engine Size",
       subtitle = "A scatter plot of highway miles per gallon vs. engine displacement",
       x = "Engine Displacement (L)",
       y = "Highway Miles per Gallon",
       colour = "Car Class",
       caption = "Data source: mpg dataset from ggplot2") +
  theme_minimal()
```


The geom_smooth() is somewhat misleading because the hwy for large engines is skewed upwards due to the inclusion of lightweight sports cars with big engines. Use your modelling tools to fit and display a better model.


```{r}
mpg %>%
  ggplot(aes(x = displ, y = cty)) +
  geom_point(aes(colour = class)) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Fuel Economy by Engine Size",
       subtitle = "A scatter plot of highway miles per gallon vs. engine displacement",
       x = "Engine Displacement (L)",
       y = "Highway Miles per Gallon",
       colour = "Car Class",
       caption = "Data source: mpg dataset from ggplot2") +
  theme_minimal()
```


Take an exploratory graphic that you’ve created in the last month, and add informative titles to make it easier for others to understand.


how to add new column ?
```{r}
mt <- rownames(mtcars)
mtcars <- cbind(mtcars, mt)
```


```{r}
ggplot(mtcars, aes(x = reorder(mt, mpg), y = mpg)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(title = "Fuel Economy by Car Model",
       subtitle = "A bar plot of miles per gallon for different car models",
       x = "Car Model",
       y = "Miles per Gallon",
       caption = "Data source: mtcars dataset") +
  theme_minimal()
```
## exercises28.3
Use geom_text() with infinite positions to place text at the four corners of the plot.

```{r}
label <- tibble(
  displ = Inf,
  hwy = Inf,
  label = "Increasing engine size is \nrelated to decreasing fuel economy."
)

plot <- ggplot(mpg, aes(displ, hwy)) +
  geom_point() 
plot +
  geom_text(aes(label = label), data = label, vjust = "top", hjust = "right")
```


```{r}
label <- tibble(
  displ = -Inf,
  hwy = -Inf,
  label = "Increasing engine size is \nrelated to decreasing fuel economy."
)

plot <- ggplot(mpg, aes(displ, hwy)) +
  geom_point() 
plot +
  geom_text(aes(label = label), data = label, vjust = "bottom", hjust = "right")
```


```{r}
label <- tibble(
  displ = -Inf,
  hwy = Inf,
  label = "Increasing engine size is \nrelated to decreasing fuel economy."
)

plot <- ggplot(mpg, aes(displ, hwy)) +
  geom_point() 
plot +
  geom_text(aes(label = label), data = label, vjust = "top", hjust = "left")
```


```{r}
label <- tibble(
  displ = Inf,
  hwy = -Inf,
  label = "Increasing engine size is \nrelated to decreasing fuel economy."
)

plot <- ggplot(mpg, aes(displ, hwy)) +
  geom_point() 
plot +
  geom_text(aes(label = label), data = label, vjust = "bottom", hjust = "right")
```



Read the documentation for annotate(). How can you use it to add a text label to a plot without having to create a tibble?

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  annotate("text", x = Inf, y = Inf,
           label = "Increasing engine size is \nrelated to decreasing fuel economy.",
           vjust = "top", hjust = "right")
```


How do labels with geom_text() interact with faceting? How can you add a label to a single facet? How can you put a different label in each facet? (Hint: think about the underlying data.)

If  label's data matches a facet, the label shows up in that facet.

If you want labels in specific facets, your label data must include the same variables used for faceting.


```{r}
label_data <- tibble(
  class = "suv",      # must match facet variable exactly
  displ = 4,
  hwy = 20,
  label = "SUV Special!"
)

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  facet_wrap(~class) +
  geom_text(data = label_data, aes(x = displ, y = hwy, label = label))

```



```{r}
label_data <- tibble(
  class = c("suv", "compact", "pickup"),
  displ = c(4, 2, 5),
  hwy = c(20, 35, 17),
  label = c("SUV Label", "Compact Label", "Pickup Label")
)

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  facet_wrap(~class) +
  geom_text(data = label_data, aes(x = displ, y = hwy, label = label), color = "blue")

```

What arguments to geom_label() control the appearance of the background box?



| Argument         | What it controls                        |
|------------------|------------------------------------------|
| `fill`           | Color **inside** the box (background color) |
| `color` or `colour` | Color of the **text** (NOT the box)         |
| `label.size`     | Thickness of the **box border** (outline)  |
| `label.r`        | How **rounded** the box corners are        |
| `label.padding`  | Space **between text and box edges**       |

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_label(
    aes(label = class),
    fill = "lightblue",        # Box background color
    color = "black",            # Text color
    label.size = 0.5,           # Border thickness
    label.r = unit(0.25, "lines"), # Round the corners
    label.padding = unit(0.5, "lines") # Padding inside box
  )

```




The question is
What are the four main arguments of arrow()?
How do those arguments control how the arrow looks?
Can you show examples of plots using different arrow settings?

| Argument       | What it controls                          |
|----------------|-------------------------------------------|
| `angle`        | The **angle (sharpness)** of the arrowhead |
| `length`       | The **length** of the arrowhead            |
| `ends`         | Where the arrows are drawn (start, end, both) |
| `type`         | Shape of the arrowhead (open or closed)    |



```{r}
library(ggplot2)
library(grid)  # for arrow()

ggplot() +
  geom_segment(aes(x = 1, y = 1, xend = 2, yend = 2),
               arrow = arrow(angle = 30, length = unit(0.25, "inches"),
                             ends = "last", type = "open")) +
  xlim(0, 3) + ylim(0, 3)

```


D:\All R projects\R2_Project\R2dataProject\Data-projects-with-R-and-GitHub



































