---
title: "hello_world"
author: "Adkhamjon"
date: "2025-04-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

| Hello | World |
|-------|-------|
| line1 | line2 |

```{r hello }
library(ggplot2)
```

```{r headers, evaluate=FALSE}
# I can write many examples codes but this  won't be evaluated
# just show don't run

```

```{r inlude=FALSE}
# this code does not show up in the final document
# run but dont show
# ggplot(mtcars, aes(x=mpg, y=disp)) +
#   geom_line()
```

```{r echo=FALSE}
# just show the result not the code 
# ggplot(mtcars, aes(x=mpg, y=disp)) +
#   geom_density_2d()
```

```{r message=FALSE, warning=FALSE}
# prevents messages or warnings from appearing in the finished file.
```

```{r error=TRUE}
# causes the render to continue even if code returns an error
ggplot(mtcars, aes(x=mpg)) +
  geom_bar(stat="identity")

```

```{r}
#mtcars[1:5,]
```

```{r}
#If you prefer that data be displayed with additional formatting you can use the knitr::kable function.

#knitr::kable(mtcars[1:5,],
#             caption = "A knitr table")
```

| Option | What it does |
|-------------------|-----------------------------------------------------|
| `cache = TRUE` | Saves chunk results to speed up future knitting |
| `dependson` | Makes sure a cached chunk reruns if another chunk it depends on changes |
| `cache.extra` | Triggers cache reset if an external file or value (e.g., CSV) changes |
| `clean_cache()` | Clears all caches manually |

you can set global options that apply to all chunks in the document using: knitr::opts_chunk\$set()

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

```

| Setting | What it controls | Use it when... |
|----------------------|---------------------------|-----------------------|
| `echo = FALSE` | Hide code but show output | Making clean reports for non-coders |
| `message = FALSE` / `warning = FALSE` | Hide messages/warnings from packages | You want cleaner output (but debug later) |
| `comment = "#>"` | Customize the symbol before printed output | For stylish code examples |
| `collapse = TRUE` | Keeps code and output close together | Makes examples cleaner in tutorials |


### 28  graphics for communication
subtitle adds additional detail in a smaller font beneath the title.

caption adds text at the bottom right of the plot, often used to describe the source of the data.

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  labs(
    title = "Fuel efficiency generally decreases with engine size",
    subtitle = "Two seaters (sports cars) are an exception because of their light weight",
    caption = "Data from fueleconomy.gov"
  )
```

----

It’s possible to use mathematical equations instead of text strings. Just switch "" out for quote() and read about the available options in ?plotmath:

```{r}
df <- tibble(
  x = runif(10),
  y = runif(10)
)
ggplot(df, aes(x, y)) +
  geom_point() +
  labs(
    x = quote(sum(x[i] ^ 2, i == 1, n)),
    y = quote(alpha + beta + frac(delta, theta))
  )
```




In addition to labelling major components of your plot, it’s often useful to label individual observations or groups of observations. The first tool you have at your disposal is geom_text(). geom_text() is similar to geom_point(), but it has an additional aesthetic: label. This makes it possible to add textual labels to your plots.

```{r}
best_in_class <- mpg %>%
  group_by(class) %>%
  filter(row_number(desc(hwy)) == 1)

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_text(aes(label = model), data = best_in_class)
```


We can make things a little better by switching to geom_label() which draws a rectangle behind the text. We also use the nudge_y parameter to move the labels slightly above the corresponding points


```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_label(aes(label = model), data = best_in_class, nudge_y = 2, alpha = 0.5)
```

we can use the ggrepel package by Kamil Slowikowski. This useful package will automatically adjust labels so that they don’t overlap:

```{r}
library(ggrepel)
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_point(size = 3, shape = 1, data = best_in_class) +
  geom_label_repel(aes(label = model), data = best_in_class)
```

----
You can sometimes use the same idea to replace the legend with labels placed directly on the plot. (theme(legend.position = "none") turns the legend off — we’ll talk about it more shortly.)


```{r}
class_avg <- mpg %>%
  group_by(class) %>%
  summarise(
    displ = median(displ),
    hwy = median(hwy)
  )

ggplot(mpg, aes(displ, hwy, colour = class)) +
  ggrepel::geom_label_repel(aes(label = class),
    data = class_avg,
    size = 6,
    label.size = 0,
    segment.color = NA
  ) +
  geom_point() +
  theme(legend.position = "none")
```
----
you might just want to add a single label to the plot, but you’ll still need to create a data frame. Often, you want the label in the corner of the plot, 
```{r}
label <- mpg %>%
  summarise(
    displ = max(displ),
    hwy = max(hwy),
    label = "Increasing engine size is \nrelated to decreasing fuel economy."
  )

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_text(aes(label = label), data = label, vjust = "top", hjust = "right")
```
----
If you want to place the text exactly on the borders of the plot, you can use +Inf and -Inf. Since we’re no longer computing the positions from mpg, we can use tibble() to create the data frame:

```{r}
label <- tibble(
  displ = Inf,
  hwy = Inf,
  label = "Increasing engine size is \nrelated to decreasing fuel economy."
)

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_text(aes(label = label), data = label, vjust = "top", hjust = "right")
```



Transiently, by calling rmarkdown::render() by hand:

rmarkdown::render("diamond-sizes.Rmd", output_format = "word_document")

----
29.5 Presentations
You can also use R Markdown to produce presentations. You get less visual control than with a tool like Keynote or PowerPoint, but automatically inserting the results of your R code into a presentation can save a huge amount of time. Presentations work by dividing your content into slides, with a new slide beginning at each first (#) or second (##) level header. You can also insert a horizontal rule (***) to create a new slide without a header.

R Markdown comes with three presentation formats built-in:

ioslides_presentation - HTML presentation with ioslides

slidy_presentation - HTML presentation with W3C Slidy

beamer_presentation - PDF presentation with LaTeX Beamer.

Two other popular formats are provided by packages:

revealjs::revealjs_presentation - HTML presentation with reveal.js. Requires the revealjs package.

rmdshower, https://github.com/MangoTheCat/rmdshower, provides a wrapper around the shower, https://github.com/shower/shower, presentation engine


