---
title: "hello_world"
author: "Adkhamjon"
date: "2025-04-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

| Hello | World |
|-------|-------|
| line1 | line2 |

```{r hello }
library(ggplot2)
```

```{r headers, evaluate=FALSE}
# I can write many examples codes but this  won't be evaluated
# just show don't run

```

```{r inlude=FALSE}
# this code does not show up in the final document
# run but dont show
# ggplot(mtcars, aes(x=mpg, y=disp)) +
#   geom_line()
```

```{r echo=FALSE}
# just show the result not the code 
# ggplot(mtcars, aes(x=mpg, y=disp)) +
#   geom_density_2d()
```

```{r message=FALSE, warning=FALSE}
# prevents messages or warnings from appearing in the finished file.
```

```{r error=TRUE}
# causes the render to continue even if code returns an error
ggplot(mtcars, aes(x=mpg)) +
  geom_bar(stat="identity")

```

```{r}
#mtcars[1:5,]
```

```{r}
#If you prefer that data be displayed with additional formatting you can use the knitr::kable function.

#knitr::kable(mtcars[1:5,],
#             caption = "A knitr table")
```

| Option | What it does |
|-------------------|-----------------------------------------------------|
| `cache = TRUE` | Saves chunk results to speed up future knitting |
| `dependson` | Makes sure a cached chunk reruns if another chunk it depends on changes |
| `cache.extra` | Triggers cache reset if an external file or value (e.g., CSV) changes |
| `clean_cache()` | Clears all caches manually |

you can set global options that apply to all chunks in the document using: knitr::opts_chunk\$set()

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

```

| Setting | What it controls | Use it when... |
|----------------------|---------------------------|-----------------------|
| `echo = FALSE` | Hide code but show output | Making clean reports for non-coders |
| `message = FALSE` / `warning = FALSE` | Hide messages/warnings from packages | You want cleaner output (but debug later) |
| `comment = "#>"` | Customize the symbol before printed output | For stylish code examples |
| `collapse = TRUE` | Keeps code and output close together | Makes examples cleaner in tutorials |
