---
title: "test.it"
author: "Adkhamjon"
date: "`r Sys.time()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#####################################
#' Create an RMD file which
#' - renders to Markdown output
#' - uses the current time within the YAML header
#' - contains a single code chunk that
#'   - loads tidyverse libraries
#'   - uses read_csv to read the file shown on the following website
#'     https://github.com/Dr-Eberle-Zentrum/Data-projects-with-R-and-GitHub/blob/main/Projects/martin-raden/Mammal-comeback-in-Europe-data-OWID-download.csv
#' - tune the code chunk such that
#'   - the code chunks name is ""
#'   - the code itself is shown, BUT
#'   - no loading information from the tidyverse library is shown
#'   - no loading details from the read_csv() call are shown
#'   consult https://yihui.org/knitr/options/ for details
#####################################


```{r test.it, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(readr)
data <- read_csv("https://github.com/Dr-Eberle-Zentrum/Data-projects-with-R-and-GitHub/blob/main/Projects/martin-raden/Mammal-comeback-in-Europe-data-OWID-download.csv")

head(data, 5)
```

```{r}
library(tidyverse)
?ToothGrowth
```


####################################################################
#' Let's practice our ggplot plotting skills. To this end, we use the
#' "ToothGrowth" data set directly available within R (check help).
#' We want to plot for the different dose levels (X) the respective
#' tooth length. The dose levels should be represented as follows:
#' - 0.5 = non-filled circle in a dark green
#' - 1 = orange filled triangles
#' - 2 = filled circles in a light bluedos
#' Use jittering to avoid overplotting of data points.
#' We don't want to have any background decorations etc. and useful
#' labels for axes and legend.
#' Ensure there is only ONE legend (showing both color and shape).
####################################################################


```{r}
ggplot(ToothGrowth, aes(x = dose, y = len)) +
  geom_jitter(aes(shape = factor(dose), color = factor(dose)), size = 3) +
  scale_shape_manual(values = c(0.5, 1, 2)) +
  scale_color_manual(values = c("darkgreen", "orange", "lightblue")) +
  labs(x = "Dose levels", y = "Tooth length") +
  theme_minimal() +
  theme(legend.position = "right")
```


