---
title: "Worldwide HIV cases (Jan's project)"
author: "Bernhard"
output: 
  html_notebook:
    theme: darkly
---

## First step: tidying up data.

Column `Location` gets renamed to `Country`, column `Period` gets renamed to `Year` and column `FactValueNumeric` will be renamed to `Cases`.

Note: I held the plots in a *mourning color* by purpose.

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(ggdark)
library(gghighlight)
data_raw <- read_csv("https://raw.githubusercontent.com/Dr-Eberle-Zentrum/Advanced-data-processing-with-R/main/Projects/jan-boethling/Data_Project_Jan.csv")
#data_raw <- read_csv("C:\\Users\\berbac\\Google Drive\\Uni\\Geographie\\R2 - Advanced R\\Project\\Advanced-data-processing-with-R\\Projects\\jan-boethling\\Data_Project_Jan.csv")

# renaming columns
data_raw |>
  rename(Country = Location, Year = Period, Cases = FactValueNumeric) |>
  select(Country, Year, Cases) |>
  mutate(
    #Country = as.factor(Country),
    #Year = as.factor(Year),
    Cases = replace_na(Cases, 0) 
    ) -> data

```

## Second step: cases by year

The worldwide HIV cases by year from 2000 to 2021 are shown in this plot. The year with the most cases is highlighted.

```{r echo=FALSE, message=FALSE, warning=FALSE}

data |>
  group_by(Year) |>
  summarise(Cases = sum(Cases)) |>
  ggplot(aes(x = as.factor(Year), y = Cases, fill = Year)) +
  geom_col() +
  guides(fill = "none") + # removes legend
  dark_theme_classic() +
  xlab("Year") +
  scale_y_continuous(labels = scales::comma) +
  ggtitle("Worldwide HIV cases by years with highlighted maximum (2000 - 2021)") +
  gghighlight(max(Cases), max_highlight = 1)

```

As we can see 2021 was the year with the most HIV cases during the time period of 2000 and 2021.

## Third step: cases by country

This chart shows the development of HIV cases from 2000 to 2021 of the 5 countries with the most cases over that time period.

```{r echo=FALSE, message=FALSE, warning=FALSE}

data |>
  group_by(Country) |>
  summarise(Cases = sum(Cases)) |>
  arrange(desc(Cases)) |>
  slice(1:5) -> top_5_countries

data |>
  filter(is.element(Country, top_5_countries$Country)) |> 
  ggplot(aes(x = Year, y = Cases, group = Country, color = Country, fill = Country)) + 
  geom_line() +
  scale_y_continuous(labels = scales::comma) +
  ggtitle("Development of HIV cases of the 5 countries with most cases\n(2000 - 2021)") +
  dark_theme_classic()

```

As the chart shows, in all the countries but South Africa the HIV cases are either decreasing or at a relative stable level. South Africa shows a steep increase in case numbers and hence must be the main cause for the rising cases worldwide that the first chart shows.

## Fourth step: 5 year interval *summary*

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

data$Rank = NA
for (i in 1:length(top_5_countries$Country)) {
    current_country <- top_5_countries$Country[i]
    data[data$Country == current_country,]$Rank = 
      as.integer(
      data[data$Country == current_country & data$Year == 2020,]$Cases - 
      data[data$Country == current_country & data$Year == 2000,]$Cases
      )
}

data |>
  filter(is.element(Country, top_5_countries$Country)) |>
  filter(is.element(Year, Year[Year %% 5 == 0])) |>
  mutate(Country = str_c(Country, " (", if_else(Rank > 0, "+", ""), scales::label_comma()(Rank), ")")) |>
  ggplot(aes(x = Year, y = Cases, fill = fct_reorder(factor(Country), Rank) )) +
  geom_col(position = "stack") +
  dark_theme_classic() +
  guides(fill = guide_legend(title = "Countries (case difference 2000 - 2020)")) +
  scale_y_continuous(labels = scales::comma) +
  ggtitle("HIV cases of the 5 countries with the highest \nHIV cases count worldwide in a 5 year interval (2000 - 2020)")

```

This chart gives an overview over the HIV cases of the five most affected countries worldwide. The cases are shown in a 5 year interval. This chart can be seen as a summary of the foregoing charts.

In addition, the stacked countriese are sorted by increase/decrease of the case numbers from 2000 to 2020 with the country nearest to the x-axis having the largest increase (South Africa) while the topmost (India) the lowest. As the foregoing chart shows India had a decrease in HIV infections during the analysed time period.

Hopefully, it's beautiful enough.
