---
title: "Bene-Klein"
output: md_document
---
```{r include = FALSE}
#Here packages
library(dplyr)
library(stringr)
library(ggplot2)
library(tidyr)
library(knitr)
setwd(
  # using RStudio library
  dirname(rstudioapi::getActiveDocumentContext()$path)
  # using additional library
  # this.path::this.dir()
)
```

# Data manipulation
Make the data set clean and tidy -> 1 column = one variable, 1 row = one observation, if there is no data -> NA, always use the same delimiter, watch out for commas that should not be read as a delimiter, watch for spaces and Quotation marks etc.

```{r echo=FALSE}
df<-read.csv('movies.csv')
df$ONE.LINE<-gsub("\n","",as.character(df$ONE.LINE))
df$STARS<-gsub(".*Stars:","",as.character(df$STARS))
df$STARS<-gsub("\n","",as.character(df$STARS))
df$STARS<-gsub(".*Star:","",as.character(df$STARS))
df$STARS<-gsub(".*Director:","",as.character(df$STARS))
df$STARS<-gsub(".*Directors:","",as.character(df$STARS))
df$GENRE<-gsub("\n","",as.character(df$GENRE))
df$GENRE <- lapply(df$GENRE, function(x) strsplit(x, ",\\s*")[[1]])
df$STARS <- lapply(df$STARS, function(x) strsplit(x, ",\\s*")[[1]])
df$VOTES <- as.numeric(gsub(",", "", df$VOTES))
df <- df %>%
  mutate(
    Start = as.numeric(str_extract(YEAR, "\\d{4}")),
    End = as.numeric(str_extract(YEAR, "(?<=-)\\d{4}")),
    End = ifelse(grepl("-\\)", YEAR), 2022, End),  # If the end is missing, set it to 2022
    End = ifelse(is.na(End), Start, End),  # If there's no end year, use start year
    Year_List = mapply(function(s, e) if (!is.na(s) & !is.na(e)) seq(s, e) else NA, Start, End, SIMPLIFY = FALSE)
    )
df[df==""]<-NA
final_table <- df %>% select(YEAR, Start, End, Year_List)
```
#Visualization

Plot the average imdb-score of each genre (group observation by genre) against the time -> which genre is the most liked by the judges, is there a trend over the years? -> Make a lineplot for each genre and plot them all in the same graphic. Give every genre a different color, make a legend. Sort the genres in the legend according to their current imdb-score (-> score for year 2022 -> use factors for that!)
```{r echo=FALSE}
#generate a simple dummy plot
data_expanded <- df %>%
  unnest(GENRE) %>%                 # Expand list-column into multiple rows
  mutate(GENRE = trimws(GENRE)) %>%  # Remove extra spaces (if any)
  mutate(GENRE = as.factor(GENRE)) 
# Create grouped bar plot
ggplot(data_expanded, aes(x = Start, y = RATING, fill = GENRE,color = GENRE))+
  geom_point(size = 1) + 
  labs(title = "Critic Feedback of media over the years",
       x = "Year",
       y = "IMDb Rating") +
  theme_minimal()
ggsave("critic_feedback.pdf")
#TODO: sort LEGEND
```
Do the same for “votes” (=audience opinion) and compare the results of the two graphics (discrepancy between “professional” and “laymen” taste).
```{r echo=FALSE}
#generate a simple dummy plot
data_expanded$VOTES <- as.numeric(data_expanded$VOTES)
# Create grouped bar plot
ggplot(data_expanded, aes(x = Start, y = VOTES, fill = GENRE,color = GENRE))+
  geom_point(size = 1) + 
  labs(title = "Audience Feedback of media over the years",
       x = "Year",
       y = "audience votes") +
  scale_y_continuous(breaks = NULL)+
    theme_minimal()
ggsave("audience_feedback.pdf")
```


```{r echo=FALSE}
#generate discrepancy data
df_sorted_desc <- df[order(df$VOTES, decreasing = TRUE), ]
max_votes<-df_sorted_desc$VOTES[1]
discr <- ifelse(is.na(df$RATING) | is.na(df$VOTES), NA, df$RATING - (df$VOTES*10 / max_votes))
# Create grouped bar plot
ggplot(df, aes(x = discr)) +
  geom_histogram(binwidth = 0.1, fill = "steelblue", color = "black", alpha = 0.7) +
  labs(title = "discrepancy of Ratings",
       x = "discrepancy of rating (<-audence only, critic only ->)",
       y = "Count") +
  theme_minimal()
ggsave("discrepancy_of_feedback.pdf")
```



```{r echo=FALSE}
data_expanded_stars <- df %>%
  unnest(STARS)
mean_RATINGS <- aggregate(RATING ~ STARS, data = data_expanded_stars, FUN = mean)
mean_RATINGS <- mean_RATINGS[order(mean_RATINGS$RATING, decreasing = TRUE), ]
mean_RATINGS <- mean_RATINGS %>%
  arrange(desc(RATING)) %>%  # Sorting by rating (optional)
  head(50)  # Take the first 50 stars

# Plot the data
ggplot(mean_RATINGS, aes(x = reorder(STARS, RATING), y = RATING)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +  # Flip to make names readable
  labs(title = "Top 50 Stars by Mean Rating",
       x = "Star Name",
       y = "Mean Rating") +
  theme_minimal()
ggsave("Top_50_stars_rating.pdf")
#temp plot
```
