---
 title: "Solution for Arlene"
 author: "Martin"
 output: md_document
---

```{r setup, echo=FALSE, show=FALSE}
#load without warnings
suppressMessages(library(tidyverse))

# set working directory to source file location
library(this.path)
setwd(this.path::here())
```

In the end I loaded the your prepared data ..

```{r import}
load("data_PISA.Rdata")

dataPerId <-
  data |>
  # slice_head(n=10) |>
  # select relevant columns
  select( CNT, ST005Q01JA, ST007Q01JA, matches("ST327Q0[1-8]JA") ) |>
  # number observations
  mutate( id = 1:n()) |>
  # best education of parents
  rowwise() |>
  mutate( minEduParents = ifelse(is.na(ST005Q01JA)&is.na(ST007Q01JA),
                                 NA,
                                 min(ST005Q01JA, ST007Q01JA, na.rm=TRUE)) ) |>
  ungroup() |>
  select(-c(ST005Q01JA, ST007Q01JA)) |>
  
  # drop missing data
  filter( minEduParents %in% 1:8) |> 
  # grouped parental education 
  mutate( eduParentGroup =
            case_when(
              minEduParents %in% 1:2 ~ "1-2",
              minEduParents %in% 3:4 ~ "3-4",
              minEduParents %in% 5:6 ~ "5-6",
              minEduParents %in% 7:8 ~ "7-8"
            )
          ) |> 

  # education goal
  # reformat goal columns into rows
  pivot_longer( cols = matches("ST327Q0[1-8]JA"), 
                names_to = "goal", 
                values_to = "goalReached" ) |>
  # process goal labels and check if reached
  mutate(goalReached = goalReached == 2,
         goal = str_remove(goal, "ST327Q0") |> str_remove("JA") |> as.numeric()) |>
  # data to keep
  group_by(id,CNT,minEduParents) |>
  # get last goal assumed reachable
  summarize( educationGoal = ifelse(sum(goalReached,na.rm=T)==0,
                          # no goal reachable
                          0,
                          # highest goal that was assumed reachable
                          max(goal*as.integer(goalReached), na.rm=T)),
           .groups ="drop") |>
  # filter for people with goals
  filter(educationGoal > 0) 
```

As you can see below, I used:

- the minimal parental education score,
- created groups "1-2" etc. for these, and 
- dropped all data where no parental education score was available.

Furthermore I 

- identified for each row the highest education goal assumed reachable.

With this I did a first visualization showing the distribution for each combination.

```{r first.glimpse}
dataPerId |>
  ggplot( aes(x=minEduParents,y=educationGoal) ) +
  geom_count() +
  geom_smooth(method = "lm") +
  facet_wrap(~CNT)
```
What comes to mind from the plot

- no parental education levels 2 and 4 in Germany
- nothing higher than 5 (or 3 in Japan)

So are the education levels comparable or only their relation?


Following your suggestion, I visualized mean values per level.
I also added error bars to visualize the standard deviation for each mean value.

```{r mean.viz}
dataPerId |>
  group_by(CNT, minEduParents) |>
  summarize( educationGoalMean = mean(educationGoal, na.rm=TRUE),
             educationGoalSd = sd(educationGoal, na.rm=TRUE),
             groupSize = n(),
             .groups="drop") |>
  ggplot(aes(x=minEduParents, y=educationGoalMean)) +
  # y ticks at 1:8
  scale_y_continuous(breaks=1:8, limits = c(1,10)) +
  # draw per country
  facet_wrap(~CNT) +
  # mean value of all countries
  geom_hline( yintercept = mean(dataPerId$educationGoal),
              linetype="dashed", col="darkgray", linewidth=2)+
  # mean values per country
  geom_point(aes(size=groupSize), fill="darkgray") +
  # respective sd interval
  geom_errorbar(aes( ymin=educationGoalMean-educationGoalSd,
                     ymax=educationGoalMean+educationGoalSd),
                width=0.2 )
```
Now it would be time to beautify, which I will drop for now ..
