---
title: "Solution for Arlene's project"
author: "Sarah Löber"
date: "2024-06-13"
output: md_document
---

### Solution for Arlene's project

Below is my solution for Arlene's project. It is not quite finished yet, at least the plot looks unfinished. I had trouble bringing it in the same form as the example plot and eventually just gave up. So if there's an easier solution, I'd love to hear it :)

```{r load data, message = FALSE, warning = FALSE} 
# libraries
library(tidyverse)

# load the data
load("~/Data Projects with R and GitHub/Projects/ArleneHohl/data_PISA.Rdata")
head(data)

# select only columns needed
clean_data <- data %>%
  select(CNT, CNTSTUID, ST005Q01JA, ST007Q01JA, starts_with("ST327")) 
head(clean_data)
```

## Data manipulation: Build parents' highest level of education for each person

```{r data cleaning, message = FALSE, warning = FALSE} 
# remove NAs 
clean_data <- clean_data %>%
  drop_na(ST005Q01JA, ST007Q01JA)

# reverse the scores by getting the max value of the column (+1) and subtracting the actual value

clean_data <- clean_data %>%
 mutate(
   ST005Q01JA = max(ST005Q01JA) +1 -ST005Q01JA,
   ST007Q01JA = max(ST007Q01JA) +1 -ST007Q01JA
 )

head(clean_data)
```
```{r socioeconomic status, message = FALSE, warning = FALSE}
# store the sum of the two columns in a new column called ED_PARENTS
clean_data <- clean_data %>%
  mutate(ED_PARENTS = ST005Q01JA + ST007Q01JA) %>%
  drop_na(ED_PARENTS)

head(clean_data)
```

## Data manipulation: Expected education qualification

```{r qualification, message = FALSE, warning = FALSE}
# find the expected education qualification of the student with a for-loop
education_columns <- c("ST327Q01JA", "ST327Q02JA", "ST327Q03JA", "ST327Q04JA", "ST327Q05JA", "ST327Q06JA", "ST327Q07JA", "ST327Q08JA")

exp_level <- function(row) {
  education_levels <- c("ISCED_2", "ISCED_3_3", "ISCED_3_4", "ISCED_4", "ISCED_5", "ISCED_6", "ISCED_7", "ISCED_8")
  for (i in length(row):1) {
    if (!is.na(row[i]) && row[i] == 1) {
      return(education_levels[i])
    }
  }
  return(NA) 
}
    
clean_data$EXP_ED <- apply(clean_data[, education_columns], 1, exp_level)

# drop the NA values and only keep relevant columns
clean_data <- clean_data[!is.na(clean_data$EXP_ED), ]
clean_data %>%
  select(CNT, CNTSTUID, EXP_ED, ED_PARENTS) 

head(clean_data)

# also map numerical values to the expected education

isced_mapping <- c(
  "ISCED_2" = 1,
  "ISCED_3_3" = 2,
  "ISCED_3_4" = 3,
  "ISCED_4" = 4,
  "ISCED_5" = 5,
  "ISCED_6" = 6,
  "ISCED_7" = 7,
  "ISCED_8" = 8
)

clean_data <- clean_data %>%
  mutate(EXP_ED_NUM = isced_mapping[EXP_ED],
         ED_PARENTS = as.numeric(ED_PARENTS))

clean_data <- clean_data %>%
  filter(!is.na(EXP_ED_NUM), !is.na(ED_PARENTS))
```

# Plot the data

This plot is far away from perfect and it took me a while to even get it into that shape (and obviously a lot of ChatGPT). What is missing are the error bars as well as the "median
line" that is shown in the original plot. Furthermore, I have commented out the scale_y_discrete part, it just removes the labels when I leave it in. 
Also interesting is the fact that the bubbles for low socioeconomic status (or parent education level) for Germany and Estonia are so high on the 2 and 4-level. Probably a mistake from my side... 

```{r df for plot, message = FALSE, warning = FALSE}

#get data ready for plotting, add mean of numeric column. Also get number of students for the bubble plot later
sum_data <- clean_data %>%
  group_by(ED_PARENTS, CNT) %>%
  summarise(
    global_avg_exp_ed = mean(EXP_ED_NUM, na.rm = TRUE),
    mean_exp_ed = mean(EXP_ED_NUM, na.rm = TRUE),
    sd_exp_ed = sd(EXP_ED_NUM, na.rm = TRUE),
    no_students = n()
  ) %>%
  ungroup()
```

```{r plot, message=FALSE, warning=FALSE}
# Plot the data
country_labels <- c("DEU" = "Germany", "EST" = "Estonia", "JPN" = "Japan")

ggplot(sum_data, aes(x = factor(ED_PARENTS), y = mean_exp_ed, color = CNT)) +
  scale_color_brewer(palette = "Accent") +
  geom_point(aes(size = no_students), alpha = 0.8) + 
  geom_hline(aes(yintercept=first(global_avg_exp_ed)), linetype = "dashed", color = "grey") +
  geom_errorbar(aes(ymin = mean_exp_ed - sd_exp_ed, ymax = mean_exp_ed + sd_exp_ed), width = 0.2) +
  scale_size(range = c(1, 15)) +
  labs(
    title = "Expected Educational Achievement by Parents' Socioeconomic Status",
    x = "Parents' Socioeconomic Status",
    y = "Expected Educational Achievement"
  ) +
  theme_minimal() +
  theme(legend.position = "none",
        strip.text = element_text(size = 10),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(cols=vars(CNT), scales = "free_x", space = "free_x", labeller = labeller(CNT = country_labels)) +
  scale_x_discrete(labels = function(labels) {
    c("Bottom 20%", rep("", length(labels) - 2), "Top 20%")
  }) +
  scale_y_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7, 8), labels = c("ISCED2", "ISCED3.3", "ISCED3.4", "ISCED4", "ISCED5", "ISCED6", "ISCED7", "ISCED8"))

```
