---
title: "Solution for Arlene's project"
author: "Sarah Löber"
date: "2024-06-13"
output: md_document
---

### Solution for Arlene's project

Below is my solution for Arlene's project. It is not quite finished yet, at least the plot looks unfinished. I had trouble bringing it in the same form as the example plot and eventually just gave up. So if there's an easier solution, I'd love to hear it :)

```{r message = FALSE, warning = FALSE} 
# libraries
library(tidyverse)
library(dbplyr)
library(ggplot2)

# load the data
load("~/Data Projects with R and GitHub/Projects/ArleneHohl/data_PISA.Rdata")
head(data)

# select only columns needed
clean_data <- data %>%
  select(CNT, CNTSTUID, ST005Q01JA, ST007Q01JA, starts_with("ST327")) 
head(clean_data)
```

## Data manipulation: Build parents' highest level of education for each person

```{r message = FALSE, warning = FALSE} 
# remove NAs 
clean_data <- clean_data %>%
  drop_na(ST005Q01JA, ST007Q01JA)

# reverse the scores by getting the max value of the column (+1) and subtracting the actual value

clean_data <- clean_data %>%
 mutate(
   ST005Q01JA = max(ST005Q01JA) +1 -ST005Q01JA,
   ST007Q01JA = max(ST007Q01JA) +1 -ST007Q01JA
 )

head(clean_data)
````
```{r message = FALSE, warning = FALSE}
# store the sum of the two columns in a new column called ED_PARENTS
clean_data <- clean_data %>%
  mutate(ED_PARENTS = ST005Q01JA + ST007Q01JA) %>%
  drop_na(ED_PARENTS)

head(clean_data)
````

## Data manipulation: Expected education qualification

```{r message = FALSE, warning = FALSE}
# find the expected education qualification of the student with a for-loop
education_columns <- c("ST327Q01JA", "ST327Q02JA", "ST327Q03JA", "ST327Q04JA", "ST327Q05JA", "ST327Q06JA", "ST327Q07JA", "ST327Q08JA")

exp_level <- function(row) {
  education_levels <- c("ISCED_2", "ISCED_3_3", "ISCED_3_4", "ISCED_4", "ISCED_5", "ISCED_6", "ISCED_7", "ISCED_8")
  for (i in length(row):1) {
    if (!is.na(row[i]) && row[i] == 1) {
      return(education_levels[i])
    }
  }
  return(NA) 
}
    
clean_data$EXP_ED <- apply(clean_data[, education_columns], 1, exp_level)

# drop the NA values and only keep relevant columns
clean_data <- clean_data[!is.na(clean_data$EXP_ED), ]
clean_data %>%
  select(CNT, CNTSTUID, EXP_ED, ED_PARENTS) 

head(clean_data)

# also map numerical values to the expected education

isced_mapping <- c(
  "ISCED_2" = 1,
  "ISCED_3_3" = 2,
  "ISCED_3_4" = 3,
  "ISCED_4" = 4,
  "ISCED_5" = 5,
  "ISCED_6" = 6,
  "ISCED_7" = 7,
  "ISCED_8" = 8
)

clean_data <- clean_data %>%
  mutate(EXP_ED_NUM = isced_mapping[EXP_ED],
         ED_PARENTS = as.numeric(ED_PARENTS))

clean_data <- clean_data %>%
  filter(!is.na(EXP_ED_NUM), !is.na(ED_PARENTS))
````

# Plot the data

This plot is far away from perfect and it took me a while to even get it into that shape (and obviously a lot of ChatGPT). What is missing are the error bars as well as the "median
line" that is shown in the original plot. Furthermore, I have commented out the scale_y_discrete part, it just removes the labels when I leave it in. 
The geom_line needs to be revised to being just one straight line, but I did not have time to find out how that would work. 
Also interesting is the fact that the bubbles for low socioeconomic status (or parent education level) for Germany and Estonia are so high on the 2 and 4-level. Probably a mistake from my side... 

```{r message = FALSE, warning = FALSE}

#get data ready for plotting, add mean of numeric column. Also get number of students for the bubble plot later
sum_data <- clean_data %>%
  group_by(ED_PARENTS, CNT) %>%
  summarise(
    mean_exp_ed = mean(EXP_ED_NUM, na.rm = TRUE),
    no_students = n()
  ) %>%
  ungroup()

# Calculating the global average for the straight trend line later
global_avg <- clean_data %>%
  group_by(ED_PARENTS) %>%
  summarise(global_avg_exp_ed = mean(EXP_ED_NUM, na.rm = TRUE))

plot_data <- sum_data %>%
  left_join(global_avg, by = "ED_PARENTS")

# Plot the data
plot <- ggplot(plot_data, aes(x = factor(ED_PARENTS), y = mean_exp_ed, color = CNT)) +
  geom_point(aes(size = no_students), alpha = 0.8) + 
  #geom_line(aes(y = global_avg_exp_ed, group = 1), linetype = "dashed", color = "grey") +
  scale_size(range = c(1, 15)) +
  labs(
    title = "Expected Educational Achievement by Parents' Socioeconomic Status",
    x = "Parents' Socioeconomic Status",
    y = "Expected Educational Achievement"
  ) +
  theme_minimal() +
  theme(legend.position = "none") +
  facet_wrap(vars(CNT)) +
  scale_x_discrete(breaks = c("2", "4", "5", "6", "7", "8", "9", "10"),
                   labels = c("Bottom 20%", "", "", "", "", "", "Top 20%", ""))
  #scale_y_discrete(breaks = c("4", "5", "6", "7"),
                   #labels = c("ISCED4", "ISCED5", "ISCED6", "ISCED7"))

print(plot)

````
