---
title: "Project solution for Johanna's project"
author: "E. Buecklein"
date: "`r Sys.Date()`"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(naniar)
library(ggplot2)
library(magrittr)
library(viridis) # for colors
```

# Read in and clean data
```{r}
data <- read.csv("Projects/johannaroever/StelaeCyprus.csv", 
                 sep = ";", 
                 header = TRUE,
                 na = c("unknown", "undefined"))
```

## Data modification
2. In `epoch` the value _antoninisch_ should be replaced by _antonine_ for using a consistent language.
```{r}
data %<>%
  mutate(epoch = str_replace(epoch, "antoninisch", "antonine"))
```

3. In the column `type`some values still include a "?" which should be deleted, since the information it was meant to convey is written in `typeCertain`.
```{r}
data %<>%
  mutate(type = str_replace_all(type, "\\?", ""))

```

4. For the stelae with the `catNo` =  _98_ and _99_ the value of `type` has to be changed to _6_.
```{r}
data %<>%
  mutate(type = if_else(catNo == 98 | catNo == 99, "6", type))
```


5. The values in `typeCertain` are to be replaced for easier handling in R: _0_ to _FALSE_, _1_ to _TRUE_.
```{r}
data %<>%
  mutate(typeCertain = (typeCertain == 0))
```

6. The column `location` has to be reviewed for only containing the values of the general regions, while more specific information can be deleted: 
    * _Amathus, Golgoi, Idalion, Kition, Marion/Paphos, Salamis, Soloi, Tamassos_
    * (1) All _unknown_ in this column can be converted to _Cyprus_. - CHECK
    * (2) All values which include "(?)" can be converted to _Cyprus_. - CHECK
    * (5) If the current value already includes one of the region names, shorten it to only that. This applies to values which include  "near", "bei"
    * (3) The value of the observation of `catNo`=_12_ which says "probably Idalion" should be changed to _Cyprus_.
    * (4) Other:
        * region of Limassol --> Amathus
        * Limassol --> Amathus
        * Mathikoloni --> Amathus  
        * Athienou --> Golgoi
        * Melousha --> Golgoi
        * Pergamon --> Golgoi  
        * Alambra (Larnaca) --> Kition
        * Pano Arodes --> Marion
        * Polis --> Marion  
        * Kotschines (neighbourhood of Lysi) --> Salamis
        * north of Lysi, district Famagusta --> Salamis    
        * Ambelia, near Morphou --> Soloi  
        * Pera (Asproji) --> Tamassos
```{r}
data %<>%
  mutate(location = if_else(location == "unknown", "Cyprus", location)) %>% # (1)
  mutate(location = if_else(str_detect(location, "\\?"), "Cyprus", location)) %>% # (2)
  mutate(location = if_else(location == "probably Idalion", "Cyprus", location)) %>% # (3)
  mutate(location = if_else(location == "region of Limassol" | location == "Limassol" | location == "Mathikoloni", "Amathus", location)) %>% # (4 ff.)
  mutate(location = if_else(location == "Athienou" | location == "Melousha" | location == "Pergamon", "Golgoi", location)) %>%
  mutate(location == if_else(location == "Alambra (Larnaca)", "Kition", location)) %>%
  mutate(location == if_else(location == "Pano Arodes" | location == "Polis", "Marion", location)) %>%
  mutate(location == if_else(location == "Kotschines (neighbourhood of Lysi)" | location == "north of Lysi, district Famagusta", "Salamis", location)) %>%
  mutate(location == if_else(location == "Ambelia, near Morphou", "Soloi", location)) %>%
  mutate(location == if_else(location == "Pera (Asproji", "Tamassos", location))

data %<>% # (5 ff.)
  mutate(location = case_when(str_detect(location, "Golgoi") ~ "Golgoi",
                              str_detect(location, "Tamassos") ~ "Tamassos",
                              str_detect(location, "Marion") ~ "Marion",
                              str_detect(location, "Amathus") ~ "Amathus",
                              str_detect(location, "Idalion") ~ "Idalion",
                              str_detect(location, "Salamis") ~ "Salamis",
                              str_detect(location, "Salamiu") ~ "Salamis")
         )
```

even after this, there are some regions that are single, or I am not sure where to put them - is this correct? Or do they belong to another region?
        
7. For easier handling of `material` the values should just be reduced to either _limestone_ or _marble_.
```{r}
data %<>%
  mutate(material = if_else(str_detect(material, "limestone"), "limestone", material)) %>%
  mutate(material = if_else(str_detect(material, "marble"), "marble", material))
```

# Visualization

## 1. Stacked bar plot epoch - type

### Table to show amount of type per epoch
```{r}
data$epoch <- factor(data$epoch, ordered = TRUE, 
                        levels = c("cypro-archaic", "cypro-classical", "hellenistic",
                                   "julian-claudian", "early_imperial", "flavian",
                                   "antonine"))
data$type <- factor(data$type)
data$epoch <- factor(data$epoch)
levels(data$epoch)[6] <- "hellenistic-roman"

knitr::kable(table(data$type, data$epoch))
```

### Stacked bar plot to show the correlation `epoch - type`:

- How can you show a correlation with a bar plot?

```{r}
typeData <- 
  distinct(data, type,typeDesc) %>% 
  replace_na(list(type = "NA", typeDesc = "unknown")) %>%
  arrange(type)

data %<>% 
  mutate( type = factor(type, levels = typeData$type, labels = typeData$typeDesc))

# make a color palette to be used
my_colors <- viridis_pal()(13)
my_colors[1] <- "#750786FF"
my_colors[2] <- "#6D378CFF"

ggplot(data, aes(x = epoch, fill = type)) +
  geom_bar() +
  labs(title = "Stacked Barplot",
       x = "epoch",
       fill = "type") +
  geom_text(stat = "count", aes(label = after_stat(count)), position = position_stack(vjust = 0.6), size = 2) +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = my_colors)
```

## 2. Stacked barplot type - material

### Table to show amounts of each material per type

```{r}
knitr::kable(table(data$type, data$material))
```

### Stacked bar plot to show the correlation type - material

Also here: how can you see a correlation with a barplot?

```{r}
ggplot(data, aes(x = type, fill = material)) +
  geom_bar() +
  labs(title = "Stacked Barplot",
       x = "type",
       fill = "material") +
  geom_text(stat = "count", aes(label = after_stat(count)), position = position_stack(vjust = 0.6), size = 2) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

- do you really want the type on the x-axis? It looks a bit weird... would look nicer if material was on the x axis I think!