---
title: "Fomula One Solution"
author: "Nicolai Schwarz"
date: "2023-12-06"
output: html_document
runtime: shiny
---



```{r setup, include=FALSE}

# get this packages

source("install_packages.R")

# To https://github.com/Dr-Eberle-Zentrum/Data-projects-with-R-and-GitHub.git
#   71850a3..a54e60c  HEAD -> Nicos_tryOut_branch
# :) :
knitr::opts_chunk$set(echo = FALSE)
```
# This is my **Formula One Solution**

### Dear Jana,

after i first took a closer look in your provided `archive.zip`and sorted out the `.csv` tables of intrest.

My first thoughts been its just some selecting/deselecting and joining and go on with the visualization part.

Honestly i even tried to cheat and prepared the `.csv`s already with **OpenRefine** `r emo::ji("grin")` [Download Link](https://github.com/OpenRefine/OpenRefine/releases/download/3.7.6/openrefine-win-with-java-3.7.6.zip)

But by accident i imported the wrong `.csv` dats and already after trying to im port the second `csv` from your provided  [archive.zip](https://github.com/Dr-Eberle-Zentrum/Data-projects-with-R-and-GitHub/blob/main/Projects/JanaKimmich/archive.zip) folder, i came to this wunderful idea of creating this [**drop_n_slice_folder**](https://github.com/Dr-Eberle-Zentrum/Data-projects-with-R-and-GitHub/tree/Nicos_tryOut_branch/Jana_Formula_One/drop_n_slice_folder) for you ! `r emo::ji("grin")`


I  wasn't actuall sure if you wanna extend this project anytime to the other `.csv` dats aswell, but i should work with some minor adjustments.
  
          I didn't fully wrapped my head around already, but even mightbe you could recycle some snippets from, for the:
[**SOP-Issue**](https://github.com/Dr-Eberle-Zentrum/Data-projects-with-R-and-GitHub/blob/main/Projects/blubbi90/StandardOperatingProcedureLayout/project_description_v0.1.md)

***

I found it prettxy intresting and useful to see the **Top-10** Winners of diffrent years as well, so i used an [Interactive RMD Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).
With the ***shiny package***',  you can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.

(*If you wanna learn more about, click the* [*Link*](http://rmarkdown.rstudio.com/authoring_shiny.html))


***

So let us come back to this fantastic [**drop_n_slice_folder**](https://github.com/Dr-Eberle-Zentrum/Data-projects-with-R-and-GitHub/tree/Nicos_tryOut_branch/Jana_Formula_One/drop_n_slice_folder)  `r emo::ji("folder")` and let's have an closer look how it actuall works and helps to solve [your **Formula One Issue:**](https://github.com/Dr-Eberle-Zentrum/Data-projects-with-R-and-GitHub/blob/main/Projects/JanaKimmich/project-description.md)

***

#  THE **Drop_n_slice_folder**

***
So let us take a look inside the folder:

***

```{r folder_content, echo=FALSE}

ui <- fluidPage(
  titlePanel("Here we can take a look into the Solution folder"),
  sidebarLayout(
    sidebarPanel(
      shinyFilesButton('files', 'Choose a file', 'Upload', FALSE)
    ),
    mainPanel(
      tableOutput('filetable')
    )
  )
)


server <- function(input, output, session) {
  shinyFileChoose(input, 'files', roots = c(wd = '.'), filetypes = c('', 'txt', 'csv'))
  
  
  output$filetable <- renderTable({
    if (!is.null(input$files$datapath)) {
      read.csv(input$files$datapath, header = TRUE)
    }
  })
}


shinyApp(ui, server)




```

## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:



Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.



