---
title: 'Second project solution - Nicolas´ project'
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(fmsb)
library(readxl)
library(tidyverse)
```
## General project topic

In this project, you are tasked with generating a spider-web diagram on the relation between social-media applications use and the personality profiles of more than 3000 users. We want to focus on the data regarding “young adults” (18 to 34) in particular. The social media platforms included are Whatsapp, Facebook and Instagram, and the personality profiles are based on the “Big-5” trait-model of personality, which measures Openness, Conscientiousness, Extraversion, Agreeableness and Neuroticism by means of a standardized questionnaire.

## The Data Set (processed)
```{r include=FALSE}
# Define the URL of the raw csv file
url <- "https://www.frontiersin.org/api/v3/articles/527189/file/Data_Sheet_2.xlsx/527189_supplementary-materials_datasheets_2_xlsx/1"

# Download and import excel file 
download.file(url, destfile = "Data_Sheet_2.xlsx", mode = "wb")
data_raw <- read_excel("Data_Sheet_2.xlsx")

# Remove file after data import
file.remove("Data_Sheet_2.xlsx")

# Data processing
data_pro <- data_raw %>%
  # Select required/interesting columns from df data_raw
  select(all_of(c("Age", "has_whatsapp", "has_facebook", "has_instagram", "BFI_Extra", "BFI_Agree", "BFI_Consc", "BFI_Neuro", "BFI_Open"))) %>%
  filter(Age %in% c(18:34)) %>%
  # Add multiple and none App users
  mutate(
    # has whatsapp and facebook
    `has_WF` = has_whatsapp & has_facebook,
    # has facebook and instagram
    `has_FI` = has_facebook & has_instagram,
    # has whatsapp and instagram
    `has_WI` = has_whatsapp & has_instagram,
    # has all
    `has_WFI` = has_whatsapp & has_facebook & has_instagram,
    # has none
    `has_none` = !(has_whatsapp | has_facebook | has_instagram),
    # add case number
    case_num = case_when(
      has_none == 1 ~ 8,
      has_WFI == 1 ~ 7,
      has_WF == 1 & has_instagram == 0 ~ 6,
      has_FI == 1 & has_whatsapp == 0 ~ 5,
      has_WI == 1 & has_facebook == 0 ~ 4,
      has_whatsapp == 1 & has_instagram == 0 & has_facebook == 0 ~3,
      has_instagram == 1 & has_whatsapp == 0 & has_facebook == 0 ~2,
      has_facebook == 1 & has_instagram == 0 & has_whatsapp == 0 ~1,
      TRUE ~ NA_real_
      )
    )
```
Find the filtered data frame below:
```{r}
# Preview of the data
head(data_pro)

# Calculate average personality trait values per case
data_means <- data_pro %>%
  group_by(case_num) %>%
  summarise(across(
    starts_with("BFI_"),
    ~ mean(.x, na.rm = TRUE)
  )) %>%
  rename_with(~ c("Extraversion", "Agreeableness", "Conscientiousness", "Neuroticism", "Openess"), starts_with("BFI_"))

#head(data_means)
# str(data_means)

# Add respective min and max values
radar_minmax <- data.frame(
  case_num = c("min", "max"),
  Extraversion = c(min(data_pro$BFI_Extra), max(data_pro$BFI_Extra)), # c(mean(data_means$Extraversion) - 1, mean(data_means$Extraversion) + 1)
  Agreeableness = c(min(data_pro$BFI_Agree), max(data_pro$BFI_Agree)),
  Conscientiousness = c(min(data_pro$BFI_Consc), max(data_pro$BFI_Consc)),
  Neuroticism = c(min(data_pro$BFI_Neuro), max(data_pro$BFI_Neuro)),
  Openess = c(min(data_pro$BFI_Open), max(data_pro$BFI_Open))
  )

# Bind rows and remove first column
data_radar <- rbind(radar_minmax, data_means)

#head(data_radar)

# Show df containing all data for plotting
knitr::kable(data_radar, caption = "Here you can find the processed data used for plotting:", format = "markdown", align = "c") 
```

## Visualisation

```{r visualisation, fig.width = 17.5, fig.height = 15}
# Define plot colors
colors_border <- rainbow(nrow(data_means))
colors_in <- adjustcolor(colors_border, alpha.f = 0.2)


# Define legend labels
leglabs <- c("FB", "Insta", "WA", "WA + Insta", "FB + Insta", "WA + FB", "WA + FB + Insta", "None")

# Create relative data for radar chart
data_radar_rel <- data_radar %>%
  mutate(across(-1, ~ . / max(.))) %>%
  select(-1)

# Show relative data
#head(data_radar_rel)

# Create spider web diagram (somehow use relative values here. U need to adjust the data properly)
# Sorry, I was not able to scale it properly. If you have any ideas on how to do this, let me know and I will implement it.
radarchart(
  data_radar_rel,
  axistype = 1,
  pcol = colors_border,
  pfcol = colors_in,
  plwd = 1.5,
  plty = 1,
  cglcol = "grey",
  cglty = 1,
  caxislabels = c(0, 0.25, 0.5, 0.75, 1),

  #caxislabels = c(0, 0.2, 0.4, 0.6, 0.7),
  calcex = 1.75,
  axislabcol = "black",
  cglwd = 3,
  vlcex = 2,
  title = "Big Five Personality Profiles by Social Media Platforms",
  cex.main = 2.5
)

# Add legend
legend(
  x = "topright",
  legend = leglabs,
  bty = "n",
  pch = 20,
  col = colors_border,
  text.col = "black",
  cex = 2,
  pt.cex = 2
)


# Add labs to data_mean
data_means$labs <- leglabs

# Crate long table format
data_long <- data_means %>%
  select(-case_num) %>%
  mutate(ID = row_number()) %>%
  gather(key = "Personality Trait", value = "Wert", -ID, -labs)

#head(data_long)

# Add barchart
ggplot(data_long, aes(x = labs, y = Wert, fill = `Personality Trait`)) +
  geom_bar(stat = "identity") +
  #theme_minimal() +
  labs(title = "Big Five Personality Profiles  by Social Media Platforms",
       x = "Social Media Platform",
       y = "Absolute values of Personality Traits") +
  theme(
    panel.background = element_rect(fill = "white", colour = "black"),
    panel.grid.major = element_line(colour = "black"),
    panel.grid.minor.y = element_line(colour = "black"),
    plot.title = element_text(size = 30, face = "bold"),         # Diagrammtitel
    axis.title.x = element_text(size = 25),                      # Titel der x-Achse
    axis.title.y = element_text(size = 25),                      # Titel der y-Achse
    axis.text.x = element_text(size = 20, angle = 45, hjust = 1, colour = "black"),# Beschriftung der x-Achse
    axis.text.y = element_text(size = 20, colour = "black"),                       # Beschriftung der y-Achse
    legend.title = element_text(size = 25),                      # Legendentitel
    legend.text = element_text(size = 20)                        # Legendentext
  )


```